---
title: 'Processing the rbcL'
author: "B. Karina Montero and Mark Gillingham"
date: "`r Sys.Date()`"
#output: pdf_document
output: html_document
---

## Loading libraries
```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
rm(list = ls())

# basics
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(ggpubr)

# phyloseq section
library(phyloseq)
library(decontam)

```

## WORKING DIRECTORIES
```{r directories, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
```


## I. LOAD & CHECK DATA

### 1. metadata
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
fringilla <- import_qiime_sample_data(mapfilename = paste0(dataDir, "/metadata_fringilla_diet_03.06.23.csv"))

```

### 2. Directory containing the fastq files
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
COI_path <- paste0(metabarcodingDir, "COI_raw_data") 
list.files(COI_path)
COI_files <- sort(list.files(COI_path, pattern="COI", full.names = TRUE))
```

### 3. Inspect read quality profiles

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
plotQualityProfile(COI_files[c(45,36,33,38,8,46,10,18,17)])
```
## II. REMOVE PRIMERS


Set patterns to discriminate your forward and reverse read files
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
file_pattern <- c("F" = "_R1_001.fastq", "R" = "_R2_001.fastq")
fas_Fs_raw <- sort(list.files(COI_path, pattern = file_pattern["F"], full.names = TRUE))
fas_Rs_raw <- sort(list.files(COI_path, pattern = file_pattern["R"], full.names = TRUE))
```

Primer sequences

For arthropoda library preparation, a fragment of the mitochondrial gene COI of around 178 (bp) was amplified using the following primers:

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
FWD_COI <- "ATTCAACCAATCATAAAGATATTGG" # LepF1 (Hebert et al. 2004).
REV_COI <- "WACTAATCAATTWCCAAAHCCHCC" #  ZBJ-ArtR2c-deg (modified by Shutt et al. (2020) from Zeale et al. (2011)).
```

To ensure we have the right primers, and the correct orientation of the primers on the reads, we will verify the presence and orientation of these primers in the data.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
allOrients <- function(primer) {
    # Create all orientations of the input sequence
    require(Biostrings)
    dna <- DNAString(primer)  # The Biostrings works w/ DNAString objects rather than character vectors
    orients <- c(Forward = dna, Complement = complement(dna), Reverse = reverse(dna), 
        RevComp = reverseComplement(dna))
    return(sapply(orients, toString))  # Convert back to character vector
}

FWD_orients_COI <- allOrients(FWD_COI)
REV_orients_COI <- allOrients(REV_COI)

```

Count the number of times the primers appear in the forward and reverse read, while considering all possible primer orientations.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
primerHits <- function(primer, fn) {
    # Counts number of reads in which the primer is found
    nhits <- vcountPattern(primer, sread(readFastq(fn)), fixed = FALSE)
    return(sum(nhits > 0))
}

primer_count <- rbind(FWD.ForwardReads_COI = sapply(FWD_orients_COI, primerHits, fn = fas_Fs_raw[[1]]), 
    FWD.ReverseReads_COI = sapply(FWD_orients_COI, primerHits, fn = fas_Rs_raw[[1]]), 
    REV.ForwardReads_COI = sapply(REV_orients_COI, primerHits, fn = fas_Fs_raw[[1]]), 
    REV.ReverseReads_COI = sapply(REV_orients_COI, primerHits, fn = fas_Rs_raw[[1]]))
primer_count
```
As expected, the FWD primer is found in the forward reads in its forward orientation, and in some of the reverse reads in its reverse-complement orientation (due to read-through when the COI region is short). Similarly the REV primer is found with its expected orientations.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
FWD <- c("ATTCAACCAATCATAAAGATATTGG")
REV <- c("WACTAATCAATTWCCAAAHCCHCC")

FWD_RC <- dada2:::rc(FWD)
REV_RC <- dada2:::rc(REV)

path_cut <- file.path(COI_path, "cutadapt")
if(!dir.exists(path_cut)) dir.create(path_cut)


#path_cut <- paste0(barcodingDir, "/cutadapt")

fas_Fs_cut <- file.path(path_cut, basename(fas_Fs_raw))
fas_Rs_cut <- file.path(path_cut, basename(fas_Rs_raw))

R1_flags <- paste(paste("-g", FWD, collapse = " "), paste("-a", REV_RC, collapse = " "))
R2_flags <- paste(paste("-G", REV, collapse = " "), paste("-A", FWD_RC, collapse = " "))

cutadapt <- "/home/fringilla/miniconda3/envs/cutadaptenv/bin/cutadapt" # Path to the executable

for(i in seq_along(fas_Fs_raw)) {
  cat("Processing", "-----------", i, "/", length(fas_Fs_raw), "-----------\n")
  system2(cutadapt, args = c(R1_flags, R2_flags,
                             "-m 50",
                             "-M 200",
                             "--discard-untrimmed",
                             "--max-n 0",
                             "--cores 0", # Use 0 to auto-detect.
                             "-n 2", # required to remove FWD and REV from read
                             # Optional strong constraint on expected length
                             #paste0("-m ", 250-nchar(FWD)[1], ":", 250-nchar(REV)[1]), 
                             #paste0("-M ", 250-nchar(FWD)[1], ":", 250-nchar(REV)[1]), 
                             "-o", fas_Fs_cut[i], "-p", fas_Rs_cut[i],
                             fas_Fs_raw[i], fas_Rs_raw[i]))
}


primer_count <- rbind(FWD.ForwardReads_COI = sapply(FWD_orients_COI, primerHits, fn = fas_Fs_cut[[1]]), 
    FWD.ReverseReads_COI = sapply(FWD_orients_COI, primerHits, fn = fas_Rs_cut[[1]]), 
    REV.ForwardReads_COI = sapply(REV_orients_COI, primerHits, fn = fas_Fs_cut[[1]]), 
    REV.ReverseReads_COI = sapply(REV_orients_COI, primerHits, fn = fas_Rs_cut[[1]]))
primer_count

```

## III. FILTER AND TRIM

Remove BPCR5_COI empty file

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
path_cut <- paste0(metabarcodingDir, "COI_raw_data/cutadapt") # If you skipped primers removal, provide the path to your sequences here
list.files(path_cut)

fas_Fs_cut <- sort(list.files(path_cut, pattern = file_pattern["F"], full.names = TRUE))
fas_Rs_cut <- sort(list.files(path_cut, pattern = file_pattern["R"], full.names = TRUE))

sample_names <- sapply(strsplit(basename(fas_Fs_cut), "_"), function(x) x[1])

COI_cut <- sort(list.files(path_cut, pattern="COI", full.names = TRUE))
```

Inspect read quality profiles

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

plotQualityProfile(COI_cut[c(45,36,33,38,8,46,10,18,17)])
#ggsave(paste(outputDir, "/02_DietMetabarcoding/01_Exploration/01_qualityProfile/COI_cut_plotQualityProfile.png", sep = "/"))

fas_Fs_filtered <- file.path(COI_path, "filtered", basename(fas_Fs_cut))
fas_Rs_filtered <- file.path(COI_path, "filtered", basename(fas_Rs_cut))
all.equal(basename(fas_Fs_raw), basename(fas_Fs_filtered))

names(fas_Fs_filtered) <- sample_names
names(fas_Rs_filtered) <- sample_names

```

Standard filtering paraments: maxN=0 (DADA2 requires sequences contain no Ns), truncQ = 2, rm.phix = TRUE and maxEE=2. The maxEE parameter sets the maximum number of “expected errors” allowed in a read, which is a better filter than simply averaging quality scores.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
filter_trim <- filterAndTrim(fas_Fs_cut, fas_Fs_filtered, fas_Rs_cut, fas_Rs_filtered,
                       #truncLen = c(170, 100), 
                       minLen = 175,
                       maxLen = 184, 
                       maxN = 0, maxEE = c(2, 2), truncQ = 2,
                       rm.phix = TRUE, multithread = TRUE, verbose = FALSE)
head(filter_trim)
```
The filter removed all reads: /home/fringilla/Documents/AllGenetics/COI_raw_data/filtered/BPCR5_COI_S1_L001_R1_001.fastq.gz and /home/fringilla/Documents/AllGenetics/COI_raw_data/filtered/BPCR5_COI_S1_L001_R2_001.fastq.gz not written.
Some input samples had no reads pass the filter.

I delete these files manually.

## IV. DADA 2

Learn error rates
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
error_F <- learnErrors(fas_Fs_filtered, multithread = TRUE, randomize = TRUE)
#101903502 total bases in 565302 reads from 112 samples will be used for learning the error rates.
saveRDS(error_F, paste0(outputDir, "/02_DietMetabarcoding/02_dada2/COI_error_F.rds"))
#err <- readRDS(paste0(outputDir, "/02_Dada2/err.rds"))

plotErrors(error_F, nominalQ=TRUE)
ggsave(paste(outputDir, "/02_DietMetabarcoding/02_dada2/plotErrors_COI_F.png", sep = "/"))

error_R <- learnErrors(fas_Rs_filtered, multithread = TRUE, randomize = TRUE)
#100633957 total bases in 560007 reads from 137 samples will be used for learning the error rates.
saveRDS(error_R, paste0(outputDir, "/02_DietMetabarcoding/02_dada2/COI_error_R.rds"))
#err <- readRDS(paste0(outputDir, "/02_Dada2/err.rds"))

plotErrors(error_R, nominalQ=TRUE)
ggsave(paste(outputDir, "/02_DietMetabarcoding/02_dada2/plotErrors_COI_R.png", sep = "/"))
```


### DEREPLICATION, SAMPLE INFERENCE & MERGE PAIRED READS

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
merged_list <- vector("list", length(sample_names))
names(merged_list) <- sample_names

for(i in sample_names){
  cat("Processing -------", which(sample_names == i), "/", length(sample_names), "-------", i, "\n")
  derep_Fs <- derepFastq(fas_Fs_filtered[[i]], verbose = TRUE)
  derep_Rs <- derepFastq(fas_Rs_filtered[[i]], verbose = TRUE)
  dds_Fs <- dada(derep_Fs, err = error_F, multithread = TRUE, verbose = TRUE)
  dds_Rs <- dada(derep_Rs, err = error_R, multithread = TRUE, verbose = TRUE)
  merged_list[[i]] <- mergePairs(dds_Fs, derep_Fs, dds_Rs, derep_Rs, verbose = TRUE)
}
```

### CONSTRUCT SEQUENCE TABLE

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
seqtab <- makeSequenceTable(merged_list)
dim(seqtab)
table(nchar(getSequences(seqtab)))
```

### REMOVE CHIMERA

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
seqtab_nochim <- removeBimeraDenovo(seqtab, method = "consensus", multithread = TRUE, verbose = TRUE)
#Identified 232 bimeras out of 3275 input sequences.
dim(seqtab_nochim)
table(nchar(getSequences(seqtab_nochim)))
```

#### TRACK READS THROUGH THE PIPELINE

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
track <- cbind(filter_trim, rowSums(seqtab), rowSums(seqtab_nochim))
colnames(track) <- c("Cutadapt", "Filtered", "Merged", "Nonchim")
rownames(track) <- sample_names
head(track)

write.csv(track, paste0(outputDir, "/02_DietMetabarcoding/02_dada2/COI_track_reads.csv"))
```

### Export files for use in QIIME2
https://forum.qiime2.org/t/exporting-dada2-r-package-results-to-work-with-qiime2/2573/3
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
write.table(t(seqtab_nochim), paste0(outputDir, "/02_DietMetabarcoding/02_dada2/COIQIIME2seqtab.txt"), sep="\t", row.names=TRUE, col.names=NA, quote=FALSE)
uniquesToFasta(seqtab_nochim, fout=paste0(outputDir, "/02_DietMetabarcoding/02_dada2/COI-rep-seqs.fna"), ids=colnames(seqtab_nochim))

```



### TAXONOMY ASSIGNMENT (CESGA cluster)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

qiime2run qiime feature-classifier classify-sklearn \
  --i-classifier /mnt/lustre/scratch/nlsas/home/csic/dde/pla/FringillaProject/06_rbcL/NCBI_rbcL_Viridiplantae_classifier_extracted_2022_09_22_derep.qza \
  --i-reads /mnt/lustre/scratch/nlsas/home/csic/dde/pla/FringillaProject/06_rbcL/rbcL-rep-seqs.qza \
  --o-classification /mnt/lustre/scratch/nlsas/home/csic/dde/pla/FringillaProject/06_rbcL/rbcL_4_derep_restricted.qza

```

# CLEAN COI data 

The result were very similar regardless of whether the classifier was dereplicated after primer timming but taxonomic assignement was slightly higher when not dereplicating so we choose this pipeline (ps_mkCOInr_trimmed_filt)


```{r psobject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
mkCOInr_trimmed <- import_biom(BIOMfilename = paste0(outputDir, "/02_DietMetabarcoding/02_dada2/mkCOInr/COInr_2022_05_06/qiime/mkCOInr_trimmed/mkCOInr_trimmed.biom"))

rep.seqs <- Biostrings::readDNAStringSet(paste0(outputDir,"/02_DietMetabarcoding/02_dada2/PhyloseqRegional/dna-sequences.fasta"), format = "fasta")

ps_mkCOInr_trimmed <- merge_phyloseq(t(otu_table(mkCOInr_trimmed)),tax_table(mkCOInr_trimmed), fringillaCOI,rep.seqs)

taxa_names(ps_mkCOInr_trimmed) <- paste0("ASVCOI", seq(ntaxa(ps_mkCOInr_trimmed)))

tax_mkCOInr_trimmed<- data.frame(tax_table(ps_mkCOInr_trimmed))

ps_mkCOInr_trimmed

names(tax_mkCOInr_trimmed)
levels(factor(tax_mkCOInr_trimmed$Rank1))
levels(factor(tax_mkCOInr_trimmed$Rank2))


tax_mkCOInr_trimmed.clean <- data.frame(row.names = row.names(tax_mkCOInr_trimmed),Kingdom = str_replace(tax_mkCOInr_trimmed[,1], "k__",""),Phylum = str_replace(tax_mkCOInr_trimmed[,2], "p__",""),Class = str_replace(tax_mkCOInr_trimmed[,3], "c__",""),Order = str_replace(tax_mkCOInr_trimmed[,4], "o__",""),Family = str_replace(tax_mkCOInr_trimmed[,5], "f__",""),Genus = str_replace(tax_mkCOInr_trimmed[,6], "g__",""),Species = str_replace(tax_mkCOInr_trimmed[,7], "s__",""),stringsAsFactors = FALSE)

tax_mkCOInr_trimmed.clean [tax_mkCOInr_trimmed.clean  == ''] <- NA

tax_mkCOInr_trimmed.clean$Kingdom<-gsub("_.*","",tax_mkCOInr_trimmed.clean$Kingdom)
tax_mkCOInr_trimmed.clean$Phylum<-gsub("_.*","",tax_mkCOInr_trimmed.clean$Phylum)
tax_mkCOInr_trimmed.clean$Class<-gsub("_.*","",tax_mkCOInr_trimmed.clean$Class)
tax_mkCOInr_trimmed.clean$Order<-gsub("_.*","",tax_mkCOInr_trimmed.clean$Order)
tax_mkCOInr_trimmed.clean$Family<-gsub("_.*","",tax_mkCOInr_trimmed.clean$Family)
tax_mkCOInr_trimmed.clean$Genus<-gsub("_.*","",tax_mkCOInr_trimmed.clean$Genus)
tax_mkCOInr_trimmed.clean$Species<-gsub("_[[:digit:]]*$","",tax_mkCOInr_trimmed.clean$Species)
tax_mkCOInr_trimmed.clean$Species<-gsub("_-[[:digit:]]*$","",tax_mkCOInr_trimmed.clean$Species)

tax_table(ps_mkCOInr_trimmed)<-as.matrix(tax_mkCOInr_trimmed.clean)

ps_mkCOInr_trimmed

ps_mkCOInr_trimmed_filt <- ps_mkCOInr_trimmed  %>%
  subset_taxa(
    Kingdom == "Metazoa" &
    Phylum != "Chordata" &
    Phylum !=  "Tardigrada" &
    Phylum !=  "Rotifera" &
    Phylum !=  "Cnidaria" &
    Phylum != "Nemertea" &
    Phylum != "Nematoda" &
    Phylum != "Acanthocephala"
    # & Class != is.na(TRUE)
  )

ps_mkCOInr_trimmed_filt


```

## Removing contaminants using PCR controls

We remove the contaminant ASVs using the DECONTAM package:

(https://bioconductor.org/packages/release/bioc/vignettes/decontam/inst/doc/decontam_intro.html)

We first look at the library sizes (i.e. the number of reads) in each sample, as a function of whether that sample was a true positive sample or a negative control

```{r librarysizes, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE }

df <- as.data.frame(sample_data(ps_mkCOInr_trimmed_filt)) 
df$LibrarySize <- sample_sums(ps_mkCOInr_trimmed_filt)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))

ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_or_Control)) + geom_point()

```

Second we investigate the prevalence of ASVs in control samples and so called "true samples". Then we would remove identified contaminants from our dataset, however no contaminants were identified

```{r prevalenceDecontam, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

sample_data(ps_mkCOInr_trimmed_filt)$is.neg <- sample_data(ps_mkCOInr_trimmed_filt)$Sample_or_Control == "Control"

contamdf.prev <- isContaminant(ps_mkCOInr_trimmed_filt, method="prevalence", neg="is.neg")
table(contamdf.prev$contaminant)

contamdf.prev05 <- isContaminant(ps_mkCOInr_trimmed_filt, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev05$contaminant)

ps_mkCOInr_trimmed_filt.pa <- transform_sample_counts(ps_mkCOInr_trimmed_filt, function(abund) 1*(abund>0))

ps_mkCOInr_trimmed_filt.pa.neg <- prune_samples(sample_data(ps_mkCOInr_trimmed_filt.pa)$Sample_or_Control == "Control", ps_mkCOInr_trimmed_filt.pa)


ps_mkCOInr_trimmed_filt.pa.pos <- prune_samples(sample_data(ps_mkCOInr_trimmed_filt.pa)$Sample_or_Control == "Sample", ps_mkCOInr_trimmed_filt.pa)

df.pa <- data.frame(pa.pos=taxa_sums(ps_mkCOInr_trimmed_filt.pa.pos), pa.neg=taxa_sums(ps_mkCOInr_trimmed_filt.pa.neg),contaminant=contamdf.prev$contaminant)

ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")

#tax_table(ps_mkCOInr_trimmed_filt)[contamdf.prev05$contaminant,]
#ps_mkCOInr_trimmed_filt.contam <- prune_taxa(contamdf.prev05$contaminant, ps_mkCOInr_trimmed_filt)
#ps_mkCOInr_trimmed_filt

ps_mkCOInr_trimmed_filt.noncontam <- prune_taxa(!contamdf.prev05$contaminant, ps_mkCOInr_trimmed_filt)
ps_mkCOInr_trimmed_filt.noncontam

ps_mkCOInr_trimmed_filt<-ps_mkCOInr_trimmed_filt.noncontam

```

We remove the PCR control samples from the phyloseq object
```{r removeControls, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE , results="hide"}

Samples_toRemove <- c("Control")
subset_samples(ps_mkCOInr_trimmed_filt, Sample_or_Control %in% Samples_toRemove)

subset_samples(ps_mkCOInr_trimmed_filt, !(Sample_or_Control %in% Samples_toRemove))

ps_mkCOInr_trimmed_filt<-subset_samples(ps_mkCOInr_trimmed_filt, !(Sample_or_Control %in% Samples_toRemove))

```


### Are there trends in the fraction of **reads** that can be assigned to taxonomic level by experimental variables?

```{r FractionAssigned, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
readsPerSample = rowSums(otu_table(ps_mkCOInr_trimmed_filt))
fractionReadsAssigned = sapply(colnames(tax_table(ps_mkCOInr_trimmed_filt)), function(x){
  rowSums(otu_table(ps_mkCOInr_trimmed_filt)[, !is.na(tax_table(ps_mkCOInr_trimmed_filt))[,x]]) / readsPerSample
})

fractionReadsAssigned = data.frame(SampleID = rownames(fractionReadsAssigned), fractionReadsAssigned)
fractionReadsAssigned.L = pivot_longer(fractionReadsAssigned, 
    cols=colnames(tax_table(ps_mkCOInr_trimmed_filt)), names_to="taxlevel", values_to="fractionReadsAssigned")
fractionReadsAssigned.L = data.frame(fractionReadsAssigned.L, sample_data(ps_mkCOInr_trimmed_filt)[fractionReadsAssigned.L$SampleID,])

fractionReadsAssigned.L$taxlevelf = factor(fractionReadsAssigned.L$taxlevel, levels=c("Kingdom","Phylum","Class","Order","Family","Genus","Species"))



pmkCOInr_trimmed_filt<-ggplot(fractionReadsAssigned.L, aes(y = fractionReadsAssigned, x = taxlevelf)) +  theme_bw(base_size = 10) +  geom_boxplot(color='black', outlier.shape=NA) +  ggtitle("NCBI&BOLD: mkCOInr (trained with primers; 1487 ASVs)") + geom_point() + ylab ("Fraction of reads assigned") + xlab ("Taxonomic rank")



pmkCOInr_trimmed_filt
```

### Replace NAs with the highest possible rank in the taxnomic names 
```{r ReplaceNAs, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
tax_mkCOInr_trimmed_filt<- data.frame(tax_table(ps_mkCOInr_trimmed_filt))


tax_mkCOInr_trimmed_filt[is.na(tax_mkCOInr_trimmed_filt)] <- ""

for (i in 1:7){ tax_mkCOInr_trimmed_filt[,i] <- as.character(tax_mkCOInr_trimmed_filt[,i])}
tax_mkCOInr_trimmed_filt[is.na(tax_mkCOInr_trimmed_filt)] <- ""


for (i in 1:nrow(tax_mkCOInr_trimmed_filt)){if (tax_mkCOInr_trimmed_filt[i,2] == "")
  {
    kingdom <- paste("Kingdom_", tax_mkCOInr_trimmed_filt[i,1], sep = "")
    tax_mkCOInr_trimmed_filt[i, 2:7] <- kingdom
  } else if (tax_mkCOInr_trimmed_filt[i,3] == ""){
    phylum <- paste("Phylum_", tax_mkCOInr_trimmed_filt[i,2], sep = "")
    tax_mkCOInr_trimmed_filt[i, 3:7] <- phylum
  } else if (tax_mkCOInr_trimmed_filt[i,4] == ""){
    class <- paste("Class_", tax_mkCOInr_trimmed_filt[i,3], sep = "")
    tax_mkCOInr_trimmed_filt[i, 4:7] <- class
  } else if (tax_mkCOInr_trimmed_filt[i,5] == ""){
    order <- paste("Order_", tax_mkCOInr_trimmed_filt[i,4], sep = "")
    tax_mkCOInr_trimmed_filt[i, 5:7] <- order
  } else if (tax_mkCOInr_trimmed_filt[i,6] == ""){
    family <- paste("Family_", tax_mkCOInr_trimmed_filt[i,5], sep = "")
    tax_mkCOInr_trimmed_filt[i, 6:7] <- family
  } else if (tax_mkCOInr_trimmed_filt[i,7] == ""){
    tax_mkCOInr_trimmed_filt$Species[i] <- paste("Genus",tax_mkCOInr_trimmed_filt$Genus[i], sep = "_")
  }
}

tax_mkCOInr_trimmed_filt

tax_table(ps_mkCOInr_trimmed_filt)<-as.matrix(tax_mkCOInr_trimmed_filt)
```


### Barplot of taxonomy at the class level according to the geographical local location of the samples (all sample types pooled):

```{r taxonomyClass, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
(sample_data(ps_mkCOInr_trimmed_filt))$Island_Mainland

Merged<-merge_samples(ps_mkCOInr_trimmed_filt,"Island_Mainland")
warnings() 

names(sample_data(Merged))

physeq2 = phyloseq::transform_sample_counts(Merged, function(x) x / sum(x) )
physeq2


ps_melt<- psmelt(physeq2)
head(ps_melt)

p8c <- ggplot(data=ps_melt, aes(x=Sample, y=Abundance, fill=Class))
p8c <-p8c + geom_bar(aes(), stat="identity", position="stack") +scale_fill_manual(values = c("darkblue", "#8b5b5b", "darkorchid", "lightskyblue", "darkgreen", "#4a3b2c", "deeppink", "khaki2","#aa6c39", "darkorange1", "#E0FFFF", "darksalmon", "#eee7e1", "black", "darkolivegreen1","pink","darkseagreen","#E41A1C","#0000FF","#00FF00","#FFA500","#E7298A","#00FFFF","#008000","#800080","#008080","#999999","#FF4500","#FFFF00","#4DAF4A","#A65628","#F781BF","#1B9E77","#808000","#377EB8", "#CBC3E3","#460046","#AA336A","#d2691e")) + guides(fill=guide_legend(title="Taxanomic assignment (Class or higher)",ncol=2)) +theme(text = element_text(size = 10),legend.position="right",axis.text.x = element_text(angle = 45, hjust = 1),panel.background = element_blank(),plot.margin = unit(c(1,1,1,1), "cm"))+ggtitle("Relative abundance of COI in fecal sample according to host species and geography \nusing the mkCOInr pipeline (trimmed)")+
  xlab("")



(sample_data(ps_mkCOInr_trimmed_derep_filt))$Island_Mainland

```

We will use this classifier for all down stream analyses. We export ps_mkCOInr_trimmed_filt and use this phyloseq object for further downstream analyses on COI.

Update metadata Fringilla project

```{r export, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Add sequencing depth column 
sample_data(ps_mkCOInr_trimmed_filt)$Seq_depth_COI <- sample_sums(ps_mkCOInr_trimmed_filt)

# remove unnecessary columns from metadata
df <-  data.frame(sample_data(ps_mkCOInr_trimmed_filt)) %>%  select("sampleid", "Lab_Code", "Plate", "Seq_depth_rbcL", "Ring","Genus" , "Species", "Subspecies", "Species_Label", "Month", "Year", "Locality", "Habitat", "Xdec", "Ydec", "Archipelago_Mainland", "Island_Mainland",  "Area_km2", "Distance_mainland", "Age_ma", "Age", "Sex", , "Weight", "Wing", "X8_P", "Tail", "Nar", "P_Cr", "Height", "Width", "Cran", "Tars")

sample_data(ps_mkCOInr_trimmed_filt) <- df

# Replace sampleid with Lab_Code to match sample names across datasets 

sample_names(ps_mkCOInr_trimmed_filt) <- paste0("X",sample_data(ps_mkCOInr_trimmed_filt)$Lab_Code)

```
Save phyloseq object
```{r exportpartII saveObjects, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_mkCOInr_trimmed_filt
ps_mkCOInr_trimmed_filt %>% sample_data %>% names

saveRDS(ps_mkCOInr_trimmed_filt, paste0(dataDir, "/ps.clean_COI.rds"))

```

## Session Info

```{r sessioninfo, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
sessioninfo::session_info()%>%
  details::details(summary = 'Current session info', open = FALSE)
```