---
title: '16S alpha rarefaction'
author: "B. Karina Montero "
date: "`r Sys.Date()`"
output: html_document
---


## LOAD LIBRARIES

```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rm(list = ls())

# basics
library(tidyverse)
library(stringr)
library(vegan)
library(magrittr)
library(tidyverse)


# plots
library(ggplot2)
library(ggpubr)
library(fantaxtic)
library(ggtext)
library(ggh4x)
library(png)
library(cowplot)
library(grid)
library(gtable)
library(rlist)
library(ggmagnify)


#phyloseq section
library(MicEco)
library(phyloseq)
library(Biostrings)
library(microbiome)
library(microViz)
library(DECIPHER)
library(speedyseq)
library(metagMisc)
library(phylosmith)
library(seqateurs)
library(mirlyn)
library(biomeUtils)
library(metagMisc) # rarefy phyloseq

# diversity 
library(hilldiv)
library(iNEXT)
library(RInSp)

#rmd
library(details)
library(knitr)
library(kableExtra)

```

## WORKING DIRECTORIES
```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/rare_function.R")

```

# MERGED DIET DATA

```{r psObject_diet, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_diet <- readRDS(paste0(dataDir, "/ps.merge_diet.rds"))
diet_fr <- ps_diet %>% subset_samples(Genus == "Fringilla")

```

## 16S DATA

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# ASV clean data 

ps_clean <- readRDS(paste0(dataDir, "/ps_clean_16S_GTDB2023.rds" ))
ps_clean

microb_fr <- ps_clean %>% subset_samples(Genus == "Fringilla")
microb_fr <- removeZeros(microb_fr)


```

Overlap microbiome and diet
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# TipId as sample names 
sample_data(microb_fr)$Species_Label <- gsub(" ", "_", sample_data(microb_fr)$Species_Label)
sample_names(microb_fr) <- paste(sample_data(microb_fr)$Lab_Code, sample_data(microb_fr)$Species_Label, sep = "_")

# TipId as sample names 
sample_data(diet_fr)$Species_Label <- gsub(" ", "_", sample_data(diet_fr)$Species_Label)
sample_names(diet_fr) <- paste(sample_data(diet_fr)$Lab_Code, sample_data(diet_fr)$Species_Label, sep = "_")

setdiff(sample_names(microb_fr), sample_names(diet_fr)) %>% print
setdiff(sample_names(diet_fr), sample_names(microb_fr)) %>% print

# overlap
microb_overlap <- prune_samples(sample_names(microb_fr) %in% sample_names(diet_fr), microb_fr) 

diet_overlap <- prune_samples(sample_names(diet_fr) %in% sample_names(microb_fr), diet_fr) 

#write_tsv(sample_data(microb_overlap), paste0(dataDir, "/microb_diet_overlap_metadata.tsv"))


```

Covariates 

```{r covariates_diet, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

meta <- diet_overlap %>% sample_data %>% group_by(Genus) %>%  summarize(sampleSize = n()) %>% data.frame

# create Age (juvenile and adult) category
sample_data(diet_overlap)$Age_cat <- ifelse(sample_data(diet_overlap)$Age > 3, "adult", "fledging")

# recode Sex covariate 
sample_data(diet_overlap)$Sex <- ifelse(sample_data(diet_overlap)$Sex == 1, "male", "female")

# relevel
sample_data(diet_overlap)$Species <- as.factor(sample_data(diet_overlap)$Species)
sample_data(diet_overlap)$Species <- relevel(sample_data(diet_overlap)$Species, "coelebs")
str(sample_data(diet_overlap)$Species)
sample_data(diet_overlap)$Species_Label <- as.factor(sample_data(diet_overlap)$Species_Label)
sample_data(diet_overlap)$Species_Label <- relevel(sample_data(diet_overlap)$Species_Label, "Mainland_F_coelebs")
str(sample_data(diet_overlap)$Species_Label)

# recode numeric variables
sample_data(diet_overlap)$Area_km2 <- as.numeric(sample_data(diet_overlap)$Area_km2)
sample_data(diet_overlap)$Distance_mainland <- as.numeric(sample_data(diet_overlap)$Distance_mainland)
sample_data(diet_overlap)$Age_ma <- as.numeric(sample_data(diet_overlap)$Age_ma)

```

## RAREFY DIET

Rarefy data and extract species richness and shannon diversity

```{r rarefy, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# clean diet assignments (weird marine arthropods likely to be misassignments and plant material unassigned to Class) 
diet_clean <- ps_diet %>% subset_taxa(Order!= "Cumacea" & Order != "Cyclopoida" & Order != "Podocopida" & Order != "Spionida" & Class!= "Klebsormidiophyceae" & Order != "Class_Magnoliopsida")

mirl_object <- mirl(merge_phyloseq(t(otu_table(diet_clean)),sample_data(diet_clean),tax_table(diet_clean), refseq(diet_clean)), libsize = 1000, rep = 100, set.seed = 120, mc.cores =16)

mirl_object[sapply(mirl_object, is.null)] <- NULL

List = list ()

for (i in 1:length(mirl_object)){
  phy <- mirl_object[[i]]
  List[[i]] <- otu_table(phy)
  }


OTU_df <-list.cbind(List)

rare_OTU_df <- t(apply(OTU_df,1, function(x) round(tapply(x,colnames(OTU_df),mean),0)))

ps_diet_rarefied <- merge_phyloseq(t(otu_table(rare_OTU_df, taxa_are_rows =TRUE)), sample_data(diet_clean), tax_table(diet_clean), refseq(diet_clean))

# Remove taxa with zero abundance after rarefaction
ps <- removeZeros(ps_diet_rarefied)
max(otu_table(ps))

diet_alpha <- estimate_richness(ps_diet_rarefied, measures=c("Observed","Shannon"))

sample_data(ps_diet_rarefied)$Observed_diet_rare <- diet_alpha$Observed
sample_data(ps_diet_rarefied)$Shannon_diet_rare <- diet_alpha$Shannon

saveRDS(ps, paste0(dataDir, "/ps_diet_rarefied.rds"))

#ps_diet_rarefied <- readRDS(paste0(dataDir, "/ps_diet_rarefied.rds" ))
```


## 16S DATA

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# ASV clean phyloseq
ps_clean <- readRDS(paste0(dataDir, "/ps_clean_16S_GTDB2023.rds" ))
ps_clean

```

Metadata

```{r covariates, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# fix Species Label
sample_data(ps_clean)$Species_Label <- gsub(" ", "_", sample_data(ps_clean)$Species_Label) %>% as.factor

sample_data(ps_clean)$TipID <- paste(sample_data(ps_clean)$WGS_ID, sample_data(ps_clean)$Species_Label, sep = "_")

(sample_data(ps_clean)) %>% data.frame %>% names

# add WGS column

sample_data(ps_clean)$WGS_data <- ifelse(is.na(sample_data(ps_clean)$WGS_ID), 0, 1 )

(sample_data(ps_clean)) %>% data.frame %>% head

# add Diet column 
ps_diet <- readRDS(paste0(dataDir, "/ps_diet_rarefied.rds"))

meta_diet <- ps_diet %>% sample_data %>% data.frame

#write_tsv(meta_OTU,paste0(dataDir, "/meta_ps_merge_diet.tsv"))

diet_data <- prune_samples(sample_names(ps_clean) %in% sample_names(ps_diet), ps_clean) %>% sample_data %>% data.frame %>% dplyr::select("Lab_Code") 
diet_data <- diet_data$Lab_Code
sample_data(ps_clean)$diet_data <- ifelse(sample_data(ps_clean)$Lab_Code %in%  diet_data, 1, 0 )

table(sample_data(ps_clean)$diet_data)

# overlap WGS, diet
overlap <- ps_clean %>% sample_data %>% data.frame %>% filter(WGS_data == 1 & diet_data == 1)
nrow(overlap)

# create Age (juvenile and adult) category
sample_data(ps_clean)$Age_cat <- ifelse(sample_data(ps_clean)$Age > 3, "adult", "fledging")

# relevel
sample_data(ps_clean)$Species <- as.factor(sample_data(ps_clean)$Species)
sample_data(ps_clean)$Species <- relevel(sample_data(ps_clean)$Species, "coelebs")
str(sample_data(ps_clean)$Species)

sample_data(ps_clean)$Species_Label <- as.factor(sample_data(ps_clean)$Species_Label)
sample_data(ps_clean)$Species_Label <- relevel(sample_data(ps_clean)$Species_Label, "Mainland_F_coelebs")

sample_data(ps_clean)$Fringilla_Other <- if_else(sample_data(ps_clean)$Genus == "Fringilla", "Fringilla", "Other")

sample_data(ps_clean)$Chaff_Blue <- ifelse(sample_data(ps_clean)$Species_Label == "Tenerife_F_teydea" | sample_data(ps_clean)$Species_Label == "Gran_Canaria_F_polatzeki", "Bluefinch", ifelse(sample_data(ps_clean)$Fringilla_Other == "Other", "Other", "Chaffinch"))


```

## Rarefy 16S

Rarefy data and extract species richness and shannon diversity

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# warning: slow! 
mirl_object <- mirl(merge_phyloseq(t(otu_table(ps_clean)),sample_data(ps_clean),tax_table(ps_clean),refseq(ps_clean)), libsize = 2000, rep = 100, set.seed = 120, mc.cores = 16)

mirl_object[sapply(mirl_object, is.null)] <- NULL

List = list ()

for (i in 1:length(mirl_object)){
  phy <- mirl_object[[i]]
  List[[i]] <- otu_table(phy)
  }

ASV_df <-list.cbind(List)

rare_ASV_df <- t(apply(ASV_df,1, function(x) round(tapply(x,colnames(ASV_df),mean), 0)))
dim(rare_ASV_df)

ps_ASV_filter_rarefied <- merge_phyloseq(t(otu_table(rare_ASV_df, taxa_are_rows =TRUE)), sample_data(ps_clean), tax_table(ps_clean), refseq(ps_clean))

# Remove taxa with zero abundance after rarefaction
ps <- removeZeros(ps_ASV_filter_rarefied)
max(otu_table(ps)[5])

microb_alpha <- estimate_richness(ps, measures=c("Observed","Shannon"))

sample_data(ps)$Observed_ASV_rare <- microb_alpha$Observed
sample_data(ps)$Shannon_ASV_rare <- microb_alpha$Shannon

saveRDS(ps, paste0(dataDir, "/ps_ASV_rarefied.rds"))
```


Fasta file for ASVs

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps <- readRDS(paste0(dataDir, "/ps_ASV_rarefied.rds" ))

ps %>%
      refseq() %>%
      Biostrings::writeXStringSet(paste0(dataDir, "/ps_ASV_rarefied.fasta"), append=FALSE, compress=FALSE, compression_level=NA, format="fasta")

```



## ESTIMATE DIET GENERALIST/SPECIALIST PER ISLAND

This was estimated using the the E measure of inter-individual variation after Araujo et al. (2008)

https://doi.org/10.1890/07-0630.1
https://rdrr.io/cran/RInSp/man/Eindex.html

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}

ps_diet_rarefied <- readRDS(paste0(dataDir, "/ps_diet_rarefied.rds" ))

diet_df <- ps_diet_rarefied %>% sample_data %>% data.frame %>% group_by(Species_Label) %>% mutate(sampleSize = n())

#diet_df <- diet_df[1:(length(diet_df)-3)]
#head(diet_df)

#diet_df <- diet_df %>% group_by(Species_Label) %>% mutate(sampleSize = n())
#head(data.frame(diet_df))

table(droplevels(as.factor(diet_df$Species_Label)))

names(diet_df)

RInSp.Mainland_F_coelebs <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Mainland_F_coelebs")))
RInSp.Hierro_F_c_ombriosa <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "El Hierro_F_c_ombriosa")))
RInSp.Flores_F_moreletti <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Flores_F_moreletti")))
RInSp.Canaria_F_c_bakeri <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Gran Canaria_F_c_bakeri")))
RInSp.Canaria_F_polatzeki <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Gran Canaria_F_polatzeki")))
RInSp.Gomera_F_c_canariensis <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "La Gomera_F_c_canariensis")))
RInSp.Palma_F_c_palmae <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "La Palma_F_c_palmae")))
RInSp.Madeira_F_maderensis <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Madeira_F_maderensis")))
RInSp.Miguel_F_moreletti <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Sao Miguel_F_moreletti")))
RInSp.Tenerife_F_c_canariensis <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Tenerife_F_c_canariensis")))
RInSp.Tenerife_F_teydea <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Tenerife_F_teydea")))
RInSp.Hierro_P_canariensis <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "El Hierro_P_canariensis")))
RInSp.Hierro_T_merula <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "El Hierro_T_merula")))
RInSp.Gomera_C_teneriffae <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "La Gomera_C_teneriffae")))
RInSp.Gomera_E_rubecula <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "La Gomera_E_rubecula")))
RInSp.Gomera_T_merula <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "La Gomera_T_merula")))
RInSp.Sao_Miguel_M_cinerea <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Sao Miguel_M_cinerea")))
RInSp.Sao_Miguel_S_atricapilla <- import.RInSp(otu_table(subset_samples(ps_diet_rarefied, Species_Label == "Sao Miguel_S_atricapilla")))


Eresult.Mainland_F_coelebs <- Eindex(RInSp.Mainland_F_coelebs, index = "saramaki", jackknife = TRUE)
Eresult.Hierro_F_c_ombriosa <- Eindex(RInSp.Hierro_F_c_ombriosa, index = "saramaki", jackknife = TRUE)
Eresult.Flores_F_moreletti <- Eindex(RInSp.Flores_F_moreletti, index = "saramaki", jackknife = TRUE)
Eresult.Canaria_F_c_bakeri <- Eindex(RInSp.Canaria_F_c_bakeri, index = "saramaki", jackknife = TRUE)
Eresult.Canaria_F_polatzeki <- Eindex(RInSp.Canaria_F_polatzeki, index = "saramaki", jackknife = TRUE)
Eresult.Gomera_F_c_canariensis <- Eindex(RInSp.Gomera_F_c_canariensis, index = "saramaki", jackknife = TRUE)
Eresult.Palma_F_c_palmae <- Eindex(RInSp.Palma_F_c_palmae, index = "saramaki", jackknife = TRUE)
Eresult.Madeira_F_maderensis <- Eindex(RInSp.Madeira_F_maderensis, index = "saramaki", jackknife = TRUE)
Eresult.Miguel_F_moreletti <- Eindex(RInSp.Miguel_F_moreletti, index = "saramaki", jackknife = TRUE)
Eresult.Tenerife_F_c_canariensis <- Eindex(RInSp.Tenerife_F_c_canariensis, index = "saramaki", jackknife = TRUE)
Eresult.Tenerife_F_teydea <- Eindex(RInSp.Tenerife_F_teydea, index = "saramaki", jackknife = TRUE)
Eresult.Hierro_P_canariensis <- Eindex(RInSp.Hierro_P_canariensis, index = "saramaki", jackknife = TRUE)
Eresult.Hierro_T_merula <- Eindex(RInSp.Hierro_T_merula, index = "saramaki", jackknife = TRUE)
Eresult.Gomera_C_teneriffae <- Eindex(RInSp.Gomera_C_teneriffae, index = "saramaki", jackknife = TRUE)
Eresult.Gomera_E_rubecula <- Eindex(RInSp.Gomera_E_rubecula, index = "saramaki", jackknife = TRUE)
Eresult.Gomera_T_merula <- Eindex(RInSp.Gomera_T_merula, index = "saramaki", jackknife = TRUE)
Eresult.Sao_Miguel_M_cinerea <- Eindex(RInSp.Sao_Miguel_M_cinerea, index = "saramaki", jackknife = TRUE)
Eresult.Sao_Miguel_S_atricapilla <- Eindex(RInSp.Sao_Miguel_S_atricapilla, index = "saramaki", jackknife = TRUE)


Edata <- data.frame(E = c(Eresult.Mainland_F_coelebs$E, 
                          Eresult.Hierro_F_c_ombriosa$E,  
                          Eresult.Hierro_P_canariensis$E, 
                          Eresult.Hierro_T_merula$E, 
                          Eresult.Flores_F_moreletti$E,
                          Eresult.Canaria_F_c_bakeri$E,
                          Eresult.Canaria_F_polatzeki$E, 
                          Eresult.Gomera_C_teneriffae$E, 
                          Eresult.Gomera_E_rubecula$E, 
                          Eresult.Gomera_F_c_canariensis$E, 
                          Eresult.Gomera_T_merula$E,
                          Eresult.Palma_F_c_palmae$E,
                          Eresult.Madeira_F_maderensis$E,
                          Eresult.Miguel_F_moreletti$E,
                          Eresult.Sao_Miguel_M_cinerea$E, 
                          Eresult.Sao_Miguel_S_atricapilla$E, 
                          Eresult.Tenerife_F_c_canariensis$E, 
                          Eresult.Tenerife_F_teydea$E),
                    VarE = c(Eresult.Mainland_F_coelebs$VarE, 
                          Eresult.Hierro_F_c_ombriosa$VarE, 
                          Eresult.Hierro_P_canariensis$VarE, 
                          Eresult.Hierro_T_merula$VarE, 
                          Eresult.Flores_F_moreletti$VarE,
                          Eresult.Canaria_F_c_bakeri$VarE,
                          Eresult.Canaria_F_polatzeki$VarE, 
                          Eresult.Gomera_C_teneriffae$VarE, 
                          Eresult.Gomera_E_rubecula$VarE, 
                          Eresult.Gomera_F_c_canariensis$VarE, 
                          Eresult.Gomera_T_merula$VarE, 
                          Eresult.Palma_F_c_palmae$VarE,
                          Eresult.Madeira_F_maderensis$VarE,
                          Eresult.Miguel_F_moreletti$VarE, 
                          Eresult.Sao_Miguel_M_cinerea$VarE, 
                          Eresult.Sao_Miguel_S_atricapilla$VarE, 
                          Eresult.Tenerife_F_c_canariensis$VarE,
                          Eresult.Tenerife_F_teydea$VarE), 
                    Species_Label = c("Mainland_F_coelebs","El Hierro_F_c_ombriosa","El Hierro_P_canariensis","El Hierro_T_merula","Flores_F_moreletti","Gran Canaria_F_c_bakeri","Gran Canaria_F_polatzeki","La Gomera_C_teneriffae","La Gomera_E_rubecula","La Gomera_F_c_canariensis","La Gomera_T_merula","La Palma_F_c_palmae","Madeira_F_maderensis","Sao Miguel_F_moreletti","Sao Miguel_M_cinerea","Sao Miguel_S_atricapilla","Tenerife_F_c_canariensis","Tenerife_F_teydea"))

head(Edata)

Edata$Species_Label <- factor(Edata$Species_Label)
Edata

diet_df <- (merge(diet_df, Edata, by = "Species_Label", all = TRUE))
tail(diet_df)
diet_df <-(merge(data.frame(sample_data(ps_diet_rarefied)),diet_df[,c(3,37,38)], by ="Lab_Code"))
tail(diet_df)

base::setdiff(sample_data(ps_diet_rarefied)$Lab_Code, diet_df$Lab_Code) %>% print
#setdiff(sample_data(ps_OTU_99_rarefied)$Lab_Code, diet_df$Lab_Code) %>% print
  
sample_data(ps_diet_rarefied)$E <- diet_df$E
sample_data(ps_diet_rarefied)$VarE <- diet_df$VarE

saveRDS(ps_diet_rarefied, paste0(dataDir, "/ps_diet_rarefied.rds"))

fringilla <- c("Mainland_F_coelebs","El Hierro_F_c_ombriosa","Flores_F_moreletti","Gran Canaria_F_c_bakeri","Gran Canaria_F_polatzeki","La Gomera_F_c_canariensis","La Palma_F_c_palmae","Madeira_F_maderensis","Sao Miguel_F_moreletti","Tenerife_F_c_canariensis","Tenerife_F_teydea")


Edata_df <- Edata[Edata$Species_Label %in% fringilla,]

write.table(Edata_df, file=paste0(dataDir, "/Edata_df.tsv"), quote=FALSE, sep='\t', col.names = NA)

Edata_df <- read_tsv(paste0(dataDir, "/Edata_df.tsv"))

```


## MERGE DIET AND MICROBIOME DIVERSITY DATA


Overlap microbiome and diet
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_ASV_rarefied <- readRDS(paste0(dataDir, "/ps_ASV_rarefied.rds"))
ps_diet_rarefied <- readRDS(paste0(dataDir, "/ps_diet_rarefied.rds"))

sample_data(ps_diet_rarefied)$E
sample_data(ps_diet_rarefied)$Lab_Code

ps_ASV_rarefied %>% sample_data %>% names
ps_diet_rarefied %>% sample_data %>% names

# TipId as sample names 
sample_data(ps_ASV_rarefied)$Species_Label <- gsub(" ", "_", sample_data(ps_ASV_rarefied)$Species_Label)
sample_names(ps_ASV_rarefied) <- paste(sample_data(ps_ASV_rarefied)$Lab_Code, sample_data(ps_ASV_rarefied)$Species_Label, sep = "_")

# TipId as sample names 
sample_data(ps_diet_rarefied)$Species_Label <- gsub(" ", "_", sample_data(ps_diet_rarefied)$Species_Label)
sample_names(ps_diet_rarefied) <- paste(sample_data(ps_diet_rarefied)$Lab_Code, sample_data(ps_diet_rarefied)$Species_Label, sep = "_")

setdiff(sample_names(ps_ASV_rarefied), sample_names(ps_diet_rarefied)) %>% print
setdiff(sample_names(ps_diet_rarefied), sample_names(ps_ASV_rarefied)) %>% print

# overlap
microb_overlap_rare <- prune_samples(sample_names(ps_ASV_rarefied) %in% sample_names(ps_diet_rarefied), ps_ASV_rarefied) 

diet_overlap_rare <- prune_samples(sample_names(ps_diet_rarefied) %in% sample_names(ps_ASV_rarefied), ps_diet_rarefied) 

#write_tsv(sample_data(microb_overlap), paste0(dataDir, "/microb_diet_overlap_metadata.tsv"))

microb_overlap_rare_fr <- microb_overlap_rare %>% subset_samples(Genus == "Fringilla")
diet_overlap_rare_fr <- diet_overlap_rare %>% subset_samples(Genus == "Fringilla")

sample_names(microb_overlap_rare_fr) == sample_names(diet_overlap_rare_fr)

sample_data(diet_overlap_rare_fr)$Observed_ASV_rare <- sample_data(microb_overlap_rare_fr)$Observed_ASV_rare
sample_data(diet_overlap_rare_fr)$Shannon_ASV_rare <- sample_data(microb_overlap_rare_fr)$Shannon_ASV_rare

sample_data(microb_overlap_rare_fr)$Observed_diet_rare <- sample_data(diet_overlap_rare_fr)$Observed_diet_rare
sample_data(microb_overlap_rare_fr)$Shannon_diet_rare <- sample_data(diet_overlap_rare_fr)$Shannon_diet_rare
sample_data(microb_overlap_rare_fr)$E <- sample_data(diet_overlap_rare_fr)$E
sample_data(microb_overlap_rare_fr)$VarE <- sample_data(diet_overlap_rare_fr)$VarE


sample_data(diet_overlap_rare)$Observed_ASV_rare <- sample_data(microb_overlap_rare)$Observed_ASV_rare
sample_data(diet_overlap_rare)$Shannon_ASV_rare <- sample_data(microb_overlap_rare)$Shannon_ASV_rare

sample_data(microb_overlap_rare)$Observed_diet_rare <- sample_data(diet_overlap_rare)$Observed_diet_rare
sample_data(microb_overlap_rare)$Shannon_diet_rare <- sample_data(diet_overlap_rare)$Shannon_diet_rare

microb_overlap_rare <- removeZeros(microb_overlap_rare)
diet_overlap_rare <- removeZeros(diet_overlap_rare)
microb_overlap_rare_fr <- removeZeros(microb_overlap_rare_fr)
diet_overlap_rare_fr <- removeZeros(diet_overlap_rare_fr)

```

write to file

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

saveRDS(microb_overlap_rare_fr, paste0(dataDir, "/microb_overlap_ASV_rare_fr.rds"))

saveRDS(diet_overlap_rare_fr, paste0(dataDir, "/diet_overlap_ASV_rare_fr.rds"))


```
