---
title: 'Figure 4 plots'
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute:
    code_folding: hide
    toc_depth: 2
---

## LOAD LIBRARIES

```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rm(list = ls())

library(dplyr)
library(tidyverse)
library(stringr)
library(vegan)
library(ggplot2)
library(grid)
library(patchwork)

# phyloseq
library(phyloseq)
library(biomeUtils)
library(speedyseq)

# distances
library(sf)
library(geosphere)

# Euler / venn
library(eulerr)

# GLMM
library(glmmTMB)
library(effects)

```

## WORKING DIRECTORIES

```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```

## LOAD STATISTICAL OUTPUTS

```{r outputs, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

varpart_out <- read.csv(paste0(dataDir, "/output_varpart.csv"))
#mantel_out <- read.csv(paste0(dataDir, "/output_mantel_weightedfst.csv"))
mantel_out <- read.csv(paste0(dataDir, "/mantel_resampled_summary.csv")) %>%
  filter(data == "all", test == "partial-mantel")
glmm_out <- read.csv(paste0(dataDir, "/glmm_fst_roh_output_miss0.5_mac3_noZ_speciesFix_100kb_1Mb.csv"))

```

## EULER PLOTS (variance partitioning)

```{r euler, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Helper that builds a named vector for euler() from a varpart output df
build_euler_set <- function(varpart_df, dist_name, taxon) {
  df <- varpart_df %>%
    filter(distance == dist_name & tax == taxon) %>%
    dplyr::select(variable, percent)
  
  c("R"               = df[8, 2],
    "R&host"          = df[1, 2],
    "R&diet"          = df[2, 2],
    "R&geo"           = df[3, 2],
    "R&host&diet"     = df[4, 2],
    "R&diet&geo"      = df[5, 2],
    "R&host&geo"      = df[6, 2],
    "R&host&diet&geo" = df[7, 2])
}

# Helper that plots an Euler diagram with the standard styling
plot_euler <- function(euler_set, title_str = "Species") {
  fit <- euler(euler_set, input = "disjoint", shape = "ellipse")
  
  set_colors <- c(R = "white", host = "#a00000",
                  diet = "#000000", geo = "#59a89c")
  
  # average two or more hex colours in RGB space
  blend <- function(cols, amount = 0.25) {
    rgbs <- col2rgb(cols) / 255
    avg  <- rowMeans(rgbs)
    lit  <- avg + (1 - avg) * amount
    rgb(lit["red"], lit["green"], lit["blue"])
  }
  
  p <- plot(fit,
       fills = list(fill = set_colors),
       quantities = FALSE,
       labels = list(labels = c("", "", "", ""), cex = 1))
  
  region_names <- names(fit$fitted.values)
  diag_grob <- p$children$canvas.grob$children$diagram.grob.1
  
  for (i in seq_along(region_names)) {
    nm <- paste0("fills.grob.", i)
    g <- diag_grob$children[[nm]]
    if (is.null(g) || is.null(g$gp)) next
    
    sets  <- strsplit(region_names[i], "&")[[1]]
    non_R <- setdiff(sets, "R")
    
    if (length(non_R) == 0) {
      g$gp$fill <- "white"
    } else if (length(non_R) == 1) {
      g$gp$fill <- unname(set_colors[non_R])
    } else {
      g$gp$fill <- blend(unname(set_colors[non_R]))
    }
    
    diag_grob$children[[nm]] <- g
  }
  p$children$canvas.grob$children$diagram.grob.1 <- diag_grob
  
  gridExtra::grid.arrange(grobs = list(p),
                          top = textGrob(title_str, gp = gpar(fontsize = 16)))
}
# At Species level (used in main figure)
species_bray_set <- build_euler_set(varpart_out, "Bray-Curtis", "Species")
species_jacc_set <- build_euler_set(varpart_out, "Jaccard",     "Species")

euler_bray <- plot_euler(species_bray_set, "Species")
euler_jacc <- plot_euler(species_jacc_set, "Species")

```

## VARPART BARPLOTS

```{r barplot, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Collapse overlap categories
varpart_out$variable2 <- gsub(".*_.*", "overlap", varpart_out$variable)

ind_sum <- varpart_out %>%
  group_by(pop_ind, distance, tax, variable2) %>%
  summarise(percent2 = sum(percent), .groups = "drop")

ind_sum$order <- factor(ind_sum$variable2,
  levels = c("host", "diet", "geo", "overlap", "Unexplained"),
  labels = c("Phylogeny", "Diet", "Geography", "Overlap", "Unexplained"))

# Colour palette tied to the levels
varpart_colors <- c(
  "Phylogeny"    = "#a00000",
  "Diet"         = "#000000",
  "Geography"    = "#59a89c",
  "Overlap"      = "lightgrey",
  "Unexplained"  = "white")

ind_sum$tax <- factor(ind_sum$tax,
  levels = c("OTU99", "Species", "Genus", "Family", "Order", "Class", "Phylum"))

(varPar <- ggplot(ind_sum, aes(y = tax, x = percent2, fill = order)) +
    geom_bar(position = "stack", stat = "identity", linewidth = 0.25, color = "black") +
    geom_bar(position = "stack", stat = "identity", linewidth = 0.5, show.legend = FALSE) +
    scale_x_continuous("% variance explained", limits = c(0, 25), expand = c(0, 0)) +
    scale_fill_manual(values = varpart_colors, name = "") +
    facet_wrap(. ~ factor(distance), strip.position = "top", ncol = 2, dir = "v") +
    barplot_theme +
    theme(legend.position = "right",
          panel.spacing = unit(2, "lines"),
          panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(colour = "#bababa", linewidth = 1),
          axis.title.x = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          axis.text.y = element_text(size = 16),
          axis.text.x = element_text(angle = 0, size = 16),
          axis.line = element_line(colour = 'black', size = 0.5),
          legend.title = element_text(size = 18),
          legend.text = element_text(size = 16),
          strip.text = element_text(size = 18)) +
    xlab("% variance explained") + ylab(""))

```

## LOAD AND PREPARE DATA FOR FST AND F_ROH PLOTS

```{r load-microb, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Use corrected microbiome ∩ diet ∩ WGS object
microb_overlap_otu <- readRDS(paste0(dataDir, "/microb_overlap_diet_corrected.rds")) %>%
  biomeUtils::removeZeros()

microb_overlap_species <- speedyseq::tax_glom(microb_overlap_otu, "Species")

# Add TipID
sample_data(microb_overlap_species)$TipID_microbiome <- paste(
  sample_data(microb_overlap_species)$Lab_Code,
  sample_data(microb_overlap_species)$Species_Label, sep = "_")

microb_overlap_meta <- data.frame(sample_data(microb_overlap_species))

```

## SUBSAMPLE MICROBIOME (equal sample size per population)

```{r subsample, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

sample_ps <- function(ps, size, FUN = sample) {
  ids <- sample_names(ps)
  prune_samples(FUN(ids, size = size), ps)
}

pop_labels <- levels(factor(sample_data(microb_overlap_species)$Species_Label))

set.seed(999)
subsampled_list <- lapply(pop_labels, function(lbl) {
  # Use prune_samples with logical vector instead of subset_samples to avoid NSE
  keep <- sample_data(microb_overlap_species)$Species_Label == lbl
  ps <- prune_samples(keep, microb_overlap_species)
  if (nrow(sample_data(ps)) >= 9) sample_ps(ps, size = 9) else ps
})

microb_fr_subsample_species <- do.call(merge_phyloseq, subsampled_list) %>%
  biomeUtils::removeZeros()
```

## MERGE BY POPULATION

```{r merge, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_merged_species <- phyloseq::merge_samples(microb_fr_subsample_species,
                                             "Species_Label", fun = sum)

```

## GEOGRAPHIC CENTROIDS

```{r centroids, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

my_points <- data.frame(
  longitude = sample_data(microb_overlap_otu)$Xdec,
  latitude = sample_data(microb_overlap_otu)$Ydec,
  Island_Mainland = sample_data(microb_overlap_otu)$Species_Label,
  row.names = sample_names(microb_overlap_otu))

df_sf <- st_as_sf(my_points, coords = c("longitude", "latitude"))

centroids_sf <- df_sf %>%
  group_by(Island_Mainland) %>%
  dplyr::summarize(geometry = st_union(geometry)) %>%
  st_centroid() %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[, 1],
                lat = sf::st_coordinates(.)[, 2])

rownames(centroids_sf) <- centroids_sf$Island_Mainland

dist_mat <- distm(cbind(centroids_sf$lon, centroids_sf$lat), fun = distGeo) / 1000
rownames(dist_mat) <- centroids_sf$Island_Mainland
colnames(dist_mat) <- centroids_sf$Island_Mainland
geo_dist <- as.dist(dist_mat)

```

## MICROBIOME DISTANCES (population level, for FST plots)

```{r microb-dist-pop, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_otu_tab <- otu_table(ps_merged_species)[rownames(dist_mat), ]
bray_dist_species    <- vegdist(microb_otu_tab, method = "bray")
jaccard_dist_species <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)

```

## LOAD FST

```{r fst, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

fst <- read.csv(paste0(dataDir, "/MeanFst_Fr_weighted_estimates.csv"),
                header = TRUE, row.names = 1) %>%
  as.matrix()
fst <- fst[rownames(dist_mat), colnames(dist_mat)]
fst_dist <- as.dist(fst)


```

## DIVERGENCE (individual-level Bray and Jaccard)

```{r divergence, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Helper to compute divergence (betadisper distance to median) per metric
compute_divergence <- function(ps, method, binary = FALSE, name) {
  tab <- otu_table(ps)
  d <- vegdist(tab, method = method, binary = binary)
  bd <- betadisper(d, sample_data(ps)$Species_Label, type = "median")$distance
  df <- data.frame(beta_median = bd) %>%
    rownames_to_column("TipID_microbiome")
  names(df)[2] <- name
  df
}

bray_div    <- compute_divergence(microb_overlap_species, "bray",    binary = FALSE, "b.beta_median")
jaccard_div <- compute_divergence(microb_overlap_species, "jaccard", binary = TRUE,  "j.beta_median")

```

## LOAD F_ROH

```{r froh, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

metadata <- read_delim(paste0(dataDir, "/WGS_Fringilla_samples.tsv"), delim = ",")

froh <- read_delim(paste0(dataDir, "/fringilla_microbiome_froh_100kb_1Mb.csv"), delim = ",")
names(froh)[names(froh) == "file_id"] <- "TipID_microbiome"


froh_meta <- left_join(froh, metadata, by = "TipID_microbiome")

roh_bray <- left_join(froh_meta, bray_div, by = "TipID_microbiome") %>%
  filter(!is.na(b.beta_median))
roh_jacc <- left_join(froh_meta, jaccard_div, by = "TipID_microbiome") %>%
  filter(!is.na(j.beta_median))

```

## PREPARE PLOT DATA FRAMES

```{r plot-data, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Long format for FST plot
dist_fst_df <- bind_rows(
  data.frame(Distance = as.vector(bray_dist_species),
             fst = as.vector(fst_dist), Distance_name = "Bray"),
  data.frame(Distance = as.vector(jaccard_dist_species),
             fst = as.vector(fst_dist), Distance_name = "Jaccard")
) %>% mutate(fst = as.numeric(fst), Distance = as.numeric(Distance))

# Long format for F_ROH plot
dist_froh_df <- bind_rows(
  data.frame(beta_median = roh_bray$b.beta_median,
             froh = roh_bray$froh, Distance_name = "Bray"),
  data.frame(beta_median = roh_jacc$j.beta_median,
             froh = roh_jacc$froh, Distance_name = "Jaccard")
) %>% mutate(froh = as.numeric(froh), beta_median = as.numeric(beta_median))

```

## SHARED THEME FOR ALL FIGURE 4 PANELS

```{r theme, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# One theme to rule them all. Use base sizes that look right at the
# final compose dimensions (21 × 30 inches at 300 dpi).

fig4_base_size <- 14

fig4_theme <- theme(
  axis.title       = element_text(size = fig4_base_size + 2),
  axis.text        = element_text(size = fig4_base_size, colour = "black"),
  axis.line        = element_line(colour = "black", linewidth = 0.5),
  legend.title     = element_text(size = fig4_base_size + 2),
  legend.text      = element_text(size = fig4_base_size),
  legend.position  = "right",
  panel.background = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border     = element_blank(),
  panel.spacing    = unit(1.5, "lines"),
  strip.background = element_blank(),
  strip.text       = element_text(size = fig4_base_size + 2, face = "bold"),
  strip.text.x     = element_blank())

fig4_themeB <- theme(
  axis.title       = element_text(size = fig4_base_size + 2),
  axis.text        = element_text(size = fig4_base_size, colour = "black"),
  axis.line        = element_line(colour = "black", linewidth = 0.5),
  legend.title     = element_text(size = fig4_base_size + 2),
  legend.text      = element_text(size = fig4_base_size),
  legend.position  = "right",
  panel.background = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border     = element_blank(),
  panel.spacing    = unit(1.5, "lines"),
  strip.background = element_rect(fill = "black"),
  strip.text       = element_text(size = fig4_base_size + 2, face = "bold", color = "white"))#,
  #strip.text.x     = element_blank())

```

## Partial MANTEL PLOT

```{r mantel-plot, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

#mantel_out <- mantel_out %>%
#  mutate(signif = case_when(
#    p < 0.0005 ~ "< 0.0005",
#    p < 0.005  ~ "< 0.005",
#    p < 0.05   ~ "< 0.05",
#    TRUE       ~ "n.s."),
#    order = factor(taxonomy,
#      levels = c("OTU99", "Species", "Genus", "Family", "Order", "Class", "Phylum")))

#mantel_fill_colors <- c("< 0.05" = "#999999", "< 0.005" = "#333333",
#                        "< 0.0005" = "#000000", "n.s." = "white")

#(p_mantel <- mantel_out %>%
#   filter(data == "all", test == "partial-mantel") %>%
#   ggplot(aes(y = r, x = order, fill = signif)) +
#     geom_bar(stat = "identity", linewidth = 0.25, color = "black") +
#     scale_fill_manual("p-value", values = mantel_fill_colors) +
#     ylab(expression("Mantel" ~ italic("r"))) + xlab("") +
#     coord_flip() +
#     facet_wrap(. ~ factor(dist), strip.position = "top", ncol = 2, dir = "v") +
#     fig4_theme)

# add significance bins (mirroring the GLMM signif coding) and order taxonomy
mantel_out$signif <- ifelse(mantel_out$p_median < 0.0005, "< 0.0005",
              ifelse(mantel_out$p_median < 0.005,  "< 0.005",
              ifelse(mantel_out$p_median < 0.05,   "< 0.05", "n.s")))

mantel_out$order <- factor(mantel_out$taxonomy,
                    levels = c("OTU99","Species","Genus","Family","Order","Class","Phylum"))


glmm_fill_colors <- c("< 0.05" = "#999999", "< 0.005" = "#333333",
                      "< 0.0005" = "#000000", "n.s" = "white")

(p_mantel_out_forest <- mantel_out %>%
   ggplot(aes(x = r_median, y = order)) +
     geom_linerange(aes(xmin = r_lo, xmax = r_hi), linewidth = 1.2, color = "#000000") +
     geom_point(aes(fill = signif), pch = 21, size = 4, color = "black") +
     geom_vline(xintercept = 0, linetype = "longdash") +
     scale_fill_manual("p-value", values = glmm_fill_colors) +
     xlab(expression("Partial Mantel"~italic(r)~"(F"[ST]*" | geo)")) + ylab("") +
     facet_wrap(. ~ factor(dist), strip.position = "top") +
     fig4_theme)




```

## FST SCATTER PLOT

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
pair_summary <- read.csv(paste0(dataDir, "/dist_fst_resampled_summary.csv"))

dist_fst_df <- bind_rows(
  data.frame(Distance      = pair_summary$bray_med,
             lower         = pair_summary$bray_lo,
             upper         = pair_summary$bray_hi,
             fst           = pair_summary$fst,
             Distance_name = "Bray"),
  data.frame(Distance      = pair_summary$jacc_med,
             lower         = pair_summary$jacc_lo,
             upper         = pair_summary$jacc_hi,
             fst           = pair_summary$fst,
             Distance_name = "Jaccard")
)

```

```{r fst-plot, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(p_fst <- ggplot(dist_fst_df, aes(y = Distance, x = fst, color = Distance_name)) +
   geom_point(size = 3, alpha = 0.3, colour = "black") +
   labs(x = expression(italic("F")[ST]), y = "Distance") +
   geom_smooth(method = "lm", alpha = 0.2, linewidth = 1.5) +
   scale_color_manual(values = c("darkred", "black")) +
   facet_wrap(. ~ factor(Distance_name), strip.position = "top") +
   fig4_theme +
   theme(legend.position = "none"))

```

## F_ROH GLMM AND PLOT

```{r froh-plot, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

glmm_bray <- glmmTMB(b.beta_median ~ froh + (1 | Species_Label),
                     data = roh_bray, na.action = na.fail, family = gaussian)
glmm_jacc <- glmmTMB(j.beta_median ~ froh + (1 | Species_Label),
                     data = roh_jacc, na.action = na.fail, family = gaussian)

effect_bray <- as.data.frame(effect("froh", glmm_bray, confint = list(alpha = .95))) %>%
  mutate(Distance_name = "Bray")
effect_jac <- as.data.frame(effect("froh", glmm_jacc, confint = list(alpha = .95))) %>%
  mutate(Distance_name = "Jaccard")
effect_df <- bind_rows(effect_bray, effect_jac)

(p_het <- ggplot(dist_froh_df, aes(y = beta_median, x = froh, color = Distance_name)) +
   geom_point(size = 3, alpha = 0.3, colour = "black") +
   scale_color_manual(values = c("darkred", "black")) +
   xlab(expression(italic("F"[ROH]))) + ylab("Divergence") +
   geom_line(data = effect_df, aes(x = froh, y = fit), linewidth = 1.5) +
   geom_ribbon(data = effect_df,
               aes(x = froh, y = fit, ymin = lower, ymax = upper),
               alpha = 0.1, colour = NA) +
   facet_wrap(. ~ factor(Distance_name), strip.position = "top") +
   fig4_theme +
   theme(legend.position = "none"))

```

## GLMM F_ROH FOREST PLOT

```{r glmm-plot, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

colnames(glmm_out) <- c("term", "Estimate", "Std_Error", "z_value", "P",
                        "lower", "upper", "data", "taxonomy", "dist")

glmm_out <- glmm_out %>%
  mutate(
    order = factor(taxonomy,
      levels = c("OTU99", "species", "genus", "family", "order", "class", "phylum"),
      labels = c("OTU99", "Species", "Genus", "Family", "Order", "Class", "Phylum")),
    signif = case_when(
      P < 0.0005 ~ "< 0.0005",
      P < 0.005  ~ "< 0.005",
      P < 0.05   ~ "< 0.05",
      TRUE       ~ "n.s"),
    term = gsub("[[:digit:]]+", "", term))

glmm_fill_colors <- c("< 0.05" = "#999999", "< 0.005" = "#333333",
                      "< 0.0005" = "#000000", "n.s" = "white")

(p_glmm <- glmm_out %>%
   filter(term == "scale(froh)", data == "all") %>%
   ggplot(aes(x = Estimate, y = order)) +
     geom_linerange(aes(xmin = lower, xmax = upper), linewidth = 1.2, color = "#000000") +
     geom_point(aes(fill = signif), pch = 21, size = 4, color = "black") +
     geom_vline(xintercept = 0, linetype = "longdash") +
     scale_fill_manual("p-value", values = glmm_fill_colors) +
     xlab("Estimate") + ylab("") +
     facet_wrap(. ~ factor(dist), strip.position = "top") +
     fig4_theme)

```

## VARPART BARPLOT

```{r barplot-rethemed, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(varPar <- ggplot(ind_sum, aes(y = tax, x = percent2, fill = order)) +
    geom_bar(position = "stack", stat = "identity", linewidth = 0.25, color = "black") +
    scale_x_continuous("% variance explained", limits = c(0, 25), expand = c(0, 0)) +
    scale_fill_manual(values = varpart_colors, name = "") +
    facet_wrap(. ~ factor(distance), strip.position = "top", ncol = 2, dir = "v") +
    xlab("% variance explained") + ylab("") +
    fig4_themeB) 
    #+
    #theme(panel.grid.major.x = element_line(colour = "#dadada", linewidth = 0.5)))

```

## COMPOSE FIGURE 4

```{r figure4, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.height=20, fig.width=14}

((varPar +
    inset_element(euler_bray, 0.3, 0.18, 0.48, 0.99) +
    inset_element(euler_jacc, 0.82, 0.18, 1, 0.99)) /
  p_fst /
  p_mantel_out_forest /
  p_het /
  p_glmm) +
  plot_annotation(tag_levels = list(c("A", "", "", "B", "", "C", ""))) &
  theme(plot.tag = element_text(size = 18, face = "bold"))

ggsave(paste0(outputDir, "/Figures/Figure4_diet_rc_roh_100kb_1Mb.png"),
       width = 10.5, height = 15, units = "in", dpi = 300)
```