---
title: 'Microbiome Viz'
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: html_document
---

## LOAD LIBRARIES



```{r loadLibraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
rm(list = ls())

# basics
library(stringr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(magrittr)
library(data.table)
library(ggplot2)
library(ggpubr)
require(ape)
library(patchwork)

# phyloseq section
library(phyloseq)
library(Biostrings)
library(DECIPHER)
library(speedyseq)
library(metagMisc)
library(phylosmith)
library(seqateurs)

# tree
require(readxl)
require(ape)
require(phylobase)
library(gplots)
library(phytools)
library(ggpattern)
library(Rcpp)
library(ggtreeExtra)
library(ggtree)
library(treeio)
library(tidytree)
library(ggstar)
library(ggplot2)
library(ggnewscale)
library(TDbook)
library(ggtext)
options(getClass.msg=FALSE)

#neutral model
library(tyRa)
library(minpack.lm)
library(Hmisc)
library(stats4)

#rmd
library(details)
library(knitr)
library(kableExtra)

```

## WORKING DIRECTORIES
```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/PNM_functions_tyralibrary.r")

```

## PANEL A

Microbiome Family TREE 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr <- readRDS(paste0(dataDir, "/ps_core_familyFix.rds" ))

ps_per_island_core <- microb_fr[[1]]
ps_across_island_core <- microb_fr[[2]]

Phylum_core <- data.frame(OTU_ID = as.data.frame(rownames(tax_table(ps_across_island_core))), Phylum_core = as.data.frame(tax_table(ps_across_island_core))$Phylum, Family_core = as.data.frame(tax_table(ps_across_island_core))$Family)

(names(Phylum_core) <- c("OTU_ID", "Phylum_core", "Family_core"))
table(Phylum_core$Phylum_core)
head(Phylum_core)
# neutral model output
pred_df <- read.table(paste0(dataDir, "/predictionNeutralModel_overallDataset_familyFix.tsv"), header = TRUE)
str(pred_df)
table(pred_df$Core_prev)
table(pred_df$Core_across)
pred_df_core <- pred_df %>% filter(Core_across == "Core_across") %>% dplyr::select(Family, fit_class)

```


Prepare tables for treeplot 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# metadata 
TAXdf <- as(tax_table(ps_per_island_core), "matrix")  %>% as_tibble()
TAXdf <- left_join(TAXdf, Phylum_core, by = "OTU_ID")
TAXdf <- left_join(TAXdf, pred_df_core, by = "Family")
TAXdf$Phylum_core[is.na(TAXdf$Phylum_core)] <- ""
TAXdf %>% data.frame %>% head
TAXdf$Prediction_core <- ifelse(!is.na(TAXdf$Family_core), TAXdf$fit_class, NA)
TAXdf$Prediction_core[is.na(TAXdf$Prediction_core)] <- ""
#TAXdf$Prediction_core <- factor(ifelse(!is.na(TAXdf$Family_core), TAXdf$fit_class, NA),labels = c("above", "neutral", "below") )


table(TAXdf$fit_class)
table(TAXdf$Prediction_core)

# Extract relative abundance matrix from the phyloseq object

ps_per_island_core <- transform_sample_counts(ps_per_island_core, function(x) x / sum(x))

OTUdf <-  as(otu_table(ps_per_island_core), "matrix") %>%   as_tibble(rownames = "sampleID") %>% pivot_longer(!sampleID, names_to = "OTU_ID", values_to = "Abundance") %>% filter(Abundance != 0) 

Prevalence <- data.frame(Prevalence = colSums(otu_table(ps_per_island_core) > 0), tax_table(ps_per_island_core)) %>%  select(Prevalence) %>% as_tibble(rownames = "OTU_ID") 

Prevalence <- Prevalence %>% filter(Prevalence != 0)
Prevalence

OTUdf <- left_join(OTUdf, Prevalence, by = "OTU_ID" )

Phylum_core_meta <- ps_across_island_core %>% sample_data %>% data.frame %>% rownames_to_column("sampleID") %>% dplyr::select(sampleID, Species_Label) 
head(Phylum_core_meta)

OTUdf <- left_join(OTUdf, Phylum_core_meta, by = "sampleID")
is.na(table(OTUdf$Species_Label))

OTUdf_summ <- OTUdf %>% dplyr::group_by(Species_Label, OTU_ID) %>% summarise(RelAbund_mean = mean(Abundance), RelAbund_median = median(Abundance), Prevalence = max(Prevalence))
table( OTUdf_summ$OTU_ID)

topOTU <- OTUdf_summ %>%  arrange(desc(Prevalence)) #%>% head(10)

# Tree
tree_16S <- phy_tree(ps_per_island_core)
treetips <- as.character(tree_16S$tip.label)

# Ladderize tree
tree_16S<-ladderize(tree_16S)

# Reorder the sample information to match the order in the tree labels
treetipsorder <- cbind(treetips,"treenames"=treetips,"indexx"=c(100:(length(treetips)+99)))
head(treetipsorder)

mergedSampleInfo <- merge(TAXdf,treetipsorder, by.x="OTU_ID",by.y="treetips",all.y=T)
head(mergedSampleInfo)
sortedSampleInfo <- mergedSampleInfo[order(mergedSampleInfo$indexx),]
sortedSampleInfo <- droplevels(sortedSampleInfo)
sortedSampleInfo <- unique(x = sortedSampleInfo)
sortedSampleInfo <- cbind(sortedSampleInfo,"group_color" =sortedSampleInfo$Phylum_core)
sortedSampleInfo$group_color <- as.factor(sortedSampleInfo$group_color)

head(sortedSampleInfo)
str(sortedSampleInfo)

table(sortedSampleInfo$Phylum, sortedSampleInfo$group_color)
table(sortedSampleInfo$group_color)
levels(sortedSampleInfo$group_color) <- c("grey80", "#8db405", "#f3c300", "#a9c7e2", "#f38400", "#a7e586")
#                 Actinomycetota       Bacillota     Myxococcota Planctomycetota 
#            182              16               2               1               1 
 #Pseudomonadota 
 #            10 

#, "#018856",  "#d596ab", "#0266a1", "#f9937a", "#be0131", "#882d18", "#8db405")

sortedSampleInfo$group_color <- as.character(sortedSampleInfo$group_color)
sortedSampleInfo$group_pch <- as.factor(sortedSampleInfo$Prediction_core)
levels(sortedSampleInfo$group_pch) <- c(21, 24, 22, 25)

sortedSampleInfo <- droplevels(sortedSampleInfo)
sortedSampleInfo <- unique(x=sortedSampleInfo)
sortedSampleInfo %>% str

# Add phylum symbol
sortedSampleInfo$PhylumPch <- sortedSampleInfo$Phylum_core
sortedSampleInfo$PhylumPch[is.na(sortedSampleInfo$Phylum_core)] <- NA
sortedSampleInfo$PhylumPch <- as.factor(sortedSampleInfo$PhylumPch)
levels(sortedSampleInfo$PhylumPch) <- rep(1:20,length.out=length(levels(sortedSampleInfo$PhylumPch)))
sortedSampleInfo$PhylumPch <- as.integer(sortedSampleInfo$PhylumPch)

# colors to predictions
sortedSampleInfo$PredColor <- sortedSampleInfo$Prediction_core
sortedSampleInfo$PredColor <- as.factor(sortedSampleInfo$PredColor)
table(sortedSampleInfo$PredColor)
levels(sortedSampleInfo$PredColor)[levels(sortedSampleInfo$PredColor)==""]<-"white"
levels(sortedSampleInfo$PredColor)[levels(sortedSampleInfo$PredColor)=="Above prediction"]<-"red"
levels(sortedSampleInfo$PredColor)[levels(sortedSampleInfo$PredColor)=="Below prediction"]<-"blue"
levels(sortedSampleInfo$PredColor)[levels(sortedSampleInfo$PredColor)=="As predicted"]<-"#f3c300"


# Add phylum colours 
sortedSampleInfo$PhylumCol <- sortedSampleInfo$Phylum
sortedSampleInfo$PhylumCol <- as.factor(sortedSampleInfo$PhylumCol)
table(sortedSampleInfo$PhylumCol)
table(Phylum_core$Phylum_core)

table(levels(as.factor(sortedSampleInfo$group_color)))

levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Actinomycetota"]<-"#f3c300"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Bacillota"]<-"#a9c7e2"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Myxococcota"]<-"#8db405"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Planctomycetota"]<-"#f9937a"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Pseudomonadota"]<-"#be0131"

levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Campylobacterota"]<-"#be0131"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Bacillota_A"]<-"#a9c7e2"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Bacillota_C"]<-"#a9c7e2"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Bacillota_D"]<-"#a9c7e2"

levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Dormibacterota"]<-"#f3c300"



levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Acidobacteriota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Bacillota_A"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Bacteroidota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Chloroflexota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Cyanobacteriota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Patescibacteria"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Spirochaetota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Verrucomicrobiota"]<-"grey80"

levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Armatimonadota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Bdellovibrionota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Deinococcota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Desulfobacterota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Dormibacterota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Eremiobacterota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Fusobacteriota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Gemmatimonadota"]<-"grey80"
#levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Pseudomonadota"]<-"grey80"
levels(sortedSampleInfo$PhylumCol)[levels(sortedSampleInfo$PhylumCol)=="Synergistota"]<-"grey80"

table(levels(sortedSampleInfo$PhylumCol))
str(sortedSampleInfo)
# Add genus colours 
sortedSampleInfo$FamilyCol <- sortedSampleInfo$Family_core
sortedSampleInfo$FamilyCol <- as.factor(sortedSampleInfo$FamilyCol)
table(levels(sortedSampleInfo$FamilyCol))
sortedSampleInfo$FamilyCol<-as.character(sortedSampleInfo$FamilyCol)

tree_16S$tip.label<-paste(sortedSampleInfo$OTU_ID,sortedSampleInfo$Phylum_core,sortedSampleInfo$Family_core)

sortedSampleInfo<-cbind(tiplab=tree_16S$tip.label,sortedSampleInfo)

# Add label line colour
sortedSampleInfo$labelCol<-ifelse(is.na(sortedSampleInfo$Family_core),"grey80","grey")
sortedSampleInfo$labelCol[sortedSampleInfo$Family_core=="grey80"]<-"grey80"

# Keep only one label per genus
# indices for labels
middle<-function(x){x[as.integer(length(x)/2+0.5)]}
a <- numeric()
d <- 0
genusList <- unique(sortedSampleInfo$Family_core[!is.na(sortedSampleInfo$Family_core)])
for(genus in genusList){
  x<-which(sortedSampleInfo$Family_core == genus)
  d=d+1
  a<-c(a,middle(x))
}
indices<-1:length(tree_16S$tip.label)
toRemove<-indices[-c(a)]

sortedSampleInfo$Family_core[toRemove] <- NA
table(table(sortedSampleInfo$Family_core))
table(sortedSampleInfo$group_color, sortedSampleInfo$Phylum_core)
table( sortedSampleInfo$Phylum_core)
# Get phylum colours for branches
table(Phylum_core$Phylum_core)


clades <- list(which(sortedSampleInfo$group_color=="#8db405"),
which(sortedSampleInfo$group_color=="#f3c300"),
which(sortedSampleInfo$group_color=="#a9c7e2"),
which(sortedSampleInfo$group_color=="#f38400"),
which(sortedSampleInfo$group_color=="#a7e586"), 
which(sortedSampleInfo$group_color=="grey80"))

str(clades)

treeInfo <- (tree_16S) %>% tidytree::groupOTU(clades) %>% as_tibble

table(treeInfo$group)

require(phylobase)
g1 <- as(tree_16S, 'phylo4')
head(g1)

d = data.frame("color"=treeInfo$group[nodeId(g1, "tip")])
rownames(d) = tree_16S$tip.label
t <- tibble::rownames_to_column(d, "group")
t <- separate(t, 1, into = c("ASV", "Phylum", "Family"), sep = " ")
str(t)
head(d)
table(t$Phylum, t$color)

g2 <- phylo4d(g1, d)
str(g2)
table(g2@label)

rNodeData <- data.frame("color" = treeInfo$group[nodeId(g1, "internal")] , row.names = nodeId(g1, "internal"))

#palGenera <-  c("#A6BDDB","#ADDD8E","#2B8CBE","#FEE391","#D95F02","grey","#276419","#018856","#E5F5F9","#F7FCFD","#EFEDF5","#DF65B0","#7FBC41","#E0F3DB","#E6AB02","#E7D4E8","#FFD92F","#d596ab","#B2ABD2","#FDBB84","#C51B7D","#4575B4","#FEB24C","#0868AC","#006D2C","#FFF7FB","#DFC27D","#FB6A4A","#F7F4F9","#3690C0","#1B9E77","#6BAED6","#8db405", "#313695")

library(RColorBrewer)
n <- length(levels(sortedSampleInfo$FamilyCol))
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'seq',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
palFamily <- col_vector
#pie(rep(1,n), col=sample(col_vector, n))

str(g2)

library(hilldiv)


hierSpLabel <- ps_per_island_core %>%
                sample_data %>% 
                data.frame %>% 
                rownames_to_column("sampleID") %>%
                filter(!is.na(Species_Label)) %>%
                dplyr::select(sampleID, Species_Label)  

head(hierSpLabel)
rownames(hierSpLabel) <- hierSpLabel$sampleID

hierSpLabel$sampleID <- paste0("X", hierSpLabel$sampleID)
str(hierSpLabel)
# ASV relative abundance table2

dfAbund <- data.frame(t(otu_table(ps_per_island_core)))
head(dfAbund)
str(dfAbund)
dfIncidence <- as.data.frame(to.incidence(dfAbund, hierSpLabel[, c(1,2)]))
head(dfIncidence )
dfIncidence$OTU_ID <- rownames(dfIncidence)
names(dfIncidence)
dfIncidence$overall_freq <- rowSums(dfIncidence[,1:11])
dfIncidence <- dfIncidence %>% pivot_longer(cols = c("Flores_F_moreletti":"Tenerife_F_teydea"), names_to = "Species_Label", values_to = "freq")
sample_totals <- data.frame(table(sample_data(hierSpLabel)$Species_Label))
names(sample_totals) <- c("Species_Label", "n_samples")
dfIncidence <- left_join(dfIncidence, sample_totals, by = "Species_Label")
dfIncidence$overall_prevalence <- dfIncidence$overall_freq/nrow(sample_data(ps_per_island_core))
dfIncidence$perSpeciesLabel_prevalence <- dfIncidence$freq/nrow(sample_data(ps_per_island_core))

dfIncidence <- dfIncidence %>% filter(perSpeciesLabel_prevalence > 0)
head(dfIncidence )

str(g2)

dfIncidence <- dfIncidence %>% mutate(HostLabels = str_replace(Species_Label, "Species_Label", "") %>%
              str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%
              str_replace("_F_", " : F.") %>%
              str_replace("c_", "c.")) 
table(dfIncidence$Species_Label)
dfIncidence %>% names
sortedSampleInfo %>% names

s <- sortedSampleInfo %>%  dplyr::select(tiplab, OTU_ID)
str(s)
dfIncidenceTips <- left_join(dfIncidence, s, by ="OTU_ID")
dfIncidenceTips <- dfIncidenceTips %>% select(tiplab, everything())
table(dfIncidenceTips$Species_Label)

META <- as(sample_data(ps_per_island_core), "matrix") %>% as_tibble(rownames = "sampleID") %>% select( sampleID, Species_Label, Archipelago_Mainland, Island_Mainland, Chaff_Blue, Species, Subspecies) %>% mutate( Labels = str_replace(Species_Label, "Species_Label", "") %>%     str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%              str_replace("_F_", " : F.") %>% str_replace("c_", "c.")) 


OTU_META <- left_join(OTUdf, META, by = "sampleID") 
OTU_META %>% names
str(sortedSampleInfo)
s <- sortedSampleInfo %>%  select(tiplab, OTU_ID)
s <- sortedSampleInfo[, c(1,2)]
head(s)
tipMeta <- left_join(s, OTU_META, by ="OTU_ID")
head(tipMeta, 2)

```



```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

toexprItal <- function(x, plain = NULL) {
  getfun <- function(x) {
    ifelse(x == plain, "plain", "italic")
  }
  as.expression(unname(Map(function(f,v) substitute(f(v), list(f=as.name(f), v=as.character(v))), getfun(x), x)))
}

toexprBold <- function(x, plain = NULL) {
  getfun <- function(x) {
    ifelse(x == plain, "plain", "bold")
  }
  as.expression(unname(Map(function(f,v) substitute(f(v), list(f=as.name(f), v=as.character(v))), getfun(x), x)))
}

levels(as.factor(tipMeta$Labels))
summary(tipMeta$Abundance)
table(sortedSampleInfo$Phylum, sortedSampleInfo$group_color)
table(sortedSampleInfo$PhylumCol)
table(Phylum_core$Phylum_core)
table(t$Phylum, t$color)
head(sortedSampleInfo, 2)


SpeciesLabel.order <- c("Tenerife_F_teydea",  "Gran Canaria_F_polatzeki", "Mainland_F_coelebs", "Sao Miguel_F_moreletti",  "Flores_F_moreletti", "Madeira_F_maderensis", "La Palma_F_c_palmae",   "El Hierro_F_c_ombriosa", "Gran Canaria_F_c_bakeri", "Tenerife_F_c_canariensis",  "La Gomera_F_c_canariensis" )

SpeciesLabel.colors <- c( "#A6CEE3",  "#1F78B4", "black","darkred", "darkred", "#E6AB02" ,  "#BAE4B3",  "#006D2C", "#31A354", "#708238",  "#708238")

SpeciesLabel.labels <-c("Tenerife : F.teydea", "Gran Canaria : F.polatzeki", "Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "La Palma : F.c.palmae", "El Hierro : F.c.ombriosa", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis")

#palGenera <- sortedSampleInfo$PredColor
p1 <- ggtree(g2,layout="circular",size=1.5,aes(col=factor(color))) %<+% sortedSampleInfo  +
 scale_color_manual(name = "Core phyla",
    values=c("#f3c300", "#a9c7e2", "#8db405", "#f9937a", "#be0131","grey80"),
     na.value = "grey40",
    labels = c("Actinomycetota","Bacillota", "Myxococcota", "Planctomycetota",  "Pseudomonadota", "< Threshold"),  guide="none") +
  new_scale_colour() +
  geom_tippoint(aes(fill=Prediction_core, shape = Prediction_core), col = "white",size=6) +
  scale_shape_manual(name = "", values = c(`Above prediction` = 24, 
            `As predicted` = 21, `Below prediction` = 25), breaks = c("Above prediction", "As predicted", 
            "Below prediction"), labels = c("","","")) +
  scale_fill_manual(name = "Prediction", values = c(`Above prediction` = "grey40", `As predicted` = "grey40", `Below prediction` = "grey40"), breaks = c("Above prediction", "As predicted", 
            "Below prediction"), labels = c("","",""), guide = "none") +
  new_scale_fill() +
   geom_tiplab2(aes(label=NA),col="lightgrey", align=T, 
               linetype = "12", linesize = 0.1, 
               size=1, offset=0.005, hjust=0)  +
  scale_fill_manual(values=sortedSampleInfo$PredColor,
    labels=toexprItal(levels(factor(sortedSampleInfo$Prediction_core)), plain = "SHVA01"),
    na.value = "white", guide = "none") +
    new_scale_fill() +
  new_scale_color() +
   geom_fruit(geom=geom_tile,width=0.04,
             mapping=aes(fill = PhylumCol),
             alpha=1,offset = 0.06,size = 0.04) +
    scale_fill_manual(breaks = c("#f3c300", "#a9c7e2", "#8db405", "#f9937a", "#be0131","grey80"), values = c("#f3c300", "#a9c7e2", "#8db405", "#f9937a","#be0131", "grey80"), labels = c("Actinomycetota","Bacillota", "Myxococcota", "Planctomycetota",  "Pseudomonadota", "< Threshold"),  guide=guide_legend(keywidth = 1, keysize=2, keyheight = 1, order=5))

p1

p2 <- open_tree(p1, 90)
p3 <- rotate_tree(p2, 0)

dfPrev <- dfIncidenceTips %>% group_by(tiplab) %>% summarise(Prev = max(overall_prevalence))

p4 <- p3  +  new_scale_fill() +
         geom_fruit(data = tipMeta, geom = geom_tile,
                  mapping=aes(y = tiplab, x = factor(Labels), alpha = Abundance, fill = factor(Labels), group = Labels),  color = "grey30", offset = 0.05, size = 0.03, width = 0.05, 
                    axis.params=list(
                         axis       = "x",
                         text.angle = 0, 
                         text.size  = 3,
                         hjust      = - 0.05,
                         vjust      = 0),
                         #nbreak     = 10),
                        
         grid.params=list(vline = FALSE, color ="white"), 
        show.legend = FALSE) +
         scale_alpha_continuous(range=c(0.004, 1),
                             guide=guide_legend(keywidth = 0.3, 
                                             keyheight = 0.3, order=1))  +
         scale_fill_manual(name = "Island / Mainland : Species",
                                   values = SpeciesLabel.colors,
                                   limits = SpeciesLabel.labels,
                                   labels = SpeciesLabel.labels, 
                                   guide = guide_legend(keywidth = 0.3, 
                                        keyheight = 0.3, order=4)
                                   ) 
p4
(p5 <- p3 + new_scale_fill() +
         geom_fruit(data = dfPrev, geom = geom_bar,
                  mapping=aes(y = tiplab, x = Prev), offset = 0.07, size = 0.01,
                    pwidth=0.3, 
                    orientation="y", 
                    stat="identity", 
                    inherit.aes = FALSE, 
                    axis.params=list(
                         axis       = "x",
                         line.size =  0.3,
                         text.angle = 90, 
                         text.size  = 4,
                         hjust      = 0.8,
                         vjust      = 0.15,
                         nbreak     = 5
                     ),
         grid.params=list(vline = FALSE, color ="grey80", linetype = 1, size = 0.5), 
        show.legend = FALSE) )


#p5 <- p4 +  guides(linetype = guide_legend(keywidth = unit(10, 'cm'))) +theme(
#              #legend.key.size = unit(9, "line"),
#              legend.key.size = unit(1, 'cm'), #change legend key size
#              legend.key.height = unit(1, 'cm'), #change legend key height
#               legend.key.width = unit(9, 'cm'),
#              legend.position=c(0.15, 0.85),
#               legend.background=element_rect(fill=NA),
#               legend.title=element_text(size=22),
#               legend.text=element_text(size=20),
#               legend.spacing.y = unit(0.02, "cm"),
#             )

p5
ggsave(paste0(outputDir, "/Figures/treeNeutralModel_familyFixCore.png"), width = 50, height = 50, units = "cm", dpi = 300, limitsize = FALSE)


```

## Neutral Model 


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
microb_fr_fix <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr_family <- speedyseq::tax_glom(microb_fr_fix, "Family",  NArm = FALSE)

tax_table(microb_fr_family) <- tax_table(microb_fr_family)[,1:5]
tax_table(microb_fr_family) <- cbind(tax_table(microb_fr_family), rownames(tax_table(microb_fr_family)))
colnames(tax_table(microb_fr_family)) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "OTU_ID")

microb_fr_meta_family <- sample_data(microb_fr_family) %>% data.frame



# core all ids
core_all <- readRDS(paste0(dataDir, "/ps_across_island_core_allTaxa.rds" ))

core_family <- core_all[[4]]
microb_tax_family <- tax_table(microb_fr_family) %>% as.data.frame
core_tax_across_family <- tax_table(core_family) %>% data.frame
core_tax_across_family$Core_across <- rep("Core_across", nrow(tax_table(core_tax_across_family)))
core_tax_across_family <- core_tax_across_family %>% dplyr::select("Family", "Core_across")
microb_tax_core_family <- left_join(microb_tax_family, core_tax_across_family, by = "Family")
microb_tax_core_family$Core_across <- ifelse(is.na(microb_tax_core_family$Core_across) == FALSE, "Core_across", "Not_core")
microb_tax_core_family <- microb_tax_core_family %>% column_to_rownames("OTU_ID") 
tax_table(microb_fr_family) <- as.matrix(microb_tax_core_family)

```


fit neutral model
#https://github.com/DanielSprockett/tyRa/blob/master/R/modeling.R

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# tyRa function OTUs in columns and samples in rows
pnm_family <- tyRa::fit_sncm(spp = otu_table(microb_fr_family)@.Data, pool=NULL, taxon=data.frame(tax_table(microb_fr_family)))
(pnm_family_plot <- plot_sncm_fit_col(pnm_family, fill = NULL) + pca_theme  + guides(shape = guide_legend(override.aes = list(size = 5)), fill = guide_legend(override.aes = list(size = 5)), size = guide_legend(override.aes = list(size = c(5, 3) , fill= c("black")))))


#png(paste0(outputDir, "/Figures/PNM_acrossCore.png") ,width=1500, height=1200,res=200)
#pnm_plot + theme( legend.position= "none")
#dev.off()

```

### PANEL B

Annotated paco output

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

paco_output <- read.csv(paste0(dataDir, "/PACo_familyFix_annotated.csv"), sep = "\t", header = TRUE)
str(paco_output)
names(paco_output)
table(paco_output$Phylum)

paco_output$prediction <- ifelse(paco_output$fit_class == "As predicted", "Neutral", "Non-neutral")
#paco_output$prediction <- ifelse(paco_output$fit_class == "As predicted", "Neutral", ifelse(paco_output$fit_class == "Above prediction", "Above", "Below"))

```

Plot PACo annotate

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


library(tidyverse)
library(glue)
library(ggtext)

df <- paco_output %>%
  mutate(
    BoldLabel = if_else(p_adj_bonferroni > 0.05, glue("<p>{Family_edit}</p>"), glue("<b>{Family_edit}*</b>")))


paco_label <- df %>%
    mutate(BoldLabel = BoldLabel %>% reorder(R2),
           BoldLabel = BoldLabel %>% reorder(Phylum %>% as.factor %>% as.numeric)
           )

paco_label$prediction 
df_txt <- data.frame(
    variable = c("Neutral", "Non-neutral"),
    x = c(2, 2, 2, 2),
    y = c(1, 1, 1, 1),
    label = paste0("Text in facet ", 1:4))

(p_tax_annot <- ggplot(paco_label) +
    geom_point(aes(BoldLabel, R2, fill=Phylum), stroke=1.2, show.legend = T, shape=22, size=4, alpha = 1) +
     xlab("") +
     scale_fill_manual(name = "Phylum", values = c(`Actinomycetota` = "#f3c300", `Bacillota` = "#a9c7e2", `Planctomycetota` = "#f9937a",`Pseudomonadota` = "#be0131", `Myxococcota` = "#8db405"), breaks = c("Actinomycetota", "Bacillota", "Myxococcota", "Planctomycetota", "Pseudomonadota"), labels = c("Actinomycetota", "Bacillota", "Myxococcota", "Planctomycetota", "Pseudomonadota"  )) +
        guides(color = guide_legend(override.aes = list(linewidth = 12))) +
        theme(legend.position="none", 
          axis.title.x = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          axis.text.x = element_text(size = 16, angle = 45, hjust=1),
          #axis.text.x = element_text(angle = 90,  size = 12), 
          axis.text.y = ggtext::element_markdown(angle = 0,  size = 16), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16), 
          panel.spacing = unit(2, "lines"),
          strip.placement = "outside",
          strip.background = element_rect(fill = "white", color = "black"),
          strip.text = element_text(size=18, color = "black"))+
        facet_grid(factor(prediction)~.,  scales = "free", space = "free", switch = "y") +
        coord_flip()+
        ylab(bquote("PACo"~ R^2)))
  

(p_sppcount <- ggplot(paco_label, aes(x = BoldLabel, y= n_Spp_overall)) +
    geom_bar(stat ="identity", fill = "grey", alpha = 0.9) +
     xlab("") +
    guides(color = guide_legend(override.aes = list(linewidth = 12))) +
      theme(legend.position="none", 
          axis.title.x = element_text(size = 18, angle = 0),
          axis.title.y = element_text(size = 18),
          axis.text.x = element_text(size = 16, angle = 45, hjust =1), 
          axis.text.y = element_blank(), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16), 
          panel.spacing = unit(2, "lines"),
          strip.placement =  "outside",
          strip.background = element_rect(fill = "white", color = "white"),
          strip.text = element_blank())+
        facet_grid(factor(prediction)~.,  scales = "free", space = "free", switch = "y") +
        coord_flip() + 
        ylab("N species"))

# Side by side
(p_combined <- p_tax_annot + p_sppcount +
  plot_layout(widths = c(3, 3))
)
png(paste0(outputDir, "/Figures/paco_taxplot_familyFix.png") ,width=720, height=1000,res=100)
p_combined + theme( legend.position= "none")
dev.off()

```



### PANEL C, D and E

Phyloseq object
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr_fix <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr <- microb_fr_fix

microb_fr <- biomeUtils::removeZeros(microb_fr)

# Add TipID microbiome to match stuff later on
sample_data(microb_fr)$TipID_microbiome <- paste(sample_data(microb_fr)$Lab_Code, sample_data(microb_fr)$Species_Label, sep = "_")

microb_fr_meta <- microb_fr%>% sample_data %>% data.frame 
nrow(microb_fr_meta)


```

core Microbacteriaceae family signif PACo

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_Microbacteriaceae <- microb_fr %>% microViz::tax_select(strict_matches= TRUE, ranks_searched = "Family", tax_list = "Microbacteriaceae")

ps_Microbacteriaceae <- speedyseq::tax_glom(ps_Microbacteriaceae, "Species",  NArm = FALSE)

tax_table(ps_Microbacteriaceae) %>% data.frame %>% head(2)
tax_table(ps_Microbacteriaceae) <- tax_table(ps_Microbacteriaceae)[,1:7]
tax_table(ps_Microbacteriaceae) <- cbind(tax_table(ps_Microbacteriaceae), 
    rownames(tax_table(ps_Microbacteriaceae)))
colnames(tax_table(ps_Microbacteriaceae)) <- 
  c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species","OTU_ID")

```

prep data 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu_wide <- otu_table(ps_Microbacteriaceae) %>% data.frame
otu_wide$sampleID <- rownames(otu_wide) 
head(otu_wide)

otu_long <- otu_wide %>% 
  pivot_longer(cols = starts_with("OTU"), names_to = "OTU_ID", values_to = "count") %>% filter(count > 0) %>% data.frame
otu_long$Species_Label <- sub('.*?[-_]', "", otu_long$sampleID)
str(otu_long)

otu_long_s <- otu_long %>%  group_by(OTU_ID, Species_Label) %>% dplyr::select(OTU_ID, Species_Label) %>% dplyr::slice(1)
unique(otu_long_s$OTU_ID)

table(otu_long$OTU_ID)

Strain_df <- data.frame(OTU_ID = as.data.frame(rownames(tax_table(ps_Microbacteriaceae))), Spp = as.data.frame(tax_table(ps_Microbacteriaceae))$Species)
head(Strain_df)
(names(Strain_df) <- c("OTU_ID", "Spp"))
table(Strain_df$Spp)

# metadata 
TAXdf <- as(tax_table(ps_Microbacteriaceae), "matrix")  %>% as_tibble()
TAXdf <- left_join(TAXdf, Strain_df, by = "OTU_ID")
TAXdf %>% data.frame %>% head

# Extract relative abundance matrix from the phyloseq object

ps_Microbacteriaceae <- transform_sample_counts(ps_Microbacteriaceae, function(x) x / sum(x))

OTUdf <-  as(otu_table(ps_Microbacteriaceae), "matrix") %>%   as_tibble(rownames = "sampleID") %>% pivot_longer(!sampleID, names_to = "OTU_ID", values_to = "Abundance") %>% filter(Abundance != 0) 

is.nan.data.frame <- function(x)
do.call(cbind, lapply(x, is.nan))

mic_df <- (otu_table(ps_Microbacteriaceae)) %>% data.frame
mic_df[is.nan(mic_df)] <- 0

Prevalence <- data.frame(Prevalence = colSums(mic_df > 0), tax_table(ps_Microbacteriaceae)) %>%  dplyr::select(Prevalence) %>% as_tibble(rownames = "OTU_ID") 


Prevalence <- Prevalence %>% filter(Prevalence != 0)
Prevalence

OTUdf <- left_join(OTUdf, Prevalence, by = "OTU_ID" )

Microbact_meta <- ps_Microbacteriaceae %>% sample_data %>% data.frame %>% rownames_to_column("sampleID") %>% dplyr::select(sampleID, Species_Label, Archipelago_Mainland, Chaff_Blue) 
head(Microbact_meta)

OTUdf <- left_join(OTUdf, Microbact_meta, by = "sampleID")
is.na(table(OTUdf$Species_Label))

OTUdf <- left_join(OTUdf, TAXdf, by = "OTU_ID")

OTUdf_summ <- OTUdf %>% dplyr::group_by(Archipelago_Mainland, Species_Label, Chaff_Blue,OTU_ID) %>% summarise(RelAbund_mean = mean(Abundance), RelAbund_median = median(Abundance), Prevalence = max(Prevalence))
table( OTUdf_summ$OTU_ID)

topOTU <- OTUdf_summ %>%  arrange(desc(Prevalence)) #%>% head(10)

SPPdf_summ <- OTUdf %>% dplyr::group_by(Species_Label, Species) %>% summarise(RelAbund_mean = mean(Abundance), RelAbund_median = median(Abundance), Prevalence = max(Prevalence))
table(SPPdf_summ$Species)
#write.csv(SPPdf_summ, paste0(outputDir, "/01_Exploration/Streptococcus_spp.csv"))

topSPP <- SPPdf_summ %>%  arrange(desc(Prevalence)) #%>% head(10)

# wide otu
hostDF <- topOTU %>% group_by(Species_Label, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Species_Label, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- hostDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Species_Label", "OTU_ID")], by = "OTU_ID")

hostDF <- left_join(hostDF, uniqueDF[, c("Species_Label", "OTU_ID")], by = "OTU_ID")
hostDF$HostLabel <- ifelse(is.na(hostDF$Species_Label), hostDF$host, hostDF$Species_Label)

table(hostDF$HostLabel)

# wide otu
archipDF <- topOTU %>% group_by(Archipelago_Mainland, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Archipelago_Mainland, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- archipDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Archipelago_Mainland", "OTU_ID")], by = "OTU_ID")
uniqueDF <- uniqueDF %>% distinct(OTU_ID, Archipelago_Mainland, .keep_all = TRUE)


archipDF <- left_join(archipDF, uniqueDF[, c("Archipelago_Mainland", "OTU_ID")], by = "OTU_ID")
archipDF$archipLabel <- ifelse(is.na(archipDF$Archipelago_Mainland), archipDF$host, archipDF$Archipelago_Mainland)

table(hostDF$HostLabel)


# wide otu
blueDF <- topOTU %>% group_by(Chaff_Blue, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Chaff_Blue, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- blueDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Chaff_Blue", "OTU_ID")], by = "OTU_ID")
uniqueDF <- uniqueDF %>% distinct(OTU_ID, Chaff_Blue, .keep_all = TRUE)


blueDF <- left_join(blueDF, uniqueDF[, c("Chaff_Blue", "OTU_ID")], by = "OTU_ID")
blueDF$blueLabel <- ifelse(is.na(blueDF$Chaff_Blue), blueDF$host, blueDF$Chaff_Blue)

table(blueDF$blueLabel)



# Tree

tree_16S <- phy_tree(ps_Microbacteriaceae)
treetips <- as.character(tree_16S$tip.label)

# Reorder the sample information to match the order in the tree labels
treetipsorder <- cbind(treetips,"treenames"=treetips,"indexx"=c(100:(length(treetips)+99)))
head(treetipsorder)


mergedSampleInfo <- merge(TAXdf,treetipsorder, by.x="OTU_ID",by.y="treetips",all.y=T)
head(mergedSampleInfo)
sortedSampleInfo <- mergedSampleInfo[order(mergedSampleInfo$indexx),]
sortedSampleInfo <- droplevels(sortedSampleInfo)
sortedSampleInfo <- unique(x = sortedSampleInfo)
library(stringr)
sortedSampleInfo$Spp <- sub("\\(.*)", "", sortedSampleInfo$Spp)
sortedSampleInfo <- cbind(sortedSampleInfo,"group_color" =sortedSampleInfo$Genus)
sortedSampleInfo$group_color <- as.factor(sortedSampleInfo$group_color)
sortedSampleInfo$Genus_color <- sortedSampleInfo$Genus
sortedSampleInfo$Genus_color <- as.factor(sortedSampleInfo$Genus_color)

head(sortedSampleInfo)
str(sortedSampleInfo)
nrow(table(sortedSampleInfo$Genus))
#41


levels(sortedSampleInfo$Genus_color) <- rep("black", nrow(table(sortedSampleInfo$Genus)))

sortedSampleInfo <- droplevels(sortedSampleInfo)
sortedSampleInfo <- unique(x=sortedSampleInfo)
sortedSampleInfo %>% str

tree_16S$tip.label <- paste(sortedSampleInfo$OTU_ID,sortedSampleInfo$Spp)
sortedSampleInfo <- cbind(tiplab=tree_16S$tip.label,sortedSampleInfo)

clades <- list(which(sortedSampleInfo$Genus_color=="black"))


treeInfo <- (tree_16S) %>% tidytree::groupOTU(clades) %>% as_tibble

require(phylobase)
g1 <- as(tree_16S, 'phylo4')
head(g1)
labels(g1)


d = data.frame("color"=treeInfo$group[nodeId(g1, "tip")])
rownames(d) = tree_16S$tip.label
t <- tibble::rownames_to_column(d, "group")
t <- separate(t, 1, into = c("OTU_ID", "Genus", "Species"), sep = " ")
head(t)
table(t$Genus, t$color)

g2 <- phylo4d(g1, d)

rNodeData <- data.frame("color" = treeInfo$group[nodeId(g1, "internal")] , row.names = nodeId(g1, "internal"))


META <- as(sample_data(ps_Microbacteriaceae), "matrix") %>% as_tibble(rownames = "sampleID") %>% dplyr::select( sampleID, Species_Label, Archipelago_Mainland, Island_Mainland, Chaff_Blue, Species, Subspecies) %>% mutate( Labels = str_replace(Species_Label, "Species_Label", "") %>%     str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%              str_replace("_F_", " : F.") %>% str_replace("c_", "c.")) 

#META$Host_order <-factor(META$Labels,levels=c("Tenerife : F.teydea", "Gran Canaria : F.polatzeki", "Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "La Palma : F.c.palmae", "El Hierro : F.c.ombriosa", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis"))
#names(META)
META <- dplyr::rename(META,c('Host_species' = 'Species'))
names(OTUdf)
names(META)

OTU_META <- left_join(OTUdf[,-c(5,6)], META, by = "sampleID") 
OTU_META %>% names
head(OTU_META)
s <- sortedSampleInfo %>%  dplyr::select(tiplab, OTU_ID)
head(s)
tipMeta <- left_join(s, OTU_META, by ="OTU_ID")
head(tipMeta, 2)

tipMeta <- left_join(tipMeta, hostDF[,c("OTU_ID", "HostLabel")], by = "OTU_ID")
tipMeta <- left_join(tipMeta, archipDF[,c("OTU_ID", "archipLabel")], by = "OTU_ID")
tipMeta <- left_join(tipMeta, blueDF[,c("OTU_ID", "blueLabel")], by = "OTU_ID")

head(tipMeta, 2)
sortedSampleInfo <- left_join(sortedSampleInfo, hostDF[,c("OTU_ID", "HostLabel")], by = "OTU_ID")
names(sortedSampleInfo)

```


tree


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


(p0 <- ggtree(g2,layout="fan",  branch.length="none", size=1,col="black") %<+% sortedSampleInfo  +
      new_scale_fill() +
        geom_tiplab2(aes(label=NA),col="lightgrey", align=T, 
               linetype = "12", linesize = 0, 
               size=1, offset=0.05, hjust=0) +
    scale_color_manual(name = "Species", values = sortedSampleInfo$Genus_color,guide = "none") +
     new_scale_colour() +
     geom_tippoint(aes(fill=HostLabel), shape =21, col = "black",size=4) +
   scale_fill_manual(name = "HostLabel", breaks = c("El_Hierro_F_c_ombriosa", "Flores_F_moreletti", "Gran_Canaria_F_c_bakeri", "Gran_Canaria_F_polatzeki", "La_Gomera_F_c_canariensis", "La_Palma_F_c_palmae", "Madeira_F_maderensis" , "Mainland_F_coelebs", "Sao_Miguel_F_moreletti", "Tenerife_F_c_canariensis", "Tenerife_F_teydea", "shared"), values = c("#006D2C", "darkred", "#31A354", "#1F78B4", "#708238", "#BAE4B3", "#E6AB02" , "black", "darkred", "#708238", "#A6CEE3", "white" ), labels = c("El Hierro: *F.c. ombriosa*","Flores: *F. moreletti*","Gran Canaria: *F.c. bakeri*","Gran Canaria: *F. polatzeki*","La Gomera: *F.c. canariensis*","La Palma: *F.c. palmae*","Madeira: *F. maderensis*","Mainland: *F. coelebs*","Sao Miguel: *F. moreletti*","Tenerife: *F.c. canariensis*","Tenerife: *F. teydea*", "Shared")) + 
  theme(legend.text = element_markdown()))




(p0 <- ggtree(g2,layout="fan",  branch.length="none", size=1,col="black") %<+% sortedSampleInfo  +
      new_scale_fill() +
        geom_tiplab2(aes(label=NA),col="lightgrey", align=T, 
               linetype = "12", linesize = 0, 
               size=1, offset=0.05, hjust=0) +
    scale_color_manual(name = "Species", values = sortedSampleInfo$Genus_color,guide = "none") +
     new_scale_colour() +
     geom_tippoint(aes(fill=HostLabel,  shape = HostLabel),  col = "black",size=4) +
      scale_shape_manual(name = "Host species", breaks = c( "Mainland_F_coelebs","Sao_Miguel_F_moreletti","Flores_F_moreletti", "Madeira_F_maderensis", "El_Hierro_F_c_ombriosa", "La_Palma_F_c_palmae", "Gran_Canaria_F_c_bakeri","Tenerife_F_c_canariensis", "La_Gomera_F_c_canariensis", "Tenerife_F_teydea", "Gran_Canaria_F_polatzeki", "shared" ), values = c(21,21, 22, 21,21,25,23,22,23,25, 21, 21), labels = c("Mainland: *F. coelebs*", "Sao Miguel: *F. moreletti*", "Flores: *F. moreletti*",  "Madeira: *F. maderensis*",  "El Hierro: *F.c. ombriosa*", "La Palma: *F.c. palmae*", "Gran Canaria: *F.c. bakeri*", "Tenerife: *F.c. canariensis*", "La Gomera: *F.c. canariensis*", "Tenerife: *F. teydea*",  "Gran Canaria: *F. polatzeki*", "Shared")) +
  scale_fill_manual(name = "Host species", breaks = c( "Mainland_F_coelebs","Sao_Miguel_F_moreletti","Flores_F_moreletti", "Madeira_F_maderensis", "El_Hierro_F_c_ombriosa", "La_Palma_F_c_palmae", "Gran_Canaria_F_c_bakeri","Tenerife_F_c_canariensis", "La_Gomera_F_c_canariensis", "Tenerife_F_teydea", "Gran_Canaria_F_polatzeki", "shared" ), values = c("black","darkred", "darkred", "#E6AB02" , "#006D2C", "#BAE4B3", "#31A354", "#708238", "#708238", "#A6CEE3", "#1F78B4", "white"), labels = c("Mainland: *F. coelebs*", "Sao Miguel: *F. moreletti*", "Flores: *F. moreletti*",  "Madeira: *F. maderensis*",  "El Hierro: *F.c. ombriosa*", "La Palma: *F.c. palmae*", "Gran Canaria: *F.c. bakeri*", "Tenerife: *F.c. canariensis*", "La Gomera: *F.c. canariensis*", "Tenerife: *F. teydea*",  "Gran Canaria: *F. polatzeki*", "Shared"))+
  theme(legend.text = element_markdown()))

p1 <- open_tree(p0, 20)
p2 <- rotate_tree(p1, 170)

(pMicrobacteriaceae <- p2  + 
   new_scale_color() + new_scale_fill() +
   geom_fruit(data = tipMeta, geom=geom_tile,width=0.6, color = "black",
             mapping=aes(y = tiplab, fill = archipLabel),
             alpha=1,offset = 0.05,size = 0.4) +
   scale_fill_manual(name = "Archipelago / Mainland", breaks = c("Azores", "Canary Islands", "Madeira", "Mainland", "shared"), values = c("darkred",  "forestgreen", "#E6AB02", "black" ,"white" ), labels = c("Azores", "Canary Islands", "Madeira", "Mainland", "Shared"))+
       new_scale_color() + new_scale_fill() +
   geom_fruit(data = tipMeta, geom=geom_tile,width=0.7, color = "black",
             mapping=aes(y = tiplab, fill = blueLabel),
             alpha=1,offset = 0.13,size = 0.4) +
   scale_fill_manual(name = "Species complex", breaks = c("Bluefinch", "Chaffinch", "shared"), values = c("#2596be",  "#898b80", "white" ), labels = c("Bluechaffinch", "Chaffinch", "Shared"))+ theme(legend.text=element_markdown(size = 18), legend.title = element_markdown(size = 18))) 




#ggsave(paste0(outputDir, "/Figures/tree_Microbacteriaceae.png"), width = 50, height = 50, units = "cm", dpi = 300, limitsize = FALSE)


```



core Burkholderiaceae family signif PACo

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_Burkholderiaceae <- microb_fr %>% tax_select(strict_matches= TRUE, ranks_searched = "Family", tax_list = "Burkholderiaceae")

ps_Burkholderiaceae <- speedyseq::tax_glom(ps_Burkholderiaceae, "Species",  NArm = FALSE)

tax_table(ps_Burkholderiaceae) %>% data.frame %>% head(2)
tax_table(ps_Burkholderiaceae) <- tax_table(ps_Burkholderiaceae)[,1:7]
tax_table(ps_Burkholderiaceae) <- cbind(tax_table(ps_Burkholderiaceae), 
    rownames(tax_table(ps_Burkholderiaceae)))
colnames(tax_table(ps_Burkholderiaceae)) <- 
  c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species","OTU_ID")

```

prep data 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu_wide <- otu_table(ps_Burkholderiaceae) %>% data.frame
otu_wide$sampleID <- rownames(otu_wide) 
head(otu_wide)

otu_long <- otu_wide %>% 
  pivot_longer(cols = starts_with("OTU"), names_to = "OTU_ID", values_to = "count") %>% filter(count > 0) %>% data.frame
otu_long$Species_Label <- sub('.*?[-_]', "", otu_long$sampleID)
str(otu_long)

otu_long_s <- otu_long %>%  group_by(OTU_ID, Species_Label) %>% dplyr::select(OTU_ID, Species_Label) %>% dplyr::slice(1)
unique(otu_long_s$OTU_ID)

table(otu_long$OTU_ID)

Strain_df <- data.frame(OTU_ID = as.data.frame(rownames(tax_table(ps_Burkholderiaceae))), Spp = as.data.frame(tax_table(ps_Burkholderiaceae))$Species)
head(Strain_df)
(names(Strain_df) <- c("OTU_ID", "Spp"))
table(Strain_df$Spp)

# metadata 
TAXdf <- as(tax_table(ps_Burkholderiaceae), "matrix")  %>% as_tibble()
TAXdf <- left_join(TAXdf, Strain_df, by = "OTU_ID")
TAXdf %>% data.frame %>% head

# Extract relative abundance matrix from the phyloseq object

ps_Burkholderiaceae <- transform_sample_counts(ps_Burkholderiaceae, function(x) x / sum(x))

OTUdf <-  as(otu_table(ps_Burkholderiaceae), "matrix") %>%   as_tibble(rownames = "sampleID") %>% pivot_longer(!sampleID, names_to = "OTU_ID", values_to = "Abundance") %>% filter(Abundance != 0) 

is.nan.data.frame <- function(x)
do.call(cbind, lapply(x, is.nan))

mic_df <- (otu_table(ps_Burkholderiaceae)) %>% data.frame
mic_df[is.nan(mic_df)] <- 0

Prevalence <- data.frame(Prevalence = colSums(mic_df > 0), tax_table(ps_Burkholderiaceae)) %>%  dplyr::select(Prevalence) %>% as_tibble(rownames = "OTU_ID") 

Prevalence <- Prevalence %>% filter(Prevalence != 0)
Prevalence

OTUdf <- left_join(OTUdf, Prevalence, by = "OTU_ID" )

Burkho_meta <- ps_Burkholderiaceae %>% sample_data %>% data.frame %>% rownames_to_column("sampleID") %>% dplyr::select(sampleID, Species_Label, Archipelago_Mainland, Chaff_Blue) 
head(Burkho_meta)

OTUdf <- left_join(OTUdf, Burkho_meta, by = "sampleID")
is.na(table(OTUdf$Species_Label))

OTUdf <- left_join(OTUdf, TAXdf, by = "OTU_ID")

OTUdf_summ <- OTUdf %>% dplyr::group_by(Archipelago_Mainland, Species_Label, Chaff_Blue,OTU_ID) %>% summarise(RelAbund_mean = mean(Abundance), RelAbund_median = median(Abundance), Prevalence = max(Prevalence))
table( OTUdf_summ$OTU_ID)

topOTU <- OTUdf_summ %>%  arrange(desc(Prevalence)) #%>% head(10)

SPPdf_summ <- OTUdf %>% dplyr::group_by(Species_Label, Species) %>% summarise(RelAbund_mean = mean(Abundance), RelAbund_median = median(Abundance), Prevalence = max(Prevalence))
table(SPPdf_summ$Species)
#write.csv(SPPdf_summ, paste0(outputDir, "/01_Exploration/Streptococcus_spp.csv"))

topSPP <- SPPdf_summ %>%  arrange(desc(Prevalence)) #%>% head(10)

# wide otu
hostDF <- topOTU %>% group_by(Species_Label, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Species_Label, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- hostDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Species_Label", "OTU_ID")], by = "OTU_ID")

hostDF <- left_join(hostDF, uniqueDF[, c("Species_Label", "OTU_ID")], by = "OTU_ID")
hostDF$HostLabel <- ifelse(is.na(hostDF$Species_Label), hostDF$host, hostDF$Species_Label)

table(hostDF$HostLabel)

# wide otu
archipDF <- topOTU %>% group_by(Archipelago_Mainland, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Archipelago_Mainland, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- archipDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Archipelago_Mainland", "OTU_ID")], by = "OTU_ID")
uniqueDF <- uniqueDF %>% distinct(OTU_ID, Archipelago_Mainland, .keep_all = TRUE)


archipDF <- left_join(archipDF, uniqueDF[, c("Archipelago_Mainland", "OTU_ID")], by = "OTU_ID")
archipDF$archipLabel <- ifelse(is.na(archipDF$Archipelago_Mainland), archipDF$host, archipDF$Archipelago_Mainland)

table(hostDF$HostLabel)


# wide otu
blueDF <- topOTU %>% group_by(Chaff_Blue, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Chaff_Blue, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- blueDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Chaff_Blue", "OTU_ID")], by = "OTU_ID")
uniqueDF <- uniqueDF %>% distinct(OTU_ID, Chaff_Blue, .keep_all = TRUE)


blueDF <- left_join(blueDF, uniqueDF[, c("Chaff_Blue", "OTU_ID")], by = "OTU_ID")
blueDF$blueLabel <- ifelse(is.na(blueDF$Chaff_Blue), blueDF$host, blueDF$Chaff_Blue)

table(blueDF$blueLabel)


# Tree

tree_16S <- phy_tree(ps_Burkholderiaceae)
treetips <- as.character(tree_16S$tip.label)

# Reorder the sample information to match the order in the tree labels
treetipsorder <- cbind(treetips,"treenames"=treetips,"indexx"=c(100:(length(treetips)+99)))
head(treetipsorder)


mergedSampleInfo <- merge(TAXdf,treetipsorder, by.x="OTU_ID",by.y="treetips",all.y=T)
head(mergedSampleInfo)
sortedSampleInfo <- mergedSampleInfo[order(mergedSampleInfo$indexx),]
sortedSampleInfo <- droplevels(sortedSampleInfo)
sortedSampleInfo <- unique(x = sortedSampleInfo)
library(stringr)
sortedSampleInfo$Spp <- sub("\\(.*)", "", sortedSampleInfo$Spp)
sortedSampleInfo <- cbind(sortedSampleInfo,"group_color" =sortedSampleInfo$Genus)
sortedSampleInfo$group_color <- as.factor(sortedSampleInfo$group_color)
sortedSampleInfo$Genus_color <- sortedSampleInfo$Genus
sortedSampleInfo$Genus_color <- as.factor(sortedSampleInfo$Genus_color)

head(sortedSampleInfo)
str(sortedSampleInfo)
nrow(table(sortedSampleInfo$Genus))
#41


levels(sortedSampleInfo$Genus_color) <- rep("black", nrow(table(sortedSampleInfo$Genus)))

sortedSampleInfo <- droplevels(sortedSampleInfo)
sortedSampleInfo <- unique(x=sortedSampleInfo)
sortedSampleInfo %>% str

tree_16S$tip.label <- paste(sortedSampleInfo$OTU_ID,sortedSampleInfo$Spp)
sortedSampleInfo <- cbind(tiplab=tree_16S$tip.label,sortedSampleInfo)

clades <- list(which(sortedSampleInfo$Genus_color=="black"))


treeInfo <- (tree_16S) %>% tidytree::groupOTU(clades) %>% as_tibble

require(phylobase)
g1 <- as(tree_16S, 'phylo4')
head(g1)
labels(g1)


d = data.frame("color"=treeInfo$group[nodeId(g1, "tip")])
rownames(d) = tree_16S$tip.label
t <- tibble::rownames_to_column(d, "group")
t <- separate(t, 1, into = c("OTU_ID", "Genus", "Species"), sep = " ")
head(t)
table(t$Genus, t$color)

g2 <- phylo4d(g1, d)

rNodeData <- data.frame("color" = treeInfo$group[nodeId(g1, "internal")] , row.names = nodeId(g1, "internal"))


META <- as(sample_data(ps_Burkholderiaceae), "matrix") %>% as_tibble(rownames = "sampleID") %>% dplyr::select( sampleID, Species_Label, Archipelago_Mainland, Island_Mainland, Chaff_Blue, Species, Subspecies) %>% mutate( Labels = str_replace(Species_Label, "Species_Label", "") %>%     str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%              str_replace("_F_", " : F.") %>% str_replace("c_", "c.")) 

#META$Host_order <-factor(META$Labels,levels=c("Tenerife : F.teydea", "Gran Canaria : F.polatzeki", "Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "La Palma : F.c.palmae", "El Hierro : F.c.ombriosa", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis"))
#names(META)
META <- dplyr::rename(META,c('Host_species' = 'Species'))
names(OTUdf)
names(META)

OTU_META <- left_join(OTUdf[,-c(5,6)], META, by = "sampleID") 
OTU_META %>% names
head(OTU_META)
s <- sortedSampleInfo %>%  dplyr::select(tiplab, OTU_ID)
head(s)
tipMeta <- left_join(s, OTU_META, by ="OTU_ID")
head(tipMeta, 2)

tipMeta <- left_join(tipMeta, hostDF[,c("OTU_ID", "HostLabel")], by = "OTU_ID")
tipMeta <- left_join(tipMeta, archipDF[,c("OTU_ID", "archipLabel")], by = "OTU_ID")
tipMeta <- left_join(tipMeta, blueDF[,c("OTU_ID", "blueLabel")], by = "OTU_ID")

head(tipMeta, 2)
sortedSampleInfo <- left_join(sortedSampleInfo, hostDF[,c("OTU_ID", "HostLabel")], by = "OTU_ID")
names(sortedSampleInfo)

```


tree


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


(p0 <- ggtree(g2,layout="fan",  branch.length="none", size=1,col="black") %<+% sortedSampleInfo  +
      new_scale_fill() +
        geom_tiplab2(aes(label=NA),col="lightgrey", align=T, 
               linetype = "12", linesize = 0, 
               size=1, offset=0.05, hjust=0) +
    scale_color_manual(name = "Species", values = sortedSampleInfo$Genus_color,guide = "none") +
     new_scale_colour() +
     geom_tippoint(aes(fill=HostLabel), shape =21, col = "black",size=4) +
   scale_fill_manual(name = "HostLabel", breaks = c("El_Hierro_F_c_ombriosa", "Flores_F_moreletti", "Gran_Canaria_F_c_bakeri", "Gran_Canaria_F_polatzeki", "La_Gomera_F_c_canariensis", "La_Palma_F_c_palmae", "Madeira_F_maderensis" , "Mainland_F_coelebs", "Sao_Miguel_F_moreletti", "Tenerife_F_c_canariensis", "Tenerife_F_teydea", "shared"), values = c("#006D2C", "darkred", "#31A354", "#1F78B4", "#708238", "#BAE4B3", "#E6AB02" , "black", "darkred", "#708238", "#A6CEE3", "white" ), labels = c("El Hierro: *F.c. ombriosa*","Flores: *F. moreletti*","Gran Canaria: *F.c. bakeri*","Gran Canaria: *F. polatzeki*","La Gomera: *F.c. canariensis*","La Palma: *F.c. palmae*","Madeira: *F. maderensis*","Mainland: *F. coelebs*","Sao Miguel: *F. moreletti*","Tenerife: *F.c. canariensis*","Tenerife: *F. teydea*", "Shared")) +
  theme(legend.text = element_markdown())) 


(p0 <- ggtree(g2,layout="fan",  branch.length="none", size=1,col="black") %<+% sortedSampleInfo  +
      new_scale_fill() +
        geom_tiplab2(aes(label=NA),col="lightgrey", align=T, 
               linetype = "12", linesize = 0, 
               size=1, offset=0.05, hjust=0) +
    scale_color_manual(name = "Species", values = sortedSampleInfo$Genus_color,guide = "none") +
     new_scale_colour() +
     geom_tippoint(aes(fill=HostLabel,  shape = HostLabel),  col = "black",size=4) +
      scale_shape_manual(name = "Host species", breaks = c( "Mainland_F_coelebs","Sao_Miguel_F_moreletti","Flores_F_moreletti", "Madeira_F_maderensis", "El_Hierro_F_c_ombriosa", "La_Palma_F_c_palmae", "Gran_Canaria_F_c_bakeri","Tenerife_F_c_canariensis", "La_Gomera_F_c_canariensis", "Tenerife_F_teydea", "Gran_Canaria_F_polatzeki", "shared" ), values = c(21,21, 22, 21,21,25,23,22,23,25, 21, 21), labels = c("Mainland: *F. coelebs*", "Sao Miguel: *F. moreletti*", "Flores: *F. moreletti*",  "Madeira: *F. maderensis*",  "El Hierro: *F.c. ombriosa*", "La Palma: *F.c. palmae*", "Gran Canaria: *F.c. bakeri*", "Tenerife: *F.c. canariensis*", "La Gomera: *F.c. canariensis*", "Tenerife: *F. teydea*",  "Gran Canaria: *F. polatzeki*", "Shared")) +
  scale_fill_manual(name = "Host species", breaks = c( "Mainland_F_coelebs","Sao_Miguel_F_moreletti","Flores_F_moreletti", "Madeira_F_maderensis", "El_Hierro_F_c_ombriosa", "La_Palma_F_c_palmae", "Gran_Canaria_F_c_bakeri","Tenerife_F_c_canariensis", "La_Gomera_F_c_canariensis", "Tenerife_F_teydea", "Gran_Canaria_F_polatzeki", "shared" ), values = c("black","darkred", "darkred", "#E6AB02" , "#006D2C", "#BAE4B3", "#31A354", "#708238", "#708238", "#A6CEE3", "#1F78B4", "white"), labels = c("Mainland: *F. coelebs*", "Sao Miguel: *F. moreletti*", "Flores: *F. moreletti*",  "Madeira: *F. maderensis*",  "El Hierro: *F.c. ombriosa*", "La Palma: *F.c. palmae*", "Gran Canaria: *F.c. bakeri*", "Tenerife: *F.c. canariensis*", "La Gomera: *F.c. canariensis*", "Tenerife: *F. teydea*",  "Gran Canaria: *F. polatzeki*", "Shared"))+
  theme(legend.text = element_markdown()))


p1 <- open_tree(p0, 20)
p2 <- rotate_tree(p1, 170)

(pBurkholderiaceae <- p2  + 
   new_scale_color() + new_scale_fill() +
   geom_fruit(data = tipMeta, geom=geom_tile,width=0.6, color = "black",
             mapping=aes(y = tiplab, fill = archipLabel),
             alpha=1,offset = 0.05,size = 0.4) +
   scale_fill_manual(name = "Archipelago / Mainland", breaks = c("Azores", "Canary Islands", "Madeira", "Mainland", "shared"), values = c("darkred",  "forestgreen", "#E6AB02", "black" ,"white" ), labels = c("Azores", "Canary Islands", "Madeira", "Mainland", "Shared"))+
       new_scale_color() + new_scale_fill() +
   geom_fruit(data = tipMeta, geom=geom_tile,width=0.7, color = "black",
             mapping=aes(y = tiplab, fill = blueLabel),
             alpha=1,offset = 0.13,size = 0.4) +
   scale_fill_manual(name = "Species complex", breaks = c("Bluefinch", "Chaffinch", "shared"), values = c("#2596be",  "#898b80", "white" ), labels = c("Bluechaffinch", "Chaffinch", "Shared"))) 


#ggsave(paste0(outputDir, "/Figures/tree_Burkholderiaceae.png"), width = 50, height = 50, units = "cm", dpi = 300, limitsize = FALSE)


```


core Lactobacillaceae family signif PACo

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_Lactobacillaceae <- microb_fr %>% tax_select(strict_matches= TRUE, ranks_searched = "Family", tax_list = "Lactobacillaceae")

ps_Lactobacillaceae <- speedyseq::tax_glom(ps_Lactobacillaceae, "Species",  NArm = FALSE)

tax_table(ps_Lactobacillaceae) %>% data.frame %>% head(2)
tax_table(ps_Lactobacillaceae) <- tax_table(ps_Lactobacillaceae)[,1:7]
tax_table(ps_Lactobacillaceae) <- cbind(tax_table(ps_Lactobacillaceae), 
    rownames(tax_table(ps_Lactobacillaceae)))
colnames(tax_table(ps_Lactobacillaceae)) <- 
  c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species","OTU_ID")

```

prep data 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu_wide <- otu_table(ps_Lactobacillaceae) %>% data.frame
otu_wide$sampleID <- rownames(otu_wide) 
head(otu_wide)

otu_long <- otu_wide %>% 
  pivot_longer(cols = starts_with("OTU"), names_to = "OTU_ID", values_to = "count") %>% filter(count > 0) %>% data.frame
otu_long$Species_Label <- sub('.*?[-_]', "", otu_long$sampleID)
str(otu_long)

otu_long_s <- otu_long %>%  group_by(OTU_ID, Species_Label) %>% dplyr::select(OTU_ID, Species_Label) %>% dplyr::slice(1)
unique(otu_long_s$OTU_ID)

table(otu_long$OTU_ID)

Strain_df <- data.frame(OTU_ID = as.data.frame(rownames(tax_table(ps_Lactobacillaceae))), Spp = as.data.frame(tax_table(ps_Lactobacillaceae))$Species)
head(Strain_df)
(names(Strain_df) <- c("OTU_ID", "Spp"))
table(Strain_df$Spp)

# metadata 
TAXdf <- as(tax_table(ps_Lactobacillaceae), "matrix")  %>% as_tibble()
TAXdf <- left_join(TAXdf, Strain_df, by = "OTU_ID")
TAXdf %>% data.frame %>% head

# Extract relative abundance matrix from the phyloseq object

ps_Lactobacillaceae <- transform_sample_counts(ps_Lactobacillaceae, function(x) x / sum(x))

OTUdf <-  as(otu_table(ps_Lactobacillaceae), "matrix") %>%   as_tibble(rownames = "sampleID") %>% pivot_longer(!sampleID, names_to = "OTU_ID", values_to = "Abundance") %>% filter(Abundance != 0) 

is.nan.data.frame <- function(x)
do.call(cbind, lapply(x, is.nan))

mic_df <- (otu_table(ps_Lactobacillaceae)) %>% data.frame
mic_df[is.nan(mic_df)] <- 0

Prevalence <- data.frame(Prevalence = colSums(mic_df > 0), tax_table(ps_Lactobacillaceae)) %>%  dplyr::select(Prevalence) %>% as_tibble(rownames = "OTU_ID") 

Prevalence <- Prevalence %>% filter(Prevalence != 0)
Prevalence

OTUdf <- left_join(OTUdf, Prevalence, by = "OTU_ID" )

Lacto_meta <- ps_Lactobacillaceae %>% sample_data %>% data.frame %>% rownames_to_column("sampleID") %>% dplyr::select(sampleID, Species_Label, Archipelago_Mainland, Chaff_Blue) 
head(Lacto_meta)

OTUdf <- left_join(OTUdf, Lacto_meta, by = "sampleID")
is.na(table(OTUdf$Species_Label))

OTUdf <- left_join(OTUdf, TAXdf, by = "OTU_ID")

OTUdf_summ <- OTUdf %>% dplyr::group_by(Archipelago_Mainland, Species_Label, Chaff_Blue,OTU_ID) %>% summarise(RelAbund_mean = mean(Abundance), RelAbund_median = median(Abundance), Prevalence = max(Prevalence))
table( OTUdf_summ$OTU_ID)

topOTU <- OTUdf_summ %>%  arrange(desc(Prevalence)) #%>% head(10)

SPPdf_summ <- OTUdf %>% dplyr::group_by(Species_Label, Species) %>% summarise(RelAbund_mean = mean(Abundance), RelAbund_median = median(Abundance), Prevalence = max(Prevalence))
table(SPPdf_summ$Species)
#write.csv(SPPdf_summ, paste0(outputDir, "/01_Exploration/Streptococcus_spp.csv"))

topSPP <- SPPdf_summ %>%  arrange(desc(Prevalence)) #%>% head(10)

# wide otu
hostDF <- topOTU %>% group_by(Species_Label, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Species_Label, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- hostDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Species_Label", "OTU_ID")], by = "OTU_ID")

hostDF <- left_join(hostDF, uniqueDF[, c("Species_Label", "OTU_ID")], by = "OTU_ID")
hostDF$HostLabel <- ifelse(is.na(hostDF$Species_Label), hostDF$host, hostDF$Species_Label)

table(hostDF$HostLabel)

# wide otu
archipDF <- topOTU %>% group_by(Archipelago_Mainland, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Archipelago_Mainland, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- archipDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Archipelago_Mainland", "OTU_ID")], by = "OTU_ID")
uniqueDF <- uniqueDF %>% distinct(OTU_ID, Archipelago_Mainland, .keep_all = TRUE)


archipDF <- left_join(archipDF, uniqueDF[, c("Archipelago_Mainland", "OTU_ID")], by = "OTU_ID")
archipDF$archipLabel <- ifelse(is.na(archipDF$Archipelago_Mainland), archipDF$host, archipDF$Archipelago_Mainland)

table(hostDF$HostLabel)


# wide otu
blueDF <- topOTU %>% group_by(Chaff_Blue, OTU_ID) %>% 
    summarise(total = sum(Prevalence)) %>% 
    pivot_wider(names_from = Chaff_Blue, values_from = total, values_fill= 0) %>%
    mutate(across(where(is.numeric), .fns = function(x) ifelse(x > 0, 1, 0))) %>% 
    mutate(total = rowSums(across(where(is.numeric)))) %>%
    mutate(host = ifelse(total > 1, "shared", "unique")) %>% 
    dplyr::select("OTU_ID", "host")

uniqueDF <- blueDF %>% dplyr::filter(host == "unique")

uniqueDF <- left_join(uniqueDF, topOTU[, c("Chaff_Blue", "OTU_ID")], by = "OTU_ID")
uniqueDF <- uniqueDF %>% distinct(OTU_ID, Chaff_Blue, .keep_all = TRUE)


blueDF <- left_join(blueDF, uniqueDF[, c("Chaff_Blue", "OTU_ID")], by = "OTU_ID")
blueDF$blueLabel <- ifelse(is.na(blueDF$Chaff_Blue), blueDF$host, blueDF$Chaff_Blue)

table(blueDF$blueLabel)


# Tree

tree_16S <- phy_tree(ps_Lactobacillaceae)
treetips <- as.character(tree_16S$tip.label)

# Reorder the sample information to match the order in the tree labels
treetipsorder <- cbind(treetips,"treenames"=treetips,"indexx"=c(100:(length(treetips)+99)))
head(treetipsorder)


mergedSampleInfo <- merge(TAXdf,treetipsorder, by.x="OTU_ID",by.y="treetips",all.y=T)
head(mergedSampleInfo)
sortedSampleInfo <- mergedSampleInfo[order(mergedSampleInfo$indexx),]
sortedSampleInfo <- droplevels(sortedSampleInfo)
sortedSampleInfo <- unique(x = sortedSampleInfo)
library(stringr)
sortedSampleInfo$Spp <- sub("\\(.*)", "", sortedSampleInfo$Spp)
sortedSampleInfo <- cbind(sortedSampleInfo,"group_color" =sortedSampleInfo$Genus)
sortedSampleInfo$group_color <- as.factor(sortedSampleInfo$group_color)
sortedSampleInfo$Genus_color <- sortedSampleInfo$Genus
sortedSampleInfo$Genus_color <- as.factor(sortedSampleInfo$Genus_color)

head(sortedSampleInfo)
str(sortedSampleInfo)
nrow(table(sortedSampleInfo$Genus))
#41


levels(sortedSampleInfo$Genus_color) <- rep("black", nrow(table(sortedSampleInfo$Genus)))

sortedSampleInfo <- droplevels(sortedSampleInfo)
sortedSampleInfo <- unique(x=sortedSampleInfo)
sortedSampleInfo %>% str

tree_16S$tip.label <- paste(sortedSampleInfo$OTU_ID,sortedSampleInfo$Spp)
sortedSampleInfo <- cbind(tiplab=tree_16S$tip.label,sortedSampleInfo)

clades <- list(which(sortedSampleInfo$Genus_color=="black"))


treeInfo <- (tree_16S) %>% tidytree::groupOTU(clades) %>% as_tibble

require(phylobase)
g1 <- as(tree_16S, 'phylo4')
head(g1)
labels(g1)


d = data.frame("color"=treeInfo$group[nodeId(g1, "tip")])
rownames(d) = tree_16S$tip.label
t <- tibble::rownames_to_column(d, "group")
t <- separate(t, 1, into = c("OTU_ID", "Genus", "Species"), sep = " ")
head(t)
table(t$Genus, t$color)

g2 <- phylo4d(g1, d)

rNodeData <- data.frame("color" = treeInfo$group[nodeId(g1, "internal")] , row.names = nodeId(g1, "internal"))


META <- as(sample_data(ps_Lactobacillaceae), "matrix") %>% as_tibble(rownames = "sampleID") %>% dplyr::select( sampleID, Species_Label, Archipelago_Mainland, Island_Mainland, Chaff_Blue, Species, Subspecies) %>% mutate( Labels = str_replace(Species_Label, "Species_Label", "") %>%     str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%              str_replace("_F_", " : F.") %>% str_replace("c_", "c.")) 

#META$Host_order <-factor(META$Labels,levels=c("Tenerife : F.teydea", "Gran Canaria : F.polatzeki", "Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "La Palma : F.c.palmae", "El Hierro : F.c.ombriosa", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis"))
#names(META)
META <- dplyr::rename(META,c('Host_species' = 'Species'))
names(OTUdf)
names(META)

OTU_META <- left_join(OTUdf[,-c(5,6)], META, by = "sampleID") 
OTU_META %>% names
head(OTU_META)
s <- sortedSampleInfo %>%  dplyr::select(tiplab, OTU_ID)
head(s)
tipMeta <- left_join(s, OTU_META, by ="OTU_ID")
head(tipMeta, 2)

tipMeta <- left_join(tipMeta, hostDF[,c("OTU_ID", "HostLabel")], by = "OTU_ID")
tipMeta <- left_join(tipMeta, archipDF[,c("OTU_ID", "archipLabel")], by = "OTU_ID")
tipMeta <- left_join(tipMeta, blueDF[,c("OTU_ID", "blueLabel")], by = "OTU_ID")

head(tipMeta, 2)
sortedSampleInfo <- left_join(sortedSampleInfo, hostDF[,c("OTU_ID", "HostLabel")], by = "OTU_ID")
names(sortedSampleInfo)
str(tipMeta)
#write.csv(tipMeta, paste0(dataDir, "/tipMeta_Lacto.csv"))

```


tree


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


(p0 <- ggtree(g2,layout="fan",  branch.length="none", size=1,col="black") %<+% sortedSampleInfo  +
      new_scale_fill() +
        geom_tiplab2(aes(label=NA),col="lightgrey", align=T, 
               linetype = "12", linesize = 0, 
               size=1, offset=0.05, hjust=0) +
    scale_color_manual(name = "Species", values = sortedSampleInfo$Genus_color,guide = "none") +
     new_scale_colour() +
     geom_tippoint(aes(fill=HostLabel), shape =21, col = "black",size=4) +
   scale_fill_manual(name = "HostLabel", breaks = c("El_Hierro_F_c_ombriosa", "Flores_F_moreletti", "Gran_Canaria_F_c_bakeri", "Gran_Canaria_F_polatzeki", "La_Gomera_F_c_canariensis", "La_Palma_F_c_palmae", "Madeira_F_maderensis" , "Mainland_F_coelebs", "Sao_Miguel_F_moreletti", "Tenerife_F_c_canariensis", "Tenerife_F_teydea", "shared"), values = c("#006D2C", "darkred", "#31A354", "#1F78B4", "#708238", "#BAE4B3", "#E6AB02" , "black", "darkred", "#708238", "#A6CEE3", "white" ), labels = c("El Hierro: *F.c. ombriosa*","Flores: *F. moreletti*","Gran Canaria: *F.c. bakeri*","Gran Canaria: *F. polatzeki*","La Gomera: *F.c. canariensis*","La Palma: *F.c. palmae*","Madeira: *F. maderensis*","Mainland: *F. coelebs*","Sao Miguel: *F. moreletti*","Tenerife: *F.c. canariensis*","Tenerife: *F. teydea*", "Shared"))+
  theme(legend.text = element_markdown())) 


(p0 <- ggtree(g2,layout="fan",  branch.length="none", size=1,col="black") %<+% sortedSampleInfo  +
      new_scale_fill() +
        geom_tiplab2(aes(label=NA),col="lightgrey", align=T, 
               linetype = "12", linesize = 0, 
               size=1, offset=0.05, hjust=0) +
    scale_color_manual(name = "Species", values = sortedSampleInfo$Genus_color,guide = "none") +
     new_scale_colour() +
     geom_tippoint(aes(fill=HostLabel,  shape = HostLabel),  col = "black",size=4) +
      scale_shape_manual(name = "Host species", breaks = c( "Mainland_F_coelebs","Sao_Miguel_F_moreletti","Flores_F_moreletti", "Madeira_F_maderensis", "El_Hierro_F_c_ombriosa", "La_Palma_F_c_palmae", "Gran_Canaria_F_c_bakeri","Tenerife_F_c_canariensis", "La_Gomera_F_c_canariensis", "Tenerife_F_teydea", "Gran_Canaria_F_polatzeki", "shared" ), values = c(21,21, 22, 21,21,25,23,22,23,25, 21, 21), labels = c("Mainland: *F. coelebs*", "Sao Miguel: *F. moreletti*", "Flores: *F. moreletti*",  "Madeira: *F. maderensis*",  "El Hierro: *F.c. ombriosa*", "La Palma: *F.c. palmae*", "Gran Canaria: *F.c. bakeri*", "Tenerife: *F.c. canariensis*", "La Gomera: *F.c. canariensis*", "Tenerife: *F. teydea*",  "Gran Canaria: *F. polatzeki*", "Shared")) +
  scale_fill_manual(name = "Host species", breaks = c( "Mainland_F_coelebs","Sao_Miguel_F_moreletti","Flores_F_moreletti", "Madeira_F_maderensis", "El_Hierro_F_c_ombriosa", "La_Palma_F_c_palmae", "Gran_Canaria_F_c_bakeri","Tenerife_F_c_canariensis", "La_Gomera_F_c_canariensis", "Tenerife_F_teydea", "Gran_Canaria_F_polatzeki", "shared" ), values = c("black","darkred", "darkred", "#E6AB02" , "#006D2C", "#BAE4B3", "#31A354", "#708238", "#708238", "#A6CEE3", "#1F78B4", "white"), labels = c("Mainland: *F. coelebs*", "Sao Miguel: *F. moreletti*", "Flores: *F. moreletti*",  "Madeira: *F. maderensis*",  "El Hierro: *F.c. ombriosa*", "La Palma: *F.c. palmae*", "Gran Canaria: *F.c. bakeri*", "Tenerife: *F.c. canariensis*", "La Gomera: *F.c. canariensis*", "Tenerife: *F. teydea*",  "Gran Canaria: *F. polatzeki*", "Shared")) + 
  theme(legend.text = element_markdown()))


p1 <- open_tree(p0, 20)
p2 <- rotate_tree(p1, 170)

(pLactobacillaceae <- p2  + 
   new_scale_color() + new_scale_fill() +
   geom_fruit(data = tipMeta, geom=geom_tile,width=0.6, color = "black",
             mapping=aes(y = tiplab, fill = archipLabel),
             alpha=1,offset = 0.05,size = 0.4) +
   scale_fill_manual(name = "Archipelago / Mainland", breaks = c("Azores", "Canary Islands", "Madeira", "Mainland", "shared"), values = c("darkred",  "forestgreen", "#E6AB02", "black" ,"white" ), labels = c("Azores", "Canary Islands", "Madeira", "Mainland", "Shared"))+
       new_scale_color() + new_scale_fill() +
   geom_fruit(data = tipMeta, geom=geom_tile,width=0.7, color = "black",
             mapping=aes(y = tiplab, fill = blueLabel),
             alpha=1,offset = 0.13,size = 0.4) +
   scale_fill_manual(name = "Species complex", breaks = c("Bluefinch", "Chaffinch", "shared"), values = c("#2596be",  "#898b80", "white" ), labels = c("Bluechaffinch", "Chaffinch", "Shared"))) 

#ggsave(paste0(outputDir, "/Figures/tree_Lactobacillaceae.png"), width = 50, height = 50, units = "cm", dpi = 300, limitsize = FALSE)


```



### Patchwork

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

library(patchwork)

(p5b <- p5 + theme(legend.text=element_text(size = 18), legend.title = element_text(size = 18)) + guides(fill = "Phylum"))


png(paste0(outputDir, "/Figures/Figure5_patchwork_a.png") ,width=5900, height=4000,res=300)
(p5b + inset_element(pnm_family_plot, left = 0.56, bottom = 0.1125, right = 0.95,top = 0.42)) + plot_layout(guides = 'collect') & theme(plot.tag = element_text(size = 36),plot.margin = unit(c(-2,-2,-2,-2), "cm"))
dev.off()


png(paste0(outputDir, "/Figures/Figure5_patchwork_b.png") ,width=2900, height=4000,res=300)
p_tax_annot | p_sppcount
dev.off()


png(paste0(outputDir, "/Figures/Figure5_patchwork_c.png") ,width=8800, height=2500,res=300)

(pBurkholderiaceae + ggtitle("Burkholderiaceae") + theme(legend.position="none",plot.title = element_text(vjust = -5,hjust = 0.5, size = 28)) + pLactobacillaceae + ggtitle("Lactobacillaceae") + theme(legend.position="none",plot.title = element_text(vjust = -5,hjust = 0.5, size = 28)) + pMicrobacteriaceae + ggtitle("Microbacteriaceae") + theme(plot.title = element_text(vjust = -5,hjust = 0.5, size = 28)) + plot_layout(guides = 'collect'))

dev.off()

```

