---
title: 'Effect of colonisation history and biogeography'
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute:
    code_folding: hide
    toc_depth: 2
---


## LOAD LIBRARIES

```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rm(list = ls())

# basics
library(tibble)
library(dplyr)
library(tidyverse)
library(data.table)
library(ggtern)
library(patchwork)

# phyloseq section
library(phyloseq)
library(Biostrings)
library(microbiome)
library(biomeUtils)
library(microViz)

# diversity 
library(adespatial)


#rmd
library(details)
library(knitr)
library(kableExtra)

#models
library(MASS)
library(MuMIn)
library(car)
library(multcomp)
library(glmmTMB)
library(ggeffects)
```

## WORKING DIRECTORIES

```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```

## SHARED LABELLING SCHEME

```{r labelling, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

order_levels <- c(
  "Mainland : F.coelebs",
  "Sao Miguel : F.moreletti",
  "Flores : F.moreletti",
  "Madeira : F.maderensis",
  "El Hierro : F.c.ombriosa",
  "La Palma : F.c.palmae",
  "Gran Canaria : F.c.bakeri",
  "Tenerife : F.c.canariensis",
  "La Gomera : F.c.canariensis",
  "Tenerife : F.teydea",
  "Gran Canaria : F.polatzeki"
)

short_labels <- c(
  "Mainland : F.coelebs"           = "Mainland",
  "Sao Miguel : F.moreletti"       = "Sao Miguel",
  "Flores : F.moreletti"           = "Flores",
  "Madeira : F.maderensis"         = "Madeira",
  "El Hierro : F.c.ombriosa"       = "El Hierro",
  "La Palma : F.c.palmae"          = "La Palma",
  "Gran Canaria : F.c.bakeri"      = "Gran Canaria",
  "Tenerife : F.c.canariensis"     = "Tenerife",
  "La Gomera : F.c.canariensis"    = "La Gomera",
  "Tenerife : F.teydea"            = "Tenerife",
  "Gran Canaria : F.polatzeki"     = "Gran Canaria"
)

species_complex_lookup <- c(
  "Mainland : F.coelebs"           = "Common chaffinch species complex",
  "Sao Miguel : F.moreletti"       = "Common chaffinch species complex",
  "Flores : F.moreletti"           = "Common chaffinch species complex",
  "Madeira : F.maderensis"         = "Common chaffinch species complex",
  "El Hierro : F.c.ombriosa"       = "Common chaffinch species complex",
  "La Palma : F.c.palmae"          = "Common chaffinch species complex",
  "Gran Canaria : F.c.bakeri"      = "Common chaffinch species complex",
  "Tenerife : F.c.canariensis"     = "Common chaffinch species complex",
  "La Gomera : F.c.canariensis"    = "Common chaffinch species complex",
  "Tenerife : F.teydea"            = "Blue finches",
  "Gran Canaria : F.polatzeki"     = "Blue finches"
)

order_to_species_label <- c(
  "Mainland : F.coelebs"           = "Mainland_F_coelebs",
  "Sao Miguel : F.moreletti"       = "Sao Miguel_F_moreletti",
  "Flores : F.moreletti"           = "Flores_F_moreletti",
  "Madeira : F.maderensis"         = "Madeira_F_maderensis",
  "El Hierro : F.c.ombriosa"       = "El Hierro_F_c_ombriosa",
  "La Palma : F.c.palmae"          = "La Palma_F_c_palmae",
  "Gran Canaria : F.c.bakeri"      = "Gran Canaria_F_c_bakeri",
  "Tenerife : F.c.canariensis"     = "Tenerife_F_c_canariensis",
  "La Gomera : F.c.canariensis"    = "La Gomera_F_c_canariensis",
  "Tenerife : F.teydea"            = "Tenerife_F_teydea",
  "Gran Canaria : F.polatzeki"     = "Gran Canaria_F_polatzeki"
)


SpeciesLabel.order_local <- c(
  "Mainland_F_coelebs",
  "Sao Miguel_F_moreletti",
  "Flores_F_moreletti",
  "Madeira_F_maderensis",
  "El Hierro_F_c_ombriosa",
  "La Palma_F_c_palmae",
  "Gran Canaria_F_c_bakeri",
  "Tenerife_F_c_canariensis",
  "La Gomera_F_c_canariensis",
  "Tenerife_F_teydea",
  "Gran Canaria_F_polatzeki"
)

SpeciesLabel.colors_local <- c(
  "Mainland_F_coelebs"           = "black",
  "Sao Miguel_F_moreletti"       = "darkred",
  "Flores_F_moreletti"           = "darkred",
  "Madeira_F_maderensis"         = "#E6AB02",
  "El Hierro_F_c_ombriosa"       = "#006D2C",
  "La Palma_F_c_palmae"          = "#BAE4B3",
  "Gran Canaria_F_c_bakeri"      = "#31A354",
  "Tenerife_F_c_canariensis"     = "#708238",
  "La Gomera_F_c_canariensis"    = "#708238",
  "Tenerife_F_teydea"            = "#A6CEE3",
  "Gran Canaria_F_polatzeki"     = "#1F78B4"
)

SpeciesLabel.fill_local <- scale_fill_manual(
  name = "Island / Mainland",
  values = SpeciesLabel.colors_local,
  limits = SpeciesLabel.order_local,
  guide = "none"
)

```

## LOAD 16S DATA

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_fringilla_rare <- readRDS(paste0(dataDir, "/microb_fr_corrected.rds"))
ps_fringilla_rare <- biomeUtils::removeZeros(ps_fringilla_rare)
ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Species")
ps_fringilla_rare <- biomeUtils::removeZeros(ps_fringilla_rare)

```


```{r helpers, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Convert a p-value to significance stars
p_to_stars <- function(p) {
  ifelse(is.na(p), "",
    ifelse(p < 0.001, "***",
      ifelse(p < 0.01, "**",
        ifelse(p < 0.05, "*", ""))))
}

# Build the significance stars vector from a glm.nb fit,
# matched to the order_levels vector
build_signif_stars <- function(model) {
  ci <- as_tibble(confint(model), rownames = 'predictor')
  coefs <- as_tibble(summary(model)$coefficients)
  
  tab <- cbind(ci, coefs)
  
  # Back-transform on log scale
  tab$CImin <- c(tab$`2.5 %`[1], tab$Estimate[1] + tab$`2.5 %`[-1])
  tab$CImax <- c(tab$`97.5 %`[1], tab$Estimate[1] + tab$`97.5 %`[-1])
  tab$Estimate.transformed <- c(tab$Estimate[1], tab$Estimate[1] + tab$Estimate[-1])
  tab$Estimate <- exp(tab$Estimate.transformed)
  tab$CImin <- exp(tab$CImin)
  tab$CImax <- exp(tab$CImax)
  
  tab %>%
    mutate(island = predictor %>%
             str_replace("\\(Intercept\\)", "Mainland : F.coelebs") %>%
             str_replace("Species_Label", "") %>%
             str_replace("_F_", " : F.") %>%
             str_replace("c_", "c.") %>%
             str_replace("_", " ")) %>%
    mutate(Order = factor(island, levels = order_levels),
           p.signif = p_to_stars(`Pr(>|z|)`)) %>%
    # Intercept (Mainland) is always blank
    mutate(p.signif = ifelse(predictor == "(Intercept)", "", p.signif)) %>%
    arrange(Order) %>%
    dplyr::select(Order, Estimate, CImin, CImax, `Pr(>|z|)`, p.signif)
}

# Build a parsed annotation string for biogeography plots
# Format: "atop(paste(~beta~' = 0.157 [0.002, 0.311];'),~italic('p-value')~'= 0.047')"
build_biogeo_annot <- function(model, term) {
  ci <- confint(model)
  est <- summary(model)$coefficients$cond[term, "Estimate"]
  p <- summary(model)$coefficients$cond[term, "Pr(>|z|)"]
  cimin <- ci[term, 1]
  cimax <- ci[term, 2]
  
  p_str <- if (p < 0.001) "'< 0.001'" else paste0("'= ", formatC(p, format = "f", digits = 3), "'")
  
  paste0("atop(paste(~beta~' = ", formatC(est, format = "f", digits = 3),
         " [", formatC(cimin, format = "f", digits = 3), ", ",
         formatC(cimax, format = "f", digits = 3), "];'),~italic('p-value')~", p_str, ")")
}

```

## TESTING THE EFFECT OF COLONISATION HISTORY

### ALPHA DIVERSITY

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

rich <- estimate_richness(ps_fringilla_rare, measures = c("Observed", "Shannon"))
pd <- calculatePD(ps_fringilla_rare, justDF = TRUE)

sample_data(ps_fringilla_rare)$Observed_sps_rare <- rich$Observed
sample_data(ps_fringilla_rare)$PD_sps_rare <- pd$PD
sample_data(ps_fringilla_rare)$PD_integer <- round(pd$PD, 0)

alpha_df <- data.frame(sample_data(ps_fringilla_rare))

mainland_fringilla <- alpha_df %>% dplyr::filter(!is.na(Age_cat) & !is.na(Sex))
mainland_fringilla$Species_Label <- as.factor(mainland_fringilla$Species_Label)
mainland_fringilla$Species_Label <- relevel(mainland_fringilla$Species_Label, "Mainland_F_coelebs")

GLM1 <- glm.nb(Observed_sps_rare ~ Age_cat + Sex + Species_Label, data = mainland_fringilla, na.action = na.fail)
summary(GLM1)
dredge(GLM1)
Anova(GLM1, type = "III", test = "F")

alpha_df$Species_Label <- as.factor(alpha_df$Species_Label)
alpha_df$Species_Label <- relevel(alpha_df$Species_Label, "Mainland_F_coelebs")

GLM2 <- glm.nb(Observed_sps_rare ~ Species_Label, data = alpha_df, na.action = na.fail)
summary(GLM2)
dredge(GLM2)
Anova(GLM2, type = "III")
mcp <- glht(GLM2, linfct = mcp(Species_Label = "Tukey"))
summary(mcp)

CI <- as_tibble(confint(GLM2), rownames = 'predictor') 
CI <- cbind(CI, as_tibble(summary(GLM2)$coefficients))
CI$CImin <- c(CI$`2.5 %`[1], CI$Estimate[1] + CI$`2.5 %`[c(-1)])
CI$CImax <- c(CI$`97.5 %`[1], CI$Estimate[1] + CI$`97.5 %`[c(-1)])
CI$Estimate.transformed <- c(CI$`Estimate`[1], CI$`Estimate`[1] + CI$`Estimate`[c(-1)])
CI$Estimate <- exp(CI$Estimate.transformed)
CI$CImin <- exp(CI$CImin)
CI$CImax <- exp(CI$CImax)
CI <- CI[c(1, 4, 5, 6, 7, 8, 9)]

write_tsv(CI, paste0(outputDir, "/07_Stats/02_analyses/glm_rare_sobs_speciesFix.tsv"))

```

### Build alpha_df with shared Order factor + sample-size labels

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

alpha_df <- alpha_df %>% 
  filter(Genus == "Fringilla") %>%
  mutate(Labels = Species_Label %>%
           str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%
           str_replace("_F_", " : F.") %>%
           str_replace("c_", "c.") %>%
           str_replace("_", " "))

alpha_df$Order <- factor(alpha_df$Labels, levels = order_levels)
alpha_df$SpeciesComplex <- factor(species_complex_lookup[as.character(alpha_df$Order)],
                                  levels = c("Common chaffinch species complex", "Blue finches"))

sample_sizes <- alpha_df %>%
  group_by(Order) %>%
  summarise(n = n(), .groups = "drop")

labels_with_n <- sapply(order_levels, function(lvl) {
  n <- sample_sizes$n[match(lvl, sample_sizes$Order)]
  paste0(short_labels[lvl], "\nn = ", n)
})

estimate_obj <- read_tsv(paste0(outputDir, "/07_Stats/02_analyses/glm_rare_sobs_speciesFix.tsv"))

model_estimates <- estimate_obj %>%
  mutate(value = Estimate + estimate_obj[[1, 2]]) %>%
  mutate(value = ifelse(predictor == '(Intercept)', Estimate, value)) %>%
  mutate(island = predictor %>%
           str_replace("as.factor", "") %>%
           str_replace("\\(Species_Label\\)", "") %>%
           str_replace("\\(Intercept\\)", "Mainland : F.coelebs") %>%
           str_replace("Sexmale", "Male") %>%
           str_replace("Age_catfledging", "Fledging") %>%
           str_replace("_F_", " : F.") %>%
           str_replace("c_", "c.") %>%
           str_replace("_", " "))

species_estimates <- model_estimates
species_estimates$Order <- factor(species_estimates$island, levels = order_levels)

# Build CI table with p-values for richness GLM
rich_ci_table <- as_tibble(confint(GLM2), rownames = 'predictor') %>%
  cbind(as_tibble(summary(GLM2)$coefficients))

signif2 <- build_signif_stars(GLM2)
signif2 <- left_join(
  distinct(alpha_df, SpeciesComplex, Order),
  signif2,
  by = "Order"
) %>% arrange(SpeciesComplex, Order)

```

### p1: Richness boxplot

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

(p1 <- ggplot(alpha_df, aes(Order, Observed_sps_rare)) +
    facet_grid(~ SpeciesComplex, scales = "free_x", space = "free_x") +
    geom_boxplot(aes(fill = Species_Label), show.legend = FALSE,
                 lwd = 0.25, fatten = 1, alpha = 0.5,
                 outlier.colour = 'transparent', width = 0.5) +
    geom_point(aes(fill = Species_Label), shape = 21,
               position = position_jitter(seed = 100, width = 0.1),
               size = 3, alpha = 0.7, stroke = 0.25) +
    SpeciesLabel.fill +
    ylab("Richness") + xlab("") +
    sl_theme + boxplot_theme +
    scale_x_discrete(breaks = names(labels_with_n), labels = labels_with_n) +
    geom_hline(aes(yintercept = signif2$Estimate[1]), linetype = 'dashed') +
    stat_summary(aes(x = Order, y = Observed_sps_rare),
                 fun.data = "mean_cl_boot", geom = "pointrange",
                 shape = 23, col = "#acacac", fill = '#acacac',
                 size = 0.5, fun.args = list(conf.int = .95)) +
    geom_text(data = signif2, aes(x = Order, y = 250, label = p.signif),
              vjust = 0.75, size = 8) +
    scale_y_continuous(breaks = seq(0, 250, by = 50)))

```

### Faith's PD analogue

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

GLM1 <- glm.nb(PD_integer ~ Age_cat + Sex + Species_Label, data = mainland_fringilla, na.action = na.fail)
summary(GLM1)
Anova(GLM1, type = "III", test = "F")

GLM2 <- glm.nb(PD_integer ~ Species_Label, data = alpha_df, na.action = na.fail)
summary(GLM2)
Anova(GLM2, type = "III", test = "F")

CI <- as_tibble(confint(GLM2), rownames = 'predictor')
CI <- cbind(CI, as_tibble(summary(GLM2)$coefficients))
CI$CImin <- c(CI$`2.5 %`[1], CI$Estimate[1] + CI$`2.5 %`[c(-1)])
CI$CImax <- c(CI$`97.5 %`[1], CI$Estimate[1] + CI$`97.5 %`[c(-1)])
CI$Estimate.transformed <- c(CI$`Estimate`[1], CI$`Estimate`[1] + CI$`Estimate`[c(-1)])
CI$Estimate <- exp(CI$Estimate.transformed)
CI$CImin <- exp(CI$CImin)
CI$CImax <- exp(CI$CImax)
CI <- CI[c(1, 4, 5, 6, 7, 8, 9)]

write_tsv(CI, paste0(outputDir, "/07_Stats/02_analyses/glm_rare_pd_speciesFix.tsv"))

estimate_obj_pd <- read_tsv(paste0(outputDir, "/07_Stats/02_analyses/glm_rare_pd_speciesFix.tsv"))

species_estimates_pd <- estimate_obj_pd %>%
  mutate(value = Estimate + estimate_obj_pd[[1, 2]]) %>%
  mutate(value = ifelse(predictor == '(Intercept)', Estimate, value)) %>%
  mutate(island = predictor %>%
           str_replace("as.factor", "") %>%
           str_replace("\\(Species_Label\\)", "") %>%
           str_replace("\\(Intercept\\)", "Mainland : F.coelebs") %>%
           str_replace("Sexmale", "Male") %>%
           str_replace("Age_catfledging", "Fledging") %>%
           str_replace("_F_", " : F.") %>%
           str_replace("c_", "c.") %>%
           str_replace("_", " "))

species_estimates_pd$Order <- factor(species_estimates_pd$island, levels = order_levels)

pd_ci_table <- as_tibble(confint(GLM2), rownames = 'predictor') %>%
  cbind(as_tibble(summary(GLM2)$coefficients))

signif2_pd <- build_signif_stars(GLM2)
signif2_pd <- left_join(
  distinct(alpha_df, SpeciesComplex, Order),
  signif2_pd,
  by = "Order"
) %>% arrange(SpeciesComplex, Order)

(p_PD <- ggplot(alpha_df, aes(Order, PD_sps_rare)) +
    facet_grid(~ SpeciesComplex, scales = "free_x", space = "free_x") +
    geom_boxplot(aes(fill = Species_Label), show.legend = FALSE,
                 lwd = 0.25, fatten = 1, alpha = 0.5,
                 outlier.colour = 'transparent', width = 0.5) +
    geom_point(aes(fill = Species_Label), shape = 21,
               position = position_jitter(seed = 100, width = 0.1),
               size = 3, alpha = 0.7, stroke = 0.25) +
    SpeciesLabel.fill +
    ylab("Faith's phylogenetic diversity") + xlab("") +
    sl_theme + boxplot_theme +
    scale_x_discrete(breaks = names(labels_with_n), labels = labels_with_n) +
    geom_hline(aes(yintercept = signif2_pd$Estimate[1]), linetype = 'dashed') +
    stat_summary(aes(x = Order, y = PD_sps_rare),
                 fun.data = "mean_cl_boot", geom = "pointrange",
                 shape = 23, col = "#acacac", fill = '#acacac',
                 size = 0.5, fun.args = list(conf.int = .95)) +
    geom_text(data = signif2_pd, aes(x = Order, y = 40, label = p.signif),
              vjust = 0.75, size = 8) +
    scale_y_continuous(breaks = seq(0, 40, by = 5)))

```

## GLMs across taxonomic levels (Phylum → OTU99)

Loads the phyloseq object fresh (without species-glomming) and runs negative binomial GLMs at each level for both richness and Faith's PD, building one combined table.

```{r glm-across-levels, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_base <- readRDS(paste0(dataDir, "/microb_fr_corrected.rds")) %>%
  biomeUtils::removeZeros()

# Helper: at one taxonomic level, compute Observed richness + Faith's PD,
# fit GLM nb against Species_Label, return a tidy CI table
glm_at_level <- function(ps_base, level) {
  ps <- if (level == "OTU99") ps_base else speedyseq::tax_glom(ps_base, level)
  ps <- biomeUtils::removeZeros(ps)
  
  rich <- estimate_richness(ps, measures = c("Observed"))
  pd <- calculatePD(ps, justDF = TRUE)
  
  sample_data(ps)$Observed_rare <- rich$Observed
  sample_data(ps)$PD_rare <- pd$PD
  sample_data(ps)$PD_integer <- round(pd$PD, 0)
  
  df <- data.frame(sample_data(ps)) %>%
    filter(Genus == "Fringilla")
  df$Species_Label <- relevel(as.factor(df$Species_Label), "Mainland_F_coelebs")
  
  # Helper to fit and extract a tidy CI table from a glm.nb
  fit_and_extract <- function(formula, df, metric_name, level_name) {
    m <- glm.nb(formula, data = df, na.action = na.fail)
    
    CI <- as_tibble(confint(m), rownames = 'predictor')
    CI <- cbind(CI, as_tibble(summary(m)$coefficients))
    CI$CImin <- c(CI$`2.5 %`[1], CI$Estimate[1] + CI$`2.5 %`[-1])
    CI$CImax <- c(CI$`97.5 %`[1], CI$Estimate[1] + CI$`97.5 %`[-1])
    CI$Estimate.transformed <- c(CI$`Estimate`[1], CI$`Estimate`[1] + CI$`Estimate`[-1])
    CI$Estimate <- exp(CI$Estimate.transformed)
    CI$CImin <- exp(CI$CImin)
    CI$CImax <- exp(CI$CImax)
    
    CI %>%
      mutate(`Alpha diversity metric` = metric_name,
             `Taxonomic level` = level_name,
             Predictor = gsub("Species_Label", "", predictor)) %>%
      dplyr::select(`Alpha diversity metric`, `Taxonomic level`, Predictor,
                    Estimate, `Std. Error`, `z value`, `Pr(>|z|)`, CImin, CImax)
  }
  
  rich_tab <- fit_and_extract(Observed_rare ~ Species_Label, df, "Richness", level)
  pd_tab <- fit_and_extract(PD_integer ~ Species_Label, df, "Faith's phylogenetic diversity", level)
  
  bind_rows(rich_tab, pd_tab)
}

# Run across all levels
tax_levels <- c("Phylum", "Class", "Order", "Family", "Genus", "Species", "OTU99")

glm_table <- bind_rows(lapply(tax_levels, function(lvl) {
  message("Running GLMs at level: ", lvl)
  glm_at_level(ps_base, lvl)
}))

# Round and format
glm_table <- glm_table %>%
  mutate(
    Estimate = round(Estimate, 3),
    `Std. Error` = round(`Std. Error`, 3),
    `z value` = round(`z value`, 3),
    `Pr(>|z|)` = ifelse(`Pr(>|z|)` < 0.0001,
                       "<0.0001",
                       formatC(`Pr(>|z|)`, format = "f", digits = 4)),
    CImin = round(CImin, 3),
    CImax = round(CImax, 3)
  )

# Save
write.csv(glm_table,
          paste0(outputDir, "/glm_alpha_diversity_across_taxlevels.csv"),
          row.names = FALSE)

```

### PARTITION BETA DIVERSITY (Bray-Curtis)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

Matrix_abund <- data.frame(otu_table(ps_fringilla_rare))

j.matrix <- beta.div.comp(Matrix_abund, coef = "BJ", quant = FALSE, save.abc = FALSE)
s.quant <- beta.div.comp(Matrix_abund, coef = "BS", quant = TRUE, save.abc = FALSE)

```

### Helper functions

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

build_pairwise_df <- function(s_obj, meta) {
  if ("repl" %in% names(s_obj)) {
    xy <- t(combn(colnames(as.matrix(s_obj$D)), 2))
    df <- data.frame(
      X1 = xy[, 1],
      X2 = xy[, 2],
      repl = as.matrix(s_obj$repl)[xy],
      rich = as.matrix(s_obj$rich)[xy],
      D = as.matrix(s_obj$D)[xy]
    )
  } else {
    dm <- as.matrix(s_obj)
    xy <- which(upper.tri(dm), arr.ind = TRUE)
    df <- data.frame(
      X1 = rownames(dm)[xy[, 1]],
      X2 = rownames(dm)[xy[, 2]],
      D = dm[xy]
    )
  }
  df$X1_X2 <- paste0(df$X1, "_", df$X2)
  
  meta_X1 <- data.frame(
    X1 = paste0(meta$Lab_Code, "_", meta$Species_Label),
    Species_Label_X1 = meta$Species_Label,
    Archipelago_Mainland_X1 = meta$Archipelago_Mainland)
  meta_X2 <- data.frame(
    X2 = paste0(meta$Lab_Code, "_", meta$Species_Label),
    Species_Label_X2 = meta$Species_Label,
    Archipelago_Mainland_X2 = meta$Archipelago_Mainland)
  
  df <- df %>% left_join(meta_X1, by = "X1") %>% left_join(meta_X2, by = "X2")
  df$X1_X2_Species_Label <- paste0(df$Species_Label_X1, ":", df$Species_Label_X2)
  df$X1_X2_Same_site <- df$Species_Label_X1 == df$Species_Label_X2
  df$X1_X2_Archipelago_Mainland <- paste0(df$Archipelago_Mainland_X1, ":", df$Archipelago_Mainland_X2)
  
  if ("repl" %in% names(s_obj)) df$S <- 1 - df$D
  df
}

prepare_dist_for_plot <- function(df_diff_mainland) {
  df_diff_mainland$site_interest <- gsub('Mainland_F_coelebs', '', df_diff_mainland$X1_X2_Species_Label)
  df_diff_mainland$site_interest <- gsub(':', '', df_diff_mainland$site_interest)
  
  df_diff_mainland <- df_diff_mainland %>%
    mutate(Labels = site_interest %>%
             str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%
             str_replace("_F_", " : F.") %>%
             str_replace("c_", "c.") %>%
             str_replace("_", " "))
  
  df_diff_mainland$Order <- factor(df_diff_mainland$Labels, levels = order_levels)
  df_diff_mainland$SpeciesComplex <- factor(species_complex_lookup[as.character(df_diff_mainland$Order)],
                                            levels = c("Common chaffinch species complex", "Blue finches"))
  df_diff_mainland$SpeciesLabelMatch <- factor(
    order_to_species_label[as.character(df_diff_mainland$Order)],
    levels = SpeciesLabel.order_local)
  df_diff_mainland
}

# Add a placeholder Mainland row so Mainland appears on the x-axis
add_mainland_placeholder <- function(df, columns_to_fill = list()) {
  mainland_row <- data.frame(
    X1 = "Mainland_F_coelebs", X2 = "Mainland_F_coelebs",
    Species_Label_X1 = "Mainland_F_coelebs", Species_Label_X2 = "Mainland_F_coelebs",
    X1_X2_Species_Label = "Mainland_F_coelebs:Mainland_F_coelebs",
    site_interest = "Mainland", Labels = "Mainland : F.coelebs",
    Order = factor("Mainland : F.coelebs", levels = order_levels),
    SpeciesComplex = factor("Common chaffinch species complex",
                            levels = c("Common chaffinch species complex", "Blue finches")),
    SpeciesLabelMatch = factor("Mainland_F_coelebs", levels = SpeciesLabel.order_local)
  )
  for (col in names(columns_to_fill)) {
    mainland_row[[col]] <- columns_to_fill[[col]]
  }
  bind_rows(df, mainland_row)
}

```

### Bray-Curtis dissimilarity vs mainland

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

Meta <- data.frame(sample_data(ps_fringilla_rare))

dist_df <- build_pairwise_df(s.quant, Meta)
dist_df_diff_site <- dist_df %>% filter(!X1_X2_Same_site)
dist_df_diff_mainland <- dist_df_diff_site %>%
  filter(grepl("Mainland_F_coelebs", X1_X2_Species_Label))

dist_df_diff_mainland <- prepare_dist_for_plot(dist_df_diff_mainland)
dist_df_diff_mainland <- add_mainland_placeholder(
  dist_df_diff_mainland,
  list(D = NA, repl = NA, rich = NA, S = NA, X1_X2 = "mainland_mainland",
       X1_X2_Same_site = TRUE, X1_X2_Archipelago_Mainland = "Mainland:Mainland",
       Archipelago_Mainland_X1 = "Mainland", Archipelago_Mainland_X2 = "Mainland")
)

# Common chaffinch only subset (used for ternary plots)
dist_df_common <- dist_df %>%
  filter(!Species_Label_X1 %in% c("Tenerife_F_teydea", "Gran_Canaria_F_polatzeki"),
         !Species_Label_X2 %in% c("Tenerife_F_teydea", "Gran_Canaria_F_polatzeki"))
dist_df_diff_common <- dist_df_common %>% filter(!X1_X2_Same_site)

dat_text <- data.frame(
  label = c("Reference", ""),
  SpeciesComplex = factor(c("Common chaffinch species complex", "Blue finches"),
                          levels = c("Common chaffinch species complex", "Blue finches")),
  x = 0.5, y = 0.35)

(D_overall_bray <- ggplot(data = dist_df_diff_mainland, aes(Order, D)) +
    geom_boxplot(aes(fill = SpeciesLabelMatch), show.legend = FALSE,
                 lwd = 0.25, fatten = 1, alpha = 0.7,
                 outlier.colour = 'transparent', width = 0.5) +
    geom_point(aes(fill = SpeciesLabelMatch), shape = 21,
               position = position_jitter(seed = 100, width = 0.1),
               size = 3, alpha = 0.5, stroke = 0.25) +
    xlab("") + ylab("Dissimilarity") +
    facet_grid(. ~ SpeciesComplex, scales = "free_x", space = "free_x") +
    scale_x_discrete(breaks = names(labels_with_n), labels = labels_with_n) +
    SpeciesLabel.fill_local + sl_theme + boxplot_theme +
    stat_summary(aes(x = Order, y = D), fun.data = "mean_cl_boot",
                 geom = "pointrange", shape = 23,
                 col = '#acacac', fill = '#acacac', size = 0.5))

(D_overall_bray_patch <- D_overall_bray +
    geom_rect(data = dist_df_diff_mainland %>% filter(SpeciesComplex == "Common chaffinch species complex"),
              aes(xmin = 0.5, xmax = 1.5, ymin = -Inf, ymax = Inf),
              alpha = 0.5, fill = 'lightgrey') +
    geom_text(data = dat_text, mapping = aes(x = x, y = y, label = label),
              hjust = -0.25, vjust = 2.5, angle = 90, size = 7) +
    theme(strip.background = element_blank(), strip.text.x = element_blank()))

```

### Ternary plots

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

(triangle.overall.diff.bray <- ggtern(data = dist_df_diff_common, aes(rich, repl, S)) +
    geom_point(size = 4, alpha = 0.05, stroke = 0.25) +
    geom_crosshair_tern(aes(mean(dist_df_diff_common$rich),
                            mean(dist_df_diff_common$repl),
                            mean(dist_df_diff_common$S)),
                        col = "darkred", lty = "dashed",
                        lineend = "round", size = 2.2) +
    geom_point(aes(mean(dist_df_diff_common$rich),
                   mean(dist_df_diff_common$repl),
                   mean(dist_df_diff_common$S)),
               col = 'darkred', size = 10) +
    geom_point(aes(x = mean(dist_df_diff_common$rich),
                   y = 1 - mean(dist_df_diff_common$rich), z = 0),
               col = "darkred", size = 10) +
    annotate(geom = 'text', x = mean(dist_df_diff_common$rich),
             y = 1 - mean(dist_df_diff_common$rich), z = 0,
             label = paste(round(mean(dist_df_diff_common$rich) * 100, 2)),
             hjust = 1, vjust = -1, size = 12, col = "darkred") +
    geom_point(aes(x = 0, y = mean(dist_df_diff_common$repl),
                   z = 1 - mean(dist_df_diff_common$repl)),
               col = "darkred", size = 10) +
    annotate(geom = 'text', x = 0, y = mean(dist_df_diff_common$repl),
             z = 1 - mean(dist_df_diff_common$repl),
             label = paste(round(mean(dist_df_diff_common$repl) * 100, 2)),
             hjust = -0.3, vjust = 1, size = 12, col = "darkred") +
    geom_point(aes(x = 1 - mean(dist_df_diff_common$S), y = 0,
                   z = mean(dist_df_diff_common$S)),
               col = "darkred", size = 8) +
    annotate(geom = 'text', x = 1 - mean(dist_df_diff_common$S), y = 0,
             z = mean(dist_df_diff_common$S),
             label = paste(round(mean(dist_df_diff_common$S) * 100, 2)),
             hjust = 0.5, vjust = 2, size = 12, col = "darkred") +
    theme_classic() + theme_hidetitles() + theme_nomask() +
    theme(axis.text = element_text(size = 28),
          tern.axis.arrow = element_line(size = 3)))

```

### Jaccard partitioning

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

j_dist_df <- build_pairwise_df(j.matrix, Meta)
j_dist_df_common <- j_dist_df %>%
  filter(!Species_Label_X1 %in% c("Tenerife_F_teydea", "Gran_Canaria_F_polatzeki"),
         !Species_Label_X2 %in% c("Tenerife_F_teydea", "Gran_Canaria_F_polatzeki"))
j_dist_df_diff_common <- j_dist_df_common %>% filter(!X1_X2_Same_site)

(triangle.overall.diff.jaccard <- ggtern(data = j_dist_df_diff_common, aes(rich, repl, S)) +
    geom_point(size = 4, alpha = 0.05, stroke = 0.25) +
    geom_crosshair_tern(aes(mean(j_dist_df_diff_common$rich),
                            mean(j_dist_df_diff_common$repl),
                            mean(j_dist_df_diff_common$S)),
                        col = "darkred", lty = "dashed",
                        lineend = "round", size = 2.2) +
    geom_point(aes(mean(j_dist_df_diff_common$rich),
                   mean(j_dist_df_diff_common$repl),
                   mean(j_dist_df_diff_common$S)),
               col = 'darkred', size = 10) +
    geom_point(aes(x = mean(j_dist_df_diff_common$rich),
                   y = 1 - mean(j_dist_df_diff_common$rich), z = 0),
               col = "darkred", size = 10) +
    annotate(geom = 'text', x = mean(j_dist_df_diff_common$rich),
             y = 1 - mean(j_dist_df_diff_common$rich), z = 0,
             label = paste(round(mean(j_dist_df_diff_common$rich) * 100, 2)),
             hjust = 1, vjust = -1, size = 12, col = "darkred") +
    geom_point(aes(x = 0, y = mean(j_dist_df_diff_common$repl),
                   z = 1 - mean(j_dist_df_diff_common$repl)),
               col = "darkred", size = 10) +
    annotate(geom = 'text', x = 0, y = mean(j_dist_df_diff_common$repl),
             z = 1 - mean(j_dist_df_diff_common$repl),
             label = paste(round(mean(j_dist_df_diff_common$repl) * 100, 2)),
             hjust = -0.3, vjust = 1, size = 12, col = "darkred") +
    geom_point(aes(x = 1 - mean(j_dist_df_diff_common$S), y = 0,
                   z = mean(j_dist_df_diff_common$S)),
               col = "darkred", size = 8) +
    annotate(geom = 'text', x = 1 - mean(j_dist_df_diff_common$S), y = 0,
             z = mean(j_dist_df_diff_common$S),
             label = paste(round(mean(j_dist_df_diff_common$S) * 100, 2)),
             hjust = 0.5, vjust = 2, size = 12, col = "darkred") +
    theme_classic() + theme_hidetitles() + theme_nomask() +
    theme(axis.text = element_text(size = 28),
          tern.axis.arrow = element_line(size = 3)))

```

### UniFrac dissimilarities (unweighted and weighted)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

build_unifrac_df <- function(unifrac_obj, meta) {
  dm <- as.matrix(unifrac_obj)
  pairs <- which(upper.tri(dm), arr.ind = TRUE)
  df <- data.frame(
    X1 = rownames(dm)[pairs[, 1]],
    X2 = rownames(dm)[pairs[, 2]],
    D = dm[pairs])
  df$Species_Label_X1 <- meta[df$X1, "Species_Label"]
  df$Species_Label_X2 <- meta[df$X2, "Species_Label"]
  df$X1_X2_Species_Label <- paste0(df$Species_Label_X1, ":", df$Species_Label_X2)
  
  df <- df %>% filter(
    (Species_Label_X1 == "Mainland_F_coelebs" & Species_Label_X2 != "Mainland_F_coelebs") |
    (Species_Label_X2 == "Mainland_F_coelebs" & Species_Label_X1 != "Mainland_F_coelebs"))
  
  df$site_interest <- ifelse(df$Species_Label_X1 == "Mainland_F_coelebs",
                             as.character(df$Species_Label_X2),
                             as.character(df$Species_Label_X1))
  
  df <- df %>%
    mutate(Labels = site_interest %>%
             str_replace("_F_", " : F.") %>%
             str_replace("c_", "c.") %>%
             str_replace("_", " "))
  
  df$Order <- factor(df$Labels, levels = order_levels)
  df$SpeciesComplex <- factor(species_complex_lookup[as.character(df$Order)],
                              levels = c("Common chaffinch species complex", "Blue finches"))
  df$SpeciesLabelMatch <- factor(
    order_to_species_label[as.character(df$Order)],
    levels = SpeciesLabel.order_local)
  
  # Add Mainland placeholder
  mainland_row <- data.frame(
    X1 = "Mainland_F_coelebs", X2 = "Mainland_F_coelebs", D = NA,
    Species_Label_X1 = "Mainland_F_coelebs", Species_Label_X2 = "Mainland_F_coelebs",
    X1_X2_Species_Label = "Mainland_F_coelebs:Mainland_F_coelebs",
    site_interest = "Mainland", Labels = "Mainland : F.coelebs",
    Order = factor("Mainland : F.coelebs", levels = order_levels),
    SpeciesComplex = factor("Common chaffinch species complex",
                            levels = c("Common chaffinch species complex", "Blue finches")),
    SpeciesLabelMatch = factor("Mainland_F_coelebs", levels = SpeciesLabel.order_local))
  bind_rows(df, mainland_row)
}

make_unifrac_panel <- function(df, ylab_text) {
  y_range <- range(df$D, na.rm = TRUE)
  y_low <- y_range[1] + (y_range[2] - y_range[1]) * 0.35   # 35% from bottom
  
  dat_text <- data.frame(
  label = c("Reference", ""),
  SpeciesComplex = factor(c("Common chaffinch species complex", "Blue finches"),
                          levels = c("Common chaffinch species complex", "Blue finches")),
  x = 0.5, y = y_low)
  
  p <- ggplot(data = df, aes(Order, D)) +
    geom_boxplot(aes(fill = SpeciesLabelMatch), show.legend = FALSE,
                 lwd = 0.25, fatten = 1, alpha = 0.7,
                 outlier.colour = 'transparent', width = 0.5) +
    geom_point(aes(fill = SpeciesLabelMatch), shape = 21,
               position = position_jitter(seed = 100, width = 0.1),
               size = 3, alpha = 0.5, stroke = 0.25) +
    xlab("") + ylab(ylab_text) +
    facet_grid(. ~ SpeciesComplex, scales = "free_x", space = "free_x") +
    scale_x_discrete(breaks = names(labels_with_n), labels = labels_with_n) +
    SpeciesLabel.fill_local + sl_theme + boxplot_theme +
    stat_summary(aes(x = Order, y = D), fun.data = "mean_cl_boot",
                 geom = "pointrange", shape = 23,
                 col = '#acacac', fill = '#acacac', size = 0.5)
  
  p +
    geom_rect(data = df %>% filter(SpeciesComplex == "Common chaffinch species complex"),
              aes(xmin = 0.5, xmax = 1.5, ymin = -Inf, ymax = Inf),
              alpha = 0.5, fill = 'lightgrey') +
    geom_text(data = dat_text, mapping = aes(x = x, y = y, label = label),
              hjust = -0.25, vjust = 2.5, angle = 90, size = 7) +
    theme(strip.background = element_blank(), strip.text.x = element_blank())
}

unifrac_uw <- UniFrac(ps_fringilla_rare, weighted = FALSE)
df_uw <- build_unifrac_df(unifrac_uw, Meta)
D_unifrac_uw_patch <- make_unifrac_panel(df_uw, "Unweighted UniFrac")
D_unifrac_uw_patch

unifrac_w <- UniFrac(ps_fringilla_rare, weighted = TRUE)
df_w <- build_unifrac_df(unifrac_w, Meta)
D_unifrac_w_patch <- make_unifrac_panel(df_w, "Weighted UniFrac")
D_unifrac_w_patch

```

# BIOGEOGRAPHY

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

nomainland_fringilla <- alpha_df %>% dplyr::filter(!is.na(Area_km2) & Genus == "Fringilla")
nrow(nomainland_fringilla)
# Area
GLMM1 <- glmmTMB(Observed_sps_rare ~ scale(Area_km2) + (1 | Species_Label),
                 data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM1); confint(GLMM1)

(pArea <- ggpredict(GLMM1, terms = "Area_km2 [all]") %>%
    plot(line_size = 1.5, color = "darkred") +
    geom_point(data = nomainland_fringilla, aes(y = Observed_sps_rare, x = Area_km2),
               shape = 21, fill = "grey",
               position = position_jitter(seed = 100, width = 50),
               size = 3, alpha = 0.9, stroke = 0.25) +
    ylab("Richness") + xlab(bquote(Area ~ (km^2))) +
    scatter_theme + ggtitle("") + sl_theme +
    annotate(geom = "text",
             label = build_biogeo_annot(GLMM1, "scale(Area_km2)"),
             parse = TRUE, x = 1000, y = 240, size = 6) +
    scale_y_continuous(breaks = seq(0, 250, by = 50)))

GLMM1_PD <- glmmTMB(PD_integer ~ scale(Area_km2) + (1 | Species_Label),
                    data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM1_PD); confint(GLMM1_PD)

(pArea_PD <- ggpredict(GLMM1_PD, terms = "Area_km2 [all]") %>%
    plot(line_size = 1.5, color = "darkred") +
    geom_point(data = nomainland_fringilla, aes(y = PD_integer, x = Area_km2),
               shape = 21, fill = "grey",
               position = position_jitter(seed = 100, width = 50),
               size = 3, alpha = 0.9, stroke = 0.25) +
    ylab("Faith's phylogenetic diversity") + xlab(bquote(Area ~ (km^2))) +
    scatter_theme + ggtitle("") + sl_theme +
    annotate(geom = "text",
             label = build_biogeo_annot(GLMM1_PD, "scale(Area_km2)"),
             parse = TRUE, x = 1000, y = 35, size = 6) +
    scale_y_continuous(breaks = seq(0, 40, by = 5)))

# Distance
GLMM2 <- glmmTMB(Observed_sps_rare ~ scale(Distance_mainland) + (1 | Species_Label),
                 data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM2); confint(GLMM2)

(pDistance <- ggpredict(GLMM2, terms = "Distance_mainland [all]") %>%
    plot(line_size = 1.5, color = "darkred") +
    geom_point(data = nomainland_fringilla, aes(y = Observed_sps_rare, x = Distance_mainland),
               shape = 21, fill = "grey",
               position = position_jitter(seed = 100, width = 50),
               size = 3, alpha = 0.9, stroke = 0.25) +
    ylab("") + xlab("Distance from mainland (km)") +
    scatter_theme + ggtitle("") + sl_theme +
    annotate(geom = "text",
             label = build_biogeo_annot(GLMM2, "scale(Distance_mainland)"),
             parse = TRUE, x = 1000, y = 240, size = 6) +
    scale_y_continuous(breaks = seq(0, 250, by = 50)))

GLMM2_PD <- glmmTMB(PD_integer ~ scale(Distance_mainland) + (1 | Species_Label),
                    data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM2_PD); confint(GLMM2_PD)

(pDistance_PD <- ggpredict(GLMM2_PD, terms = "Distance_mainland [all]") %>%
    plot(line_size = 1.5, color = "darkred") +
    geom_point(data = nomainland_fringilla, aes(y = PD_integer, x = Distance_mainland),
               shape = 21, fill = "grey",
               position = position_jitter(seed = 100, width = 50),
               size = 3, alpha = 0.9, stroke = 0.25) +
    ylab("") + xlab("Distance from mainland (km)") +
    scatter_theme + ggtitle("") + sl_theme +
    annotate(geom = "text",
             label = build_biogeo_annot(GLMM2_PD, "scale(Distance_mainland)"),
             parse = TRUE, x = 1000, y = 35, size = 6) +
    scale_y_continuous(breaks = seq(0, 40, by = 5)))

# Age
GLMM3 <- glmmTMB(Observed_sps_rare ~ scale(Age_ma) + (1 | Species_Label),
                 data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM3); confint(GLMM3)

(pAge_ma <- ggpredict(GLMM3, terms = "Age_ma [all]") %>%
    plot(line_size = 1.5, color = "darkred") +
    geom_point(data = nomainland_fringilla, aes(y = Observed_sps_rare, x = Age_ma),
               shape = 21, fill = "grey",
               position = position_jitter(seed = 100, width = 0.3),
               size = 3, alpha = 0.9, stroke = 0.25) +
    ylab("") + xlab("Island age (mya)") +
    scatter_theme + ggtitle("") + sl_theme +
    annotate(geom = "text",
             label = build_biogeo_annot(GLMM3, "scale(Age_ma)"),
             parse = TRUE, x = 7.5, y = 240, size = 6) +
    scale_y_continuous(breaks = seq(0, 250, by = 50)))

GLMM3_PD <- glmmTMB(PD_integer ~ scale(Age_ma) + (1 | Species_Label),
                    data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM3_PD); confint(GLMM3_PD)

(pAge_ma_PD <- ggpredict(GLMM3_PD, terms = "Age_ma [all]") %>%
    plot(line_size = 1.5, color = "darkred") +
    geom_point(data = nomainland_fringilla, aes(y = PD_integer, x = Age_ma),
               shape = 21, fill = "grey",
               position = position_jitter(seed = 100, width = 0.3),
               size = 3, alpha = 0.9, stroke = 0.25) +
    ylab("") + xlab("Island age (mya)") +
    scatter_theme + ggtitle("") + sl_theme +
    annotate(geom = "text",
             label = build_biogeo_annot(GLMM3_PD, "scale(Age_ma)"),
             parse = TRUE, x = 7.5, y = 35, size = 6) +
    scale_y_continuous(breaks = seq(0, 40, by = 5)))

```
## Biogeography effects across taxonomic levels

```{r biogeo-across-levels, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Helper: fit biogeography GLMMs at one taxonomic level
biogeo_at_level <- function(ps_base, level, nomainland_meta) {
  
  # Aggregate to taxonomic level

ps <- if (level == "OTU99") ps_base else speedyseq::tax_glom(ps_base, level)
  ps <- biomeUtils::removeZeros(ps)
  
  # Calculate richness
  rich <- estimate_richness(ps, measures = "Observed")
  sample_data(ps)$Observed_rare <- rich$Observed
  
  # Build dataframe and merge with island metadata
  df <- data.frame(sample_data(ps)) %>%
    filter(Genus == "Fringilla", !is.na(Area_km2))
  
  # Fit the three biogeography models
  fit_and_extract <- function(predictor_name, formula_str) {
    m <- glmmTMB(as.formula(formula_str), 
                 data = df, na.action = na.fail, family = nbinom2)
    
    coefs <- summary(m)$coefficients$cond
    
    # Extract predictor row (not intercept)
    pred_row <- coefs[2, ]
    int_row <- coefs[1, ]
    
    data.frame(
      `Taxonomic Level` = level,
      Predictor = c("Intercept", predictor_name),
      Estimate = round(c(int_row["Estimate"], pred_row["Estimate"]), 5),
      `Std. Error` = round(c(int_row["Std. Error"], pred_row["Std. Error"]), 5),
      `z value` = round(c(int_row["z value"], pred_row["z value"]), 2),
      `Pr(>|z|)` = c(int_row["Pr(>|z|)"], pred_row["Pr(>|z|)"]),
      check.names = FALSE
    )
  }
  
  area_tab <- fit_and_extract("Island area", 
    "Observed_rare ~ scale(Area_km2) + (1 | Species_Label)")
  dist_tab <- fit_and_extract("Distance to mainland", 
    "Observed_rare ~ scale(Distance_mainland) + (1 | Species_Label)")
  age_tab <- fit_and_extract("Island age", 
    "Observed_rare ~ scale(Age_ma) + (1 | Species_Label)")
  
  bind_rows(area_tab, dist_tab, age_tab)
}

# Run across all taxonomic levels
tax_levels <- c("Phylum", "Class", "Order", "Family", "Genus", "Species", "OTU99")

biogeo_table <- bind_rows(lapply(tax_levels, function(lvl) {
  message("Running biogeography GLMMs at level: ", lvl)
  biogeo_at_level(ps_base, lvl, nomainland_fringilla)
}))

# Format p-values
biogeo_table <- biogeo_table %>%
  mutate(`Pr(>|z|)` = ifelse(`Pr(>|z|)` < 0.0001,
                             "<0.0001",
                             formatC(`Pr(>|z|)`, format = "f", digits = 4)))

# Display
kable(biogeo_table,
      col.names = c("Taxonomic Level", "Predictor", "Estimate",
                    "Std. Error", "z value", "p-value"),
      caption = "Effect of biogeography on microbiome richness across taxonomic levels") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

# Save
write.csv(biogeo_table,
          paste0(outputDir, "/glm_biogeography_richness_across_taxlevels.csv"),
          row.names = FALSE)
```


## Patchwork

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.width=14, fig.height=7}

png(paste0(outputDir, "/Figures/Colonisation_history_speciesFix.png"),
    width = 300, height = 240, units = "mm", res = 350)
((p1 + theme(axis.title.x = element_blank(), axis.text.x = element_blank(),
             axis.line.x = element_blank(), axis.ticks.x = element_blank(),
             strip.text.x = element_text(size = 18))) /
   D_overall_bray_patch) + plot_layout(axes = 'collect')
dev.off()


png(paste0(outputDir, "/Figures/triangle.plots.baselga.sorensen.quant.overall.matrix_genus_noarrows_2panels_speciesFix.png"),
    width = 8000, height = 8000, res = 650)
(triangle.overall.diff.bray <- triangle.overall.diff.bray +
    theme(axis.title.x = element_blank(), axis.text.x = element_blank(),
          axis.ticks.x = element_blank(), axis.title.y = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank()))
dev.off()


png(paste0(outputDir, "/Figures/triangle.plots.baselga.j.overall.matrix_genus_2panels_speciesFix.png"),
    width = 8000, height = 8000, res = 650)
(triangle.overall.diff.jaccard <- triangle.overall.diff.jaccard +
    theme(axis.title.x = element_blank(), axis.text.x = element_blank(),
          axis.ticks.x = element_blank(), axis.title.y = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank()))
dev.off()


png(paste0(outputDir, "/Figures/Biogeography_speciesFix.png"),
    width = 500, height = 200, units = "mm", res = 350)
pArea + pDistance + pAge_ma + plot_layout(axes = 'collect')
dev.off()


png(paste0(outputDir, "/Figures/Colonisation_history_speciesFix_PDUniFrac.png"),
    width = 300, height = 360, units = "mm", res = 350)
(p_PD + theme(axis.title.x = element_blank(), axis.text.x = element_blank(),
              axis.line.x = element_blank(), axis.ticks.x = element_blank(),
              strip.text.x = element_text(size = 18))) /
  D_unifrac_uw_patch /
  D_unifrac_w_patch +
  plot_layout(axes = 'collect') +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 24, face = 'bold'))
dev.off()


png(paste0(outputDir, "/Figures/Biogeography_speciesFix_PD.png"),
    width = 420, height = 198, units = "mm", res = 300)
pArea_PD + pDistance_PD + pAge_ma_PD +
  plot_layout(axes = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size = 32, face = "bold"))
dev.off()


png(paste0(outputDir, "/Figures/SupplFigS1_FaithsPDUniFrac.png"),
    width = 400, height = 480, units = "mm", res = 350)
(p_PD + theme(axis.title.x = element_blank(), axis.text.x = element_blank(),
              axis.line.x = element_blank(), axis.ticks.x = element_blank(),
              strip.text.x = element_text(size = 18))) /
  D_unifrac_uw_patch /
  D_unifrac_w_patch /
  (pArea_PD + pDistance_PD + pAge_ma_PD + plot_layout(axes = "collect")) +
  plot_layout(axes = 'collect') +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 24, face = 'bold'))
dev.off()

```

## DIET 

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

diet_fr <- readRDS(paste0(dataDir, "/microb_overlap_diet_corrected.rds"))

```

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

tt <- tax_table(diet_fr) %>% data.frame
order_filt <- subset(tt, !grepl("Class", Order) & !grepl("Phylum", Order)) %>% rownames_to_column("ASV")
diet_fr <- prune_taxa(order_filt$ASV, diet_fr)
alpha_diet_fr <- data.frame(sample_data(diet_fr))

mainland_fringilla_diet <- merge(
  alpha_df,
  alpha_diet_fr[, c("Shannon_diet_rare", "Observed_diet_rare"), drop = FALSE],
  by = "row.names", all.x = TRUE)
rownames(mainland_fringilla_diet) <- mainland_fringilla_diet$Row.names
mainland_fringilla_diet$Row.names <- NULL
alpha_diet_fr <- mainland_fringilla_diet[!is.na(mainland_fringilla_diet$Observed_diet_rare), ]

```

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

GLMM1 <- glmmTMB(Observed_sps_rare ~ Observed_diet_rare + (1 | Species_Label),
                 data = alpha_diet_fr, na.action = na.fail, family = nbinom2)
summary(GLMM1)

GLMM1b <- glmmTMB(Observed_sps_rare ~ scale(Observed_diet_rare) + (1 | Species_Label),
                  data = alpha_diet_fr, na.action = na.fail, family = nbinom2)
summary(GLMM1b); confint(GLMM1b)

scatter_theme <- theme(panel.border = element_blank(),
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank(),
                       panel.background = element_blank(),
                       axis.line = element_line(colour = "black"),
                       axis.text.x = element_text(size = 16),
                       axis.text.y = element_text(size = 16),
                       axis.title = element_text(size = 20))

(pDiet_richness <- ggpredict(GLMM1, terms = "Observed_diet_rare [all]") %>%
    plot(line_size = 1.5) +
    geom_point(data = alpha_diet_fr,
               aes(y = Observed_sps_rare, x = Observed_diet_rare),
               shape = 21, fill = "grey",
               position = position_jitter(seed = 100, width = 1),
               size = 3, alpha = 0.9, stroke = 0.25) +
    ylab("Microbiome richness") + xlab("Diet richness") +
    scatter_theme + ggtitle("") + sl_theme +
    annotate(geom = "text",
             label = build_biogeo_annot(GLMM1b,"scale(Observed_diet_rare)"),
             parse = TRUE, x = 15, y = 240, size = 6) +
    scale_y_continuous(breaks = seq(0, 250, by = 50)))

GLMMPD <- glmmTMB(PD_integer ~ Observed_diet_rare + (1 | Species_Label),
                  data = alpha_diet_fr, na.action = na.fail, family = nbinom2)
summary(GLMMPD)

GLMMPDb <- glmmTMB(PD_integer ~ scale(Observed_diet_rare) + (1 | Species_Label),
                   data = alpha_diet_fr, na.action = na.fail, family = nbinom2)
summary(GLMMPDb); confint(GLMMPDb)

```