---
title: 'Ancom diet'
author: "B. Karina Montero and Mark Gillingham"
date: "`r Sys.Date()`"
#output: pdf_document
output: html_document
---
## Loading libraries


```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
rm(list = ls())

# basics
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(ggpubr)
library(tidyverse)
library(ggtext)
library(ggrepel)
library(purrr)
library(patchwork)
library(ggtext)

# phyloseq section
library(phyloseq)
library(decontam)
library(biomeUtils)
library(vegan)

#devtools::install_github("gmteunisse/fantaxtic")
library(fantaxtic)
library(ggtext)

# models
library(mgcv)
library(glmmTMB) 
library(lme4)
library(MuMIn)
library(MASS)
library(nlme)
library(ggeffects)
library(multcomp)
library(car)

# ancom
#if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("ANCOMBC")
library(ANCOMBC)
```



## WORKING DIRECTORIES

```{r workingDir, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```

## LOAD DATA

Rarefied data

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

#microb_overlap_rare <- readRDS(paste0(dataDir, "/microb_overlap_ASV_rare.rds"))
#diet_overlap_rare <- readRDS(paste0(dataDir, "/diet_overlap_ASV_rare.rds"))
microb_fr <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))

tax_table(microb_fr)
microb_fr <- speedyseq::tax_glom(microb_fr, "Species")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Genus")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Family")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Order")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Class")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Phylum")

#microb_fr <- speedyseq::tax_glom(microb_fr, "Genus")
diet_fr <- readRDS(paste0(dataDir, "/diet_overlap_OTU99_fr.rds"))


```

## DIET Merged

remove unclassified taxa to Order level  

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

tt <- tax_table(diet_fr) %>% data.frame  
table(tt$Order)
order_filt <- subset(tt, !grepl("Class", Order) & !grepl("Phylum", Order)) %>% rownames_to_column("ASV")
order_filt <- order_filt$ASV

diet_fr <- prune_taxa(order_filt, diet_fr)

ps_genus <- tax_glom(diet_fr, taxrank = "Genus", NArm = FALSE)
tax_table(ps_genus)
```


## Are microbial taxa more or less abundant in the presence of the common dietary items  

# Differential abundance ANCOMBC


Relative abundance of pinus

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

diet_fr_r <- transform_sample_counts(ps_genus, function(x) x/sum(x))

ps_genus_pinus <- subset_taxa(diet_fr_r , Genus %in% c("Pinus"))

genus_df <- psmelt(ps_genus_pinus)
head(genus_df)

pinesumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(pine_abund = mean(Abundance, na.rm=TRUE)) 
head(pinesumm)
tail(pinesumm)

pinesumm <- pinesumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), pinesumm, by = "sampleid")
head(DF)
```

Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$pine_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$pine_abund <- DF$pine_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "pine_abund + Species_Label",
p_adj_method = "holm", prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
str(out_res)
out_res_lfc <- out_res$lfc[, c("taxon","pine_abund")]
out_res_se <- out_res$se[, c("taxon","pine_abund")]
out_res_W <- out_res$W[, c("taxon","pine_abund")]
out_res_q <- out_res$q_val[, c("taxon","pine_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","pine_abund")]
head(out_res_diff)

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_pine <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_pine) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_pine$diet <- paste("pine")
head(out_pine)

# check which taxa 
table(out_res_diff$pine_abund)
diff_df <- out_pine %>% filter(diff_abund == TRUE)
diff_df





```


Relative abundance of Morella

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_morella <- subset_taxa(diet_fr_r, Genus %in% c("Morella"))

genus_df <- psmelt(ps_genus_morella)
head(genus_df)

morellasumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(morella_abund = mean(Abundance, na.rm=TRUE)) 
head(morellasumm)
tail(morellasumm)

morellasumm <- morellasumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), morellasumm, by = "sampleid")
head(DF)


```

Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$morella_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$morella_abund <- DF$morella_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "morella_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","morella_abund")]
out_res_se <- out_res$se[, c("taxon","morella_abund")]
out_res_W <- out_res$W[, c("taxon","morella_abund")]
out_res_q <- out_res$q_val[, c("taxon","morella_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","morella_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_morella <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_morella) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_morella$diet <- paste("morella")
head(out_morella)

# check which taxa 
table(out_res_diff$morella_abund)
diff_df <- out_morella %>% filter(diff_abund == TRUE)

```


Relative abundance of Ericaceae

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_ericaceae <- subset_taxa(diet_fr_r, Genus %in% c("Family_Ericaceae"))

genus_df <- psmelt(ps_genus_ericaceae)
head(genus_df)

ericaceaesumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(ericaceae_abund = mean(Abundance, na.rm=TRUE)) 
head(ericaceaesumm)
tail(ericaceaesumm)

ericaceaesumm <- ericaceaesumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), ericaceaesumm, by = "sampleid")
head(DF)

```

Ericaceae Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$ericaceae_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$ericaceae_abund <- DF$ericaceae_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "ericaceae_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","ericaceae_abund")]
out_res_se <- out_res$se[, c("taxon","ericaceae_abund")]
out_res_W <- out_res$W[, c("taxon","ericaceae_abund")]
out_res_q <- out_res$q_val[, c("taxon","ericaceae_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","ericaceae_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_ericaceae <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_ericaceae) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_ericaceae$diet <- paste("ericaceae")
head(out_ericaceae)

# check which taxa 
table(out_res_diff$ericaceae_abund)
diff_df <- out_ericaceae %>% filter(diff_abund == TRUE)

```


Relative abundance of Geranium

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_geranium <- subset_taxa(diet_fr_r, Genus %in% c("Geranium"))

genus_df <- psmelt(ps_genus_geranium)
head(genus_df)

geraniumsumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(geranium_abund = mean(Abundance, na.rm=TRUE)) 
head(geraniumsumm)
tail(geraniumsumm)

geraniumsumm <- geraniumsumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), geraniumsumm, by = "sampleid")
head(DF)

```

Ericaceae Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$geranium_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$geranium_abund <- DF$geranium_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "geranium_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","geranium_abund")]
out_res_se <- out_res$se[, c("taxon","geranium_abund")]
out_res_W <- out_res$W[, c("taxon","geranium_abund")]
out_res_q <- out_res$q_val[, c("taxon","geranium_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","geranium_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_geranium <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_geranium) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_geranium$diet <- paste("geranium")
head(out_geranium)

# check which taxa 
table(out_res_diff$geranium_abund)
diff_df <- out_geranium %>% filter(diff_abund == TRUE)

```

Relative abundance of Rubus

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_rubus <- subset_taxa(diet_fr_r, Genus %in% c("Rubus"))

genus_df <- psmelt(ps_genus_rubus)
head(genus_df)

rubussumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(rubus_abund = mean(Abundance, na.rm=TRUE)) 
head(rubussumm)
tail(rubussumm)

rubussumm <- rubussumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), rubussumm, by = "sampleid")
head(DF)

```

Rubus Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$rubus_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$rubus_abund <- DF$rubus_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "rubus_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","rubus_abund")]
out_res_se <- out_res$se[, c("taxon","rubus_abund")]
out_res_W <- out_res$W[, c("taxon","rubus_abund")]
out_res_q <- out_res$q_val[, c("taxon","rubus_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","rubus_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_rubus <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_rubus) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_rubus$diet <- paste("rubus")
head(out_rubus)

# check which taxa 
table(out_res_diff$rubus_abund)
diff_df <- out_rubus %>% filter(diff_abund == TRUE)

```


merge Viri df

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

df_viri_ancom_topgenus <- rbind(out_pine, out_morella, out_ericaceae, out_geranium, out_rubus)

df_viri_ancom_topgenus <- df_viri_ancom_topgenus 
#%>% filter(diff_abund == TRUE)
df_viri_ancom_topgenus$Kingdom <- paste0("Viridiplantae")
str(df_viri_ancom_topgenus)

df_overview <- (table(df_viri_ancom_topgenus$taxon, df_viri_ancom_topgenus$diet))
table(df_viri_ancom_topgenus$taxon)

#df_viri_ancom_top10 <- rbind(out_pine, out_morella, out_ericaceae, out_geranium, out_lamiaceae, out_lauraceae, out_juglandaceae, out_poaceae, out_rubus, out_aster)

#df_viri_ancom_top10 <- df_viri_ancom_top10 %>% filter(diff_abund == TRUE)
#df_viri_ancom_top10$Kingdom <- paste0("Viridiplantae")


```

# Metazoa 

Relative abundance of Episarius

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_episauris <- subset_taxa(diet_fr_r, Genus %in% c("Episauris"))

genus_df <- psmelt(ps_genus_episauris)
head(genus_df)

episumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(epi_abund = mean(Abundance, na.rm=TRUE)) 
head(episumm)
tail(episumm)

episumm <- episumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), episumm, by = "sampleid")
head(DF)


```

Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$epi_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$epi_abund <- DF$epi_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "epi_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","epi_abund")]
out_res_se <- out_res$se[, c("taxon","epi_abund")]
out_res_W <- out_res$W[, c("taxon","epi_abund")]
out_res_q <- out_res$q_val[, c("taxon","epi_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","epi_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_epi <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_epi) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_epi$diet <- paste("episauris")
head(out_epi)

# check which taxa 
table(out_res_diff$epi_abund)
diff_df<- out_epi %>% filter(diff_abund == TRUE)

```

Relative abundance of Brachyderes

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_brachyderes <- subset_taxa(diet_fr_r, Genus %in% c("Brachyderes"))

genus_df <- psmelt(ps_genus_brachyderes)
head(genus_df)

brachysumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(brachy_abund = mean(Abundance, na.rm=TRUE)) 
head(brachysumm)
tail(brachysumm)

brachysumm <- brachysumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), brachysumm, by = "sampleid")
head(DF)


```

Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$brachy_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$brachy_abund <- DF$brachy_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "brachy_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","brachy_abund")]
out_res_se <- out_res$se[, c("taxon","brachy_abund")]
out_res_W <- out_res$W[, c("taxon","brachy_abund")]
out_res_q <- out_res$q_val[, c("taxon","brachy_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","brachy_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_brachy <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_brachy) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_brachy$diet <- paste("brachyderes")
head(out_brachy)

# check which taxa 
table(out_res_diff$brachy_abund)
diff_df<- out_brachy %>% filter(diff_abund == TRUE)

```


Relative abundance of Xysticus

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
ps_genus_xyst <- subset_taxa(diet_fr_r, Genus %in% c("Xysticus"))

genus_df <- psmelt(ps_genus_xyst)
head(genus_df)

xystsumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(xyst_abund = mean(Abundance, na.rm=TRUE)) 
head(xystsumm)
tail(xystsumm)

xystsumm <- xystsumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), xystsumm, by = "sampleid")
head(DF)

```

Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$xyst_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$xyst_abund <- DF$xyst_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "xyst_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","xyst_abund")]
out_res_se <- out_res$se[, c("taxon","xyst_abund")]
out_res_W <- out_res$W[, c("taxon","xyst_abund")]
out_res_q <- out_res$q_val[, c("taxon","xyst_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","xyst_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_xyst <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_xyst) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_xyst$diet <- paste("Xysticus")
head(out_xyst)

# check which taxa 
table(out_res_diff$xyst_abund)
diff_df<- out_xyst %>% filter(diff_abund == TRUE)

```


Relative abundance of Mythimna

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_myth <- subset_taxa(diet_fr_r, Genus %in% c("Mythimna"))

genus_df <- psmelt(ps_genus_myth)
head(genus_df)

mythsumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(myth_abund = mean(Abundance, na.rm=TRUE)) 
head(mythsumm)
tail(mythsumm)

mythsumm <- mythsumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), mythsumm, by = "sampleid")
head(DF)

```


Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$myth_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$myth_abund <- DF$myth_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "myth_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","myth_abund")]
out_res_se <- out_res$se[, c("taxon","myth_abund")]
out_res_W <- out_res$W[, c("taxon","myth_abund")]
out_res_q <- out_res$q_val[, c("taxon","myth_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","myth_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_myth <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_myth) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_myth$diet <- paste("mythimna")
head(out_myth)

# check which taxa 
table(out_res_diff$myth_abund)
diff_df<- out_myth %>% filter(diff_abund == TRUE)

```


Relative abundance of udea

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_udea <- subset_taxa(diet_fr_r, Genus %in% c("Udea"))

genus_df <- psmelt(ps_genus_udea)
head(genus_df)

udeasumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(udea_abund = mean(Abundance, na.rm=TRUE)) 
head(udeasumm)
tail(udeasumm)

udeasumm <- udeasumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), udeasumm, by = "sampleid")
head(DF)

```

Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$udea_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$udea_abund <- DF$udea_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "udea_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","udea_abund")]
out_res_se <- out_res$se[, c("taxon","udea_abund")]
out_res_W <- out_res$W[, c("taxon","udea_abund")]
out_res_q <- out_res$q_val[, c("taxon","udea_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","udea_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_udea <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_udea) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_udea$diet <- paste("udea")
head(out_udea)

# check which taxa 
table(out_res_diff$udea_abund)
diff_df<- out_udea %>% filter(diff_abund == TRUE)

```


Relative abundance of talitroides

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_genus_talitroides <- subset_taxa(diet_fr_r, Genus %in% c("Talitroides"))

genus_df <- psmelt(ps_genus_talitroides)
head(genus_df)

talitroidessumm <- genus_df %>%
  group_by(sampleid, Genus) %>%
  summarize(talitroides_abund = mean(Abundance, na.rm=TRUE)) 
head(talitroidessumm)
tail(talitroidessumm)

talitroidessumm <- talitroidessumm[,c(1,3)]

DF <- merge(data.frame(sample_data(diet_fr)), talitroidessumm, by = "sampleid")
head(DF)

```

Running ancom

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

DF$talitroides_abund
#sample_data(microb_fr)$Lab_Code == DF$Lab_Code
sample_data(microb_fr)$talitroides_abund <- DF$talitroides_abund
names(sample_data(microb_fr))

# run ancombc function
set.seed(123)
out = ancombc(data = microb_fr, tax_level = "Species",
formula = "talitroides_abund + Species_Label",
p_adj_method = "holm",prv_cut = 0.1, lib_cut = 0,
group = "Species_Label", struc_zero = FALSE, neg_lb = FALSE,
tol = 1e-5, max_iter = 100, conserve = TRUE,
alpha = 0.05, global = TRUE, n_cl = 1, verbose = TRUE)

str(out)
out_res <- out$res
out_res_lfc <- out_res$lfc[, c("taxon","talitroides_abund")]
out_res_se <- out_res$se[, c("taxon","talitroides_abund")]
out_res_W <- out_res$W[, c("taxon","talitroides_abund")]
out_res_q <- out_res$q_val[, c("taxon","talitroides_abund")]
out_res_diff <- out_res$diff_abn[, c("taxon","talitroides_abund")]

l <- list(out_res_lfc, out_res_se, out_res_W,out_res_q, out_res_diff)
str(l)
out_talitroides <- purrr::reduce(.x = l, merge, by = ("taxon"), all = T)
names(out_talitroides) <- c("taxon", "lfc", "se", "W",  "q_val", "diff_abund")
out_talitroides$diet <- paste("talitroides")
head(out_talitroides)

# check which taxa 
table(out_res_diff$talitroides_abund)
diff_df<- out_talitroides %>% filter(diff_abund == TRUE)

```

merge metazoa df

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

df_metaz_ancom_topsp <- rbind(out_brachy, out_epi,  out_myth, out_talitroides, out_udea, out_xyst)

df_metaz_ancom_topsp <- df_metaz_ancom_topsp 
#%>% filter(diff_abund == TRUE)
df_metaz_ancom_topsp$Kingdom <- paste0("Metazoa")
str(df_metaz_ancom_topsp)
head(df_metaz_ancom_topsp)
df_overview <- (table(df_metaz_ancom_topsp$taxon, df_metaz_ancom_topsp$diet))
str(df_overview)
table(df_metaz_ancom_topsp$taxon)


#df_metaz_ancom_top10 <- rbind(out_brachy, out_epi,  out_myth, out_talitroides, out_udea, out_xyst, out_manto, out_colept, out_thysan, out_erio)

#df_metaz_ancom_top10 <- df_metaz_ancom_top10 %>% filter(diff_abund == TRUE)
#df_metaz_ancom_top10$Kingdom <- paste0("Metazoa")

```

# Merge Viri and Metaz

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

str(df_metaz_ancom_topsp)
str(df_viri_ancom_topgenus)

out_ancom_all <- rbind(df_viri_ancom_topgenus, df_metaz_ancom_topsp)
head(out_ancom_all)
str(out_ancom_all)
table(out_ancom_all$diet, out_ancom_all$Kingdom)

write.csv(out_ancom_all, paste0(dataDir, "/ancomdiet_toptaxa.csv"), row.names = FALSE)
```











