---
title: 'Effect of colonisation history and biogeography'
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: html_document

---


## LOAD LIBRARIES

```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rm(list = ls())

# basics
library(tibble)
library(dplyr)
library(tidyverse)
library(data.table)
library(ggtern)
library(patchwork)

# phyloseq section
library(phyloseq)
library(Biostrings)
library(microbiome)
library(biomeUtils)
library(microViz)

# diversity 
library(adespatial)


#rmd
library(details)
library(knitr)
library(kableExtra)

#models
library(MASS)
library(MuMIn)
library(car)
library(multcomp)
library(glmmTMB)
library(ggeffects)

```
## WORKING DIRECTORIES

```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```

## LOAD 16S DATA
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_fringilla_rare <- readRDS(paste0(dataDir, "/ps_OTU99_wAlpha_fr_taxFix.rds" ))

ps_fringilla_rare <- biomeUtils::removeZeros(ps_fringilla_rare)

ps_fringilla_rare

ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Species")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Genus")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Family")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Order")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Class")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Phylum")


ps_fringilla_rare <- biomeUtils::removeZeros(ps_fringilla_rare)

```

## TESTING THE EFFECT OF COLONISATION HISTORY

### ALPHA DIVERSITY

Is there a sequential reduction of alpha diversity as predicted by colonization history? (i.e. mainland > Azores > Madeira > Canary Islands) 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rich <- estimate_richness(ps_fringilla_rare, measures=c("Observed", "Shannon"))
pd <- calculatePD(ps_fringilla_rare, justDF = TRUE)

sample_data(ps_fringilla_rare)$Observed_sps_rare <- rich$Observed
sample_data(ps_fringilla_rare)$PD_sps_rare <- pd$PD
sample_data(ps_fringilla_rare)$PD_integer <- round(pd$PD,0)

summary(sample_data(ps_fringilla_rare)$Observed_sps_rare )
alpha_df <- data.frame(sample_data(ps_fringilla_rare))

mainland_fringilla <- alpha_df %>% dplyr::filter(!is.na(Age_cat) & !is.na(Sex))  

mainland_fringilla$Species_Label <- as.factor(mainland_fringilla$Species_Label)
mainland_fringilla$Species_Label <- relevel(mainland_fringilla$Species_Label, "Mainland_F_coelebs")

names(mainland_fringilla)

GLM1 <- glm.nb(Observed_sps_rare ~ Age_cat + Sex + Species_Label, data = mainland_fringilla,na.action = na.fail)
summary(GLM1)
dredge(GLM1)
Anova(GLM1, type = "III")
#Analysis of Deviance Table (Type III tests)

#Response: Observed_sps_rare
#              LR Chisq Df Pr(>Chisq)    
#Age_cat          0.029  1     0.8640    
#Sex              1.527  1     0.2165    
#Species_Label   44.707 10  2.456e-06 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


GLM2 <- glm.nb(Observed_sps_rare ~ Species_Label, data = mainland_fringilla,na.action = na.fail)
summary(GLM2)
dredge(GLM2)
Anova(GLM2, type = "III")
mcp <- glht(GLM2, linfct = mcp(Species_Label = "Tukey"))
summary(mcp)


CI <- as_tibble(confint(GLM2), rownames = 'predictor') 

CI <- cbind(CI,as_tibble(summary(GLM2)$coefficients))

CI$CImin <- c(CI$`2.5 %`[1],CI$Estimate[1]+CI$`2.5 %`[c(-1)])
CI$CImax <- c(CI$`97.5 %`[1],CI$Estimate[1]+CI$`97.5 %`[c(-1)])
CI$Estimate.transformed <- c(CI$`Estimate`[1],CI$`Estimate`[1]+CI$`Estimate`[c(-1)])

CI$Estimate <- exp(CI$Estimate.transformed)
CI$CImin <- exp(CI$CImin)
CI$CImax <- exp(CI$CImax)

CI <- CI[c(1,4,5,6,7,8,9)]

estimate_obj <- CI
write_tsv(CI,paste0(outputDir, "/07_Stats/02_analyses/glm_rare_sobs_speciesFix.tsv"))

```

No evidence of alpha diversity mirroring the colonization history of the chaffinch spp complex. However, there is an effect (Chisq = 46.276, df = 10, P < 0.001) of species within islands but no effect of age or sex. We should then control for Species Label in subsequent models.

Plot richness according to species within islands:

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

mainland_fringilla <- mainland_fringilla %>% filter(Genus == "Fringilla") %>%    mutate( Labels = Species_Label %>% str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>% str_replace("_F_", " : F.") %>% str_replace("c_", "c.") %>% str_replace("_", " ")) 

mainland_fringilla$Order <- factor(mainland_fringilla$Labels,levels=c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))

estimate_obj <- read_tsv(paste0(outputDir, "/07_Stats/02_analyses/glm_rare_sobs_speciesFix.tsv"))

model_estimates <- estimate_obj %>% 
    mutate( value = Estimate + estimate_obj[[1,2]]) %>% 
    mutate( value = ifelse(predictor == '(Intercept)', Estimate, value))  %>% 
    mutate( island = predictor %>%
              str_replace("as.factor", "") %>%
              str_replace("\\(Species_Label\\)", "") %>%
              str_replace("\\(Intercept\\)", "Mainland : F.coelebs") %>%
              str_replace("Sexmale", "Male") %>%
              str_replace("Age_catfledging", "Fledging") %>%
              str_replace("_F_", " : F.") %>%
              str_replace("c_", "c.") %>%
              str_replace("_", " ")) 

species_estimates <- model_estimates
species_estimates$Order <- factor(species_estimates$island, levels = c("Mainland : F.coelebs", "Species LabelSao_Miguel : F.moreletti", "Species LabelFlores : F.moreletti", "Species LabelMadeira : F.maderensis", "Species LabelEl_Hierro : F.c.ombriosa", "Species LabelLa_Palma : F.c.palmae", "Species LabelGran_Canaria : F.c.bakeri", "Species LabelTenerife : F.c.canariensis", "Species LabelLa_Gomera : F.c.canariensis", "Species LabelTenerife : F.teydea", "Species LabelGran_Canaria : F.polatzeki"), labels = c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))
species_estimates$Order <- fct_rev(species_estimates$Order)

mainland_fringilla$Chaff_Blue2 <- as.character(mainland_fringilla$Chaff_Blue)

mainland_fringilla$Chaff_Blue2 <- ifelse(mainland_fringilla$Chaff_Blue2 == "Chaffinch", "Common chaffinch", "Blue chaffinch")

mainland_fringilla$Chaff_Blue2  <- as.factor(mainland_fringilla$Chaff_Blue2)

mainland_fringilla$Chaff_Blue2 <- forcats::fct_relevel(mainland_fringilla$Chaff_Blue2,"Common chaffinch","Blue chaffinch")
table(mainland_fringilla$Chaff_Blue2)

signif2 = distinct(mainland_fringilla, Chaff_Blue2,Order) %>%
    arrange(Chaff_Blue2,Order)

signif2<-left_join(signif2,species_estimates,by="Order")

data.frame(signif2)

#signif2$Estimate
signif2$p.signif = c("","","","","","","*","","*","**","*")

mainland_fringilla$Chaff_Blue2

levels(mainland_fringilla$Species_Label)

(p1 <- ggplot(mainland_fringilla, aes(Order, Observed_sps_rare)) + 
        facet_grid(~Chaff_Blue2,scales = "free_x",space = "free_x") +
        geom_boxplot(aes(fill = Species_Label), show.legend = F, lwd = 0.25,fatten = 1, alpha = 0.5, outlier.colour = 'transparent',width=0.5) + 
        geom_point( aes(fill = Species_Label),shape = 21, position = position_jitter(seed = 100, width = 0.1), size = 3,  alpha = 0.7, stroke = 0.25) + 
        SpeciesLabel.fill + 
        ylab("Richness") + 
        xlab("") + sl_theme + 
        boxplot_theme +
        scale_x_discrete(breaks = c("Tenerife : F.teydea","Gran Canaria : F.polatzeki","Mainland : F.coelebs","Sao Miguel : F.moreletti","Flores : F.moreletti","Madeira : F.maderensis","La Palma : F.c.palmae","El Hierro : F.c.ombriosa","Gran Canaria : F.c.bakeri","Tenerife : F.c.canariensis","La Gomera : F.c.canariensis"),labels = c("Tenerife","Gran Canaria","Mainland","Sao Miguel","Flores","Madeira","La Palma","El Hierro","Gran Canaria","Tenerife","La Gomera")) + 
        geom_hline(aes(yintercept = signif2$Estimate[1]), linetype='dashed') +  
        #geom_point(data = signif2, aes( x = Order, y = Estimate),fill = '#acacac', shape = 23, size = 2, stroke = 0.25) + 
        stat_summary(aes(x = Order, y = Observed_sps_rare),fun.data="mean_cl_boot", geom="pointrange",shape = 23, col = "#acacac", fill = '#acacac',size=0.5,fun.args=list(conf.int=.95)) + 
        #geom_linerange(data = signif2, aes( x = Order,ymin = CImin, ymax = CImax), inherit.aes = F, col = '#acacac', size = 0.75) + 
        geom_text(data = signif2, aes(x = Order, y = 250, label = p.signif), vjust = 0.75, size = 8) + scale_y_continuous(breaks = seq(0, 250, by=50)
    #+scale_y_log10()
    ))



```

To account for phylogenetic relationships between bacterial taxa we repeated the above analysis using Faith's PD.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

GLM1 <- glm.nb(PD_integer ~ Age_cat + Sex + Species_Label, data = mainland_fringilla,na.action = na.fail)
summary(GLM1)
dredge(GLM1)
Anova(GLM1, type = "III")



GLM2 <- glm.nb(PD_integer ~ Species_Label, data = mainland_fringilla,na.action = na.fail)
summary(GLM2)
dredge(GLM2)
Anova(GLM2, type = "III")
mcp <- glht(GLM2, linfct = mcp(Species_Label = "Tukey"))
summary(mcp)



CI <- as_tibble(confint(GLM2), rownames = 'predictor') 

CI <- cbind(CI,as_tibble(summary(GLM2)$coefficients))

CI$CImin <- c(CI$`2.5 %`[1],CI$Estimate[1]+CI$`2.5 %`[c(-1)])
CI$CImax <- c(CI$`97.5 %`[1],CI$Estimate[1]+CI$`97.5 %`[c(-1)])
CI$Estimate.transformed <- c(CI$`Estimate`[1],CI$`Estimate`[1]+CI$`Estimate`[c(-1)])

CI$Estimate <- exp(CI$Estimate.transformed)
CI$CImin <- exp(CI$CImin)
CI$CImax <- exp(CI$CImax)

CI <- CI[c(1,4,5,6,7,8,9)]

estimate_obj <- CI

write_tsv(CI,paste0(outputDir, "/07_Stats/02_analyses/glm_rare_pd_speciesFix.tsv"))


mainland_fringilla <- mainland_fringilla %>% filter(Genus == "Fringilla") %>%    mutate( Labels = Species_Label %>% str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>% str_replace("_F_", " : F.") %>% str_replace("c_", "c.") %>% str_replace("_", " ")) 

mainland_fringilla$Order <- factor(mainland_fringilla$Labels,levels=c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))


estimate_obj <- read_tsv(paste0(outputDir, "/07_Stats/02_analyses/glm_rare_pd_speciesFix.tsv"))

model_estimates <- estimate_obj %>% 
    mutate( value = Estimate + estimate_obj[[1,2]]) %>% 
    mutate( value = ifelse(predictor == '(Intercept)', Estimate, value))  %>% 
    mutate( island = predictor %>%
              str_replace("as.factor", "") %>%
              str_replace("\\(Species_Label\\)", "") %>%
              str_replace("\\(Intercept\\)", "Mainland : F.coelebs") %>%
              str_replace("Sexmale", "Male") %>%
              str_replace("Age_catfledging", "Fledging") %>%
              str_replace("_F_", " : F.") %>%
              str_replace("c_", "c.") %>%
              str_replace("_", " ")) 

species_estimates <- model_estimates
species_estimates$Order <- factor(species_estimates$island, levels = c("Mainland : F.coelebs", "Species LabelSao_Miguel : F.moreletti", "Species LabelFlores : F.moreletti", "Species LabelMadeira : F.maderensis", "Species LabelEl_Hierro : F.c.ombriosa", "Species LabelLa_Palma : F.c.palmae", "Species LabelGran_Canaria : F.c.bakeri", "Species LabelTenerife : F.c.canariensis", "Species LabelLa_Gomera : F.c.canariensis", "Species LabelTenerife : F.teydea", "Species LabelGran_Canaria : F.polatzeki"), labels = c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))
species_estimates$Order <- fct_rev(species_estimates$Order)

mainland_fringilla$Chaff_Blue2 <- as.character(mainland_fringilla$Chaff_Blue)

mainland_fringilla$Chaff_Blue2 <- ifelse(mainland_fringilla$Chaff_Blue2 == "Chaffinch", "Common chaffinch", "Blue chaffinch")

mainland_fringilla$Chaff_Blue2  <- as.factor(mainland_fringilla$Chaff_Blue2)

mainland_fringilla$Chaff_Blue2 <- forcats::fct_relevel(mainland_fringilla$Chaff_Blue2,"Common chaffinch","Blue chaffinch")
table(mainland_fringilla$Chaff_Blue2)

signif2_pd = distinct(mainland_fringilla, Chaff_Blue2,Order) %>%
    arrange(Chaff_Blue2,Order)

signif2_pd<-left_join(signif2_pd,species_estimates,by="Order")

data.frame(signif2_pd)

#signif2_pd$Estimate
signif2_pd$p.signif = c("","","","","","","**","","**","***","**")


(p_PD <- ggplot(mainland_fringilla, aes(Order, PD_sps_rare)) + 
        facet_grid(~Chaff_Blue2,scales = "free_x",space = "free_x") +
        geom_boxplot(aes(fill = Species_Label), show.legend = F, lwd = 0.25,fatten = 1, alpha = 0.5, outlier.colour = 'transparent',width=0.5) + 
        geom_point( aes(fill = Species_Label),shape = 21, position = position_jitter(seed = 100, width = 0.1), size = 3,  alpha = 0.7, stroke = 0.25) + 
        SpeciesLabel.fill + 
        ylab("Faith's phylogenetic diversity") + 
        xlab("") + sl_theme + 
        boxplot_theme +
        scale_x_discrete(breaks = c("Tenerife : F.teydea","Gran Canaria : F.polatzeki","Mainland : F.coelebs","Sao Miguel : F.moreletti","Flores : F.moreletti","Madeira : F.maderensis","La Palma : F.c.palmae","El Hierro : F.c.ombriosa","Gran Canaria : F.c.bakeri","Tenerife : F.c.canariensis","La Gomera : F.c.canariensis"),labels = c("Tenerife","Gran Canaria","Mainland","Sao Miguel","Flores","Madeira","La Palma","El Hierro","Gran Canaria","Tenerife","La Gomera")) + 
        geom_hline(aes(yintercept = signif2_pd$Estimate[1]), linetype='dashed') +  
        #geom_point(data = signif2_pd, aes( x = Order, y = Estimate),fill = '#acacac', shape = 23, size = 2, stroke = 0.25) + 
        stat_summary(aes(x = Order, y = PD_sps_rare),fun.data="mean_cl_boot", geom="pointrange",shape = 23, col = "#acacac", fill = '#acacac',size=0.5,fun.args=list(conf.int=.95)) + 
        #geom_linerange(data = signif2_pd, aes( x = Order,ymin = CImin, ymax = CImax), inherit.aes = F, col = '#acacac', size = 0.75) + 
        geom_text(data = signif2_pd, aes(x = Order, y = 40, label = p.signif), vjust = 0.75, size = 8) + scale_y_continuous(breaks = seq(0, 40, by=5)
    #+scale_y_log10()
    ))

```

We get equivalent results with Faith's PD.

### PARTITION BETA DIVERSITY

#### Distance metric Baselga family

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

Matrix_abund <- data.frame(otu_table(ps_fringilla_rare))
str(Matrix_abund)

#Baselga family

# Compute and partition a matrix of Jaccard indices (presence-absence data)
j.matrix <- beta.div.comp(Matrix_abund, coef = "BJ", quant = FALSE, save.abc = FALSE)

# Percentage difference, (quantitative form of Sorensen index)
#s.quant <- beta.div.comp(Matrix_abund, coef = "S", quant = TRUE, save.abc = FALSE)

#the percentage difference index (aka Bray-Curtis)
s.quant <- beta.div.comp(Matrix_abund, coef = "BS", quant = TRUE, save.abc = FALSE)


```

#### (Percentage difference, (quantitative form of Sorensen index) (a.k.a. Bray-Curtis))

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

##Replacement values
# dataframe of all pairwise comparisons
xy_repl <- t(combn(colnames(as.matrix(s.quant$repl)), 2))

# add replacement values to df
dist_repl_df <- (data.frame(xy_repl, repl=as.matrix(s.quant$repl)[xy_repl]))

# concatenate labels
dist_repl_df$X1_X2 <- paste0(dist_repl_df$X1,"_",dist_repl_df$X2)

##Richness difference values
# dataframe of all pairwise comparisons
xy_rich <- t(combn(colnames(as.matrix(s.quant$rich)), 2))

# add richness difference values to df
dist_rich_df <- (data.frame(xy_rich, rich=as.matrix(s.quant$rich)[xy_rich]))

# concatenate labels
dist_rich_df$X1_X2 <- paste0(dist_rich_df$X1,"_",dist_rich_df$X2)


## Dissimilarity values

# dataframe of all pairwise comparisons 
xy_D <- t(combn(colnames(as.matrix(s.quant$D)), 2))

# add dissimilarity values to df
dist_D_df <- (data.frame(xy_D, D=as.matrix(s.quant$D)[xy_D]))

# concatenate labels
dist_D_df$X1_X2 <- paste0(dist_D_df$X1,"_",dist_D_df$X2)

## merge all df
dist_df <- (cbind(dist_repl_df,dist_rich_df,dist_D_df)[,c(4,1:3,7,11)])

## Metadata microbiome
Meta <- data.frame(sample_data(ps_fringilla_rare))

# create df for Species Label
SpeciesLabels_X1 <- data.frame(X1 = paste0(Meta$Lab_Code,"_",Meta$Species_Label)
,Species_Label_X1 = Meta$Species_Label)

SpeciesLabels_X2 <- data.frame(X2 = paste0(Meta$Lab_Code,"_",Meta$Species_Label)
,Species_Label_X2 = Meta$Species_Label)

# creating df for Archipelago_Mainland

Archipelago_Mainland_X1 <- data.frame(X1 = paste0(Meta$Lab_Code,"_",Meta$Species_Label),Archipelago_Mainland_X1 = Meta$Archipelago_Mainland)
Archipelago_Mainland_X2 <- data.frame(X2 = paste0(Meta$Lab_Code,"_",Meta$Species_Label),Archipelago_Mainland_X2 = Meta$Archipelago_Mainland)

# Now give labels to distance matrix based on metadata to be able to subset data

# Species label 
dist_df_X1_site <- (left_join(dist_df,SpeciesLabels_X1,by = "X1"))
dist_df_X2_site <- (left_join(dist_df_X1_site,SpeciesLabels_X2,by = "X2"))

# archipelago mainland
dist_df_X2_site <- (left_join(dist_df_X2_site,Archipelago_Mainland_X1,by = "X1"))
dist_df_X2_site <- (left_join(dist_df_X2_site,Archipelago_Mainland_X2,by = "X2"))

# format separator
dist_df_X2_site$X1_X2_Species_Label <- paste0(dist_df_X2_site$Species_Label_X1,":",dist_df_X2_site$Species_Label_X2)
dist_df_X2_site$X1_X2_Same_site <- dist_df_X2_site$Species_Label_X1 == dist_df_X2_site$Species_Label_X2

dist_df_X2_site$X1_X2_Archipelago_Mainland <- paste0(dist_df_X2_site$Archipelago_Mainland_X1,":",dist_df_X2_site$Archipelago_Mainland_X2)

dist_df_diff_site <- dist_df_X2_site[which(dist_df_X2_site$X1_X2_Same_site==FALSE),]
dist_df_same_site <- dist_df_X2_site[which(dist_df_X2_site$X1_X2_Same_site==TRUE),]


# distance matrix where only comparisons with mainland are kept
dist_df_diff_mainland <- dist_df_diff_site[dist_df_diff_site$X1_X2_Species_Label %like% "Mainland_F_coelebs", ]


## Create subsets
dist_df_common <- dist_df_X2_site[which(!dist_df_X2_site$Species_Label_X1=="Tenerife_F_teydea" & !dist_df_X2_site$Species_Label_X1=="Gran_Canaria_F_polatzeki" & !dist_df_X2_site$Species_Label_X2=="Tenerife_F_teydea" & !dist_df_X2_site$Species_Label_X2=="Gran_Canaria_F_polatzeki") ,]


# common chaffinch subset 
dist_df_diff_common <- dist_df_common[which(dist_df_common$X1_X2_Same_site==FALSE),]
head(dist_df_diff_common)

## Make dissimilarity a similarity vector
dist_df_diff_site$S <- (1-dist_df_diff_site$D)
dist_df_diff_mainland$S <- (1-dist_df_diff_mainland$D)
dist_df_diff_common$S <- (1-dist_df_diff_common$D)

# everything sums up to 1 ;) 
dist_df_diff_site$S+dist_df_diff_site$repl+dist_df_diff_site$rich

```



Plot per site variation of beta components
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
dist_df_diff_mainland$site_interest <- gsub('Mainland_F_coelebs','', dist_df_diff_mainland$X1_X2_Species_Label)

dist_df_diff_mainland$site_interest <- gsub(':','', dist_df_diff_mainland$site_interest)


dist_df_diff_mainland <-dist_df_diff_mainland %>%
    mutate( Labels = str_replace(site_interest, "site_interest", "")         %>%
              str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%
              str_replace("_F_", " : F.") %>%
              str_replace("c_", "c.") %>%
              str_replace("_", " "))   



dist_df_diff_mainland$Order <-factor(dist_df_diff_mainland$Labels,levels=c("Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"), labels = c("Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')","Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')","La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')","Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')","Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))


dist_df_diff_mainland$SpeciesComplex<-factor(dist_df_diff_mainland$Order,labels=c("Common chaffinch species complex","Common chaffinch species complex","Common chaffinch species complex","Common chaffinch species complex","Common chaffinch species complex","Common chaffinch species complex","Common chaffinch species complex","Common chaffinch species complex","Blue finches","Blue finches")) 



head(dist_df_diff_mainland)
length(dist_df_diff_mainland)
names(dist_df_diff_mainland)

dist_df_diff_mainland_test <- rbind(dist_df_diff_mainland,data.frame(X1_X2 = "mainland_mainland", X1 = "Mainland_F_coelebs", X2 = "Mainland_F_coelebs", repl = NA, rich = NA, D = NA, Species_Label_X1 = "Mainland_F_coelebs", Species_Label_X2 = "Mainland_F_coelebs", "Archipelago_Mainland_X1" = "Mainland", Archipelago_Mainland_X2 = "Mainland", X1_X2_Species_Label = "Mainland_F_coelebs:Mainland_F_coelebs", X1_X2_Same_site = TRUE, X1_X2_Archipelago_Mainland = "Mainland:Mainland", S = "", site_interest = "Mainland", Labels = "Mainland:F.coelebs", Order = "Mainland:~italic('F. coelebs')", SpeciesComplex = "Common chaffinch species complex"))

tail(dist_df_diff_mainland_test)

dist_df_diff_mainland_test$Order <-factor(dist_df_diff_mainland_test$Order,levels=c("Mainland:~italic('F. coelebs')", "Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')","Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')","La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')","Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')","Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"), labels = c("Mainland:~italic('F. coelebs')", "Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')","Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')","La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')","Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')","Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))



levels(dist_df_diff_mainland_test$Order)


(D_overall_bray <- ggplot(data = dist_df_diff_mainland_test, aes(Order, D)) + 
        geom_boxplot(aes(fill = Order), show.legend = F, lwd = 0.25,fatten = 1, alpha = 0.7, outlier.colour = 'transparent',width=0.5) + 
        geom_point ( aes(fill = Order),shape = 21, position = position_jitter(seed = 100, width = 0.1), size = 3,  alpha = 0.5, stroke = 0.25) + 
        xlab("") + ylab("Dissimilarity") + facet_grid(.~SpeciesComplex,scales = "free_x",space = "free_x") + theme_classic(base_size=15) +
        #scale_x_discrete(labels = function(l) parse(text=l)) +
        scale_x_discrete(breaks = c("Mainland:~italic('F. coelebs')", "Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')","Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')","La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')","Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')","Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"),labels = c("Mainland","Sao Miguel","Flores","Madeira","El Hierro","La Palma","Gran Canaria","Tenerife","La Gomera","Tenerife","Gran Canaria")) + 
        SpeciesLabel.fill2 + sl_theme + 
        boxplot_theme +
        stat_summary(aes(x = Order, y = D),fun.data="mean_cl_boot", geom = "pointrange",shape = 23,col = '#acacac', fill = '#acacac',size=0.5)
        + theme(axis.text.x = element_text(angle = 45, hjust=1)))



dat_text <- data.frame(
  label = c("Reference", ""),
  SpeciesComplex = factor(c("Common chaffinch species complex", "Blue finches"), levels = c("Common chaffinch species complex", "Blue finches")),
  x = c(0.5),
  y = c(0.35))

rect_data <- data.frame(
  maxs = c(0.5),
  mins = c(1.5),
  SpeciesComplex = c("Common chaffinch species complex"))



(D_overall_bray_patch <- D_overall_bray + geom_rect(data = dist_df_diff_mainland_test[which(dist_df_diff_mainland_test$SpeciesComplex == "Common chaffinch species complex"), ], aes(xmin=0.5, xmax=1.5, ymin=-Inf, ymax=Inf), alpha = 0.5, fill = 'lightgrey') + geom_text(data = dat_text, mapping = aes(x = x, y = y, label = label), hjust = -0.25, vjust = 2.5, angle = 90, size = 7) + theme(strip.background = element_blank(), strip.text.x = element_blank()))


```







Triangle Plots 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(triangle.overall.diff.bray <-ggtern(data=dist_df_diff_common,aes(rich,repl,S)) +  geom_point(size = 4,  alpha = 0.05, stroke = 0.25) + geom_crosshair_tern(aes(mean(dist_df_diff_common$rich),mean( dist_df_diff_common$repl), mean(dist_df_diff_common$S)),col="darkred",lty="dashed", lineend = "round", size = 2.2) + 
    geom_point(aes(mean(dist_df_diff_common$rich),mean( dist_df_diff_common$repl), mean(dist_df_diff_common$S)),col='darkred',size=10) +
    geom_point(aes(x = mean(dist_df_diff_common$rich), y = 1-mean(dist_df_diff_common$rich),z = 0),col="darkred",size=10) + 
    annotate(size=12,geom  = 'text',x = mean(dist_df_diff_common$rich), y = 1-mean(dist_df_diff_common$rich),z = 0, label = paste(round(mean(dist_df_diff_common$rich)*100,2)), hjust = 1, vjust = -1, size = 5, col = "darkred") +  
    geom_point(aes(x = 0, y = mean(dist_df_diff_common$repl), z = 1-mean(dist_df_diff_common$repl)),col="darkred",size=10) + 
    annotate(size=12,geom  = 'text', x = 0, y = mean(dist_df_diff_common$repl), z = 1-mean(dist_df_diff_common$repl), label = paste(round(mean(dist_df_diff_common$repl)*100,2)), hjust = -0.3, vjust = 1, size = 5, col = "darkred") +  
    geom_point(aes(x = 1-mean(dist_df_diff_common$S), y = 0, z = mean(dist_df_diff_common$S)),col="darkred",size=8) + 
    annotate(size=12,geom  = 'text', x = 1-mean(dist_df_diff_common$S), y = 0, z = mean(dist_df_diff_common$S), label = paste(round(mean(dist_df_diff_common$S)*100,2)), hjust = 0.5, vjust = 2, size = 5, col = "darkred")+
    theme_classic() + theme_hidetitles() + theme_nomask() + theme(axis.text = element_text(size = 28), tern.axis.arrow = element_line(size = 3)))

```


### Beta partition using a matrix of Jaccard indices (presence-absence data)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

str(j.matrix)

##Replacement values
# dataframe of all pairwise comparisons
xy_repl <- t(combn(colnames(as.matrix(j.matrix$repl)), 2))

# add replacement values to df
dist_repl_df <- (data.frame(xy_repl, repl=as.matrix(j.matrix$repl)[xy_repl]))

# concatenate labels
dist_repl_df$X1_X2 <- paste0(dist_repl_df$X1,"_",dist_repl_df$X2)

##Richness difference values
# dataframe of all pairwise comparisons
xy_rich <- t(combn(colnames(as.matrix(j.matrix$rich)), 2))

# add richness difference values to df
dist_rich_df <- (data.frame(xy_rich, rich=as.matrix(j.matrix$rich)[xy_rich]))

# concatenate labels
dist_rich_df$X1_X2 <- paste0(dist_rich_df$X1,"_",dist_rich_df$X2)


## Dissimilarity values

# dataframe of all pairwise comparisons 
xy_D <- t(combn(colnames(as.matrix(j.matrix$D)), 2))

# add dissimilarity values to df
dist_D_df <- (data.frame(xy_D, D=as.matrix(j.matrix$D)[xy_D]))

# concatenate labels
dist_D_df$X1_X2 <- paste0(dist_D_df$X1,"_",dist_D_df$X2)

## merge all df
dist_df <- (cbind(dist_repl_df,dist_rich_df,dist_D_df)[,c(4,1:3,7,11)])

xy_repl <- t(combn(colnames(as.matrix(j.matrix$repl)), 2))
dist_repl_df <- (data.frame(xy_repl, repl=as.matrix(j.matrix$repl)[xy_repl]))
dist_repl_df$X1_X2 <- paste0(dist_repl_df$X1,"_",dist_repl_df$X2)

xy_rich <- t(combn(colnames(as.matrix(j.matrix$rich)), 2))
dist_rich_df <- (data.frame(xy_rich, rich=as.matrix(j.matrix$rich)[xy_rich]))
dist_rich_df$X1_X2 <- paste0(dist_rich_df$X1,"_",dist_rich_df$X2)


xy_D <- t(combn(colnames(as.matrix(j.matrix$D)), 2))
dist_D_df <- (data.frame(xy_D, D=as.matrix(j.matrix$D)[xy_D]))
dist_D_df$X1_X2 <- paste0(dist_D_df$X1,"_",dist_D_df$X2)

dist_df <- (cbind(dist_repl_df,dist_rich_df,dist_D_df)[,c(4,1:3,7,11)])

## Metadata microbiome
Meta <- data.frame(sample_data(ps_fringilla_rare))

# create df for Species Label
SpeciesLabels_X1 <- data.frame(X1 = paste0(Meta$Lab_Code,"_",Meta$Species_Label)
,Species_Label_X1 = Meta$Species_Label)

SpeciesLabels_X2 <- data.frame(X2 = paste0(Meta$Lab_Code,"_",Meta$Species_Label)
,Species_Label_X2 = Meta$Species_Label)

# creating df for Archipelago_Mainland

Archipelago_Mainland_X1 <- data.frame(X1 = paste0(Meta$Lab_Code,"_",Meta$Species_Label),Archipelago_Mainland_X1 = Meta$Archipelago_Mainland)
Archipelago_Mainland_X2 <- data.frame(X2 = paste0(Meta$Lab_Code,"_",Meta$Species_Label),Archipelago_Mainland_X2 = Meta$Archipelago_Mainland)

# Now give labels to distance matrix based on metadata to be able to subset data

# Species label 

j_dist_df_X1_site <- (left_join(dist_df,SpeciesLabels_X1,by = "X1"))
j_dist_df_X2_site <- (left_join(j_dist_df_X1_site,SpeciesLabels_X2,by = "X2"))

# archipelago mainland

j_dist_df_X2_site <- (left_join(j_dist_df_X2_site,Archipelago_Mainland_X1,by = "X1"))
j_dist_df_X2_site <- (left_join(j_dist_df_X2_site,Archipelago_Mainland_X2,by = "X2"))

# format separator

j_dist_df_X2_site$X1_X2_Species_Label <- paste0(j_dist_df_X2_site$Species_Label_X1,":",j_dist_df_X2_site$Species_Label_X2)
j_dist_df_X2_site$X1_X2_Same_site <- j_dist_df_X2_site$Species_Label_X1 == j_dist_df_X2_site$Species_Label_X2

j_dist_df_X2_site$X1_X2_Archipelago_Mainland <- paste0(j_dist_df_X2_site$Archipelago_Mainland_X1,":",j_dist_df_X2_site$Archipelago_Mainland_X2)

dist_df_diff_site.j <- j_dist_df_X2_site[which(j_dist_df_X2_site$X1_X2_Same_site==FALSE),]
dist_df_same_site.j <- j_dist_df_X2_site[which(j_dist_df_X2_site$X1_X2_Same_site==TRUE),]

## Create subsets

dist_df_common <- j_dist_df_X2_site[which(!j_dist_df_X2_site$Species_Label_X1=="Tenerife_F_teydea" & !j_dist_df_X2_site$Species_Label_X1=="Gran_Canaria_F_polatzeki" & !j_dist_df_X2_site$Species_Label_X2=="Tenerife_F_teydea" & !j_dist_df_X2_site$Species_Label_X2=="Gran_Canaria_F_polatzeki") ,]

## For all dataframes, create two subsets, one of pairwise comparisons of individuals of the same population (within pops), and on of pairwise comparisons of individuals from different sites (among pops)

# common chaffinch subset 

j_dist_df_diff_common <- dist_df_common[which(dist_df_common$X1_X2_Same_site==FALSE),]
dist_df_same_common <- dist_df_common[which(dist_df_common$X1_X2_Same_site==TRUE),]

## Make dissimilarity a similarity vector

dist_df_diff_site.j$S <- (1-dist_df_diff_site.j$D)
dist_df_same_site.j$S <- (1-dist_df_same_site.j$D)
dist_df_common$S <- (1-dist_df_common$D)
j_dist_df_diff_common$S <- (1-j_dist_df_diff_common$D)
dist_df_same_common$S <- (1-dist_df_same_common$D)


dist_df_diff_site.j$S+dist_df_diff_site.j$repl+dist_df_diff_site.j$rich


```



```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(triangle.overall.diff.jaccard<-ggtern(data=j_dist_df_diff_common,aes(rich,repl,S)) +  geom_point(size = 4,  alpha = 0.05, stroke = 0.25) + geom_crosshair_tern(aes(mean(j_dist_df_diff_common$rich),mean( j_dist_df_diff_common$repl), mean(j_dist_df_diff_common$S)),col="darkred",lty="dashed", lineend = "round", size = 2.2) + 
    geom_point(aes(mean(j_dist_df_diff_common$rich),mean( j_dist_df_diff_common$repl), mean(j_dist_df_diff_common$S)),col='darkred',size=10) +
    geom_point(aes(x = mean(j_dist_df_diff_common$rich), y = 1-mean(j_dist_df_diff_common$rich),z = 0),col="darkred",size=10) + 
    annotate(size=12,geom  = 'text',x = mean(j_dist_df_diff_common$rich), y = 1-mean(j_dist_df_diff_common$rich),z = 0, label = paste(round(mean(j_dist_df_diff_common$rich)*100,2)), hjust = 1, vjust = -1, size = 5, col = "darkred") +  
    geom_point(aes(x = 0, y = mean(j_dist_df_diff_common$repl), z = 1-mean(j_dist_df_diff_common$repl)),col="darkred",size=10) + 
    annotate(size=12,geom  = 'text', x = 0, y = mean(j_dist_df_diff_common$repl), z = 1-mean(j_dist_df_diff_common$repl), label = paste(round(mean(j_dist_df_diff_common$repl)*100,2)), hjust = -0.3, vjust = 1, size = 5, col = "darkred") +  
    geom_point(aes(x = 1-mean(j_dist_df_diff_common$S), y = 0, z = mean(j_dist_df_diff_common$S)),col="darkred",size=8) + 
    annotate(size=12,geom  = 'text', x = 1-mean(j_dist_df_diff_common$S), y = 0, z = mean(j_dist_df_diff_common$S), label = paste(round(mean(j_dist_df_diff_common$S)*100,2)), hjust = 0.5, vjust = 2, size = 5, col = "darkred")+
    theme_classic() + theme_hidetitles() + theme_nomask() + theme(axis.text = element_text(size = 28), tern.axis.arrow = element_line(size = 3)))

```

I repeat the overall dissimilarity plot but weighted and unweighted unifrac:

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# UNWEIGHTED UNIFRAC

unifrac_uw <- UniFrac(ps_fringilla_rare, weighted = FALSE)

dm_uw <- as.matrix(unifrac_uw)
sdata <- data.frame(sample_data(ps_fringilla_rare))

pairs_uw <- which(upper.tri(dm_uw), arr.ind = TRUE)

df_uw <- data.frame(
  X1 = rownames(dm_uw)[pairs_uw[,1]],
  X2 = rownames(dm_uw)[pairs_uw[,2]],
  D  = dm_uw[pairs_uw]
)

df_uw$Species_Label_X1 <- sdata[df_uw$X1, "Species_Label"]
df_uw$Species_Label_X2 <- sdata[df_uw$X2, "Species_Label"]
df_uw$X1_X2_Species_Label <- paste0(df_uw$Species_Label_X1, ":", df_uw$Species_Label_X2)

df_uw <- df_uw %>% filter(
  (Species_Label_X1 == "Mainland_F_coelebs" & Species_Label_X2 != "Mainland_F_coelebs") |
  (Species_Label_X2 == "Mainland_F_coelebs" & Species_Label_X1 != "Mainland_F_coelebs")
)

df_uw$site_interest <- ifelse(df_uw$Species_Label_X1 == "Mainland_F_coelebs",
                              as.character(df_uw$Species_Label_X2),
                              as.character(df_uw$Species_Label_X1))

df_uw <- df_uw %>%
  mutate(Labels = site_interest %>%
           str_replace("_F_", " : F.") %>%
           str_replace("c_", "c.") %>%
           str_replace("_", " "))

df_uw$Order <- factor(df_uw$Labels,
  levels = c("Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis",
             "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri",
             "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis",
             "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"),
  labels = c("Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')",
             "Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')",
             "La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')",
             "Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')",
             "Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))

df_uw$SpeciesComplex <- factor(df_uw$Order,
  labels = c("Common chaffinch species complex","Common chaffinch species complex",
             "Common chaffinch species complex","Common chaffinch species complex",
             "Common chaffinch species complex","Common chaffinch species complex",
             "Common chaffinch species complex","Common chaffinch species complex",
             "Blue finches","Blue finches"))

mainland_row_uw <- data.frame(
  X1 = "Mainland_F_coelebs", X2 = "Mainland_F_coelebs", D = NA,
  Species_Label_X1 = "Mainland_F_coelebs", Species_Label_X2 = "Mainland_F_coelebs",
  X1_X2_Species_Label = "Mainland_F_coelebs:Mainland_F_coelebs",
  site_interest = "Mainland", Labels = "Mainland:F.coelebs",
  Order = "Mainland:~italic('F. coelebs')",
  SpeciesComplex = "Common chaffinch species complex")

df_uw <- bind_rows(df_uw, mainland_row_uw)

df_uw$SpeciesComplex <- factor(df_uw$SpeciesComplex, levels = c("Common chaffinch species complex", "Blue finches"))

df_uw$Order <- factor(df_uw$Order,
  levels = c("Mainland:~italic('F. coelebs')",
             "Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')",
             "Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')",
             "La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')",
             "Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')",
             "Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))

# Plot unweighted UniFrac
dat_text_uw <- data.frame(
  label = c("Reference", ""),
  SpeciesComplex = factor(c("Common chaffinch species complex", "Blue finches"),
                          levels = c("Common chaffinch species complex", "Blue finches")),
  x = c(0.5), y = c(max(df_uw$D, na.rm = TRUE) * 0.9))

D_unifrac_uw <- ggplot(data = df_uw, aes(Order, D)) +
  geom_boxplot(aes(fill = Order), show.legend = F, lwd = 0.25, fatten = 1,
               alpha = 0.7, outlier.colour = 'transparent', width = 0.5) +
  geom_point(aes(fill = Order), shape = 21,
             position = position_jitter(seed = 100, width = 0.1),
             size = 3, alpha = 0.5, stroke = 0.25) +
  xlab("") + ylab("Unweighted UniFrac") +
  facet_grid(. ~ SpeciesComplex, scales = "free_x", space = "free_x") +
  theme_classic(base_size = 15) +
  scale_x_discrete(
    breaks = c("Mainland:~italic('F. coelebs')",
               "Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')",
               "Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')",
               "La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')",
               "Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')",
               "Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"),
    labels = c("Mainland","Sao Miguel","Flores","Madeira","El Hierro",
               "La Palma","Gran Canaria","Tenerife","La Gomera","Tenerife","Gran Canaria")) +
  SpeciesLabel.fill2 + sl_theme + boxplot_theme +
  stat_summary(aes(x = Order, y = D), fun.data = "mean_cl_boot",
               geom = "pointrange", shape = 23, col = '#acacac',
               fill = '#acacac', size = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

D_unifrac_uw_patch <- D_unifrac_uw +
  geom_rect(data = df_uw[which(df_uw$SpeciesComplex == "Common chaffinch species complex"), ],
            aes(xmin = 0.5, xmax = 1.5, ymin = -Inf, ymax = Inf),
            alpha = 0.5, fill = 'lightgrey') +
  geom_text(data = dat_text_uw, mapping = aes(x = x, y = y, label = label),
            hjust = 1, vjust = 2.5, angle = 90, size = 7) +
  theme(strip.background = element_blank(), strip.text.x = element_blank())

D_unifrac_uw_patch


# WEIGHTED UNIFRAC


unifrac_w <- UniFrac(ps_fringilla_rare, weighted = TRUE)

dm_w <- as.matrix(unifrac_w)

pairs_w <- which(upper.tri(dm_w), arr.ind = TRUE)

df_w <- data.frame(
  X1 = rownames(dm_w)[pairs_w[,1]],
  X2 = rownames(dm_w)[pairs_w[,2]],
  D  = dm_w[pairs_w]
)

df_w$Species_Label_X1 <- sdata[df_w$X1, "Species_Label"]
df_w$Species_Label_X2 <- sdata[df_w$X2, "Species_Label"]
df_w$X1_X2_Species_Label <- paste0(df_w$Species_Label_X1, ":", df_w$Species_Label_X2)

df_w <- df_w %>% filter(
  (Species_Label_X1 == "Mainland_F_coelebs" & Species_Label_X2 != "Mainland_F_coelebs") |
  (Species_Label_X2 == "Mainland_F_coelebs" & Species_Label_X1 != "Mainland_F_coelebs")
)

df_w$site_interest <- ifelse(df_w$Species_Label_X1 == "Mainland_F_coelebs",
                             as.character(df_w$Species_Label_X2),
                             as.character(df_w$Species_Label_X1))

df_w <- df_w %>%
  mutate(Labels = site_interest %>%
           str_replace("_F_", " : F.") %>%
           str_replace("c_", "c.") %>%
           str_replace("_", " "))

df_w$Order <- factor(df_w$Labels,
  levels = c("Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis",
             "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri",
             "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis",
             "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"),
  labels = c("Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')",
             "Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')",
             "La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')",
             "Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')",
             "Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))

df_w$SpeciesComplex <- factor(df_w$Order,
  labels = c("Common chaffinch species complex","Common chaffinch species complex",
             "Common chaffinch species complex","Common chaffinch species complex",
             "Common chaffinch species complex","Common chaffinch species complex",
             "Common chaffinch species complex","Common chaffinch species complex",
             "Blue finches","Blue finches"))

mainland_row_w <- data.frame(
  X1 = "Mainland_F_coelebs", X2 = "Mainland_F_coelebs", D = NA,
  Species_Label_X1 = "Mainland_F_coelebs", Species_Label_X2 = "Mainland_F_coelebs",
  X1_X2_Species_Label = "Mainland_F_coelebs:Mainland_F_coelebs",
  site_interest = "Mainland", Labels = "Mainland:F.coelebs",
  Order = "Mainland:~italic('F. coelebs')",
  SpeciesComplex = "Common chaffinch species complex")

df_w <- bind_rows(df_w, mainland_row_w)

df_w$SpeciesComplex <- factor(df_w$SpeciesComplex, levels = c("Common chaffinch species complex", "Blue finches"))

df_w$Order <- factor(df_w$Order,
  levels = c("Mainland:~italic('F. coelebs')",
             "Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')",
             "Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')",
             "La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')",
             "Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')",
             "Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))

# Plot weighted UniFrac
dat_text_w <- data.frame(
  label = c("Reference", ""),
  SpeciesComplex = factor(c("Common chaffinch species complex", "Blue finches"),
                          levels = c("Common chaffinch species complex", "Blue finches")),
  x = c(0.5), y = c(max(df_w$D, na.rm = TRUE) * 0.9))

D_unifrac_w <- ggplot(data = df_w, aes(Order, D)) +
  geom_boxplot(aes(fill = Order), show.legend = F, lwd = 0.25, fatten = 1,
               alpha = 0.7, outlier.colour = 'transparent', width = 0.5) +
  geom_point(aes(fill = Order), shape = 21,
             position = position_jitter(seed = 100, width = 0.1),
             size = 3, alpha = 0.5, stroke = 0.25) +
  xlab("") + ylab("Weighted UniFrac") +
  facet_grid(. ~ SpeciesComplex, scales = "free_x", space = "free_x") +
  theme_classic(base_size = 15) +
  scale_x_discrete(
    breaks = c("Mainland:~italic('F. coelebs')",
               "Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')",
               "Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')",
               "La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')",
               "Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')",
               "Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"),
    labels = c("Mainland","Sao Miguel","Flores","Madeira","El Hierro",
               "La Palma","Gran Canaria","Tenerife","La Gomera","Tenerife","Gran Canaria")) +
  SpeciesLabel.fill2 + sl_theme + boxplot_theme +
  stat_summary(aes(x = Order, y = D), fun.data = "mean_cl_boot",
               geom = "pointrange", shape = 23, col = '#acacac',
               fill = '#acacac', size = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

D_unifrac_w_patch <- D_unifrac_w +
  geom_rect(data = df_w[which(df_w$SpeciesComplex == "Common chaffinch species complex"), ],
            aes(xmin = 0.5, xmax = 1.5, ymin = -Inf, ymax = Inf),
            alpha = 0.5, fill = 'lightgrey') +
  geom_text(data = dat_text_w, mapping = aes(x = x, y = y, label = label),
            hjust = 1.25, vjust = 2.5, angle = 90, size = 7) +
  theme(strip.background = element_blank(), strip.text.x = element_blank())

D_unifrac_w_patch

```

# BIOGEOGRAPHY


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

nomainland_fringilla <- mainland_fringilla %>% dplyr::filter(!is.na(Area_km2) & Genus == "Fringilla")
str(nomainland_fringilla)

GLMM1 <- glmmTMB(Observed_sps_rare ~  scale(Area_km2) + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM1)

#Conditional model:
#                Estimate Std. Error z value Pr(>|z|)    
#(Intercept)      4.28862    0.08358   51.31   <2e-16 ***
#scale(Area_km2)  0.15658    0.07880    1.99   0.0469 *  
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


confint(GLMM1)

    
(pArea <- ggpredict(GLMM1, terms = "Area_km2 [all]") %>% plot(line_size = 1.5, color = "darkred") +
        geom_point(data = nomainland_fringilla, aes(y = Observed_sps_rare, x = Area_km2),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 50), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(alpha = 1, size = 0.4, stoke = 0.01) + 
        ylab("Richness") + 
        xlab(bquote(Area~(km^2))) +
       scatter_theme + ggtitle("") + sl_theme
        + annotate(geom="text",label="atop(paste(~beta~' = 0.157 [0.002, 0.311];'),~italic('p-value')~'= 0.047')",parse=T, x = 1000,  y = 240, size = 6)
        + scale_y_continuous(breaks = seq(0, 250, by=50)))

GLMM2 <- glmmTMB(Observed_sps_rare ~ scale(Distance_mainland) + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)

summary(GLMM2)
#Conditional model:
#                         Estimate Std. Error z value Pr(>|z|)    
#(Intercept)               4.28778    0.06611   64.86  < 2e-16 ***
#scale(Distance_mainland) -0.22620    0.06831   -3.31 0.000929 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


confint(GLMM2)

#GLMM2 <- glmmTMB(Observed_sps_rare ~ Distance_mainland + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
#summary(GLMM2)

(pDistance <- ggpredict(GLMM2, terms = "Distance_mainland [all]") %>% plot(line_size = 1.5, color = "darkred") +
        geom_point(data = nomainland_fringilla, aes(y = Observed_sps_rare, x = Distance_mainland),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 50), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(alpha = 1, size = 0.4, stoke = 0.01) + 
        ylab("") +
        xlab("Distance from mainland (km)") + 
        scatter_theme + ggtitle("") + sl_theme
        + annotate(geom="text",label="atop(paste(~beta~' = -0.226 [-0.360, -0.092];'),~italic('p-value')~'< 0.001')",parse=T, x = 1000,  y = 240, size = 6)
        + scale_y_continuous(breaks = seq(0, 250, by=50)))

GLMM3 <- glmmTMB(Observed_sps_rare ~  scale(Age_ma) + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM3)

#Conditional model:
#              Estimate Std. Error z value Pr(>|z|)    
#(Intercept)    4.29167    0.05827   73.65  < 2e-16 ***
#scale(Age_ma)  0.24298    0.05877    4.13 3.56e-05 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


confint(GLMM3)

#GLMM3 <- glmmTMB(Observed_sps_rare ~  Age_ma + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
#summary(GLMM3)

#slopes(GLMM3)

(pAge_ma <- ggpredict(GLMM3, terms = "Age_ma [all]") %>% plot(line_size = 1.5, color = "darkred") +
        geom_point(data = nomainland_fringilla, aes(y = Observed_sps_rare, x = Age_ma),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 0.3), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(col = "darkred", alpha = 1, size = 0.4, stoke = 0.01) +
        ylab("") +
        xlab("Island age (mya)") + 
        scatter_theme + ggtitle("") + sl_theme
        + annotate(geom="text",label="atop(paste(~beta~' = 0.243 [0.128, 0.3358];'),~italic('p-value')~'< 0.001')",parse=T, x = 7.5,  y = 240, size = 6)
        + scale_y_continuous(breaks = seq(0, 250, by=50)))

```

## Patchwork
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

png(paste0(outputDir, "/Figures/Colonisation_history_speciesFix.png"), width = 300, height = 240, units = "mm", res = 350)
((p1 + theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank(), strip.text.x = element_text(size = 18))) / D_overall_bray_patch) + plot_layout(axes = 'collect')

dev.off()



png("~/Dropbox/FringillaMicrobiome/08_Output/Figures/triangle.plots.baselga.sorensen.quant.overall.matrix_genus_noarrows_2panels_speciesFix.png" ,width=8000, height=8000,res=650)

(triangle.overall.diff.bray <- triangle.overall.diff.bray + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()))

dev.off()


png("~/Dropbox/FringillaMicrobiome/08_Output/Figures/triangle.plots.baselga.j.overall.matrix_genus_2panels_speciesFix.png" ,width=8000, height=8000,res=650)

(triangle.overall.diff.jaccard <- triangle.overall.diff.jaccard + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()))

dev.off()


png(paste0(outputDir, "/Figures/Biogeography_speciesFix.png"), width = 500, height = 200, units = "mm", res = 350)
pArea + pDistance + pAge_ma 

dev.off()

png(paste0(outputDir, "/Figures/Colonisation_history_speciesFix_PDUniFrac.png"), width = 300, height = 360, units = "mm", res = 350)
(p_PD + theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank(), strip.text.x = element_text(size = 18))) / 
  D_unifrac_uw_patch /
  D_unifrac_w_patch + plot_layout(axes = 'collect')
dev.off()

```


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
(((p1 +  theme(axis.title.x = element_blank(), axis.text.x =element_blank())) / D_overall_bray_patch) | ( triangle.overall.diff.bray  / triangle.overall.diff.jaccard)) / (pArea + pDistance + pAge_ma) + plot_annotation(tag_levels = 'A') &  theme(plot.tag = element_text(size = 24, face = "bold"))
```
