---
title: 'Diet plots'
author: "B. Karina Montero and Mark Gillingham"
date: "`r Sys.Date()`"
#output: pdf_document
output: html_document
---
## Loading libraries


```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
rm(list = ls())

# basics
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(ggpubr)
library(tidyverse)
library(ggtext)
library(ggrepel)
library(purrr)
library(patchwork)
library(ggtext)
library(lemon)

# phyloseq section
library(phyloseq)
library(decontam)
library(biomeUtils)
library(vegan)

#devtools::install_github("gmteunisse/fantaxtic")
library(fantaxtic)
library(ggtext)

# models
library(mgcv)
library(glmmTMB) 
library(lme4)
library(MuMIn)
library(MASS)
library(nlme)
library(ggeffects)
library(multcomp)
library(car)

# ancom
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("ANCOMBC")
library(ANCOMBC)
```


## WORKING DIRECTORIES

```{r workingDir, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```

## LOAD DATA

Rarefied data

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr <- speedyseq::tax_glom(microb_fr, "Species")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Genus")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Family")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Order")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Class")
#microb_fr <- speedyseq::tax_glom(microb_fr, "Phylum")

diet_fr <- readRDS(paste0(dataDir, "/diet_overlap_OTU99_fr.rds"))


```

## DIET Merged

remove unclassified taxa to Order level  

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

tt <- tax_table(diet_fr) %>% data.frame  
table(tt$Order)
order_filt <- subset(tt, !grepl("Class", Order) & !grepl("Phylum", Order)) %>% rownames_to_column("ASV")
order_filt <- order_filt$ASV

diet_fr <- prune_taxa(order_filt, diet_fr)

```


Relative abudance of plant vs animal

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

diet_fr_meta <- diet_fr %>% sample_data %>% data.frame

head(diet_fr_meta,2)
ps_kingdom <- tax_glom(diet_fr, taxrank = "Kingdom", NArm = FALSE)

tax_table(ps_kingdom)

ps_kingdom <- merge_samples(ps_kingdom,"Species_Label")

#names(sample_data(Merged))

ps_kingdom <- phyloseq::transform_sample_counts(ps_kingdom, function(x) x / sum(x) )
ps_kingdom

tax_table(ps_kingdom) <- gsub("Class_", NA, tax_table(ps_kingdom))
tax_table(ps_kingdom) <- gsub("Order_", NA, tax_table(ps_kingdom))

kingdom_df <- psmelt(ps_kingdom)
head(kingdom_df)

kingdom_df$Sample

kingdom_df <- kingdom_df  %>%    mutate(Labels = Sample %>% str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>% str_replace("_F_", " : F.") %>% str_replace("c_", "c.") %>% str_replace("_", " ")) 
table(kingdom_df$Labels)

kingdom_df$Order <- factor(kingdom_df$Labels,levels=c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"), labels = c("Mainland:~italic('F.coelebs')","Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')","Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')","La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')","Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')","Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))
kingdom_df$Order <- fct_rev(kingdom_df$Order)

kingdom_df$Abundance <- (kingdom_df$Abundance*100)


(p1 <- ggplot(kingdom_df, aes(x = Order , y = Abundance, fill = Kingdom, label = round(Abundance,2))) + geom_bar(stat = "identity") + 
                theme(text = element_text(size = 22), legend.direction = "horizontal", panel.background = element_blank(),
                    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 18)) +
                xlab("") + 
                ylab("Relative abundance (%)") +
                coord_flip() +
        scale_x_discrete(labels = function(l) parse(text=l))+
        scale_fill_manual(values=c("darkgrey",'lightgrey'),guide = guide_legend(title.position = "top",reverse=T, title = "Kingdom abundance (%)")))

```

rbcL

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

#rbcL dataset
ps_rbcL <- readRDS(paste0(dataDir, "/ps_rbcL_OTU99_rarefied.rds" ))
ps_rbcL

tt <- tax_table(ps_rbcL) %>% data.frame  
table(tt$Order)
order_filt <- subset(tt, !grepl("Phylum", Order)) %>% rownames_to_column("ASV")
order_filt <- order_filt$ASV

ps_rbcL <- prune_taxa(order_filt, ps_rbcL)

sample_names(ps_rbcL)
sample_names(microb_fr)

sample_names(ps_rbcL) <- paste(sample_data(ps_rbcL)$Lab_Code, sample_data(ps_rbcL)$Species_Label, sep = "_")

# TipId as sample names 
sample_data(ps_rbcL)$Species_Label <- gsub(" ", "_", sample_data(ps_rbcL)$Species_Label)
sample_names(ps_rbcL) <- paste(sample_data(ps_rbcL)$Lab_Code, sample_data(ps_rbcL)$Species_Label, sep = "_")

setdiff(sample_names(microb_fr), sample_names(ps_rbcL)) %>% print
setdiff(sample_names(ps_rbcL), sample_names(microb_fr)) %>% print

# overlap

rbcl_overlap <- prune_samples(sample_names(ps_rbcL) %in% sample_names(microb_fr), ps_rbcL) 
rbcl_overlap <- biomeUtils::removeZeros(rbcl_overlap)
rbcl_overlap <- prune_taxa(taxa_sums(rbcl_overlap) > 10, rbcl_overlap)
rbcl_overlap  <- phyloseq::transform_sample_counts(rbcl_overlap, function(x) x / sum(x) )
head(tax_table(rbcl_overlap))
rbcl_genus <- tax_glom(rbcl_overlap, "Genus")

```
Per island prevalence

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

core <- rbcl_genus

# sequences df 
sequences <- Biostrings::DNAStringSet(refseq(core))
dss2df <- function(dss) data.frame(width=Biostrings::width(dss), seq=as.character(dss), OTU=names(dss))
seq_df <- dss2df(sequences)
seq_df <- seq_df[,c("OTU", "seq")]
head(seq_df)

# prev df
prevalenceDF <- data.frame(Prevalence = colSums(otu_table(core) > 0 ), RelAbundance = colSums(otu_table(core)), tax_table(core)) 

prevalenceDF <- left_join(rownames_to_column(prevalenceDF, "OTU"), seq_df, by = "OTU")
head(prevalenceDF)
range(prevalenceDF$RelAbund)

topplants <- prevalenceDF %>% filter(Kingdom == "Viridiplantae") %>% arrange(desc(Prevalence)) %>% head(10) %>% dplyr::select(!seq)

topplants 

# Plot prevalent taxa

# metadata 
core_meta <- core %>% sample_data %>% data.frame %>% rownames_to_column("TipID") %>% dplyr::select(TipID, Species_Label)
head(core_meta)

# otu data 
otu_df <- core %>% otu_table %>% as.data.frame %>% rownames_to_column("TipID")
head(otu_df)
otu_df_long <- otu_df %>% gather(key = "OTU", value = "RelAbund", -TipID)
otu_df_long <- left_join(otu_df_long, core_meta, by = "TipID")
head(otu_df_long)
str(otu_df_long)

otu_df_long_summ <- otu_df_long %>% dplyr::group_by(Species_Label, OTU) %>% summarise(RelAbund_mean = mean(RelAbund) * 100, 
  RelAbund_median = median(RelAbund) * 100)

plants_df <- inner_join(otu_df_long_summ, topplants, by = "OTU") %>% as.data.frame
str(plants_df)
head(plants_df)
table(plants_df$OTU)

plants_df$RelAbund_mean <- ifelse(plants_df$RelAbund_mean == 0,  NA, plants_df$RelAbund_mean)
str(plants_df$RelAbund_mean)
range(plants_df$RelAbund_mean, na.rm = TRUE)
summary(plants_df$RelAbund_mean, na.rm = TRUE)

plants_df <- plants_df  %>%    mutate(Labels = Species_Label %>% str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>% str_replace("_F_", " : F.") %>% str_replace("c_", "c.") %>% str_replace("_", " ")) 
table(plants_df$Labels)

plants_df$Order <- factor(plants_df$Labels,levels=c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"), labels = c("Mainland:~italic('F.coelebs')","Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')","Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')","La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')","Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')","Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))
plants_df$Order <- fct_rev(plants_df$Order)

plants_df$Genus_order <- factor(plants_df$Genus, levels=c("Rubus", "Pinus", "Morella", "Geranium", "Family_Ericaceae", "Family_Poaceae", "Family_Lauraceae", "Family_Lamiaceae", "Family_Juglandaceae", "Family_Asteraceae"), labels = c("~italic('Rubus')", "~italic(`Pinus`)", "~italic('Morella')", "~italic('Geranium')", "~italic(Erica~canariensis)", "f_Poaceae", "f_Lauraceae", "f_Lamiaceae", "f_Juglandaceae", "f_Asteraceae"))


(p2 <- plants_df %>% ggplot(aes(y = Order, x = Genus_order)) +
  geom_point(aes(size = RelAbund_mean, color = RelAbund_mean), na.rm = TRUE) +
    scale_color_distiller( palette = "PiYG",
    breaks = c(0.01, 1, 10, 60, 80, 100), 
    labels = c("< 0.1", "  1", "  10", "  60", "  80", "100")) +
    guides(color= guide_legend( title = "Viridiplantae abundance (%) ",title.position = "top"), size=guide_legend( title = "Viridiplantae abundance (%) ",title.position = "top")) +
  ylab("") + xlab("") +
  scale_size_continuous(range = c(0.1, 10), 
    breaks = c(0.01, 1, 10, 60, 80, 100), 
    labels = c("< 0.1", "  1", "  10", "  60", "  80", "100")) +
  #scale_x_discrete(labels = c("~italic('Rubus')", "Pinus", "Morella", "Geranium", "Erica canariensis", "f_Poaceae", "f_Lauraceae", "f_Lamiaceae", "f_Juglandaceae", "f_Asteraceae")) +
  theme_bw() + 
  theme(text = element_text(size = 22),axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 18), legend.direction = "horizontal")+
        scale_x_discrete(labels = function(l) parse(text=l))+
        scale_y_discrete(labels = function(l) parse(text=l)))

   
```


COI

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


ps_COI <- readRDS(paste0(dataDir, "/ps_COI_OTU99_rarefied.rds" ))
ps_COI <- subset_samples(ps_COI, Seq_depth_COI > 0)
ps_COI

tax_table(ps_COI)


#COI dataset
tt <- tax_table(ps_COI) %>% data.frame  
table(tt$Order)
order_filt <- subset(tt, !grepl("Class", Order) & !grepl("Phylum", Order)) %>% rownames_to_column("ASV")
order_filt <- order_filt$ASV

ps_COI <- prune_taxa(order_filt, ps_COI)

sample_names(ps_COI)
sample_names(microb_fr)

sample_names(ps_COI) <- paste(sample_data(ps_COI)$Lab_Code, sample_data(ps_COI)$Species_Label, sep = "_")


# TipId as sample names 
sample_data(ps_COI)$Species_Label <- gsub(" ", "_", sample_data(ps_COI)$Species_Label)
sample_names(ps_COI) <- paste(sample_data(ps_COI)$Lab_Code, sample_data(ps_COI)$Species_Label, sep = "_")

setdiff(sample_names(microb_fr), sample_names(ps_COI)) %>% print
setdiff(sample_names(ps_COI), sample_names(microb_fr)) %>% print

# overlap

coi_overlap <- prune_samples(sample_names(ps_COI) %in% sample_names(microb_fr), ps_COI) 
coi_overlap <- biomeUtils::removeZeros(coi_overlap)
coi_overlap <- prune_taxa(taxa_sums(coi_overlap) > 10, coi_overlap)
coi_overlap  <- phyloseq::transform_sample_counts(coi_overlap, function(x) x / sum(x) )
coi_spp <- tax_glom(coi_overlap, "Species")

```

Core

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

core <- coi_spp

# sequences df 
sequences <- Biostrings::DNAStringSet(refseq(core))
dss2df <- function(dss) data.frame(width=Biostrings::width(dss), seq=as.character(dss), OTU=names(dss))
seq_df <- dss2df(sequences)
seq_df <- seq_df[,c("OTU", "seq")]
head(seq_df)

# prev df
prevalenceDF <- data.frame(Prevalence = colSums(otu_table(core) > 0 ), RelAbundance = colSums(otu_table(core)), tax_table(core)) 

prevalenceDF <- left_join(rownames_to_column(prevalenceDF, "OTU"), seq_df, by = "OTU")
head(prevalenceDF)
range(prevalenceDF$RelAbund)

topbugs <- prevalenceDF %>%  arrange(desc(Prevalence)) %>% head(10) %>% dplyr::select(!seq)
topbugs

#write.csv(prevalenceDF, paste0(dataDir, "/topbugs.csv"))

# Plot prevalent taxa

# metadata 
core_meta <- core %>% sample_data %>% data.frame %>% rownames_to_column("TipID") %>% dplyr::select(TipID, Species_Label)
head(core_meta)

# otu data 
otu_df <- core %>% otu_table %>% as.data.frame %>% rownames_to_column("TipID")
head(otu_df)
otu_df_long <- otu_df %>% gather(key = "OTU", value = "RelAbund", -TipID)
otu_df_long <- left_join(otu_df_long, core_meta, by = "TipID")
head(otu_df_long)
str(otu_df_long)

otu_df_long_summ <- otu_df_long %>% dplyr::group_by(Species_Label, OTU) %>% summarise(RelAbund_mean = mean(RelAbund) * 100, 
  RelAbund_median = median(RelAbund) * 100)

bugs_df <- inner_join(otu_df_long_summ, topbugs, by = "OTU") %>% as.data.frame
str(bugs_df)
head(bugs_df)
table(bugs_df$OTU)

bugs_df$RelAbund_mean <- ifelse(bugs_df$RelAbund_mean == 0,  NA, bugs_df$RelAbund_mean)
str(bugs_df$RelAbund_mean)
range(bugs_df$RelAbund_mean, na.rm = TRUE)
summary(bugs_df$RelAbund_mean, na.rm = TRUE)
range(plants_df$RelAbund_mean, na.rm = TRUE)
summary(plants_df$RelAbund_mean, na.rm = TRUE)

bugs_df <- bugs_df  %>%    mutate(Labels = Species_Label %>% str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>% str_replace("_F_", " : F.") %>% str_replace("c_", "c.") %>% str_replace("_", " ")) 
table(bugs_df$Labels)

bugs_df$Order <- factor(bugs_df$Labels,levels=c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"), labels = c("Mainland:~italic('F.coelebs')","Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')","Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')","La~Palma:~italic('F.c. palmae')","Gran~Canaria:~italic('F.c. bakeri')","Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')","Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))
bugs_df$Order <- fct_rev(bugs_df$Order)

levels(as.factor((bugs_df$Species)))
bugs_df$Sp_order <- factor(bugs_df$Species, levels=c("Brachyderes_rugatus", "Episauris_kiliani", "Mythimna_unipuncta",  "Talitroides_topitotum", "Udea_ferrugalis", "Xysticus_erraticus", "Family_Eriophyidae", "Order_Coleoptera", "Order_Mantodea","Order_Thysanoptera"), labels=c("~italic(Brachyderes~rugatus)", "~italic(Episauris~kiliani)", "~italic(Mythimna~unipuncta)", "~italic(Talitroides~topitotum)", "~italic(Udea~ferrugalis)", "~italic(Xysticus~erraticus)",  "f_Eriophyidae",  "o_Coleoptera", "o_Mantodea","o_Thysanoptera"))

        
(p3 <- bugs_df %>% ggplot(aes(y = Order, x = Sp_order)) +
  geom_point(aes(size = RelAbund_mean, color = RelAbund_mean), na.rm = TRUE) +
    scale_color_distiller( palette = "BrBG",
    breaks = c(0.01, 1, 10, 60, 80, 100), 
    labels = c("< 0.1", "  1", "  10", "  60", "  80", "100")) +
    guides(color= guide_legend( title = "Metazoa abundance (%) ",title.position = "top"), size=guide_legend( title = "Metazoa abundance (%) ",title.position = "top")) +
  ylab("") + xlab("") +
  scale_size_continuous(range = c(0.1, 10), 
    breaks = c(0.01, 1, 10, 60, 80, 100), 
    labels = c("< 0.1", "  1", "  10", "  60", "  80", "100")) +
  #scale_x_discrete(labels = c("Brachyderes rugatus", "Campylaspis sulcata", "Episauris kiliani", "Mythimna unipuncta",  "Xysticus erraticus",  "g_Acalitus",  "o_Coleoptera", "o_Lepidoptera", "o_Mantodea","o_Thysanoptera")) +
  theme_bw() + 
  theme(text = element_text(size = 22),axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 18), legend.direction = "horizontal")+
        scale_x_discrete(labels = function(l) parse(text=l))+
        scale_y_discrete(labels = function(l) parse(text=l)))


```

## E value

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


Edata_df <- read_tsv(paste0(dataDir, "/Edata_df.tsv"))

Edata_df$E_inv <- 1-(Edata_df$E)
#Edata_df <- alpha_df %>% group_by(Species_Label, Genus) %>% summarize(E = max(E), VarE = max(VarE))
Edata_df$Species_Label <- gsub(" ", "_", Edata_df$Species_Label)

Edata_df$fSpecies_Label <- factor(Edata_df$Species_Label,levels=c("Mainland_F_coelebs","Sao_Miguel_F_moreletti","Flores_F_moreletti","Madeira_F_maderensis","El_Hierro_F_c_ombriosa", "La Palma F_c_palmae", "Gran_Canaria_F_c_bakeri","Tenerife_F_c_canariensis","La_Gomera_F_c_canariensis", "Tenerife_F_teydea","Gran_Canaria_F_polatzeki"),labels=c("Mainland:~italic('F. coelebs')","Sao~Miguel:~italic('F. moreletti')","Flores:~italic('F. moreletti')","Madeira:~italic('F. maderensis')","El~Hierro:~italic('F.c. ombriosa')", "La~Palma:~italic('F.c. palmae')", "Gran~Canaria:~italic('F.c. bakeri')","Tenerife:~italic('F.c. canariensis')","La~Gomera:~italic('F.c. canariensis')", "Tenerife:~italic('F. teydea')","Gran~Canaria:~italic('F. polatzeki')"))

levels(Edata_df$fSpecies_Label)

library(paletteer)
(Eplot <- Edata_df %>% dplyr::filter(!is.na(E_inv) & !is.na(fSpecies_Label))   %>% 
            ggplot(aes(x=fSpecies_Label, y=E_inv, color = E_inv, fill = E_inv)) + 
            geom_dotplot(binaxis='y', stackdir='center', dotsize = 2) + theme_bw(base_size = 7) +
            xlab("") + ylab("1- E index") +
          scale_x_discrete(limits=rev) +
                coord_flip() +
                scale_fill_gradient(low="lightgrey",high='#C04657',na.value="yellow", breaks = c(0.2, 0.4, 0.6, 0.8),guide = guide_legend(title.position = "top", title = "Diet specialization (1 - E)")) +
                scale_color_gradient(low="lightgrey",high='#C04657',na.value="yellow", breaks = c(0.2, 0.4, 0.6, 0.8),guide = guide_legend(title.position = "top", title = "Diet specialization (1 - E)")) +
                theme(text = element_text(size = 22), legend.direction = "horizontal", panel.background = element_blank(),
                    axis.text.y = element_blank(),
                    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 18)))
Eplot


#png(paste0(outputDir, "/02_DietMetabarcoding/01_Exploration/Eplot.png"),width=1400, height=2200,res=300)
Eplot
#dev.off()

```

### Test effect of diet

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

alpha_diet_fr <- data.frame(sample_data(diet_fr))
names(alpha_diet_fr)

summary(alpha_diet_fr$Shannon_diet_rare)

Observed_genus_rare <- estimate_richness(microb_fr, measures=c("Observed"))
rownames(Observed_genus_rare) <- gsub('X', '', rownames(Observed_genus_rare))
names(Observed_genus_rare) <- c("Observed_genus_rare")
alpha_diet_fr <- merge(alpha_diet_fr, Observed_genus_rare, by = "row.names")

head(alpha_diet_fr,3)


# correlation between diet richness and shannon?

cor(alpha_diet_fr$Observed_diet_rare, alpha_diet_fr$Shannon_diet_rare) 


GLMM1 <- glmmTMB(Observed_genus_rare ~  Observed_diet_rare + (1|Species_Label), data = alpha_diet_fr, na.action = na.fail, family = nbinom2)
summary(GLMM1)

#Conditional model:
#                   Estimate Std. Error z value Pr(>|z|)    
#(Intercept)        4.241727   0.118683   35.74   <2e-16 ***
#Observed_diet_rare 0.008192   0.008394    0.98    0.329    
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

GLMM1b <- glmmTMB(Observed_genus_rare ~  scale(Observed_diet_rare) + (1|Species_Label), data = alpha_diet_fr, na.action = na.fail, family = nbinom2)

summary(GLMM1b)

#Conditional model:
#                          Estimate Std. Error z value Pr(>|z|)    
#(Intercept)                4.31914    0.08844   48.84   <2e-16 ***
#scale(Observed_diet_rare)  0.04362    0.04470    0.98    0.329    
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

confint(GLMM1b)

#                                        2.5 %    97.5 %   Estimate
#(Intercept)                        4.14579304 4.4924819 4.31913749
#scale(Observed_diet_rare)         -0.04398624 0.1312288 0.04362129
#Std.Dev.(Intercept)|Species_Label  0.14809212 0.4459598 0.25698859


max(alpha_diet_fr$Observed_genus_rare)

scatter_theme <- theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),  axis.line =element_line(colour = "black"), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), axis.title = element_text(size = 20))

(pDiet_richness <- ggpredict(GLMM1, terms = "Observed_diet_rare [all]") %>% plot(line_size = 1.5) +
        geom_point(data = alpha_diet_fr, aes(y = Observed_genus_rare, x = Observed_diet_rare),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 1), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(alpha = 1, size = 0.4, stoke = 0.01) + 
        ylab("Microbiome richness") + 
        xlab("Diet richness") +
        scatter_theme + ggtitle("") + sl_theme + 
        annotate(geom="text",label=paste0("~beta~' = 0.044 [-0.044, 0.131];'~italic('p-value')~'= 0.329'"),parse=T,x = 15, y = 240,size = 6) + scale_y_continuous(breaks = seq(0, 250, by=50)
        #+ scale_y_continuous(breaks = seq(0, 190, by=25), limits=c(0,190))
        ))

GLMM2 <- glmmTMB(Observed_genus_rare ~  Shannon_diet_rare + (1|Species_Label), data = alpha_diet_fr, na.action = na.fail, family = nbinom2)
summary(GLMM2)

GLMM2b <- glmmTMB(Observed_genus_rare ~  scale(Shannon_diet_rare) + (1|Species_Label), data = alpha_diet_fr, na.action = na.fail, family = nbinom2)

summary(GLMM2b)
#Conditional model:
#                         Estimate Std. Error z value Pr(>|z|)    
#(Intercept)               4.31437    0.08872   48.63   <2e-16 ***
#scale(Shannon_diet_rare)  0.09054    0.04260    2.13   0.0335 *  
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


confint(GLMM2b)

(pDiet_shannon <- ggpredict(GLMM2, terms = "Shannon_diet_rare [all]") %>% plot(line_size = 1.5, color = "darkred") +
        geom_point(data = alpha_diet_fr, aes(y = Observed_genus_rare, x = Shannon_diet_rare),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 0.1), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(alpha = 1, size = 0.4, stoke = 0.01) + 
        ylab("") + 
        xlab("Diet Shannon diversity") +
        scatter_theme + ggtitle("") + sl_theme + 
        annotate(geom="text",label=paste0("~beta~' = 0.091 [0.007, 0.174];'~italic('p-value')~'= 0.034'"),parse=T,x = 1.25, y = 240,size = 6)
       # + scale_y_continuous(breaks = seq(0, 190, by=25), limits=c(0,190))
         + scale_y_continuous(breaks = seq(0, 250, by=50)))


#alpha_diet_fr <- merge(Edata_df[c(2,4)], alpha_diet_fr, by = 'Species_Label')
names(alpha_diet_fr)
names(Edata_df)

alpha_diet_fr <- alpha_diet_fr %>% dplyr::rows_update(Edata_df[, c("E", "VarE", "Species_Label")], by = 'Species_Label')

alpha_diet_fr$E_inv <- 1-(alpha_diet_fr$E)

GLMM3 <- glmmTMB(Observed_genus_rare ~  E_inv + (1|Species_Label), data = alpha_diet_fr, na.action = na.fail, family = nbinom2)
summary(GLMM3)

GLMM3b <- glmmTMB(Observed_genus_rare ~  scale(E_inv) + (1|Species_Label), data = alpha_diet_fr, na.action = na.fail, family = nbinom2)

summary(GLMM3b)

#Conditional model:
#             Estimate Std. Error z value Pr(>|z|)    
#(Intercept)   4.29773    0.06293   68.29  < 2e-16 ***
#scale(E_inv)  0.19344    0.05891    3.28  0.00103 ** 
#---

confint(GLMM3b)

#                                       2.5 %    97.5 %  Estimate
#(Intercept)                       4.17438306 4.4210724 4.2977277
#scale(E_inv)                      0.07797358 0.3089159 0.1934448
#Std.Dev.(Intercept)|Species_Label 0.07542817 0.3091395 0.1527018

(pDiet_E <- ggpredict(GLMM3, terms = "E_inv [all]") %>% plot(line_size = 1.5, color = "darkred") +
        geom_point(data = alpha_diet_fr, aes(y = Observed_genus_rare, x = (1-E)),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 0.01), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(alpha = 1, size = 0.4, stoke = 0.01) + 
        ylab("") + 
        xlab("E index") +
        scatter_theme + ggtitle("") + sl_theme + 
        annotate(geom="text",label=paste0("~beta~' = 0.193 [0.078, 0.309];'~italic('p-value')~'= 0.001'"),parse=T,x = 0.35, y = 240,size = 6)
       # + scale_y_continuous(breaks = seq(0, 190, by=25), limits=c(0,190))
         + scale_y_continuous(breaks = seq(0, 250, by=50)))


(diet_plots <- ggarrange(pDiet_richness, pDiet_shannon, pDiet_E, ncol = 3))

plot(alpha_diet_fr$E_inv, alpha_diet_fr$Shannon_diet_rare)
abline(alpha_diet_fr$E_inv, alpha_diet_fr$Shannon_diet_rare, col="red")
cor.test(alpha_diet_fr$E_inv, alpha_diet_fr$Shannon_diet_rare)


alpha_diet_fr_chaff <- alpha_diet_fr[which(!alpha_diet_fr$Species_Label == "Gran_Canaria_F_polatzeki"),]
alpha_diet_fr_chaff <- alpha_diet_fr_chaff[which(!alpha_diet_fr_chaff$Species_Label == "Tenerife_F_teydea"),]

GLMM4b <- glmmTMB(Observed_genus_rare ~  scale(E_inv) + (1|Species_Label), data = alpha_diet_fr_chaff, na.action = na.fail, family = nbinom2)

summary(GLMM4b)

```


## ANCOMbc results

ancom output

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# ancom results
out_ancom <- read.csv(paste0(dataDir, "/ancomdiet_toptaxa.csv"))
out_ancom$log10p <- -log10(as.numeric(out_ancom$q_val))
out_ancom$log10p <- ifelse(out_ancom$log10p == Inf,  211, out_ancom$log10p)

# add 16S tax table
tax_micr <- microb_fr %>% tax_table %>% data.frame %>% dplyr::select(!Kingdom)
out_ancom$Species <- out_ancom$taxon
out_ancom <- left_join(out_ancom, tax_micr, by = "Species")
str(out_ancom)

# plot variables

out_ancom$taxPh <-ifelse(out_ancom$q_val < 0.05, out_ancom$Phylum,"< Threshold")
out_ancom$diff_abund2 <- ifelse(out_ancom$q_val < 0.005, TRUE, FALSE)
out_ancom$taxPh2 <-ifelse(out_ancom$q_val < 0.005, out_ancom$Phylum,"< Threshold")

table(out_ancom$taxPh)

# subset diet items with significan associations
table(out_ancom$diet, out_ancom$diff_abund)
table(out_ancom$diet, out_ancom$diff_abund2)

out_ancom_s <- out_ancom %>% dplyr::filter(diet != "ericaceae" & diet != "morella" & diet != "rubus")

out_ancom_xs <- out_ancom %>% dplyr::filter(diet != "ericaceae" & diet != "geranium" & diet != "morella" & diet != "rubus" & diet != "pine" & diet != "episauris")
table(out_ancom_xs$diet, out_ancom_xs$diff_abund2)

```

volcano plot 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

levels(as.factor(out_ancom$diet))
diet_names <- c("rubus" = "Rubus",
                "pine" = "Pinus",
                "morella" = "Morella", 
                "geranium" = "Geranium",
                "ericaceae" = "Erica canariensis",
                "brachyderes" = "Brachyderes rugatus", 
                "episauris" = "Episauris kiliani", 
                "mythimna" = "Mythimna unipuncta",
                "talitroides" = "Talitroides topitotum",
                "udea" = "Udea ferrugalis",
                "Xysticus" = "Xysticus erraticus")



# Add Phylum colors 
out_ancom$PhylumCol <- factor(out_ancom$taxPh)
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)=="Actinomycetota"]<-"#973500"
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)=="Bacillota"]<-"#0064ce"
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)=="Bacillota_A"]<-"#a7e583"
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)=="Pseudomonadota"]<-"#17c9ff"
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)=="Spirochaetota"]<-"#c3197f"
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)=="Bacteroidota"]<-"gold"
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)=="< Threshold"]<-"grey25"
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)== "Acidobacteriota"]<-"#ffb279"
levels(out_ancom$PhylumCol)[levels(out_ancom$PhylumCol)== "Planctomycetota"]<-"#006e66"

out_ancom$Species

(volcano <- ggplot(out_ancom_xs, aes(y = log10p, x = lfc, col = taxPh2)) +
         geom_jitter(width = 0.1, height = 0.1, size = 3) +
         facet_wrap(.~factor(diet, levels = c("pine", "brachyderes", "episauris", "mythimna", "talitroides", "udea", "Xysticus")), scales = "free", nrow = 1, labeller = as_labeller(diet_names)) +
        theme_linedraw() + 
       # scale_y_log10(labels = scales::comma) +
       geom_text_repel(data=out_ancom_xs[which(out_ancom_xs$diff_abund2!=FALSE),],aes(y = log10p,x = lfc, label = Genus), size=rel(5.5),nudge_x = .05,force = 1,segment.size = 1, parse=FALSE, show.legend= FALSE, max.overlaps = Inf)       +
        xlab(substitute(paste("Log 2 Fold Change")))+
        ylab(substitute("-log10(adj p-value)"))+
        #geom_vline(xintercept=0, linetype="dotted")+
        geom_hline(yintercept=2,show.legend=TRUE,linetype="dashed", col="black") +
        scale_colour_manual(values = levels(out_ancom$PhylumCol), limits=names(out_ancom$PhylumCol), name="Phylum")+
      volc_theme+theme(legend.position="right")+ guides(color = guide_legend(override.aes = list(size = 3))))

(volcano_suppl <- ggplot(out_ancom, aes(y = log10p, x = lfc, col = taxPh)) +
         geom_jitter(width = 0, height = 0, size = 3) +
         facet_wrap(.~factor(diet, levels = c("rubus", "pine", "morella", "geranium", "ericaceae", "brachyderes", "episauris", "mythimna", "talitroides", "udea", "Xysticus")), scales = "free", nrow = 3, labeller = as_labeller(diet_names)) +
        theme_linedraw()+ 
       # scale_y_log10(labels = scales::comma) +
       geom_text_repel(data=out_ancom_s[which(out_ancom_s$diff_abund!=FALSE),],aes(y = log10p,x = lfc, label = Genus), size=rel(5.5),nudge_x = .05,force = 10,segment.size = 1, parse=FALSE, show.legend= FALSE, max.overlaps = Inf)+
        xlab(substitute(paste("Log 2 Fold Change")))+
        ylab(substitute("-log10(adj p-value)"))+
        #geom_vline(xintercept=0, linetype="dotted")+
        geom_hline(yintercept = 1.30103,show.legend=TRUE,linetype="dashed", col="black") +
        scale_colour_manual(values = levels(out_ancom$PhylumCol), limits=names(out_ancom$PhylumCol), name="Phylum")+
      volc_theme+theme(legend.position="right")+ guides(color = guide_legend(override.aes = list(size = 3))))

table(out_ancom_s$Phylum)


shift_legend3 <- function(p) {
    pnls <- cowplot::plot_to_gtable(p) %>% gtable::gtable_filter("panel") %>%
      with(setNames(grobs, layout$name)) %>% purrr::keep(~identical(.x,zeroGrob()))

    if( length(pnls) == 0 ) stop( "No empty facets in the plot" )

    lemon::reposition_legend( p, "center", panel=names(pnls) )
}


png(paste0(outputDir, "/Figures/dietPlots_ancomsuppl_all_speciesFix.png"), width = 600, height = 550, units = "mm", res = 700)
shift_legend3(volcano_suppl)
dev.off()

```


## Patchwork

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

p1b <- p1 + theme(axis.title = element_blank(),legend.title=element_text(size=22),legend.position = "top")

p2b <- p2 + theme(axis.text = element_blank(),legend.title=element_text(size=22),legend.position = "top")

p3b <- p3 + theme(axis.text = element_blank(),legend.title=element_text(size=22),legend.position = "top")

Eplotb <- Eplot + theme(axis.text.y = element_blank(),axis.title = element_blank(),legend.title=element_text(size=22),legend.position = "top")

png(paste0(outputDir, "/Figures/dietPlots_speciesFix.png"), width = 661, height = 450, units = "mm", res = 700)

(pwdiet <- (free(p1b | p2b | p3b | Eplotb) /free( pDiet_richness | pDiet_shannon | pDiet_E )))

#png(paste0(outputDir, "/Figures/dietPlots.png"), width = 600, height = 600, units = "mm", res = 700)

#(pwdiet <- (free(p1b | p2b | p3b | Eplotb) / plot_spacer() / free( pDiet_richness | pDiet_shannon | pDiet_E ) / plot_spacer() / volcano))  + plot_layout(heights = unit(c(220,-15, 170,15, 125), c("mm", "mm", "mm", "mm", "mm")))

dev.off()

#(pwdiet <- (free(p1b + ggtitle("A") + theme(plot.title = element_text(size = 30)) | p2b + ggtitle("B") + theme(plot.title = element_text(size = 30)) | p3b + ggtitle("C") + theme(plot.title = element_text(size = 30)) | Eplotb + ggtitle("D") + theme(plot.title = element_text(size = 30)) )) / free(pDiet_richness + ggtitle("E") + theme(plot.title = element_text(size = 30)) | pDiet_shannon + ggtitle("F") + theme(plot.title = element_text(size = 30)) | pDiet_E + ggtitle("G") + theme(plot.title = element_text(size = 30)))) 




#layout <- 
#"ABB
#ABB"
#
#(Jaccard_PCOA_diet) + free(p1b  | p2b  | #p3b  | Eplotb ) / free(pDiet_richness | #pDiet_shannon | pDiet_E) + plot_layout(#design = layout)
#
#dev.off()


```

Richness (use estimates within phyloseq objects!)

## 1. COLONISATION HISTORY

Are there differences in diet richness according to Island? 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

names(alpha_diet_fr)

# create Age (juvenile and adult) category
alpha_diet_fr$Age_cat <- ifelse(alpha_diet_fr$Age > 3, "adult", "fledging")

alpha_diet_fr$Chaff_Blue <- ifelse(alpha_diet_fr$Species_Label == "Tenerife_F_teydea" | alpha_diet_fr$Species_Label == "Gran_Canaria_F_polatzeki", "Bluefinch", "Chaffinch")

alpha_diet_fr <- alpha_diet_fr %>% dplyr::filter(Genus == "Fringilla" & !is.na(Observed_diet_rare) & !is.na(Age_cat) & !is.na(Sex))

alpha_diet_fr$Species_Label <- as.factor(alpha_diet_fr$Species_Label)
alpha_diet_fr$Species_Label <- relevel(alpha_diet_fr$Species_Label, "Mainland_F_coelebs")
levels(alpha_diet_fr$Species_Label)

GLM1 <- glm.nb(Observed_diet_rare ~ Species_Label + Age_cat + Sex, data = alpha_diet_fr,na.action = na.fail)
summary(GLM1)
dredge(GLM1)
Anova(GLM1, type = "III")

#There is no effect of sex or age therefore to plot we do a model with only Species_Label

GLM1b <- glm.nb(Observed_diet_rare ~ Species_Label, data = alpha_diet_fr,na.action = na.fail)

CI <- as_tibble(confint(GLM1b), rownames = 'predictor') 

CI <- cbind(CI,as_tibble(summary(GLM1b)$coefficients))

CI$CImin <- c(CI$`2.5 %`[1],CI$Estimate[1]+CI$`2.5 %`[c(-1)])
CI$CImax <- c(CI$`97.5 %`[1],CI$Estimate[1]+CI$`97.5 %`[c(-1)])
CI$Estimate.transformed <- c(CI$`Estimate`[1],CI$`Estimate`[1]+CI$`Estimate`[c(-1)])

CI$Estimate <- exp(CI$Estimate.transformed)
CI$CImin <- exp(CI$CImin)
CI$CImax <- exp(CI$CImax)

CI <- CI[c(1,4,5,6,7,8,9)]

write_tsv(CI,paste0(outputDir, "/07_Stats/02_analyses/glm_rare_obs_diet_species.tsv"))


GLM2 <- glm(Shannon_diet_rare ~  Species_Label + Age_cat + Sex, data = alpha_diet_fr, family = "gaussian", na.action = na.fail)
summary(GLM2)
dredge(GLM2)
Anova(GLM2, type = "III")
mcp <- glht(GLM2, linfct = mcp(Species_Label = "Tukey"))
summary(mcp)


#There is no effect of sex or age therefore to plot we do a model with only Species_Label

GLM2b <- glm(Shannon_diet_rare ~ Species_Label, data = alpha_diet_fr,na.action = na.fail, family = "gaussian")

CI <- as_tibble(confint(GLM2b), rownames = 'predictor') 

CI <- cbind(CI,as_tibble(summary(GLM2b)$coefficients))

## only used below if exponential family
CI$CImin <- c(CI$`2.5 %`[1],CI$Estimate[1]+CI$`2.5 %`[c(-1)])
CI$CImax <- c(CI$`97.5 %`[1],CI$Estimate[1]+CI$`97.5 %`[c(-1)])
CI$Estimate.transformed <- c(CI$`Estimate`[1],CI$`Estimate`[1]+CI$`Estimate`[c(-1)])

CI$Estimate <- CI$Estimate.transformed
CI$CImin <- CI$CImin
CI$CImax <- CI$CImax


CI <- CI[c(1,4,5,6,7,8,9)]

write_tsv(CI,paste0(outputDir, "/07_Stats/02_analyses/glm_rare_shannon_diet_species.tsv"))

```


Boxplot

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

levels(alpha_diet_fr$Species_Label)

alpha_diet_fr <- alpha_diet_fr %>% filter(Genus == "Fringilla") %>%
    mutate( Labels = Species_Label %>%
              str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>%
              str_replace("_F_", " : F.") %>%
              str_replace("c_", "c.") %>%
              str_replace("_", " ")) 
table(alpha_diet_fr$Labels)

alpha_diet_fr$Order <- factor(alpha_diet_fr$Labels,levels=c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))
alpha_diet_fr$Order <- fct_rev(alpha_diet_fr$Order)

estimate_obj <- read_tsv(paste0(outputDir, "/07_Stats/02_analyses/glm_rare_obs_diet_species.tsv"))

model_estimates <- estimate_obj %>% 
    mutate( value = Estimate + estimate_obj[[1,2]]) %>% 
    mutate( value = ifelse(predictor == '(Intercept)', Estimate, value))  %>% 
    mutate( island = predictor %>%
              str_replace("as.factor", "") %>%
              str_replace("\\(Species_Label\\)", "") %>%
              str_replace("\\(Intercept\\)", "Mainland : F.coelebs") %>%
              str_replace("Sexmale", "Male") %>%
              str_replace("Age_catfledging", "Fledging") %>%
              str_replace("_F_", " : F.") %>%
              str_replace("c_", "c.") %>%
              str_replace("_", " ")) 

model_estimates$island

#species_estimates <- model_estimates[-c(2,3),] ## only use this if age and sex are in the model
species_estimates <- model_estimates ## only use this if age and sex are not in the model

species_estimates$island
levels(factor(species_estimates$island))

species_estimates$Order <- factor(species_estimates$island, levels = c("Mainland : F.coelebs", "Species LabelSao_Miguel : F.moreletti", "Species LabelFlores : F.moreletti", "Species LabelMadeira : F.maderensis", "Species LabelEl_Hierro : F.c.ombriosa", "Species LabelLa_Palma : F.c.palmae", "Species LabelGran_Canaria : F.c.bakeri", "Species LabelTenerife : F.c.canariensis", "Species LabelLa_Gomera : F.c.canariensis", "Species LabelTenerife : F.teydea", "Species LabelGran_Canaria : F.polatzeki"), labels = c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))

species_estimates$Order <- fct_rev(species_estimates$Order)
levels(species_estimates$Order)

alpha_diet_fr %>% names

alpha_diet_fr$Chaff_Blue2 <- as.character(alpha_diet_fr$Chaff_Blue)

alpha_diet_fr$Chaff_Blue2 <- ifelse(alpha_diet_fr$Chaff_Blue2 == "Chaffinch", "Common chaffinch", "Blue chaffinch")

alpha_diet_fr$Chaff_Blue2  <- as.factor(alpha_diet_fr$Chaff_Blue2)

alpha_diet_fr$Chaff_Blue2 <- forcats::fct_relevel(alpha_diet_fr$Chaff_Blue2,"Common chaffinch","Blue chaffinch")
table(alpha_diet_fr$Chaff_Blue2)

rm(signif2)
signif2 = distinct(alpha_diet_fr, Chaff_Blue2,Order) %>%
    arrange(Chaff_Blue2,Order)

#signif2$Order
#species_estimates$Order

signif2<-left_join(signif2,species_estimates,by="Order")

data.frame(signif2)
signif2 %>% names
signif2[7]
#signif2$Estimate
signif2$p.signif = c("**","","*","*","***","","","","","*","*")

alpha_diet_fr %>% names

signif2 = distinct(alpha_diet_fr, Chaff_Blue,Order) %>%
    arrange(Chaff_Blue,Order)

signif2$Order
species_estimates$Order

signif2<-left_join(signif2,species_estimates,by="Order")

data.frame(signif2) %>% names
colnames(signif2)[7] <- "p.signif"
signif2$Estimate
signif2$p.signif <- round(signif2$p.signif, 6)

signif2$p.symbol = c("*","*","**","","*","*","***","","","","")

alpha_diet_fr$Order <- factor(alpha_diet_fr$Order,rev(levels(alpha_diet_fr$Order)))
signif2$Order <- factor(signif2$Order,rev(levels(signif2$Order)))

#alpha_diet_fr$Chaff_Blue3 <- factor(alpha_diet_fr$Chaff_Blue, levels = rev(levels(factor(alpha_diet_fr$Chaff_Blue))))

alpha_diet_fr$Chaff_Blue2

signif2$Chaff_Blue2 <- factor(signif2$Chaff_Blue,levels = c("Bluefinch","Chaffinch"),labels = c("Blue chaffinch","Common chaffinch"))

(boxp1 <- ggplot(alpha_diet_fr, aes(Order, Observed_diet_rare)) + 
        facet_grid(.~Chaff_Blue2,scales = "free_x",space = "free_x") +
        geom_boxplot(aes(fill = Species_Label), show.legend = F, lwd = 0.25,fatten = 1, alpha = 0.7, outlier.colour = 'transparent',width=0.5) + 
        geom_point( aes(fill = Species_Label),shape = 21, position = position_jitter(seed = 100, width = 0.1), size = 1,  alpha = 0.7, stroke = 0.25) + 
        SpeciesLabel.fill + 
        ylab("Richness") + 
        xlab("") + sl_theme + 
        boxplot_theme +
        scale_x_discrete(breaks = c("Tenerife : F.teydea","Gran Canaria : F.polatzeki","Mainland : F.coelebs","Sao Miguel : F.moreletti","Flores : F.moreletti","Madeira : F.maderensis","La Palma : F.c.palmae","El Hierro : F.c.ombriosa","Gran Canaria : F.c.bakeri","Tenerife : F.c.canariensis","La Gomera : F.c.canariensis"),labels = c("Tenerife","Gran Canaria","Mainland","Sao Miguel","Flores","Madeira","La Palma","El Hierro","Gran Canaria","Tenerife","La Gomera")) + 
        geom_hline(aes(yintercept = signif2$Estimate[11]), linetype='dashed')+  
        #geom_point(data = signif2, aes( x = Order, y = Estimate),fill = '#acacac', shape = 23, size = 2, stroke = 0.25) + 
        stat_summary(aes(x = Order, y = Observed_diet_rare),fun.data="mean_cl_boot", geom="pointrange",shape = 23, col = "#acacac", fill = '#acacac',size=0.5,fun.args=list(conf.int=.95)) + 
        #geom_linerange(data = signif2, aes( x = Order,ymin = CImin, ymax = CImax), inherit.aes = F, col = '#acacac', size = 0.75) + 
        geom_text(data = signif2, aes(x = Order, y = 35, label = p.symbol), vjust = 0.75, size = 8) #+scale_y_log10()
        )
(boxp1 <- boxp1+ theme(axis.text.x = element_blank ()))

 ###Shannon


estimate_obj_shannon <- read_tsv(paste0(outputDir, "/07_Stats/02_analyses/glm_rare_shannon_diet_species.tsv"))

model_estimates_shannon <- estimate_obj_shannon %>% 
    mutate( value = Estimate + estimate_obj_shannon[[1,2]]) %>% 
    mutate( value = ifelse(predictor == '(Intercept)', Estimate, value))  %>% 
    mutate( island = predictor %>%
              str_replace("as.factor", "") %>%
              str_replace("\\(Species_Label\\)", "") %>%
              str_replace("\\(Intercept\\)", "Mainland : F.coelebs") %>%
              str_replace("Sexmale", "Male") %>%
              str_replace("Age_catfledging", "Fledging") %>%
              str_replace("_F_", " : F.") %>%
              str_replace("c_", "c.") %>%
              str_replace("_", " ")) 

model_estimates_shannon$island

#species_estimates_shannon <- model_estimates_shannon[-c(2,3),] ## only use this if age and sex are in the model
species_estimates_shannon <- model_estimates_shannon ## only use this if age and sex are not in the model

species_estimates_shannon$island
levels(factor(species_estimates_shannon$island))

species_estimates_shannon$Order <- factor(species_estimates_shannon$island, levels = c("Mainland : F.coelebs", "Species LabelSao_Miguel : F.moreletti", "Species LabelFlores : F.moreletti", "Species LabelMadeira : F.maderensis", "Species LabelEl_Hierro : F.c.ombriosa", "Species LabelLa_Palma : F.c.palmae", "Species LabelGran_Canaria : F.c.bakeri", "Species LabelTenerife : F.c.canariensis", "Species LabelLa_Gomera : F.c.canariensis", "Species LabelTenerife : F.teydea", "Species LabelGran_Canaria : F.polatzeki"), labels = c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))

species_estimates_shannon$Order <- fct_rev(species_estimates_shannon$Order)
levels(species_estimates_shannon$Order)

alpha_diet_fr %>% names

alpha_diet_fr$Chaff_Blue2 <- as.character(alpha_diet_fr$Chaff_Blue)

alpha_diet_fr$Chaff_Blue2 <- ifelse(alpha_diet_fr$Chaff_Blue2 == "Chaffinch", "Common chaffinch", "Blue chaffinch")

alpha_diet_fr$Chaff_Blue2  <- as.factor(alpha_diet_fr$Chaff_Blue2)

alpha_diet_fr$Chaff_Blue2 <- forcats::fct_relevel(alpha_diet_fr$Chaff_Blue2,"Common chaffinch","Blue chaffinch")
table(alpha_diet_fr$Chaff_Blue2)

signif2_shannon = distinct(alpha_diet_fr, Chaff_Blue2,Order) %>%
    arrange(Chaff_Blue2,Order)

#signif2_shannon$Order
#species_estimates_shannon$Order

signif2_shannon<-left_join(signif2_shannon,species_estimates_shannon,by="Order")

data.frame(signif2_shannon)
signif2_shannon %>% names
signif2_shannon[7]
#signif2_shannon$Estimate
signif2_shannon$p.signif = c("","","","","","","","","","","")

alpha_diet_fr %>% names

signif2_shannon = distinct(alpha_diet_fr, Chaff_Blue,Order) %>%
    arrange(Chaff_Blue,Order)

signif2_shannon$Order
species_estimates_shannon$Order

signif2_shannon<-left_join(signif2_shannon,species_estimates_shannon,by="Order")

data.frame(signif2_shannon) %>% names
colnames(signif2_shannon)[7] <- "p.signif"
signif2_shannon$Estimate
signif2_shannon$p.signif <- round(signif2_shannon$p.signif, 6)

signif2_shannon$p.symbol = c("","","","","","","","","","","")

#alpha_diet_fr$Order <- factor(alpha_diet_fr$Order,rev(levels(alpha_diet_fr$Order)))
#signif2_shannon$Order <- factor(signif2_shannon$Order,rev(levels(signif2_shannon$Order)))

#alpha_diet_fr$Chaff_Blue3 <- factor(alpha_diet_fr$Chaff_Blue, levels = rev(levels(factor(alpha_diet_fr$Chaff_Blue))))

alpha_diet_fr$Chaff_Blue2

signif2_shannon$Chaff_Blue2 <- factor(signif2_shannon$Chaff_Blue,levels = c("Bluefinch","Chaffinch"),labels = c("Blue chaffinch","Common chaffinch"))

(boxp2 <- ggplot(alpha_diet_fr, aes(Order, Shannon_diet_rare)) + 
        facet_grid(.~Chaff_Blue2,scales = "free_x",space = "free_x") +
        geom_boxplot(aes(fill = Species_Label), show.legend = F, lwd = 0.25,fatten = 1, alpha = 0.7, outlier.colour = 'transparent',width=0.5) + 
        geom_point( aes(fill = Species_Label),shape = 21, position = position_jitter(seed = 100, width = 0.1), size = 1,  alpha = 0.7, stroke = 0.25) + 
        SpeciesLabel.fill + 
        ylab("Shannon") + 
        xlab("") + sl_theme + 
        boxplot_theme +
        scale_x_discrete(breaks = c("Tenerife : F.teydea","Gran Canaria : F.polatzeki","Mainland : F.coelebs","Sao Miguel : F.moreletti","Flores : F.moreletti","Madeira : F.maderensis","La Palma : F.c.palmae","El Hierro : F.c.ombriosa","Gran Canaria : F.c.bakeri","Tenerife : F.c.canariensis","La Gomera : F.c.canariensis"),labels = c("Tenerife","Gran Canaria","Mainland","Sao Miguel","Flores","Madeira","La Palma","El Hierro","Gran Canaria","Tenerife","La Gomera")) + 
        geom_hline(aes(yintercept = signif2_shannon$Estimate[3]), linetype='dashed')+  
        #geom_point(data = signif2_shannon, aes( x = Order, y = Estimate),fill = '#acacac', shape = 23, size = 2, stroke = 0.25) + 
        stat_summary(aes(x = Order, y = Shannon_diet_rare),fun.data="mean_cl_boot", geom="pointrange",shape = 23, col = "#acacac", fill = '#acacac',size=0.5,fun.args=list(conf.int=.95)) + 
        #geom_linerange(data = signif2_shannon, aes( x = Order,ymin = CImin, ymax = CImax), inherit.aes = F, col = '#acacac', size = 0.75) + 
        geom_text(data = signif2_shannon, aes(x = Order, y = 3, label = p.symbol), vjust = 0.75, size = 8) #+scale_y_log10()
        )
boxp2 <- boxp2 + theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)


## patchwork
library(patchwork)
PW_boxplots <- boxp1/boxp2

png(paste0(outputDir, "/Figures/SupplDietDiversity.png"),width = 4000, height = 3000, units = 'px', res = 300)
PW_boxplots + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size =24))
dev.off()

```



