---
title: 'Plot x'
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: html_document



---


## LOAD LIBRARIES

```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rm(list = ls())

# basics
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(vegan)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(MicEco)
library(rlist) 
library(grid)
library(tsiR) 
library(ape)
library(ggh4x)
# phyloseq section
library(phyloseq)
library(microViz)
library(biomeUtils)
library(patchwork)

# dist
library(treedataverse)
library(geosphere)
library(treedata.table)


# MRM 
library(sf)
library(ecodist)
library(modEvA)
library(eulerr)

#glmm
library(glmmTMB)
library(effects)



```


## WORKING DIRECTORIES
```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```

## LOAD OUTPUTS

Variance partitioning 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

varpart_out <- read.csv(paste0(dataDir, "/output_varpart_speciesFix.csv"))

```

Mantel test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

mantel_out <- read.csv(paste0(dataDir, "/output_mantel_weightedfst_speciesFix.csv"))

```

euler plots

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


otu_bray_df <- varpart_out %>% filter(distance == "Bray-Curtis" & tax == "OTU99") %>% dplyr::select(variable, percent)
otu_bray_df
bray_set <- c("R" = otu_bray_df[8,2], "R&host" = otu_bray_df[1,2], "R&diet" = otu_bray_df[2,2], "R&geo" = otu_bray_df[3,2],
    "R&host&diet" = otu_bray_df[4,2], "R&diet&geo" = otu_bray_df[5,2], "R&host&geo" = otu_bray_df[6,2],
    "R&host&diet&geo" = otu_bray_df[7,2])

otu_jaccard_df <- varpart_out %>% filter(distance == "Jaccard" & tax == "OTU99") %>% dplyr::select(variable, percent)
otu_jaccard_df
jacc_set <- c("R" = otu_jaccard_df[8,2], "R&host" = otu_jaccard_df[1,2], "R&diet" = otu_jaccard_df[2,2], "R&geo" = otu_jaccard_df[3,2],
    "R&host&diet" = otu_jaccard_df[4,2], "R&diet&geo" = otu_jaccard_df[5,2], "R&host&geo" = otu_jaccard_df[6,2],
    "R&host&diet&geo" = otu_jaccard_df[7,2])

genus_bray_df <- varpart_out %>% filter(distance == "Bray-Curtis" & tax == "Genus") %>% dplyr::select(variable, percent)
genus_bray_df
bray_set <- c("R" = genus_bray_df[8,2], "R&host" = genus_bray_df[1,2], "R&diet" = genus_bray_df[2,2], "R&geo" = genus_bray_df[3,2],
    "R&host&diet" = genus_bray_df[4,2], "R&diet&geo" = genus_bray_df[5,2], "R&host&geo" = genus_bray_df[6,2],
    "R&host&diet&geo" = genus_bray_df[7,2])

genus_jaccard_df <- varpart_out %>% filter(distance == "Jaccard" & tax == "Genus") %>% dplyr::select(variable, percent)
genus_jaccard_df
jacc_set <- c("R" = genus_jaccard_df[8,2], "R&host" = genus_jaccard_df[1,2], "R&diet" = genus_jaccard_df[2,2], "R&geo" = genus_jaccard_df[3,2],
    "R&host&diet" = genus_jaccard_df[4,2], "R&diet&geo" = genus_jaccard_df[5,2], "R&host&geo" = genus_jaccard_df[6,2],
    "R&host&diet&geo" = genus_jaccard_df[7,2])


species_bray_df <- varpart_out %>% filter(distance == "Bray-Curtis" & tax == "Species") %>% dplyr::select(variable, percent)
species_bray_df
bray_set <- c("R" = species_bray_df[8,2], "R&host" = species_bray_df[1,2], "R&diet" = species_bray_df[2,2], "R&geo" = species_bray_df[3,2],
    "R&host&diet" = species_bray_df[4,2], "R&diet&geo" = species_bray_df[5,2], "R&host&geo" = species_bray_df[6,2],
    "R&host&diet&geo" = species_bray_df[7,2])

species_jaccard_df <- varpart_out %>% filter(distance == "Jaccard" & tax == "Species") %>% dplyr::select(variable, percent)
species_jaccard_df
jacc_set <- c("R" = species_jaccard_df[8,2], "R&host" = species_jaccard_df[1,2], "R&diet" = species_jaccard_df[2,2], "R&geo" = species_jaccard_df[3,2],
    "R&host&diet" = species_jaccard_df[4,2], "R&diet&geo" = species_jaccard_df[5,2], "R&host&geo" = species_jaccard_df[6,2],
    "R&host&diet&geo" = species_jaccard_df[7,2])



```

# Variance partitioning plots 

plot euler

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(euler_bray <- plot(euler(bray_set, input = c("disjoint"),shape =  "ellipse"),
    fills = list(fill = c("white","#a00000", "#000000", "#59a89c"), alpha = 0.8), quantities = FALSE, edges = list(lwd = c(3,2,2,2)), labels = list(labels = c("","","",""), cex = 1), lwd = 1))
euler_bray <- gridExtra::grid.arrange(grobs = list(euler_bray), top = textGrob("Species", gp=gpar(fontsize=16)))

(euler_jacc <- plot(euler(jacc_set, input = c("disjoint"),shape =  "ellipse"),
    fills = list(fill = c("white","#a00000", "#000000", "#59a89c"), alpha = 0.8), quantities = FALSE, edges = list(lwd = c(3,2,2,2)), labels = list(labels = c("","","",""), cex = 1), lwd = 1))

euler_jacc <- gridExtra::grid.arrange(grobs = list(euler_jacc), top = textGrob("Species", gp=gpar(fontsize=16)))


```

VarPart barplots

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# calculate overlap 
names(varpart_out)
varpart_out$variable2 <- gsub(".*_.*", "overlap", varpart_out$variable)
ind_sum <- varpart_out %>% group_by(pop_ind, distance, tax, variable2) %>%
                        summarise(percent2 = sum(percent)) 

ind_sum$order <- factor(ind_sum$variable2, levels = c("host", "diet", "geo", "overlap", "Unexplained"), labels = c("Phylogeny", "Diet","Geography", "Overlap","Unexplained"))

# Add colors 
ind_sum$varFill <- ind_sum$order
levels(ind_sum$varFill)[levels(ind_sum$varFill)=="Phylogeny"] <- "#a00000"
levels(ind_sum$varFill)[levels(ind_sum$varFill)=="Diet"] <- "#000000"
levels(ind_sum$varFill)[levels(ind_sum$varFill)=="Geography"] <- "#59a89c"
levels(ind_sum$varFill)[levels(ind_sum$varFill)=="Overlap"] <- "lightgrey"
levels(ind_sum$varFill)[levels(ind_sum$varFill)=="Unexplained"] <- "white"
 
ind_sum %>% filter(tax == "Genus" & variable2 == "overlap") 

ind_sum$tax <- factor(ind_sum$tax,levels=c("OTU99","Species","Genus","Family","Order","Class","Phylum"))

(varPar <- ggplot(ind_sum, aes(y = tax, x = percent2, fill = order, color = order)) + 
        geom_bar(position = "stack", stat = "identity", linewidth = 0.25, color = "black") + 
        geom_bar(position = "stack", stat = "identity", linewidth = 0.5, show_guide = FALSE) +
        scale_x_continuous("% variance explained",  limits = c(0, 25), expand = c(0,0)) +
        scale_fill_manual(values = levels(ind_sum$varFill), limits=names(ind_sum$varFill), name="")+ 
        scale_color_manual(values = c("white", "white","white", "white","white"), limits=names(ind_sum$varFill), name="") + 
        facet_wrap(.~factor(distance), strip.position = "top", ncol= 2, dir = "v") + 
        barplot_theme + theme(legend.position="right", panel.spacing = unit(2, "lines"),   panel.grid.major.y = element_blank(), panel.grid.major.x = element_line(colour = "#bababa", linewidth = 1), 
          axis.title.x = element_text(size = 18), 
          axis.title.y = element_text(size = 18),
          axis.text.y = element_text(size = 16),
          axis.text.x = element_text(angle = 0,  size = 16),
          axis.line = element_line(colour = 'black', size = 0.5), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16),
          strip.text = element_text(size=18)) +
        xlab("% variance explained") + ylab("")
        #+ guides(color = guide_legend(override.aes = list(linewidth = 2)))
        )

```

Data for scatterplot


Rarefied 16S

```{r import, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr_otu <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr_species <- speedyseq::tax_glom(microb_fr_otu, "Species")


# TipId 
sample_data(microb_fr_species)$TipID_microbiome <- paste(sample_data(microb_fr_species)$Lab_Code, sample_data(microb_fr_species)$Species_Label, sep = "_")

# meta
microb_fr_meta_species <- sample_data(microb_fr_species) %>% data.frame

```


Diet data

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

diet_fr <- readRDS(paste0(dataDir, "/diet_overlap_OTU99_fr.rds"))

```


Overlap all datasets

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_overlap_species <- microb_fr_species %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")

microb_overlap_meta <- microb_overlap_species %>% sample_data %>% data.frame 
names(microb_overlap_meta)


```

subsample microbiome equal sample size


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

sample_ps <- function(ps, FUN = sample, ...){
  ids <- sample_names(ps)
  sampled_ids <- FUN(ids, ...)
  ps <- prune_samples(sampled_ids, ps)
  return(ps)
}

levels(factor(sample_data(microb_fr_otu)$Species_Label))

# species
(ElHierro <- microb_fr_species %>% subset_samples(Species_Label=="El_Hierro_F_c_ombriosa" ))
(Flores <- microb_fr_species %>% subset_samples(Species_Label=="Flores_F_moreletti" ))
(GranCanaria <- microb_fr_species %>% subset_samples(Species_Label=="Gran_Canaria_F_c_bakeri" ))
(Polatzeki <- microb_fr_species %>% subset_samples(Species_Label=="Gran_Canaria_F_polatzeki" ))
(LaGomera <- microb_fr_species %>% subset_samples(Species_Label=="La_Gomera_F_c_canariensis" ))
(LaPalma <- microb_fr_species %>% subset_samples(Species_Label=="La_Palma_F_c_palmae" ))
(Madeira <- microb_fr_species %>% subset_samples(Species_Label=="Madeira_F_maderensis" ))
(Mainland <- microb_fr_species %>% subset_samples(Species_Label=="Mainland_F_coelebs" ))
(SaoMiguel <- microb_fr_species %>% subset_samples(Species_Label=="Sao_Miguel_F_moreletti" ))
(Tenerife <- microb_fr_species %>% subset_samples(Species_Label=="Tenerife_F_c_canariensis" ))
(Teydea <- microb_fr_species %>% subset_samples(Species_Label=="Tenerife_F_teydea" ))

(ElHierro <- sample_ps(ElHierro, size = 9))
(Flores <- sample_ps(Flores, size = 9))
(GranCanaria <- sample_ps(GranCanaria, size = 9))
(Polatzeki <- sample_ps(Polatzeki, size = 9))
(LaGomera <- sample_ps(LaGomera, size = 9))
(LaPalma <- sample_ps(LaPalma, size = 9))
(Madeira <- sample_ps(Madeira, size = 9))
(Mainland <- sample_ps(Mainland, size = 9))
(SaoMiguel <- sample_ps(SaoMiguel, size = 9))
(Tenerife <- sample_ps(Tenerife, size = 9))
(Teydea <- sample_ps(Teydea, size = 9))


(microb_fr_subsample_species <- merge_phyloseq(ElHierro, Flores, GranCanaria, Polatzeki, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife, Teydea) %>% removeZeros())
```

merge
 
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# merge samples per island / mainland : species
ps_merged_species <- phyloseq::merge_samples(microb_fr_subsample_species, "Species_Label", fun = sum)
```


Centroids of geographic distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# all finches
# lat long data
my_points <- data.frame(longitude = sample_data(microb_fr_otu)$Xdec, latitude = sample_data(microb_fr_otu)$Ydec, Island_Mainland = sample_data(microb_fr_otu)$Species_Label, row.names = sample_names(microb_fr_otu))
str(my_points)
table(my_points$Island_Mainland)
df_sf <- st_as_sf(my_points, coords = c('longitude', 'latitude'))

# get centroids per island / mainland 
centroids_sf <- df_sf %>%
  group_by(Island_Mainland) %>% 
  summarize(geometry = st_union(geometry)) %>% 
  st_centroid %>%   
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])
str(centroids_sf)
rownames(centroids_sf) <- centroids_sf$Island_Mainland

# dist
dist_mat <- distm(cbind(centroids_sf$lon, centroids_sf$lat), fun = distGeo)  # Vector of distances in meters
dist_mat <-dist_mat/1000 #Vector of distances in km
rownames(dist_mat) <- rownames(centroids_sf)
colnames(dist_mat) <- rownames(centroids_sf)
geo_dist <- as.dist(dist_mat)
str(dist_mat)

```

microbiome distance matrix

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
## species
# Bray
microb_otu_tab <- otu_table(ps_merged_species)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_species <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_species)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_species <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)
```


Fst values

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

fst <- read.csv(paste0(dataDir, "/MeanFst_Fr_weighted_estimates.csv"), header = T, row.names = 1)
str(fst)

fst <- fst %>% as.matrix
fst1 <- fst[rownames(dist_mat),colnames(dist_mat)] %>% as.matrix
rownames(fst1)
fst_dist <- as.dist(fst1)
str(fst_dist)

```


Calculate divergence, microbiome distance matrix

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

## species
# Bray
microb_otu_tab <- otu_table(microb_overlap_species)
bray_dist_ind_species <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_species <- betadisper(bray_dist_ind_species, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_species)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_species)
jaccard_dist_ind_species <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_species <- betadisper(jaccard_dist_ind_species, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_species)[2] <- "j.beta_median"

```


Inbreeding coefficients

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

metadata <- read_delim(paste0(dataDir, "/WGS_Fringilla_samples.tsv"), delim = ",")

ind_het <- read_delim(paste0(dataDir, "/chaffinch_bluefinch_miss0.5_mac3_noZ.het"), delim = "\t",
           col_names = c("WGS_ID","ho", "he", "nsites", "f"), skip = 1)
ind_het

ind_het_meta <- left_join(ind_het, metadata, by ="WGS_ID")
ind_het_meta %>% data.frame %>% head
```

Join df

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ind_het_meta %>% data.frame %>% head(2)
ind_het_meta$Region_Species

# species
het_bray_species <- left_join(ind_het_meta, bray_divergence_species, by = "TipID_microbiome")
het_bray_species <- subset(het_bray_species, subset = !is.na(het_bray_species$b.beta_median))

het_jacc_species <- left_join(ind_het_meta, jaccard_divergence_species, by = "TipID_microbiome")
het_jacc_species <- subset(het_jacc_species, subset = !is.na(het_jacc_species$j.beta_median))
```

# FST and inbreeding plots

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Fst scatter plot
bray = as.vector(bray_dist_species)
jacc = as.vector(jaccard_dist_species)
fst = as.vector(fst_dist)
geo = as.vector(geo_dist)

#new data frame with vectorized distance matrices
mat = data.frame(bray, jacc, fst, geo)

# Mantel r 
str(mantel_out)
mantel_out$signif <- ifelse(mantel_out$p < 0.05 & mantel_out$p > 0.005, "< 0.05", ifelse(mantel_out$p < 0.005 & mantel_out$p > 0.0005 , "< 0.005", ifelse(mantel_out$p < 0.0005, "< 0.0005", "n.s.")))
table(mantel_out$signif)
check <- mantel_out %>% filter(p < 0.05)

mantel_out$order <- factor(mantel_out$taxonomy, levels = c("OTU99", "Species", "Genus", "Family", "Order", "Class", "Phylum"))


```



```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(p_bray_fst <- ggplot(mat, aes(y = bray, x = fst)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Fst", y = "Bray-Curtis distance") +
        geom_smooth(method = "lm", colour = "darkred", alpha = 0.2,lwd=2) +scatter_theme + ggtitle("") + sl_theme +theme(legend.position="top", 
          axis.text.x = element_text(angle = 0,  size = 16), axis.text.y = element_text(angle = 0,  size = 16),
          axis.title = element_text(size = 18),
          axis.line = element_line(colour = 'black', size = 0.5), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))

(p_jacc_fst <- ggplot(mat, aes(y = jacc, x = fst)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Fst", y = "Jaccard distance") +
       geom_smooth(method = "lm", colour = "black", alpha = 0.2,lwd=2) +
       scatter_theme + 
       sl_theme +
       theme(legend.position="top", 
          axis.text.x = element_text(angle = 0,  size = 16), axis.text.y = element_text(angle = 0,  size = 16),
          axis.title = element_text(size = 18), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16),
          axis.line = element_line(colour = 'black', size = 0.5), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))

Jaccard_dist_df <- data.frame(cbind(Distance = mat$jacc,fst =mat$fst,Distance_name="Jaccard"))
Bray_dist_df <- data.frame(cbind(Distance = mat$bray,fst =mat$fst,Distance_name="Bray"))

dist_fst_df <- rbind(Bray_dist_df,Jaccard_dist_df)

dist_fst_df$fst <- as.numeric(dist_fst_df$fst)
dist_fst_df$Distance <- as.numeric(dist_fst_df$Distance)



mantel_out$signifFill <- as.factor(mantel_out$signif)
levels(mantel_out$signifFill)[levels(mantel_out$signifFill)=="< 0.05"] <- "#999999"
levels(mantel_out$signifFill)[levels(mantel_out$signifFill)=="< 0.005"] <- "#333333"
levels(mantel_out$signifFill)[levels(mantel_out$signifFill)=="n.s."] <- "white"

(p_mantel <- mantel_out %>% filter(data == "all", test == "partial-mantel") %>% 
    ggplot(aes(y = r, x = order, fill = signif, color = signif)) + 
        geom_bar(stat = "identity", linewidth = 0.25) +
        scale_fill_manual("p-value", values = levels(mantel_out$signifFill), limits=names(mantel_out$signifFill))+ 
        scale_color_manual("p-value", values = c("black", "black","black", "black","black"), limits=names(mantel_out$signifFill)) +
        ylab(expression("Mantel"~italic("r"))) + xlab("") + coord_flip()+
        barplot_theme + 
         facet_wrap(.~factor(dist), strip.position = "top", ncol= 2, dir = "v") +
        #guides(color = guide_legend(override.aes = list(linewidth = 2))) +
        theme(legend.position="right", 
          axis.title.x = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          axis.text.y = element_text(size = 16),
          axis.text.x = element_text(angle = 0,  size = 16), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16),
          axis.line = element_line(colour = 'black', size = 0.5), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))

(p_bray_het <- ggplot(het_bray_species, aes(y = b.beta_median, x = f)) + 
        geom_point(size = 5, alpha = 0.2) + 
        xlab(expression("Inbreeding coef"~italic((f)))) +
        ylab("Divergence (Bray-Curtis)") + 
        geom_smooth(method = "lm", colour = "darkred", alpha = 0.2,lwd=2) + scatter_theme + ggtitle("") + sl_theme +
        theme(legend.position="top", 
          axis.text.x = element_text(angle = 0,  size = 16), axis.text.y = element_text(angle = 0,  size = 16),
          axis.title = element_text(size = 18), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16),
          axis.line = element_line(colour = 'black', size = 0.5), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))

(p_jacc_het <- ggplot(het_jacc_species, aes(y = j.beta_median, x = f)) + 
        geom_point(size = 5, alpha = 0.2) + 
        xlab(expression("Inbreeding coef"~italic((f)))) +
        ylab("Divergence (Jaccard)") + 
        geom_smooth(method = "lm", colour = "black", alpha = 0.2,lwd=2) + scatter_theme + ggtitle("") + 
        sl_theme +
        theme(legend.position="top", 
          axis.text.x = element_text(angle = 0,  size = 16), axis.text.y = element_text(angle = 0,  size = 16),
          axis.title = element_text(size = 18), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16),
          axis.line = element_line(colour = 'black', size = 0.5), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))


Bray_dist_f <- data.frame(cbind(beta_median = het_bray_species$b.beta_median,f =het_bray_species$f,Distance_name="Bray"))
Jaccard_dist_f <- data.frame(cbind(beta_median = het_jacc_species$j.beta_median,f =het_jacc_species$f,Distance_name="Jaccard"))

dist_f_df <- rbind(Bray_dist_f,Jaccard_dist_f)

dist_f_df$f <- as.numeric(dist_f_df$f)
dist_f_df$beta_median <- as.numeric(dist_f_df$beta_median)



het_bray_species$sf <- scale(het_bray_species$f)

glmm_bray_species <- glmmTMB(b.beta_median ~  f + (1|Species_Label), data = het_bray_species, na.action = na.fail, family = gaussian)
confint(glmm_bray_species)

glmm_jacc_species <- glmmTMB(j.beta_median ~  f + (1|Species_Label), data = het_jacc_species, na.action = na.fail, family = gaussian)
confint(glmm_jacc_species)


(effect_bray <- as.data.frame(effect('f', glmm_bray_species, confint=list(alpha = .95))))

(effect_jac <- as.data.frame(effect('f', glmm_jacc_species, confint=list(alpha = .95))))

(fit_Bray <- ggplot(het_bray_species, aes(y = b.beta_median, x = f))+ 
        geom_point(size = 5, alpha = 0.2) + 
        xlab(expression("Inbreeding coef"~italic((f)))) +
        ylab("Divergence (Bray-Curtis)") +
        geom_line(data = effect_bray, aes(x = f, y = fit), colour = "darkred",lwd=2) +
        geom_ribbon(data = effect_bray,aes(x = f, y = fit,ymin = lower, ymax = upper), alpha = .2)+ scatter_theme + ggtitle("") + sl_theme +
        theme(legend.position="top", 
          axis.text.x = element_text(angle = 0,  size = 16), axis.text.y = element_text(angle = 0,  size = 16),
          axis.title = element_text(size = 18), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16),
          axis.line = element_line(colour = 'black', size = 0.5), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))



(p_jacc_het <- ggplot(het_jacc_species, aes(y = j.beta_median, x = f)) + 
        geom_point(size = 5, alpha = 0.2) + 
        xlab(expression("Inbreeding coef"~italic((f)))) +
        ylab("Divergence (Jaccard)") +
        geom_line(data = effect_jac, aes(x = f, y = fit), colour = "black",lwd=2) +
        geom_ribbon(data = effect_jac,aes(x = f, y = fit,ymin = lower, ymax = upper), alpha = .2) + scatter_theme + ggtitle("") + 
        sl_theme +
        theme(legend.position="top", 
          axis.text.x = element_text(angle = 0,  size = 16), axis.text.y = element_text(angle = 0,  size = 16),
          axis.title = element_text(size = 18), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16),
          axis.line = element_line(colour = 'black', size = 0.5), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))


effect_bray$Distance_name <- print("Bray") 
effect_jac$Distance_name <- print("Jaccard")

effect_df <- rbind(effect_bray, effect_jac)





(p_het <- ggplot(dist_f_df, aes(y = beta_median, x = f, color = Distance_name)) + 
        geom_point(size = 5, alpha = 0.2, colour = 'black') + 
       scale_color_manual(values = c("darkred","black")) + 
        xlab(expression("Inbreeding coef"~italic((f)))) +
        ylab("Divergence") + 
        geom_line(data = effect_df, aes(x = f, y = fit),lwd=2) +
        geom_ribbon(data = effect_df,aes(x = f, y = fit,ymin = lower, ymax = upper), alpha = 0.1, colour = NA) + scatter_theme + ggtitle("") + 
        sl_theme + 
         facet_wrap(.~factor(Distance_name), strip.position = "top") + 
        theme(legend.position="none", 
          axis.text.x = element_text(angle = 0,  size = 16), axis.text.y = element_text(angle = 0,  size = 16),
          axis.title = element_text(size = 18),
          axis.line = element_line(colour = 'black', size = 0.5), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))




(p_fst <- ggplot(dist_fst_df, aes(y = Distance, x = fst, color=Distance_name)) + 
        geom_point(size = 5, alpha = 0.2, colour = 'black') + 
        labs(x = expression(italic("F")[ST]), y = "Distance") +
       geom_smooth(method = "lm", alpha = 0.2,lwd=2) + 
       scale_color_manual(values = c("darkred","black"))+
       scatter_theme + 
       sl_theme + 
         facet_wrap(.~factor(Distance_name), strip.position = "top") + 
       theme(legend.position="none", 
          axis.text.x = element_text(angle = 0,  size = 16), axis.text.y = element_text(angle = 0,  size = 16),
          axis.title = element_text(size = 18),
          axis.line = element_line(colour = 'black', size = 0.5), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()
          ))






glmm_out <- read.csv(paste0(dataDir, "/glmm_fst_het_output_miss0.5_mac3_noZ_speciesFix.csv"))
head(glmm_out)
colnames(glmm_out) <- c("term","Estimate","Std_Error","z_value","P","lower","upper","data","taxonomy","dist")
head(glmm_out)


glmm_out$order <- factor(glmm_out$taxonomy, levels = c("OTU99", "species" ,"genus", "family", "order", "class", "phylum"), labels = c("OTU99", "Species", "Genus", "Family", "Order", "Class", "Phylum"))

glmm_out$signif <- ifelse(glmm_out$P < 0.05 & glmm_out$P > 0.005, "< 0.05", ifelse(glmm_out$P < 0.005 & glmm_out$P > 0.0005 , "< 0.005", ifelse(glmm_out$P < 0.0005, "< 0.0005", "n.s.")))
table(glmm_out$signif)
check <- glmm_out %>% filter(P < 0.05)

glmm_out$signifFill <- as.factor(glmm_out$signif)
levels(glmm_out$signifFill)[levels(glmm_out$signifFill)=="< 0.05"] <- "#999999"
levels(glmm_out$signifFill)[levels(glmm_out$signifFill)=="< 0.005"] <- "#333333"
levels(glmm_out$signifFill)[levels(glmm_out$signifFill)=="< 0.0005"] <- "#000000"
levels(glmm_out$signifFill)[levels(glmm_out$signifFill)=="n.s."] <- "white"

glmm_out %>% filter(term == "scale(f)", data == "all")

head(glmm_out)
head(tail)


(p_glmm <- glmm_out %>% filter(term == "scale(f)", data == "all") %>% 
    ggplot(aes(x = Estimate, y = order)) + 
      geom_linerange(aes(xmin = lower, xmax = upper), linewidth = 1.5, color = "#000000") +
      geom_point(aes(fill = signif, color = signif), pch =21, size = 5) +
      geom_vline(xintercept = 0, linetype = "longdash") + 
      scale_fill_manual("p-value", values = levels(glmm_out$signifFill), limits=names(glmm_out$signifFill))+ 
        scale_color_manual("p-value", values = c("black", "black","black", "black","black"), limits=names(glmm_out$signifFill)) +
        xlab("Estimate") + ylab("")+
        barplot_theme + 
        facet_wrap(.~factor(dist), strip.position = "top") +
        guides(color = guide_legend(override.aes = list(linewidth = 1))) +
        theme(legend.position="right", 
          axis.title.y = element_text(size=18),
          axis.title.x = element_text(size=18),   
          axis.text.y = element_text(angle = 0,  size = 16),   
          axis.text.x = element_text(angle = 0,  size = 16),
          axis.line = element_line(colour = 'black', size = 0.5), 
          legend.title=element_text(size=18), 
          legend.text=element_text(size=16), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))

```


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, fig.height = 29.7, fig.width = 21}

((varPar + inset_element(euler_bray, 0.3, 0.18, 0.48, 0.99) + inset_element(euler_jacc, 0.82, 0.18, 1, 0.99)) / (p_fst) / p_mantel / (p_het) / p_glmm ) + plot_annotation(tag_levels= list(c("A", "", "", "B", "", "C", "")))  &  theme(plot.tag = element_text(size = 24))

ggsave(paste0(outputDir, "/Figures/Figure4.png"),width=3000, height=4369.2, units = "px",dpi=300)
```