---
title: 'Effect of colonisation history and biogeography'
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: html_document

---


## LOAD LIBRARIES

```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rm(list = ls())

# basics
library(tibble)
library(dplyr)
library(tidyverse)
library(data.table)
library(ggtern)

# phyloseq section
library(phyloseq)
library(Biostrings)
library(microbiome)
library(biomeUtils)
library(microViz)

# diversity 
library(adespatial)


#rmd
library(details)
library(knitr)
library(kableExtra)

#models
library(MASS)
library(MuMIn)
library(car)
library(multcomp)
library(glmmTMB)
library(ggeffects)

```
## WORKING DIRECTORIES

```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```

## LOAD 16S DATA
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_fringilla_rare <- readRDS(paste0(dataDir, "/ps_OTU99_wAlpha_fr_taxFix.rds" ))

ps_fringilla_rare <- biomeUtils::removeZeros(ps_fringilla_rare)

ps_fringilla_rare

ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Species")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Genus")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Family")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Order")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Class")
#ps_fringilla_rare <- speedyseq::tax_glom(ps_fringilla_rare, "Phylum")


ps_fringilla_rare <- biomeUtils::removeZeros(ps_fringilla_rare)

```

## TESTING THE EFFECT OF COLONISATION HISTORY

### ALPHA DIVERSITY

Is there a sequential reduction of alpha diversity as predicted by colonization history? (i.e. mainland > Azores > Madeira > Canary Islands) 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rich <- estimate_richness(ps_fringilla_rare, measures=c("Observed", "Shannon"))

sample_data(ps_fringilla_rare)$Observed_genus_rare <- rich$Observed
summary(sample_data(ps_fringilla_rare)$Observed_genus_rare )
alpha_df <- data.frame(sample_data(ps_fringilla_rare))

mainland_fringilla <- alpha_df %>% dplyr::filter(!is.na(Age_cat) & !is.na(Sex))  

mainland_fringilla$Species_Label <- as.factor(mainland_fringilla$Species_Label)
mainland_fringilla$Species_Label <- relevel(mainland_fringilla$Species_Label, "Mainland_F_coelebs")

names(mainland_fringilla)

GLM1 <- glm.nb(Observed_genus_rare ~ Age_cat + Sex + Species_Label, data = mainland_fringilla,na.action = na.fail)
summary(GLM1)
dredge(GLM1)
Anova(GLM1, type = "III")
#Analysis of Deviance Table (Type III tests)

#Response: Observed_genus_rare
#              LR Chisq Df Pr(>Chisq)    
#Age_cat          0.029  1     0.8640    
#Sex              1.527  1     0.2165    
#Species_Label   44.707 10  2.456e-06 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


GLM2 <- glm.nb(Observed_genus_rare ~ Species_Label, data = mainland_fringilla,na.action = na.fail)
summary(GLM2)
dredge(GLM2)
Anova(GLM2, type = "III")
mcp <- glht(GLM2, linfct = mcp(Species_Label = "Tukey"))
summary(mcp)


CI <- as_tibble(confint(GLM2), rownames = 'predictor') 

CI <- cbind(CI,as_tibble(summary(GLM2)$coefficients))

CI$CImin <- c(CI$`2.5 %`[1],CI$Estimate[1]+CI$`2.5 %`[c(-1)])
CI$CImax <- c(CI$`97.5 %`[1],CI$Estimate[1]+CI$`97.5 %`[c(-1)])
CI$Estimate.transformed <- c(CI$`Estimate`[1],CI$`Estimate`[1]+CI$`Estimate`[c(-1)])

CI$Estimate <- exp(CI$Estimate.transformed)
CI$CImin <- exp(CI$CImin)
CI$CImax <- exp(CI$CImax)

CI <- CI[c(1,4,5,6,7,8,9)]

estimate_obj <- CI

```


No evidence of alpha diversity mirroring the colonization history of the chaffinch spp complex. However, there is an effect (Chisq = 46.276, df = 10, P < 0.001) of species within islands but no effect of age or sex. We should then control for Species Label in subsequent models.

Plot richness according to species within islands:

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

mainland_fringilla <- mainland_fringilla %>% filter(Genus == "Fringilla") %>%    mutate( Labels = Species_Label %>% str_replace("Mainland_F_coelebs", "Mainland : F.coelebs") %>% str_replace("_F_", " : F.") %>% str_replace("c_", "c.") %>% str_replace("_", " ")) 

mainland_fringilla$Order <- factor(mainland_fringilla$Labels,levels=c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))

estimate_obj <- read_tsv(paste0(outputDir, "/07_Stats/02_analyses/glm_rare_sobs_speciesFix.tsv"))

model_estimates <- estimate_obj %>% 
    mutate( value = Estimate + estimate_obj[[1,2]]) %>% 
    mutate( value = ifelse(predictor == '(Intercept)', Estimate, value))  %>% 
    mutate( island = predictor %>%
              str_replace("as.factor", "") %>%
              str_replace("\\(Species_Label\\)", "") %>%
              str_replace("\\(Intercept\\)", "Mainland : F.coelebs") %>%
              str_replace("Sexmale", "Male") %>%
              str_replace("Age_catfledging", "Fledging") %>%
              str_replace("_F_", " : F.") %>%
              str_replace("c_", "c.") %>%
              str_replace("_", " ")) 

species_estimates <- model_estimates
species_estimates$Order <- factor(species_estimates$island, levels = c("Mainland : F.coelebs", "Species LabelSao_Miguel : F.moreletti", "Species LabelFlores : F.moreletti", "Species LabelMadeira : F.maderensis", "Species LabelEl_Hierro : F.c.ombriosa", "Species LabelLa_Palma : F.c.palmae", "Species LabelGran_Canaria : F.c.bakeri", "Species LabelTenerife : F.c.canariensis", "Species LabelLa_Gomera : F.c.canariensis", "Species LabelTenerife : F.teydea", "Species LabelGran_Canaria : F.polatzeki"), labels = c("Mainland : F.coelebs", "Sao Miguel : F.moreletti", "Flores : F.moreletti", "Madeira : F.maderensis", "El Hierro : F.c.ombriosa", "La Palma : F.c.palmae", "Gran Canaria : F.c.bakeri", "Tenerife : F.c.canariensis", "La Gomera : F.c.canariensis", "Tenerife : F.teydea", "Gran Canaria : F.polatzeki"))
species_estimates$Order <- fct_rev(species_estimates$Order)

mainland_fringilla$Chaff_Blue2 <- as.character(mainland_fringilla$Chaff_Blue)

mainland_fringilla$Chaff_Blue2 <- ifelse(mainland_fringilla$Chaff_Blue2 == "Chaffinch", "Common chaffinch", "Blue chaffinch")

mainland_fringilla$Chaff_Blue2  <- as.factor(mainland_fringilla$Chaff_Blue2)

mainland_fringilla$Chaff_Blue2 <- forcats::fct_relevel(mainland_fringilla$Chaff_Blue2,"Common chaffinch","Blue chaffinch")
table(mainland_fringilla$Chaff_Blue2)

signif2 = distinct(mainland_fringilla, Chaff_Blue2,Order) %>%
    arrange(Chaff_Blue2,Order)

signif2<-left_join(signif2,species_estimates,by="Order")

data.frame(signif2)

#signif2$Estimate
signif2$p.signif = c("","","","","","","*","","*","**","*")

mainland_fringilla$Chaff_Blue2

(p1 <- ggplot(mainland_fringilla, aes(Order, Observed_genus_rare)) + 
        facet_grid(~Chaff_Blue2,scales = "free_x",space = "free_x") +
        geom_boxplot(aes(fill = Species_Label), show.legend = F, lwd = 0.25,fatten = 1, alpha = 0.5, outlier.colour = 'transparent',width=0.5) + 
        geom_point( aes(fill = Species_Label),shape = 21, position = position_jitter(seed = 100, width = 0.1), size = 3,  alpha = 0.7, stroke = 0.25) + 
        SpeciesLabel.fill + 
        ylab("Richness") + 
        xlab("") + sl_theme + 
        boxplot_theme +
        scale_x_discrete(breaks = c("Tenerife : F.teydea","Gran Canaria : F.polatzeki","Mainland : F.coelebs","Sao Miguel : F.moreletti","Flores : F.moreletti","Madeira : F.maderensis","La Palma : F.c.palmae","El Hierro : F.c.ombriosa","Gran Canaria : F.c.bakeri","Tenerife : F.c.canariensis","La Gomera : F.c.canariensis"),labels = c("Tenerife","Gran Canaria","Mainland","Sao Miguel","Flores","Madeira","La Palma","El Hierro","Gran Canaria","Tenerife","La Gomera")) + 
        geom_hline(aes(yintercept = signif2$Estimate[1]), linetype='dashed') +  
        #geom_point(data = signif2, aes( x = Order, y = Estimate),fill = '#acacac', shape = 23, size = 2, stroke = 0.25) + 
        stat_summary(aes(x = Order, y = Observed_genus_rare),fun.data="mean_cl_boot", geom="pointrange",shape = 23, col = "#acacac", fill = '#acacac',size=0.5,fun.args=list(conf.int=.95)) + 
        #geom_linerange(data = signif2, aes( x = Order,ymin = CImin, ymax = CImax), inherit.aes = F, col = '#acacac', size = 0.75) + 
        geom_text(data = signif2, aes(x = Order, y = 250, label = p.signif), vjust = 0.75, size = 8) + scale_y_continuous(breaks = seq(0, 250, by=50)
    #+scale_y_log10()
    ))

```

### PARTITION BETA DIVERSITY

#### Distance metric Baselga family

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

Matrix_abund <- data.frame(otu_table(ps_fringilla_rare))
str(Matrix_abund)

#Baselga family

# Compute and partition a matrix of Jaccard indices (presence-absence data)
j.matrix <- beta.div.comp(Matrix_abund, coef = "BJ", quant = FALSE, save.abc = FALSE)

# Percentage difference, (quantitative form of Sorensen index)
#s.quant <- beta.div.comp(Matrix_abund, coef = "S", quant = TRUE, save.abc = FALSE)

#the percentage difference index (aka Bray-Curtis)
s.quant <- beta.div.comp(Matrix_abund, coef = "BS", quant = TRUE, save.abc = FALSE)


```

#### (Percentage difference, (quantitative form of Sorensen index) (a.k.a. Bray-Curtis))

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

##Replacement values
# dataframe of all pairwise comparisons
xy_repl <- t(combn(colnames(as.matrix(s.quant$repl)), 2))

# add replacement values to df
dist_repl_df <- (data.frame(xy_repl, repl=as.matrix(s.quant$repl)[xy_repl]))

# concatenate labels
dist_repl_df$X1_X2 <- paste0(dist_repl_df$X1,"_",dist_repl_df$X2)

##Richness difference values
# dataframe of all pairwise comparisons
xy_rich <- t(combn(colnames(as.matrix(s.quant$rich)), 2))

# add richness difference values to df
dist_rich_df <- (data.frame(xy_rich, rich=as.matrix(s.quant$rich)[xy_rich]))

# concatenate labels
dist_rich_df$X1_X2 <- paste0(dist_rich_df$X1,"_",dist_rich_df$X2)


## Dissimilarity values

# dataframe of all pairwise comparisons 
xy_D <- t(combn(colnames(as.matrix(s.quant$D)), 2))

# add dissimilarity values to df
dist_D_df <- (data.frame(xy_D, D=as.matrix(s.quant$D)[xy_D]))

# concatenate labels
dist_D_df$X1_X2 <- paste0(dist_D_df$X1,"_",dist_D_df$X2)

## merge all df
dist_df <- (cbind(dist_repl_df,dist_rich_df,dist_D_df)[,c(4,1:3,7,11)])

## Metadata microbiome
Meta <- data.frame(sample_data(ps_fringilla_rare))

# create df for Species Label
SpeciesLabels_X1 <- data.frame(X1 = paste0(Meta$Lab_Code,"_",Meta$Species_Label)
,Species_Label_X1 = Meta$Species_Label)

SpeciesLabels_X2 <- data.frame(X2 = paste0(Meta$Lab_Code,"_",Meta$Species_Label)
,Species_Label_X2 = Meta$Species_Label)

# creating df for Archipelago_Mainland

Archipelago_Mainland_X1 <- data.frame(X1 = paste0(Meta$Lab_Code,"_",Meta$Species_Label),Archipelago_Mainland_X1 = Meta$Archipelago_Mainland)
Archipelago_Mainland_X2 <- data.frame(X2 = paste0(Meta$Lab_Code,"_",Meta$Species_Label),Archipelago_Mainland_X2 = Meta$Archipelago_Mainland)

# Now give labels to distance matrix based on metadata to be able to subset data

# Species label 
dist_df_X1_site <- (left_join(dist_df,SpeciesLabels_X1,by = "X1"))
dist_df_X2_site <- (left_join(dist_df_X1_site,SpeciesLabels_X2,by = "X2"))

# archipelago mainland
dist_df_X2_site <- (left_join(dist_df_X2_site,Archipelago_Mainland_X1,by = "X1"))
dist_df_X2_site <- (left_join(dist_df_X2_site,Archipelago_Mainland_X2,by = "X2"))

# format separator
dist_df_X2_site$X1_X2_Species_Label <- paste0(dist_df_X2_site$Species_Label_X1,":",dist_df_X2_site$Species_Label_X2)
dist_df_X2_site$X1_X2_Same_site <- dist_df_X2_site$Species_Label_X1 == dist_df_X2_site$Species_Label_X2

dist_df_X2_site$X1_X2_Archipelago_Mainland <- paste0(dist_df_X2_site$Archipelago_Mainland_X1,":",dist_df_X2_site$Archipelago_Mainland_X2)

dist_df_diff_site <- dist_df_X2_site[which(dist_df_X2_site$X1_X2_Same_site==FALSE),]
dist_df_same_site <- dist_df_X2_site[which(dist_df_X2_site$X1_X2_Same_site==TRUE),]


# distance matrix where only comparisons with mainland are kept
dist_df_diff_mainland <- dist_df_diff_site[dist_df_diff_site$X1_X2_Species_Label %like% "Mainland_F_coelebs", ]


## Create subsets
dist_df_common <- dist_df_X2_site[which(!dist_df_X2_site$Species_Label_X1=="Tenerife_F_teydea" & !dist_df_X2_site$Species_Label_X1=="Gran_Canaria_F_polatzeki" & !dist_df_X2_site$Species_Label_X2=="Tenerife_F_teydea" & !dist_df_X2_site$Species_Label_X2=="Gran_Canaria_F_polatzeki") ,]


# common chaffinch subset 
dist_df_diff_common <- dist_df_common[which(dist_df_common$X1_X2_Same_site==FALSE),]
head(dist_df_diff_common)

## Make dissimilarity a similarity vector
dist_df_diff_site$S <- (1-dist_df_diff_site$D)
dist_df_diff_mainland$S <- (1-dist_df_diff_mainland$D)
dist_df_diff_common$S <- (1-dist_df_diff_common$D)

# everything sums up to 1 ;) 
dist_df_diff_site$S+dist_df_diff_site$repl+dist_df_diff_site$rich

```


Plot per site variation of beta components
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}




```

Triangle Plots 


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(triangle.overall.diff.bray <-ggtern(data=dist_df_diff_common,aes(rich,repl,S)) +  geom_point(size = 4,  alpha = 0.05, stroke = 0.25) + geom_crosshair_tern(aes(mean(dist_df_diff_common$rich),mean( dist_df_diff_common$repl), mean(dist_df_diff_common$S)),col="darkred",lty="dashed", lineend = "round", size = 2.2) + 
    geom_point(aes(mean(dist_df_diff_common$rich),mean( dist_df_diff_common$repl), mean(dist_df_diff_common$S)),col='darkred',size=10) +
    geom_point(aes(x = mean(dist_df_diff_common$rich), y = 1-mean(dist_df_diff_common$rich),z = 0),col="darkred",size=10) + 
    annotate(size=12,geom  = 'text',x = mean(dist_df_diff_common$rich), y = 1-mean(dist_df_diff_common$rich),z = 0, label = paste(round(mean(dist_df_diff_common$rich)*100,2)), hjust = 1, vjust = -1, size = 5, col = "darkred") +  
    geom_point(aes(x = 0, y = mean(dist_df_diff_common$repl), z = 1-mean(dist_df_diff_common$repl)),col="darkred",size=10) + 
    annotate(size=12,geom  = 'text', x = 0, y = mean(dist_df_diff_common$repl), z = 1-mean(dist_df_diff_common$repl), label = paste(round(mean(dist_df_diff_common$repl)*100,2)), hjust = -0.3, vjust = 1, size = 5, col = "darkred") +  
    geom_point(aes(x = 1-mean(dist_df_diff_common$S), y = 0, z = mean(dist_df_diff_common$S)),col="darkred",size=8) + 
    annotate(size=12,geom  = 'text', x = 1-mean(dist_df_diff_common$S), y = 0, z = mean(dist_df_diff_common$S), label = paste(round(mean(dist_df_diff_common$S)*100,2)), hjust = 0.5, vjust = 2, size = 5, col = "darkred")+
    theme_classic() + theme_hidetitles() + theme_nomask() + theme(axis.text = element_text(size = 28), tern.axis.arrow = element_line(size = 3)))
```


### Beta partition using a matrix of Jaccard indices (presence-absence data)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

str(j.matrix)

##Replacement values
# dataframe of all pairwise comparisons
xy_repl <- t(combn(colnames(as.matrix(j.matrix$repl)), 2))

# add replacement values to df
dist_repl_df <- (data.frame(xy_repl, repl=as.matrix(j.matrix$repl)[xy_repl]))

# concatenate labels
dist_repl_df$X1_X2 <- paste0(dist_repl_df$X1,"_",dist_repl_df$X2)

##Richness difference values
# dataframe of all pairwise comparisons
xy_rich <- t(combn(colnames(as.matrix(j.matrix$rich)), 2))

# add richness difference values to df
dist_rich_df <- (data.frame(xy_rich, rich=as.matrix(j.matrix$rich)[xy_rich]))

# concatenate labels
dist_rich_df$X1_X2 <- paste0(dist_rich_df$X1,"_",dist_rich_df$X2)


## Dissimilarity values

# dataframe of all pairwise comparisons 
xy_D <- t(combn(colnames(as.matrix(j.matrix$D)), 2))

# add dissimilarity values to df
dist_D_df <- (data.frame(xy_D, D=as.matrix(j.matrix$D)[xy_D]))

# concatenate labels
dist_D_df$X1_X2 <- paste0(dist_D_df$X1,"_",dist_D_df$X2)

## merge all df
dist_df <- (cbind(dist_repl_df,dist_rich_df,dist_D_df)[,c(4,1:3,7,11)])

xy_repl <- t(combn(colnames(as.matrix(j.matrix$repl)), 2))
dist_repl_df <- (data.frame(xy_repl, repl=as.matrix(j.matrix$repl)[xy_repl]))
dist_repl_df$X1_X2 <- paste0(dist_repl_df$X1,"_",dist_repl_df$X2)

xy_rich <- t(combn(colnames(as.matrix(j.matrix$rich)), 2))
dist_rich_df <- (data.frame(xy_rich, rich=as.matrix(j.matrix$rich)[xy_rich]))
dist_rich_df$X1_X2 <- paste0(dist_rich_df$X1,"_",dist_rich_df$X2)


xy_D <- t(combn(colnames(as.matrix(j.matrix$D)), 2))
dist_D_df <- (data.frame(xy_D, D=as.matrix(j.matrix$D)[xy_D]))
dist_D_df$X1_X2 <- paste0(dist_D_df$X1,"_",dist_D_df$X2)

dist_df <- (cbind(dist_repl_df,dist_rich_df,dist_D_df)[,c(4,1:3,7,11)])

## Metadata microbiome
Meta <- data.frame(sample_data(ps_fringilla_rare))

# create df for Species Label
SpeciesLabels_X1 <- data.frame(X1 = paste0(Meta$Lab_Code,"_",Meta$Species_Label)
,Species_Label_X1 = Meta$Species_Label)

SpeciesLabels_X2 <- data.frame(X2 = paste0(Meta$Lab_Code,"_",Meta$Species_Label)
,Species_Label_X2 = Meta$Species_Label)

# creating df for Archipelago_Mainland

Archipelago_Mainland_X1 <- data.frame(X1 = paste0(Meta$Lab_Code,"_",Meta$Species_Label),Archipelago_Mainland_X1 = Meta$Archipelago_Mainland)
Archipelago_Mainland_X2 <- data.frame(X2 = paste0(Meta$Lab_Code,"_",Meta$Species_Label),Archipelago_Mainland_X2 = Meta$Archipelago_Mainland)

# Now give labels to distance matrix based on metadata to be able to subset data

# Species label 

j_dist_df_X1_site <- (left_join(dist_df,SpeciesLabels_X1,by = "X1"))
j_dist_df_X2_site <- (left_join(j_dist_df_X1_site,SpeciesLabels_X2,by = "X2"))

# archipelago mainland

j_dist_df_X2_site <- (left_join(j_dist_df_X2_site,Archipelago_Mainland_X1,by = "X1"))
j_dist_df_X2_site <- (left_join(j_dist_df_X2_site,Archipelago_Mainland_X2,by = "X2"))

# format separator

j_dist_df_X2_site$X1_X2_Species_Label <- paste0(j_dist_df_X2_site$Species_Label_X1,":",j_dist_df_X2_site$Species_Label_X2)
j_dist_df_X2_site$X1_X2_Same_site <- j_dist_df_X2_site$Species_Label_X1 == j_dist_df_X2_site$Species_Label_X2

j_dist_df_X2_site$X1_X2_Archipelago_Mainland <- paste0(j_dist_df_X2_site$Archipelago_Mainland_X1,":",j_dist_df_X2_site$Archipelago_Mainland_X2)

dist_df_diff_site.j <- j_dist_df_X2_site[which(j_dist_df_X2_site$X1_X2_Same_site==FALSE),]
dist_df_same_site.j <- j_dist_df_X2_site[which(j_dist_df_X2_site$X1_X2_Same_site==TRUE),]

## Create subsets

dist_df_common <- j_dist_df_X2_site[which(!j_dist_df_X2_site$Species_Label_X1=="Tenerife_F_teydea" & !j_dist_df_X2_site$Species_Label_X1=="Gran_Canaria_F_polatzeki" & !j_dist_df_X2_site$Species_Label_X2=="Tenerife_F_teydea" & !j_dist_df_X2_site$Species_Label_X2=="Gran_Canaria_F_polatzeki") ,]

## For all dataframes, create two subsets, one of pairwise comparisons of individuals of the same population (within pops), and on of pairwise comparisons of individuals from different sites (among pops)

# common chaffinch subset 

j_dist_df_diff_common <- dist_df_common[which(dist_df_common$X1_X2_Same_site==FALSE),]
dist_df_same_common <- dist_df_common[which(dist_df_common$X1_X2_Same_site==TRUE),]

## Make dissimilarity a similarity vector

dist_df_diff_site.j$S <- (1-dist_df_diff_site.j$D)
dist_df_same_site.j$S <- (1-dist_df_same_site.j$D)
dist_df_common$S <- (1-dist_df_common$D)
j_dist_df_diff_common$S <- (1-j_dist_df_diff_common$D)
dist_df_same_common$S <- (1-dist_df_same_common$D)


dist_df_diff_site.j$S+dist_df_diff_site.j$repl+dist_df_diff_site.j$rich


```



```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(triangle.overall.diff.jaccard<-ggtern(data=j_dist_df_diff_common,aes(rich,repl,S)) +  geom_point(size = 4,  alpha = 0.05, stroke = 0.25) + geom_crosshair_tern(aes(mean(j_dist_df_diff_common$rich),mean( j_dist_df_diff_common$repl), mean(j_dist_df_diff_common$S)),col="darkred",lty="dashed", lineend = "round", size = 2.2) + 
    geom_point(aes(mean(j_dist_df_diff_common$rich),mean( j_dist_df_diff_common$repl), mean(j_dist_df_diff_common$S)),col='darkred',size=10) +
    geom_point(aes(x = mean(j_dist_df_diff_common$rich), y = 1-mean(j_dist_df_diff_common$rich),z = 0),col="darkred",size=10) + 
    annotate(size=12,geom  = 'text',x = mean(j_dist_df_diff_common$rich), y = 1-mean(j_dist_df_diff_common$rich),z = 0, label = paste(round(mean(j_dist_df_diff_common$rich)*100,2)), hjust = 1, vjust = -1, size = 5, col = "darkred") +  
    geom_point(aes(x = 0, y = mean(j_dist_df_diff_common$repl), z = 1-mean(j_dist_df_diff_common$repl)),col="darkred",size=10) + 
    annotate(size=12,geom  = 'text', x = 0, y = mean(j_dist_df_diff_common$repl), z = 1-mean(j_dist_df_diff_common$repl), label = paste(round(mean(j_dist_df_diff_common$repl)*100,2)), hjust = -0.3, vjust = 1, size = 5, col = "darkred") +  
    geom_point(aes(x = 1-mean(j_dist_df_diff_common$S), y = 0, z = mean(j_dist_df_diff_common$S)),col="darkred",size=8) + 
    annotate(size=12,geom  = 'text', x = 1-mean(j_dist_df_diff_common$S), y = 0, z = mean(j_dist_df_diff_common$S), label = paste(round(mean(j_dist_df_diff_common$S)*100,2)), hjust = 0.5, vjust = 2, size = 5, col = "darkred")+
    theme_classic() + theme_hidetitles() + theme_nomask() + theme(axis.text = element_text(size = 28), tern.axis.arrow = element_line(size = 3)))

```



# BIOGEOGRAPHY


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

nomainland_fringilla <- mainland_fringilla %>% dplyr::filter(!is.na(Area_km2) & Genus == "Fringilla")
str(nomainland_fringilla)

GLMM1 <- glmmTMB(Observed_genus_rare ~  scale(Area_km2) + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM1)

#Conditional model:
#                Estimate Std. Error z value Pr(>|z|)    
#(Intercept)      4.28862    0.08358   51.31   <2e-16 ***
#scale(Area_km2)  0.15658    0.07880    1.99   0.0469 *  
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


confint(GLMM1)

    
(pArea <- ggpredict(GLMM1, terms = "Area_km2 [all]") %>% plot(line_size = 1.5, color = "darkred") +
        geom_point(data = nomainland_fringilla, aes(y = Observed_genus_rare, x = Area_km2),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 50), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(alpha = 1, size = 0.4, stoke = 0.01) + 
        ylab("Richness") + 
        xlab(bquote(Area~(km^2))) +
       scatter_theme + ggtitle("") + sl_theme
        + annotate(geom="text",label="atop(paste(~beta~' = 0.157 [0.002, 0.311];'),~italic('p-value')~'= 0.047')",parse=T, x = 1000,  y = 240, size = 6)
        + scale_y_continuous(breaks = seq(0, 250, by=50)))

GLMM2 <- glmmTMB(Observed_genus_rare ~ scale(Distance_mainland) + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)

summary(GLMM2)
#Conditional model:
#                         Estimate Std. Error z value Pr(>|z|)    
#(Intercept)               4.28778    0.06611   64.86  < 2e-16 ***
#scale(Distance_mainland) -0.22620    0.06831   -3.31 0.000929 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


confint(GLMM2)

#GLMM2 <- glmmTMB(Observed_genus_rare ~ Distance_mainland + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
#summary(GLMM2)

(pDistance <- ggpredict(GLMM2, terms = "Distance_mainland [all]") %>% plot(line_size = 1.5, color = "darkred") +
        geom_point(data = nomainland_fringilla, aes(y = Observed_genus_rare, x = Distance_mainland),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 50), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(alpha = 1, size = 0.4, stoke = 0.01) + 
        ylab("") +
        xlab("Distance from mainland (km)") + 
        scatter_theme + ggtitle("") + sl_theme
        + annotate(geom="text",label="atop(paste(~beta~' = -0.226 [-0.360, -0.092];'),~italic('p-value')~'< 0.001')",parse=T, x = 1000,  y = 240, size = 6)
        + scale_y_continuous(breaks = seq(0, 250, by=50)))

GLMM3 <- glmmTMB(Observed_genus_rare ~  scale(Age_ma) + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
summary(GLMM3)

#Conditional model:
#              Estimate Std. Error z value Pr(>|z|)    
#(Intercept)    4.29167    0.05827   73.65  < 2e-16 ***
#scale(Age_ma)  0.24298    0.05877    4.13 3.56e-05 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


confint(GLMM3)

#GLMM3 <- glmmTMB(Observed_genus_rare ~  Age_ma + (1|Species_Label), data = nomainland_fringilla, na.action = na.fail, family = nbinom2)
#summary(GLMM3)

#slopes(GLMM3)

(pAge_ma <- ggpredict(GLMM3, terms = "Age_ma [all]") %>% plot(line_size = 1.5, color = "darkred") +
        geom_point(data = nomainland_fringilla, aes(y = Observed_genus_rare, x = Age_ma),shape = 21, fill = "grey", position = position_jitter(seed = 100, width = 0.3), size = 3,  alpha = 0.9, stroke = 0.25) + 
        #geom_line(col = "darkred", alpha = 1, size = 0.4, stoke = 0.01) +
        ylab("") +
        xlab("Island age (mya)") + 
        scatter_theme + ggtitle("") + sl_theme
        + annotate(geom="text",label="atop(paste(~beta~' = 0.243 [0.128, 0.3358];'),~italic('p-value')~'< 0.001')",parse=T, x = 7.5,  y = 240, size = 6)
        + scale_y_continuous(breaks = seq(0, 250, by=50)))

```

## Patchwork
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

png(paste0(outputDir, "/Figures/Colonisation_history_speciesFix.png"), width = 300, height = 240, units = "mm", res = 350)
((p1 +  theme(axis.title.x = element_blank(), axis.text.x =element_blank(),strip.text.x = element_text(size = 18))) / D_overall_bray_patch)

dev.off()



png("~/Dropbox/FringillaMicrobiome/08_Output/Figures/triangle.plots.baselga.sorensen.quant.overall.matrix_genus_noarrows_2panels_speciesFix.png" ,width=8000, height=8000,res=650)

(triangle.overall.diff.bray <- triangle.overall.diff.bray + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()))

dev.off()


png("~/Dropbox/FringillaMicrobiome/08_Output/Figures/triangle.plots.baselga.j.overall.matrix_genus_2panels_speciesFix.png" ,width=8000, height=8000,res=650)

(triangle.overall.diff.jaccard <- triangle.overall.diff.jaccard + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()))

dev.off()


png(paste0(outputDir, "/Figures/Biogeography_speciesFix.png"), width = 500, height = 200, units = "mm", res = 350)
pArea + pDistance + pAge_ma 

dev.off()

```


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}




(((p1 +  theme(axis.title.x = element_blank(), axis.text.x =element_blank())) / D_overall_bray_patch) | ( triangle.overall.diff.bray  / triangle.overall.diff.jaccard)) / (pArea + pDistance + pAge_ma) + plot_annotation(tag_levels = 'A') &  theme(plot.tag = element_text(size = 24, face = "bold"))


```
