---
title: 'MRM microbiome '
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute:
    code_folding: hide
    toc_depth: 2

---


## LOAD LIBRARIES

```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rm(list = ls())

# basics
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(vegan)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(MicEco)
library(rlist) 
library(grid)
library(tsiR) 
library(ape)
library(ggh4x)
# phyloseq section
library(phyloseq)
library(microViz)
library(biomeUtils)
library(patchwork)

# dist
library(treedataverse)
library(geosphere)
library(treedata.table)


# MRM 
library(sf)
library(ecodist)
library(modEvA)
library(eulerr)


```


## WORKING DIRECTORIES
```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```


## LOAD DATA

Rarefied 16S

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr_otu <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr_species <- speedyseq::tax_glom(microb_fr_otu, "Species")
microb_fr_genus <- speedyseq::tax_glom(microb_fr_otu, "Genus")
microb_fr_family <- speedyseq::tax_glom(microb_fr_otu, "Family")
microb_fr_order <- speedyseq::tax_glom(microb_fr_otu, "Order")
microb_fr_class <- speedyseq::tax_glom(microb_fr_otu, "Class")
microb_fr_phylum <- speedyseq::tax_glom(microb_fr_otu, "Phylum")

# check for NAs
tax_micro <- tax_table(microb_fr_otu) %>% data.frame
table(is.na(tax_micro$Genus))

# TipId 
sample_data(microb_fr_otu)$TipID_microbiome <- paste(sample_data(microb_fr_otu)$Lab_Code, sample_data(microb_fr_otu)$Species_Label, sep = "_")
sample_data(microb_fr_species)$TipID_microbiome <- paste(sample_data(microb_fr_species)$Lab_Code, sample_data(microb_fr_species)$Species_Label, sep = "_")
sample_data(microb_fr_genus)$TipID_microbiome <- paste(sample_data(microb_fr_genus)$Lab_Code, sample_data(microb_fr_genus)$Species_Label, sep = "_")
sample_data(microb_fr_family)$TipID_microbiome <- paste(sample_data(microb_fr_family)$Lab_Code, sample_data(microb_fr_family)$Species_Label, sep = "_")
sample_data(microb_fr_order)$TipID_microbiome <- paste(sample_data(microb_fr_order)$Lab_Code, sample_data(microb_fr_order)$Species_Label, sep = "_")
sample_data(microb_fr_class)$TipID_microbiome <- paste(sample_data(microb_fr_class)$Lab_Code, sample_data(microb_fr_class)$Species_Label, sep = "_")
sample_data(microb_fr_phylum)$TipID_microbiome <- paste(sample_data(microb_fr_phylum)$Lab_Code, sample_data(microb_fr_phylum)$Species_Label, sep = "_")

# meta
microb_fr_meta_otu <- sample_data(microb_fr_otu) %>% data.frame
microb_fr_meta_species <- sample_data(microb_fr_species) %>% data.frame
microb_fr_meta_genus <- sample_data(microb_fr_genus) %>% data.frame
microb_fr_meta_family <- sample_data(microb_fr_family) %>% data.frame
microb_fr_meta_order <- sample_data(microb_fr_order) %>% data.frame
microb_fr_meta_class <- sample_data(microb_fr_class) %>% data.frame
microb_fr_meta_phylum <- sample_data(microb_fr_phylum) %>% data.frame

```


Diet data

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

diet_fr <- readRDS(paste0(dataDir, "/diet_overlap_OTU99_fr.rds"))

```


## Overlap all datasets

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_overlap_otu <- microb_fr_otu %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_species <- microb_fr_species %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_genus <- microb_fr_genus %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_family <- microb_fr_family %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_class <- microb_fr_class %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_order <- microb_fr_order %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_phylum <- microb_fr_phylum %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")

microb_overlap_meta <- microb_overlap_otu %>% sample_data %>% data.frame 
nrow(microb_overlap_meta)


```


## Distance matrix

Geographic distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# lat long data
my_points <- data.frame(longitude = sample_data(microb_overlap_otu)$Xdec, latitude = sample_data(microb_overlap_otu)$Ydec, Island_Mainland = sample_data(microb_overlap_otu)$Island_Mainland, row.names = sample_names(microb_overlap_otu))
head(my_points)

# distance based Moran's eigenvectors
#cartesian <- cbind(my_points$longitude, my_points$latitude)
#geo_MEM<-dbmem(cartesian,  thresh = NULL,  MEM.autocor = c("positive"), store.listw = TRUE,silent = TRUE )
#summary(geo_MEM)
dist_mat <- distm(cbind(my_points$longitude, my_points$latitude), fun = distGeo)  # Vector of distances in meters
dist_mat <-dist_mat/1000 #Vector of distances in km
rownames(dist_mat) <- rownames(my_points)
colnames(dist_mat) <- rownames(my_points)
geo_dist <- as.dist(dist_mat)
str(dist_mat)

```


Microbiome distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

## OTU
# Bray
microb_otu_tab <- otu_table(microb_overlap_otu)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_otu <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_otu)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_otu <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)


## species
# Bray
microb_species_tab <- otu_table(microb_overlap_species)
microb_species_tab_order <- microb_species_tab[rownames(dist_mat),]
bray_dist_species <- vegdist(microb_species_tab_order, method = "bray")

# Jaccard 
microb_species_tab <- otu_table(microb_overlap_species)
microb_species_tab_order <- microb_species_tab[rownames(dist_mat),]
jaccard_dist_species <- vegdist(microb_species_tab_order, method = "jaccard", binary = TRUE)

## genus
# Bray
microb_otu_tab <- otu_table(microb_overlap_genus)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_genus <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_genus)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_genus <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## family
# Bray
microb_otu_tab <- otu_table(microb_overlap_family)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_family <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_family)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_family <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## order
# Bray
microb_otu_tab <- otu_table(microb_overlap_order)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_order <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_order)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_order <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## class
# Bray
microb_otu_tab <- otu_table(microb_overlap_class)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_class <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_class)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_class <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## phylum
# Bray
microb_otu_tab <- otu_table(microb_overlap_phylum)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_phylum <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_phylum)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_phylum <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)
```

Diet distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

diet_overlap <-  prune_samples(sample_names(diet_fr) %in% sample_names(microb_overlap_otu), diet_fr) 


# OTU bray
diet_otu_tab <- otu_table(diet_overlap)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_bray_dist <- vegdist(diet_otu_tab_order, method = "bray")
str(diet_bray_dist)

# OTU bray
diet_otu_tab <- otu_table(diet_overlap)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_jaccard_dist <- vegdist(diet_otu_tab_order, method = "jaccard", binary = TRUE)
#str(diet_jaccard_dist)

#diet_ruzicka_dist <- vegdist(diet_otu_tab_order, method = "jaccard", binary = FALSE)


# Raup-Crick
diet_otu_tab <- otu_table(diet_overlap)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_rc_dist <- raupcrick(diet_otu_tab_order) 


```

Host distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


#nuclear genome

fringilla_tree <-  read.tree(file = paste0(dataDir, "/variants_microbiome_subset_miss0.5_mac3_noZ_subsample.phy.contree"))


annot <- read_delim(paste0(dataDir, "/WGS_Fringilla_samples.tsv"))

#annot <- read_tsv(paste0(dataDir, "/annot_tree_Fringilla_WGSmt_JCI.tsv"))

annot %>% data.frame %>% tail

table(annot$Microb)

#rename row and col names to match microbiome and diet matrices 
annot$TipID_microbiome

fringilla_tree$tip.label <- annot[match(fringilla_tree$tip.label, annot$WGS_ID),]$TipID_microbiome
length(setdiff(fringilla_tree$tip.label, microb_overlap_meta$TipID_microbiome))


# subset tree matching IDs with microbiome data 
fringilla_tree_subset <- as.treedata.table(tree = fringilla_tree, data = microb_overlap_meta, name_column = "TipID_microbiome")

fringilla_tree_subset <- fringilla_tree_subset$phy
str(fringilla_tree_subset)

fringilla_rootedtree = root(fringilla_tree_subset, outgroup = "149_Tenerife_F_teydea", resolve.root = TRUE)

microb_overlap_meta <- annot %>% dplyr::filter(Microb == 1) %>% as.data.frame
microb_overlap_meta$label <- microb_overlap_meta$TipID_microbiome 
microb_overlap_meta %>% names
microb_overlap_meta %>% head
class(microb_overlap_meta$label)

tree_microb_overlap_meta <- full_join(fringilla_rootedtree, microb_overlap_meta, by = "label")

plot(fringilla_rootedtree)

host_tree <- list(fringilla_tree_subset)
summary(host_tree)

# Host distance matrix
host_dist <- fringilla_tree_subset %>% cophenetic.phylo 
host_dist_order <- host_dist[rownames(dist_mat), colnames(dist_mat)]
head(host_dist)
host_dist %>% dim

head(host_dist)

host_dist <- as.dist(host_dist_order)

```

### Mantel tests

Mantel Bray-Curtis host distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_bray_phylog_otu <- vegan::mantel(bray_dist_otu, host_dist, method = "pearson", permutations = 9999))
str(mantel_bray_phylog_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_phylog_otu$statistic, p=mantel_bray_phylog_otu$signif)


(mantel_bray_geo_otu  <- vegan::mantel.partial(bray_dist_otu, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))
df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)



vegan::mantel(host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999)

#Call:
#vegan::mantel(xdis = host_dist, ydis = as.dist(log10(geo_dist +#      1)), method = "pearson", permutations = 9999) 
#
#Mantel statistic r: 0.2439 
#      Significance: 1e-04 
#
#Upper quantiles of permutations (null model):
#   90%    95%  97.5%    99% 
#0.0542 0.0730 0.0900 0.1127 
#Permutation: free
#Number of permutations: 9999



bray = as.vector(bray_dist_otu)
jacc = as.vector(jaccard_dist_otu)
host = as.vector(host_dist)
geo = as.vector(geo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(bray, jacc, host, geo)
#mat_edit <- read.csv(paste0(dataDir, "/host_dist_d


(p1 <- ggplot(mat, aes(y = bray, x = host)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Host phylogenetic distance", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_phylog_bray$statistic, 3), ";p-value=", round(mantel_phylog_bray$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

(p2 <- ggplot(mat, aes(y = host, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "Host phylogenetic distance") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))

# Species
(mantel_bray_phylog_species <- vegan::mantel(bray_dist_species, host_dist, method = "pearson", permutations = 9999))

#Mantel statistic r: 0.1828 
#      Significance: 1e-04 

#Upper quantiles of permutations (null model):
#   90%    95%  97.5%    99% 
#0.0473 0.0615 0.0747 0.0896 
#Permutation: free
#Number of permutations: 9999



(mantel_bray_geo_species  <- vegan::mantel.partial(bray_dist_species, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

#Mantel statistic r: 0.1506 
#      Significance: 2e-04 

#Upper quantiles of permutations (null model):
#   90%    95%  97.5%    99% 
#0.0495 0.0639 0.0769 0.0927 
#Permutation: free
#Number of permutations: 9999


# genus
(mantel_bray_phylog_genus <- vegan::mantel(bray_dist_genus, host_dist, method = "pearson", permutations = 9999))


(mantel_bray_geo_genus  <- vegan::mantel.partial(bray_dist_genus, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# family
(mantel_bray_phylog_family <- vegan::mantel(bray_dist_family, host_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_family  <- vegan::mantel.partial(bray_dist_family, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# order
(mantel_bray_phylog_order <- vegan::mantel(bray_dist_order, host_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_order  <- vegan::mantel.partial(bray_dist_order, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# class
(mantel_bray_phylog_class <- vegan::mantel(bray_dist_class, host_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_class  <- vegan::mantel.partial(bray_dist_class, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# phylum
(mantel_bray_phylog_phylum <- vegan::mantel(bray_dist_phylum, host_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_phylum  <- vegan::mantel.partial(bray_dist_phylum, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))


# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_phylog_otu$statistic, p=mantel_bray_phylog_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)


# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_phylog_species$statistic, p=mantel_bray_phylog_species$signif)

df_species_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_geo_species$statistic, p=mantel_bray_geo_species$signif)


# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_phylog_genus$statistic, p=mantel_bray_phylog_genus$signif)

df_genus_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_geo_genus$statistic, p=mantel_bray_geo_genus$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_phylog_family$statistic, p=mantel_bray_phylog_family$signif)

df_family_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_geo_family$statistic, p=mantel_bray_geo_family$signif)


# order
df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_phylog_order$statistic, p=mantel_bray_phylog_order$signif)

df_order_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_geo_order$statistic, p=mantel_bray_geo_order$signif)

# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_phylog_class$statistic, p=mantel_bray_phylog_class$signif)

df_class_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_geo_class$statistic, p=mantel_bray_geo_class$signif)

# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_phylog_phylum$statistic, p=mantel_bray_phylog_phylum$signif)

df_phylum_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_geo_phylum$statistic, p=mantel_bray_geo_phylum$signif)

mantel_out_bray <- rbind(df_otu, df_otu_partial, df_species, df_species_partial, df_genus, df_genus_partial, df_family, df_family_partial,  df_order, df_order_partial,df_class, df_class_partial, df_phylum, df_phylum_partial)

```


Mantel Jaccard host distance 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_jaccard_phylog_otu <- vegan::mantel(jaccard_dist_otu, host_dist, method = "pearson", permutations = 9999))
str(mantel_jaccard_phylog_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_phylog_otu$statistic, p=mantel_jaccard_phylog_otu$signif)

(mantel_jaccard_geo_otu  <- vegan::mantel.partial(jaccard_dist_otu, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))
df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

jacc = as.vector(jaccard_dist_otu)
host = as.vector(host_dist)
geo = as.vector(geo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(jacc, host, geo)

(p1 <- ggplot(mat, aes(y = jacc, x = host)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "host", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_phylog_jaccard$statistic, 3), ";p-value=", round(mantel_phylog_jaccard$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

(p2 <- ggplot(mat, aes(y = host, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "host") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))


# species
(mantel_jaccard_phylog_species <- vegan::mantel(jaccard_dist_species, host_dist, method = "pearson", permutations = 9999))

#Mantel statistic r: 0.09905 
#      Significance: 0.0406 

#Upper quantiles of permutations (null model):
#   90%    95%  97.5%    99% 
#0.0708 0.0931 0.1133 0.1363 
#Permutation: free
#Number of permutations: 9999


(mantel_jaccard_geo_species  <- vegan::mantel.partial(jaccard_dist_species, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))



# genus
(mantel_jaccard_phylog_genus <- vegan::mantel(jaccard_dist_genus, host_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_genus  <- vegan::mantel.partial(jaccard_dist_genus, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))


# family
(mantel_jaccard_phylog_family <- vegan::mantel(jaccard_dist_family, host_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_family  <- vegan::mantel.partial(jaccard_dist_family, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# class
(mantel_jaccard_phylog_class <- vegan::mantel(jaccard_dist_class, host_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_class  <- vegan::mantel.partial(jaccard_dist_class, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# order
(mantel_jaccard_phylog_order <- vegan::mantel(jaccard_dist_order, host_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_order  <- vegan::mantel.partial(jaccard_dist_order, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# phylum
(mantel_jaccard_phylog_phylum <- vegan::mantel(jaccard_dist_phylum, host_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_phylum  <- vegan::mantel.partial(jaccard_dist_phylum, host_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_phylog_otu$statistic, p=mantel_jaccard_phylog_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)


# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_phylog_species$statistic, p=mantel_jaccard_phylog_species$signif)

df_species_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_geo_species$statistic, p=mantel_jaccard_geo_species$signif)


# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_phylog_genus$statistic, p=mantel_jaccard_phylog_genus$signif)

df_genus_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_geo_genus$statistic, p=mantel_jaccard_geo_genus$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_phylog_family$statistic, p=mantel_jaccard_phylog_family$signif)

df_family_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_geo_family$statistic, p=mantel_jaccard_geo_family$signif)

# order

df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_phylog_order$statistic, p=mantel_jaccard_phylog_order$signif)

df_order_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_geo_order$statistic, p=mantel_jaccard_geo_order$signif)

# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_phylog_class$statistic, p=mantel_jaccard_phylog_class$signif)

df_class_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_geo_class$statistic, p=mantel_jaccard_geo_class$signif)

# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_phylog_phylum$statistic, p=mantel_jaccard_phylog_phylum$signif)

df_phylum_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_geo_phylum$statistic, p=mantel_jaccard_geo_phylum$signif)

mantel_out_jaccard <- rbind(df_otu, df_otu_partial, df_species,  df_species_partial, df_genus,  df_genus_partial, df_family, df_family_partial,  df_order, df_order_partial, df_class, df_class_partial, df_phylum, df_phylum_partial)



```

Join bray and jaccard output

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

mantel_out <- rbind(mantel_out_bray, mantel_out_jaccard)
head(mantel_out)

write.csv(mantel_out, paste0(dataDir, "/output_mantel_phylogDistance_nuclearTree.csv"), row.names=FALSE)

```






Mantel Bray-Curtis diet

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_bray_diet_otu <- vegan::mantel(bray_dist_otu, diet_bray_dist, method = "pearson", permutations = 9999))

str(mantel_bray_diet_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_diet_otu$statistic, p=mantel_bray_diet_otu$signif)


(mantel_bray_diet_otu  <- vegan::mantel.partial(bray_dist_otu, diet_bray_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))



(mantel_bray_geo_otu  <- vegan::mantel.partial(bray_dist_otu, diet_bray_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))
df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)


df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_diet_otu$statistic, p=mantel_bray_diet_otu$signif)


bray = as.vector(bray_dist_otu)
jacc = as.vector(jaccard_dist_otu)
diet = as.vector(diet_bray_dist)
geo = as.vector(geo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(bray, jacc, diet, geo)
#mat_edit <- read.csv(paste0(dataDir, "/diet_bray_dist_d


(p1 <- ggplot(mat, aes(y = bray, x = diet)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Diet distance", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_phylog_bray$statistic, 3), ";p-value=", round(mantel_phylog_bray$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

(p2 <- ggplot(mat, aes(y = diet, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "Diet distance") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))

# species
(mantel_bray_diet_species <- vegan::mantel(bray_dist_species, diet_bray_dist, method = "pearson", permutations = 9999))

#Mantel statistic r: 0.06307 
#      Significance: 0.0189 

#Upper quantiles of permutations (null model):
#   90%    95%  97.5%    99% 
#0.0321 0.0443 0.0577 0.0711 
#Permutation: free
#Number of permutations: 9999


(mantel_bray_geo_species  <- vegan::mantel.partial(bray_dist_species, diet_bray_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

#Mantel statistic r: 0.004969 
#      Significance: 0.3858 

#Upper quantiles of permutations (null model):
#   90%    95%  97.5%    99% 
#0.0300 0.0400 0.0511 0.0650 
#Permutation: free
#Number of permutations: 9999



# genus
(mantel_bray_diet_genus <- vegan::mantel(bray_dist_genus, diet_bray_dist, method = "pearson", permutations = 9999))


(mantel_bray_geo_genus  <- vegan::mantel.partial(bray_dist_genus, diet_bray_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# family
(mantel_bray_diet_family <- vegan::mantel(bray_dist_family, diet_bray_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_family  <- vegan::mantel.partial(bray_dist_family, diet_bray_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# order
(mantel_bray_diet_order <- vegan::mantel(bray_dist_order, diet_bray_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_order  <- vegan::mantel.partial(bray_dist_order, diet_bray_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# class
(mantel_bray_diet_class <- vegan::mantel(bray_dist_class, diet_bray_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_class  <- vegan::mantel.partial(bray_dist_class, diet_bray_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# phylum
(mantel_bray_diet_phylum <- vegan::mantel(bray_dist_phylum, diet_bray_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_phylum  <- vegan::mantel.partial(bray_dist_phylum, diet_bray_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))


# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_diet_otu$statistic, p=mantel_bray_diet_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)

# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_diet_species$statistic, p=mantel_bray_diet_species$signif)

df_species_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_geo_species$statistic, p=mantel_bray_geo_species$signif)

# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_diet_genus$statistic, p=mantel_bray_diet_genus$signif)

df_genus_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_geo_genus$statistic, p=mantel_bray_geo_genus$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_diet_family$statistic, p=mantel_bray_diet_family$signif)

df_family_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_geo_family$statistic, p=mantel_bray_geo_family$signif)


# order
df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_diet_order$statistic, p=mantel_bray_diet_order$signif)

df_order_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_geo_order$statistic, p=mantel_bray_geo_order$signif)

# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_diet_class$statistic, p=mantel_bray_diet_class$signif)

df_class_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_geo_class$statistic, p=mantel_bray_geo_class$signif)

# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_diet_phylum$statistic, p=mantel_bray_diet_phylum$signif)

df_phylum_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_geo_phylum$statistic, p=mantel_bray_geo_phylum$signif)

mantel_out_bray <- rbind(df_otu, df_otu_partial, df_species, df_species_partial, df_genus, df_genus_partial, df_family, df_family_partial,  df_order, df_order_partial,df_class, df_class_partial, df_phylum, df_phylum_partial)

```


Mantel Jaccard
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_jaccard_diet_otu <- vegan::mantel(jaccard_dist_otu, diet_jaccard_dist, method = "pearson", permutations = 9999))
str(mantel_jaccard_diet_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_diet_otu$statistic, p=mantel_jaccard_diet_otu$signif)

(mantel_jaccard_geo_otu  <- vegan::mantel.partial(jaccard_dist_otu, diet_jaccard_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))
df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

jacc = as.vector(jaccard_dist_otu)
diet = as.vector(diet_jaccard_dist)
geo = as.vector(geo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(jacc, diet, geo)

(p1 <- ggplot(mat, aes(y = jacc, x = diet)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Diet distance", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_diet_jaccard$statistic, 3), ";p-value=", round(mantel_diet_jaccard$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

(p2 <- ggplot(mat, aes(y = diet, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "Diet distance") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))


# species
(mantel_jaccard_diet_species <- vegan::mantel(jaccard_dist_species, diet_jaccard_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_species  <- vegan::mantel.partial(jaccard_dist_species, diet_jaccard_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))


# genus
(mantel_jaccard_diet_genus <- vegan::mantel(jaccard_dist_genus, diet_jaccard_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_genus  <- vegan::mantel.partial(jaccard_dist_genus, diet_jaccard_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))


# family
(mantel_jaccard_diet_family <- vegan::mantel(jaccard_dist_family, diet_jaccard_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_family  <- vegan::mantel.partial(jaccard_dist_family, diet_jaccard_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# class
(mantel_jaccard_diet_class <- vegan::mantel(jaccard_dist_class, diet_jaccard_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_class  <- vegan::mantel.partial(jaccard_dist_class, diet_jaccard_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# order
(mantel_jaccard_diet_order <- vegan::mantel(jaccard_dist_order, diet_jaccard_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_order  <- vegan::mantel.partial(jaccard_dist_order, diet_jaccard_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# phylum
(mantel_jaccard_diet_phylum <- vegan::mantel(jaccard_dist_phylum, diet_jaccard_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_phylum  <- vegan::mantel.partial(jaccard_dist_phylum, diet_jaccard_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_diet_otu$statistic, p=mantel_jaccard_diet_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_diet_species$statistic, p=mantel_jaccard_diet_species$signif)

df_species_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_geo_species$statistic, p=mantel_jaccard_geo_species$signif)

# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_diet_genus$statistic, p=mantel_jaccard_diet_genus$signif)

df_genus_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_geo_genus$statistic, p=mantel_jaccard_geo_genus$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_diet_family$statistic, p=mantel_jaccard_diet_family$signif)

df_family_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_geo_family$statistic, p=mantel_jaccard_geo_family$signif)

# order

df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_diet_order$statistic, p=mantel_jaccard_diet_order$signif)

df_order_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_geo_order$statistic, p=mantel_jaccard_geo_order$signif)

# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_diet_class$statistic, p=mantel_jaccard_diet_class$signif)

df_class_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_geo_class$statistic, p=mantel_jaccard_geo_class$signif)

# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_diet_phylum$statistic, p=mantel_jaccard_diet_phylum$signif)

df_phylum_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_geo_phylum$statistic, p=mantel_jaccard_geo_phylum$signif)

mantel_out_jaccard <- rbind(df_otu, df_otu_partial, df_species,  df_species_partial, df_genus,  df_genus_partial, df_family, df_family_partial,  df_order, df_order_partial, df_class, df_class_partial, df_phylum, df_phylum_partial)



```

Join bray and jaccard output

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

mantel_out_diet <- rbind(mantel_out_bray, mantel_out_jaccard)
head(mantel_out_diet)

write.csv(mantel_out_diet, paste0(dataDir, "/output_mantel_diet.csv"), row.names=FALSE)

```





Mantel Raup-Crick diet - Bray-Curtis microbiome

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_rc_otu <- vegan::mantel(bray_dist_otu, diet_rc_dist, method = "pearson", permutations = 9999))

str(mantel_rc_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_rc_otu$statistic, p=mantel_rc_otu$signif)


(mantel_rc_otu  <- vegan::mantel.partial(bray_dist_otu, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))



(mantel_bray_geo_otu  <- vegan::mantel.partial(bray_dist_otu, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))
df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)


df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_rc_otu$statistic, p=mantel_rc_otu$signif)


bray = as.vector(bray_dist_otu)
jacc = as.vector(jaccard_dist_otu)
diet = as.vector(diet_rc_dist)
geo = as.vector(geo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(bray, jacc, diet, geo)
#mat_edit <- read.csv(paste0(dataDir, "/diet_rc_dist_d


(p1 <- ggplot(mat, aes(y = bray, x = diet)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Diet distance (Raup-Crick)", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_phylog_bray$statistic, 3), ";p-value=", round(mantel_phylog_bray$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

(p2 <- ggplot(mat, aes(y = diet, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "Diet distance (Raup-Crick)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))

# species
(mantel_rc_species <- vegan::mantel(bray_dist_species, diet_rc_dist, method = "pearson", permutations = 9999))


(mantel_bray_geo_species  <- vegan::mantel.partial(bray_dist_species, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# genus
(mantel_rc_genus <- vegan::mantel(bray_dist_genus, diet_rc_dist, method = "pearson", permutations = 9999))


(mantel_bray_geo_genus  <- vegan::mantel.partial(bray_dist_genus, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# family
(mantel_rc_family <- vegan::mantel(bray_dist_family, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_family  <- vegan::mantel.partial(bray_dist_family, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# order
(mantel_rc_order <- vegan::mantel(bray_dist_order, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_order  <- vegan::mantel.partial(bray_dist_order, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# class
(mantel_rc_class <- vegan::mantel(bray_dist_class, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_class  <- vegan::mantel.partial(bray_dist_class, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# phylum
(mantel_rc_phylum <- vegan::mantel(bray_dist_phylum, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_bray_geo_phylum  <- vegan::mantel.partial(bray_dist_phylum, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))


# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_rc_otu$statistic, p=mantel_rc_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)

# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_rc_species$statistic, p=mantel_rc_species$signif)

df_species_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_geo_species$statistic, p=mantel_bray_geo_species$signif)

# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_rc_genus$statistic, p=mantel_rc_genus$signif)

df_genus_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_geo_genus$statistic, p=mantel_bray_geo_genus$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_rc_family$statistic, p=mantel_rc_family$signif)

df_family_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_geo_family$statistic, p=mantel_bray_geo_family$signif)


# order
df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_rc_order$statistic, p=mantel_rc_order$signif)

df_order_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_geo_order$statistic, p=mantel_bray_geo_order$signif)

# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_rc_class$statistic, p=mantel_rc_class$signif)

df_class_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_geo_class$statistic, p=mantel_bray_geo_class$signif)

# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_rc_phylum$statistic, p=mantel_rc_phylum$signif)

df_phylum_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_geo_phylum$statistic, p=mantel_bray_geo_phylum$signif)

mantel_out_bray <- rbind(df_otu, df_otu_partial, df_species, df_species_partial, df_genus, df_genus_partial, df_family, df_family_partial,  df_order, df_order_partial,df_class, df_class_partial, df_phylum, df_phylum_partial)

```




Mantel Raup-Crick diet - Jaccard microbiome
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_rc_diet_otu <- vegan::mantel(jaccard_dist_otu, diet_rc_dist, method = "pearson", permutations = 9999))
str(mantel_rc_diet_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_rc_diet_otu$statistic, p=mantel_rc_diet_otu$signif)

(mantel_jaccard_geo_otu  <- vegan::mantel.partial(jaccard_dist_otu, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))
df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

jacc = as.vector(jaccard_dist_otu)
diet = as.vector(diet_rc_dist)
geo = as.vector(geo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(jacc, diet, geo)

(p1 <- ggplot(mat, aes(y = jacc, x = diet)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Diet distance (Raup-Crick)", y = "Beta diversity (Jaccard distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_diet_jaccard$statistic, 3), ";p-value=", round(mantel_diet_jaccard$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

(p2 <- ggplot(mat, aes(y = diet, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "Diet distance (Raup-Crick)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))


# species
(mantel_rc_diet_species <- vegan::mantel(jaccard_dist_species, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_species  <- vegan::mantel.partial(jaccard_dist_species, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))


# genus
(mantel_rc_diet_genus <- vegan::mantel(jaccard_dist_genus, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_genus  <- vegan::mantel.partial(jaccard_dist_genus, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))


# family
(mantel_rc_diet_family <- vegan::mantel(jaccard_dist_family, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_family  <- vegan::mantel.partial(jaccard_dist_family, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# class
(mantel_rc_diet_class <- vegan::mantel(jaccard_dist_class, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_class  <- vegan::mantel.partial(jaccard_dist_class, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# order
(mantel_rc_diet_order <- vegan::mantel(jaccard_dist_order, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_order  <- vegan::mantel.partial(jaccard_dist_order, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# phylum
(mantel_rc_diet_phylum <- vegan::mantel(jaccard_dist_phylum, diet_rc_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_phylum  <- vegan::mantel.partial(jaccard_dist_phylum, diet_rc_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_rc_diet_otu$statistic, p=mantel_rc_diet_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_rc_diet_species$statistic, p=mantel_rc_diet_species$signif)

df_species_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_geo_species$statistic, p=mantel_jaccard_geo_species$signif)

# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_rc_diet_genus$statistic, p=mantel_rc_diet_genus$signif)

df_genus_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_geo_genus$statistic, p=mantel_jaccard_geo_genus$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_rc_diet_family$statistic, p=mantel_rc_diet_family$signif)

df_family_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_geo_family$statistic, p=mantel_jaccard_geo_family$signif)

# order

df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_rc_diet_order$statistic, p=mantel_rc_diet_order$signif)

df_order_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_geo_order$statistic, p=mantel_jaccard_geo_order$signif)

# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_rc_diet_class$statistic, p=mantel_rc_diet_class$signif)

df_class_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_geo_class$statistic, p=mantel_jaccard_geo_class$signif)

# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_rc_diet_phylum$statistic, p=mantel_jaccard_diet_phylum$signif)

df_phylum_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_geo_phylum$statistic, p=mantel_jaccard_geo_phylum$signif)

mantel_out_jaccard <- rbind(df_otu, df_otu_partial, df_species,  df_species_partial, df_genus,  df_genus_partial, df_family, df_family_partial,  df_order, df_order_partial, df_class, df_class_partial, df_phylum, df_phylum_partial)



```



Join bray and jaccard output

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

mantel_out_diet <- rbind(mantel_out_bray, mantel_out_jaccard)
head(mantel_out_diet)

write.csv(mantel_out_diet, paste0(dataDir, "/output_mantel_rc_diet.csv"), row.names=FALSE)

```







Mantel Bray-Curtis geographical distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

lgeo_dist <- as.dist(log10(geo_dist+1))
(mantel_bray_geo_otu <- vegan::mantel(bray_dist_otu, lgeo_dist, method = "pearson", permutations = 9999))

str(mantel_bray_geo_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)


bray = as.vector(bray_dist_otu)
jacc = as.vector(jaccard_dist_otu)
geo = as.vector(lgeo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(bray, jacc, geo)
#mat_edit <- read.csv(paste0(dataDir, "/geo_dist_d


(p1 <- ggplot(mat, aes(y = bray, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographical distance", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_phylog_bray$statistic, 3), ";p-value=", round(mantel_phylog_bray$signif, 3)), parse=F,x=0.4,y=0.4,size=10))


# species
(mantel_bray_geo_species <- vegan::mantel(bray_dist_species, lgeo_dist, method = "pearson", permutations = 9999))


# genus
(mantel_bray_geo_genus <- vegan::mantel(bray_dist_genus, lgeo_dist, method = "pearson", permutations = 9999))


# family
(mantel_bray_geo_family <- vegan::mantel(bray_dist_family, lgeo_dist, method = "pearson", permutations = 9999))

# order
(mantel_bray_geo_order <- vegan::mantel(bray_dist_order, lgeo_dist, method = "pearson", permutations = 9999))

# class
(mantel_bray_geo_class <- vegan::mantel(bray_dist_class, lgeo_dist, method = "pearson", permutations = 9999))

# phylum
(mantel_bray_geo_phylum <- vegan::mantel(bray_dist_phylum, lgeo_dist, method = "pearson", permutations = 9999))


# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)

# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_geo_species$statistic, p=mantel_bray_geo_species$signif)

# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_geo_genus$statistic, p=mantel_bray_geo_genus$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_geo_family$statistic, p=mantel_bray_geo_family$signif)


# order
df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_geo_order$statistic, p=mantel_bray_geo_order$signif)

# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_geo_class$statistic, p=mantel_bray_geo_class$signif)

# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_geo_phylum$statistic, p=mantel_bray_geo_phylum$signif)

mantel_out_bray <- rbind(df_otu, df_species, df_genus, df_family,  df_order, df_class, df_phylum)

```


Mantel Jaccard
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_jaccard_geo_otu <- vegan::mantel(jaccard_dist_otu, lgeo_dist, method = "pearson", permutations = 9999))
str(mantel_jaccard_geo_otu)

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

jacc = as.vector(jaccard_dist_otu)
geo = as.vector(lgeo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(jacc, geo)

(p1 <- ggplot(mat, aes(y = jacc, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "geo distance", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_geo_jaccard$statistic, 3), ";p-value=", round(mantel_geo_jaccard$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

# species
(mantel_jaccard_geo_species <- vegan::mantel(jaccard_dist_species, lgeo_dist, method = "pearson", permutations = 9999))

# genus
(mantel_jaccard_geo_genus <- vegan::mantel(jaccard_dist_genus, lgeo_dist, method = "pearson", permutations = 9999))


# family
(mantel_jaccard_geo_family <- vegan::mantel(jaccard_dist_family, lgeo_dist, method = "pearson", permutations = 9999))

# class
(mantel_jaccard_geo_class <- vegan::mantel(jaccard_dist_class, lgeo_dist, method = "pearson", permutations = 9999))

# order
(mantel_jaccard_geo_order <- vegan::mantel(jaccard_dist_order, lgeo_dist, method = "pearson", permutations = 9999))

# phylum
(mantel_jaccard_geo_phylum <- vegan::mantel(jaccard_dist_phylum, lgeo_dist, method = "pearson", permutations = 9999))

# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_geo_species$statistic, p=mantel_jaccard_geo_species$signif)


# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_geo_genus$statistic, p=mantel_jaccard_geo_genus$signif)


# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_geo_family$statistic, p=mantel_jaccard_geo_family$signif)

# order

df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_geo_order$statistic, p=mantel_jaccard_geo_order$signif)

# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_geo_class$statistic, p=mantel_jaccard_geo_class$signif)

# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_geo_phylum$statistic, p=mantel_jaccard_geo_phylum$signif)


mantel_out_jaccard <- rbind(df_otu, df_species, df_genus, df_family,  df_order, df_class, df_phylum)



```

Join bray and jaccard output

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

mantel_out_geo <- rbind(mantel_out_bray, mantel_out_jaccard)
mantel_out_geo

write.csv(mantel_out_geo, paste0(dataDir, "/output_mantel_geo.csv"), row.names=FALSE)

```

## MRM test 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

#OTU
#test bray
(out_bray_ind <- MRM(bray_dist_otu ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
BP = out_bray_ind$r.squared[1]

out_host_ind <- MRM(bray_dist_otu ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(bray_dist_otu ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(bray_dist_otu ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(bray_dist_otu ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(bray_dist_otu ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(bray_dist_otu ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 

(var_bray_ind_otu <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = BP, A.name = "host", B.name = "diet", C.name = "geo"))
var_bray_ind_otu$pop_ind <- paste("Individual")
var_bray_ind_otu$distance <- paste("Bray-Curtis")
var_bray_ind_otu$tax <- paste("OTU99")
var_bray_ind_otu$variable <- rownames(var_bray_ind_otu)
var_bray_ind_otu 

#test jaccard
(out_jaccard_ind <- MRM(jaccard_dist_otu ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
JP = out_jaccard_ind$r.squared[1]

out_host_ind <- MRM(jaccard_dist_otu ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(jaccard_dist_otu ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(jaccard_dist_otu ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(jaccard_dist_otu ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(jaccard_dist_otu ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(jaccard_dist_otu ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_jaccard_ind_otu <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = JP, A.name = "host", B.name = "diet", C.name = "geo"))
var_jaccard_ind_otu$pop_ind <- paste("Individual")
var_jaccard_ind_otu$distance <- paste("Jaccard")
var_jaccard_ind_otu$tax <- paste("OTU99")
var_jaccard_ind_otu$variable <- rownames(var_jaccard_ind_otu)
var_jaccard_ind_otu

#species
#test bray
(out_bray_ind <- MRM(bray_dist_species ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
BP = out_bray_ind$r.squared[1]

out_host_ind <- MRM(bray_dist_species ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(bray_dist_species ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(bray_dist_species ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(bray_dist_species ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(bray_dist_species ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(bray_dist_species ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 

(var_bray_ind_species <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = BP, A.name = "host", B.name = "diet", C.name = "geo"))
var_bray_ind_species$pop_ind <- paste("Individual")
var_bray_ind_species$distance <- paste("Bray-Curtis")
var_bray_ind_species$tax <- paste("Species")
var_bray_ind_species$variable <- rownames(var_bray_ind_species)
var_bray_ind_species 

#test jaccard
(out_jaccard_ind <- MRM(jaccard_dist_species ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
JP = out_jaccard_ind$r.squared[1]

out_host_ind <- MRM(jaccard_dist_species ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(jaccard_dist_species ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(jaccard_dist_species ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(jaccard_dist_species ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(jaccard_dist_species ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(jaccard_dist_species ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_jaccard_ind_species <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = JP, A.name = "host", B.name = "diet", C.name = "geo"))
var_jaccard_ind_species$pop_ind <- paste("Individual")
var_jaccard_ind_species$distance <- paste("Jaccard")
var_jaccard_ind_species$tax <- paste("Species")
var_jaccard_ind_species$variable <- rownames(var_jaccard_ind_species)

var_jaccard_ind_species




#Genus
#test bray
(out_bray_ind <- MRM(bray_dist_genus ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
BP = out_bray_ind$r.squared[1]

out_host_ind <- MRM(bray_dist_genus ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(bray_dist_genus ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(bray_dist_genus ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(bray_dist_genus ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(bray_dist_genus ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(bray_dist_genus ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 

(var_bray_ind_genus <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = BP, A.name = "host", B.name = "diet", C.name = "geo"))
var_bray_ind_genus$pop_ind <- paste("Individual")
var_bray_ind_genus$distance <- paste("Bray-Curtis")
var_bray_ind_genus$tax <- paste("Genus")
var_bray_ind_genus$variable <- rownames(var_bray_ind_genus)
var_bray_ind_genus 

#test jaccard
(out_jaccard_ind <- MRM(jaccard_dist_genus ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
JP = out_jaccard_ind$r.squared[1]

out_host_ind <- MRM(jaccard_dist_genus ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(jaccard_dist_genus ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(jaccard_dist_genus ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(jaccard_dist_genus ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(jaccard_dist_genus ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(jaccard_dist_genus ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_jaccard_ind_genus <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = JP, A.name = "host", B.name = "diet", C.name = "geo"))
var_jaccard_ind_genus$pop_ind <- paste("Individual")
var_jaccard_ind_genus$distance <- paste("Jaccard")
var_jaccard_ind_genus$tax <- paste("Genus")
var_jaccard_ind_genus$variable <- rownames(var_jaccard_ind_genus)

var_jaccard_ind_genus


#Family
#test bray
(out_bray_ind <- MRM(bray_dist_family ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
BP = out_bray_ind$r.squared[1]

out_host_ind <- MRM(bray_dist_family ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(bray_dist_family ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(bray_dist_family ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(bray_dist_family ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(bray_dist_family ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(bray_dist_family ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_bray_ind_family <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = BP, A.name = "host", B.name = "diet", C.name = "geo"))
var_bray_ind_family$pop_ind <- paste("Individual")
var_bray_ind_family$distance <- paste("Bray-Curtis")
var_bray_ind_family$tax <- paste("Family")
var_bray_ind_family$variable <- rownames(var_bray_ind_family)
var_bray_ind_family 

#test jaccard
(out_jaccard_ind <- MRM(jaccard_dist_family ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
JP = out_jaccard_ind$r.squared[1]

out_host_ind <- MRM(jaccard_dist_family ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(jaccard_dist_family ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(jaccard_dist_family ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(jaccard_dist_family ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(jaccard_dist_family ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(jaccard_dist_family ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_jaccard_ind_family <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = JP, A.name = "host", B.name = "diet", C.name = "geo"))
var_jaccard_ind_family$pop_ind <- paste("Individual")
var_jaccard_ind_family$distance <- paste("Jaccard")
var_jaccard_ind_family$tax <- paste("Family")
var_jaccard_ind_family$variable <- rownames(var_jaccard_ind_family)
var_jaccard_ind_family


# Order
#test bray
(out_bray_ind <- MRM(bray_dist_order ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
BP = out_bray_ind$r.squared[1]

out_host_ind <- MRM(bray_dist_order ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(bray_dist_order ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(bray_dist_order ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(bray_dist_order ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(bray_dist_order ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(bray_dist_order ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_bray_ind_order <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = BP, A.name = "host", B.name = "diet", C.name = "geo"))
var_bray_ind_order$pop_ind <- paste("Individual")
var_bray_ind_order$distance <- paste("Bray-Curtis")
var_bray_ind_order$tax <- paste("Order")
var_bray_ind_order$variable <- rownames(var_bray_ind_order)
var_bray_ind_order 

#test jaccard
(out_jaccard_ind <- MRM(jaccard_dist_order ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
JP = out_jaccard_ind$r.squared[1]

out_host_ind <- MRM(jaccard_dist_order ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(jaccard_dist_order ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(jaccard_dist_order ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(jaccard_dist_order ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(jaccard_dist_order ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(jaccard_dist_order ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_jaccard_ind_order <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = JP, A.name = "host", B.name = "diet", C.name = "geo"))
var_jaccard_ind_order$pop_ind <- paste("Individual")
var_jaccard_ind_order$distance <- paste("Jaccard")
var_jaccard_ind_order$tax <- paste("Order")
var_jaccard_ind_order$variable <- rownames(var_jaccard_ind_order)
var_jaccard_ind_order


#Class
#test bray
(out_bray_ind <- MRM(bray_dist_class ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
BP = out_bray_ind$r.squared[1]

out_host_ind <- MRM(bray_dist_class ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(bray_dist_class ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(bray_dist_class ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(bray_dist_class ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(bray_dist_class ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(bray_dist_class ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_bray_ind_class <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = BP, A.name = "host", B.name = "diet", C.name = "geo"))
var_bray_ind_class$pop_ind <- paste("Individual")
var_bray_ind_class$distance <- paste("Bray-Curtis")
var_bray_ind_class$tax <- paste("Class")
var_bray_ind_class$variable <- rownames(var_bray_ind_class)
var_bray_ind_class 

#test jaccard
(out_jaccard_ind <- MRM(jaccard_dist_class ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
JP = out_jaccard_ind$r.squared[1]

out_host_ind <- MRM(jaccard_dist_class ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(jaccard_dist_class ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(jaccard_dist_class ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(jaccard_dist_class ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(jaccard_dist_class ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(jaccard_dist_class ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 
(var_jaccard_ind_class <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = JP, A.name = "host", B.name = "diet", C.name = "geo"))
var_jaccard_ind_class$pop_ind <- paste("Individual")
var_jaccard_ind_class$distance <- paste("Jaccard")
var_jaccard_ind_class$tax <- paste("Class")
var_jaccard_ind_class$variable <- rownames(var_jaccard_ind_class)
var_jaccard_ind_class


#Phylum
#test bray
(out_bray_ind <- MRM(bray_dist_phylum ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
BP = out_bray_ind$r.squared[1]

out_host_ind <- MRM(bray_dist_phylum ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(bray_dist_phylum ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(bray_dist_phylum ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(bray_dist_phylum ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(bray_dist_phylum ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(bray_dist_phylum ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]


(var_bray_ind_phylum <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = BP, A.name = "host", B.name = "diet", C.name = "geo"))
var_bray_ind_phylum$pop_ind <- paste("Individual")
var_bray_ind_phylum$distance <- paste("Bray-Curtis")
var_bray_ind_phylum$tax <- paste("Phylum")
var_bray_ind_phylum$variable <- rownames(var_bray_ind_phylum)
var_bray_ind_phylum 

#test jaccard
(out_jaccard_ind <- MRM(jaccard_dist_phylum ~ host_dist+  diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999))
JP = out_jaccard_ind$r.squared[1]

out_host_ind <- MRM(jaccard_dist_phylum ~ host_dist,  nperm=9999)
H = out_host_ind$r.squared[1]

out_geo_ind <- MRM(jaccard_dist_phylum ~ as.dist(log10(geo_dist+1)),  nperm=9999)
G = out_geo_ind$r.squared[1]

out_diet_ind <- MRM(jaccard_dist_phylum ~ diet_rc_dist,  nperm=9999)
D = out_diet_ind$r.squared[1]

out_host_diet <- MRM(jaccard_dist_phylum ~ host_dist + diet_rc_dist,  nperm=9999)
HD = out_host_diet$r.squared[1]

out_host_geo <- MRM(jaccard_dist_phylum ~ host_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
HG = out_host_geo$r.squared[1]

out_diet_geo <- MRM(jaccard_dist_phylum ~ diet_rc_dist + as.dist(log10(geo_dist+1)),  nperm=9999)
DG = out_diet_geo$r.squared[1]

# relative contribution 

#https://modtools.wordpress.com/2012/11/06/varpart/
#The outputs are the amounts of variance explained exclusively by each factor, the amounts explained exclusively by the overlapping effects of each pair of factors, and the amount explained by the overlap of the 3 factors if this is the case (e.g. Real et al. 2003)

(var_jaccard_ind_phylum <- varPart(A = H, B = D, C = G, AB= HD, AC = HG, BC = DG, ABC = JP, A.name = "host", B.name = "diet", C.name = "geo"))
var_jaccard_ind_phylum$pop_ind <- paste("Individual")
var_jaccard_ind_phylum$distance <- paste("Jaccard")
var_jaccard_ind_phylum$tax <- paste("Phylum")
var_jaccard_ind_phylum$variable <- rownames(var_jaccard_ind_phylum)
var_jaccard_ind_phylum

```

merge output

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

out_varpart_ind <- rbind(var_bray_ind_otu,var_bray_ind_species, var_bray_ind_genus, var_bray_ind_family, var_bray_ind_order, var_bray_ind_class, var_bray_ind_phylum, var_jaccard_ind_otu,var_jaccard_ind_species,var_jaccard_ind_genus, var_jaccard_ind_family, var_jaccard_ind_order, var_jaccard_ind_class, var_jaccard_ind_phylum, make.row.names = FALSE)
out_varpart_ind$Proportion <- ifelse(out_varpart_ind$Proportion < 0, 0, out_varpart_ind$Proportion) 
out_varpart_ind$percent <- round(out_varpart_ind$Proportion * 100,  2)
head(out_varpart_ind)
str(out_varpart_ind)
out_varpart_ind$tax <- factor(out_varpart_ind$tax, levels = c("OTU99", "Species" ,"Genus", "Family", "Order", "Class", "Phylum"))
table(out_varpart_ind$pop_ind)
var_bray_ind_otu;var_jaccard_ind_otu
var_bray_ind_genus;var_jaccard_ind_genus
var_bray_ind_family;var_jaccard_ind_family

check <- out_varpart_ind %>% filter(variable == "diet")

write.csv(out_varpart_ind, paste0(dataDir, "/output_varpart_nuclearTree.csv"), row.names=FALSE)

```







