---
title: 'PACo Test'
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: html_document
---

## LOAD LIBRARIES

```{r loadLibraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=(width.cutoff=45), paged.print=FALSE}


rm(list = ls())

#general
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(vegan)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(MicEco)
library(rlist) 
library(grid)
library(tsiR) 
library(ape)
library(treedataverse)
library(ggh4x)

# phyloseq section
library(phyloseq)
library(Biostrings)
library(DECIPHER)
library(speedyseq)
library(microbiome)
library(metagMisc)
library(phylosmith)
library(seqateurs)
library(treedata.table)
library(microViz)

#rmd
library(details)
library(knitr)
library(kableExtra)

# PACo 
library(ape)
library(paco)
library(future)
library(doParallel)
library(future)
library(future.apply)
library(multcomp)

```

## WORKING DIRECTORIES

```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```

## PREPARE DATA

Microbiome 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr_fix <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr <- speedyseq::tax_glom(microb_fr_fix, "OTU_ID",  NArm = FALSE)
#microb_fr <- speedyseq::tax_glom(microb_fr_fix, "Species",  NArm = FALSE)
#microb_fr <- speedyseq::tax_glom(microb_fr_fix, "Genus",  NArm = FALSE)
#microb_fr <- speedyseq::tax_glom(microb_fr_fix, "Family",  NArm = FALSE)


# Add TipID microbiome to match stuff later on
sample_data(microb_fr)$TipID_microbiome <- paste(sample_data(microb_fr)$Lab_Code, sample_data(microb_fr)$Species_Label, sep = "_")

microb_fr_meta <- microb_fr%>% sample_data %>% data.frame 
nrow(microb_fr_meta)


```

# Overlap WGS data

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_overlap <- microb_fr %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap <- biomeUtils::removeZeros(microb_overlap)
microb_overlap_meta <- microb_overlap%>% sample_data %>% data.frame 

```

Host distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# mitogenome tree
fringilla_tree <-  read.tree(file = paste0(dataDir, "/mt_Fringilla_aln.fasta.contree"))

str(fringilla_tree)

fringilla_tree$tip.label <- gsub("\\_mtDNA","", fringilla_tree$tip.label)

annot <- read_delim(paste0(dataDir, "/WGS_Fringilla_samples.tsv"))


table(annot$Microb)

#rename row and col names to match microbiome and diet matrices 
annot$TipID_microbiome

fringilla_tree$tip.label <- annot[match(fringilla_tree$tip.label, annot$WGS_ID),]$TipID_microbiome

# subset tree matching IDs with microbiome data 
fringilla_tree_subset <- as.treedata.table(tree = fringilla_tree, data = microb_overlap_meta, name_column = "TipID_microbiome")

fringilla_tree_subset <- fringilla_tree_subset$phy
str(fringilla_tree_subset)

fringilla_rootedtree = root(fringilla_tree_subset, outgroup = "149_Tenerife_F_teydea", resolve.root = TRUE)

meta_subset <- annot %>% dplyr::filter(Microb == 1) %>% data.frame
meta_subset$label <- meta_subset$TipID_microbiome 
meta_subset %>% names
meta_subset %>% head
class(meta_subset$label)

tree_meta_microb <- full_join(fringilla_rootedtree, meta_subset, by = "label")

(t1 <- ggplot(tree_meta_microb, aes(x, y)) + geom_tree() + theme_tree() + geom_tiplab(aes(color=Species_Label, fill=Species_Label), size = 3) + SpeciesLabel.color + SpeciesLabel.fill + theme(legend.position="bottom", legend.direction = "horizontal", legend.title=element_blank()))

host_tree <- list(fringilla_rootedtree)
summary(host_tree)

# Host distance matrix
host_dist <- fringilla_rootedtree %>% cophenetic 

host_dist %>% dim
str(host_dist)
#host_dist <- as.dist(host_dist)

```


random host tree

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

random_host_dist <- host_dist

sample_host <- sample(colnames(random_host_dist))
colnames(random_host_dist) <- sample_host
rownames(random_host_dist) <- sample_host

```

#subset based on neutral expectations (neutral vs non-neutral)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


neutral <- readRDS(paste0(dataDir, "/ps_neutral_otuFix.rds" ))

#otu_table()   OTU Table:          [ 10459 taxa and 209 samples ]:
#sample_data() Sample Data:        [ 209 samples by 50 sample variables ]:

#neutral <- readRDS(paste0(dataDir, "/ps_overlap_neutral_speciesFix.rds" ))
#otu_table()   OTU Table:          [ 1883 taxa and 209 samples ]:
#sample_data() Sample Data:        [ 209 samples by 50 sample variables ]:

nonneutral <- readRDS(paste0(dataDir, "/ps_nonneutral_otu.rds" ))
#otu_table()   OTU Table:          [ 1509 taxa and 209 samples ]:
#sample_data() Sample Data:        [ 209 samples by 50 sample variables ]:

nonneutral <- readRDS(paste0(dataDir, "/ps_overlap_nonneutral_speciesFix.rds" ))
#otu_table()   OTU Table:          [ 558 taxa and 209 samples ]:
#sample_data() Sample Data:        [ 209 samples by 50 sample variables ]:


tax_table(neutral)
neutral_otus <- neutral %>% tax_table %>% data.frame
keep_neutral <- rownames(neutral_otus)
keep_neutral_otu <- neutral_otus %>% dplyr::select(OTU_ID)
keep_neutral_otu$prediction <- paste("neutral") 
ps_neutral_otu <- prune_taxa(keep_neutral, microb_overlap)
neutral_meta <- ps_neutral_otu %>% sample_data %>% data.frame 

nonneutral_otus <- nonneutral %>% tax_table %>% data.frame
keep_nonneutral <- rownames(nonneutral_otus)
keep_nonneutral_otu <- nonneutral_otus %>% dplyr::select(OTU_ID)
keep_nonneutral_otu$prediction <- paste("non-neutral") 
ps_nonneutral_otu <- prune_taxa(keep_nonneutral, microb_overlap)
nonneutral_meta <- ps_nonneutral_otu %>% sample_data %>% data.frame 

df_otu <- rbind(keep_neutral_otu, keep_nonneutral_otu) 
str(df_otu)

```


Distance matrix

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# overlap
microb_dist <- microb_overlap %>% phy_tree %>% cophenetic %>%  as.matrix
str(microb_dist)

# neutral
microb_dist_neutral <- ps_neutral_otu %>% phy_tree %>% cophenetic %>%  as.matrix

# nonneutral
microb_dist_nonneutral <- ps_nonneutral_otu %>% phy_tree %>% cophenetic %>%  as.matrix

```

# PACO test core across overlap otus

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- microb_overlap %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix

setdiff(rownames(otu), rownames(host_dist))
setdiff(rownames(core_across_meta), rownames(otu))
setdiff(rownames(core_across_meta), rownames(host_dist))

paco_data <- prepare_paco_data(H = host_dist, P = microb_dist, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names

```

PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

overlap_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
overlap_test$gof

# Proc
proc_obj_across <- overlap_test$proc
str(proc_obj_across)
sqrt(1-proc_obj_across$ss)

# goodness of fit

H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo

y_overlap <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)

#Procrustes Sum of Squares (m12 squared):        0.9974 
#Correlation in a symmetric Procrustes rotation: 0.05114 
#Significance:  0.001 



overlap_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
overlap_random$gof

# Proc
proc_obj_overlap_random <- overlap_random$proc
str(proc_obj_overlap_random)
sqrt(1-proc_obj_overlap_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_overlap_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9986 
#Correlation in a symmetric Procrustes rotation: 0.0376 
#Significance:  0.001 

```

# PACO test neutral otus

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- ps_neutral_otu %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix


setdiff(rownames(otu), rownames(host_dist))
setdiff(rownames(microb_overlap_meta), rownames(otu))
setdiff(rownames(microb_overlap_meta), rownames(host_dist))

paco_data <- prepare_paco_data(H = host_dist, P = microb_dist_neutral, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist_neutral, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names

```


PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# r0 method => the assumption that the column group tracks the evolution of the row group; therefore, maintaining the degree of the species in the row group

neutral_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
neutral_test$gof

# Proc
proc_obj_neutral <- neutral_test$proc
str(proc_obj_neutral)
sqrt(1-proc_obj_neutral$ss)


# goodness of fit
H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo
++++++++++++++++
y_neutral <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.985 
#Correlation in a symmetric Procrustes rotation: 0.1223 
#Significance:  0.001 


#

neutral_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
neutral_random$gof

# Proc
proc_obj_neutral_random <- neutral_random$proc
str(proc_obj_neutral_random)
sqrt(1-proc_obj_neutral_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_neutral_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9878 
#Correlation in a symmetric Procrustes rotation: 0.1106 
#Significance:  0.037 



```


# PACO test non-neutral otus

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- ps_nonneutral_genus %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix


paco_data <- prepare_paco_data(H = host_dist, P = microb_dist_nonneutral, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist_nonneutral, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names


```

PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# r0 method => the assumption that the column group tracks the evolution of the row group; therefore, maintaining the degree of the species in the row group

nonneutral_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
nonneutral_test$gof

# Proc
proc_obj_nonneutral <- nonneutral_test$proc
str(proc_obj_nonneutral)
sqrt(1-proc_obj_nonneutral$ss)

# goodness of fit
H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo

y_nonneutral <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9952 
#Correlation in a symmetric Procrustes rotation: 0.06906 
#Significance:  0.001 


nonneutral_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
nonneutral_random$gof

# Proc
proc_obj_nonneutral_random <- nonneutral_random$proc
str(proc_obj_nonneutral_random)
sqrt(1-proc_obj_nonneutral_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_nonneutral_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)

#Procrustes Sum of Squares (m12 squared):        0.997 
#Correlation in a symmetric Procrustes rotation: 0.05513 
#Significance:  1 


```


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

procr_p_host <- c(y_overlap$signif, y_neutral$signif, y_nonneutral$signif) 
p_adj_bonferroni <- round(p.adjust(as.numeric(procr_p),method = "bonferroni"),3)

procr_p_host <- as.data.frame(cbind(procr_p_host, p_adj_bonferroni))
rownames(procr_p_host) <- c("all", "neutral", "non-neutral")

procr_p_random <- c(y_overlap_random$signif, y_neutral_random$signif, y_nonneutral_random$signif) 
p_adj_bonferroni <- round(p.adjust(as.numeric(procr_p_random),method = "bonferroni"),3)

procr_p_random <- as.data.frame(cbind(procr_p_random, p_adj_bonferroni))
rownames(procr_p_random) <- c("all", "neutral", "non-neutral")


```

test on species level 


Microbiome 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr_fix <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr <- speedyseq::tax_glom(microb_fr_fix, "Species",  NArm = FALSE)

# Add TipID microbiome to match stuff later on
sample_data(microb_fr)$TipID_microbiome <- paste(sample_data(microb_fr)$Lab_Code, sample_data(microb_fr)$Species_Label, sep = "_")

microb_fr_meta <- microb_fr%>% sample_data %>% data.frame 
nrow(microb_fr_meta)


```

# Overlap WGS data

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_overlap <- microb_fr %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap <- biomeUtils::removeZeros(microb_overlap)
microb_overlap_meta <- microb_overlap%>% sample_data %>% data.frame 

```

#subset based on neutral expectations (neutral vs non-neutral)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


neutral <- readRDS(paste0(dataDir, "/ps_overlap_neutral_speciesFix.rds" ))
#otu_table()   OTU Table:          [ 1883 taxa and 209 samples ]:
#sample_data() Sample Data:        [ 209 samples by 50 sample variables ]:

nonneutral <- readRDS(paste0(dataDir, "/ps_overlap_nonneutral_speciesFix.rds" ))
#otu_table()   OTU Table:          [ 558 taxa and 209 samples ]:
#sample_data() Sample Data:        [ 209 samples by 50 sample variables ]:


tax_table(neutral)
neutral_species <- neutral %>% tax_table %>% data.frame
keep_neutral <- rownames(neutral_species)
keep_neutral_species <- neutral_species %>% dplyr::select(Species)
keep_neutral_species$prediction <- paste("neutral") 
ps_neutral_species <- prune_taxa(keep_neutral, microb_overlap)
neutral_meta <- ps_neutral_species %>% sample_data %>% data.frame 

nonneutral_species <- nonneutral %>% tax_table %>% data.frame
keep_nonneutral <- rownames(nonneutral_species)
keep_nonneutral_species <- nonneutral_species %>% dplyr::select(Species)
keep_nonneutral_species$prediction <- paste("non-neutral") 
ps_nonneutral_species <- prune_taxa(keep_nonneutral, microb_overlap)
nonneutral_meta <- ps_nonneutral_species %>% sample_data %>% data.frame 

df_species <- rbind(keep_neutral_species, keep_nonneutral_species) 
str(df_species)

```


Distance matrix

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# overlap
microb_dist <- microb_overlap %>% phy_tree %>% cophenetic %>%  as.matrix
str(microb_dist)

# neutral
microb_dist_neutral <- ps_neutral_species %>% phy_tree %>% cophenetic %>%  as.matrix

# nonneutral
microb_dist_nonneutral <- ps_nonneutral_species %>% phy_tree %>% cophenetic %>%  as.matrix

```

# PACO test core across overlap species

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- microb_overlap %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix

setdiff(rownames(otu), rownames(host_dist))
setdiff(rownames(core_across_meta), rownames(otu))
setdiff(rownames(core_across_meta), rownames(host_dist))

paco_data <- prepare_paco_data(H = host_dist, P = microb_dist, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names

```

PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

overlap_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
overlap_test$gof

# Proc
proc_obj_across <- overlap_test$proc
str(proc_obj_across)
sqrt(1-proc_obj_across$ss)

# goodness of fit

H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo

y_overlap <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
++++++++++++++++++
#Procrustes Sum of Squares (m12 squared):        0.9974 
#Correlation in a symmetric Procrustes rotation: 0.05114 
#Significance:  0.001 



overlap_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
overlap_random$gof

# Proc
proc_obj_overlap_random <- overlap_random$proc
str(proc_obj_overlap_random)
sqrt(1-proc_obj_overlap_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_overlap_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9986 
#Correlation in a symmetric Procrustes rotation: 0.0376 
#Significance:  0.001 

```

# PACO test neutral otus

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- ps_neutral_otu %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix


setdiff(rownames(otu), rownames(host_dist))
setdiff(rownames(microb_overlap_meta), rownames(otu))
setdiff(rownames(microb_overlap_meta), rownames(host_dist))

paco_data <- prepare_paco_data(H = host_dist, P = microb_dist_neutral, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist_neutral, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names

```


PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# r0 method => the assumption that the column group tracks the evolution of the row group; therefore, maintaining the degree of the species in the row group

neutral_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
neutral_test$gof

# Proc
proc_obj_neutral <- neutral_test$proc
str(proc_obj_neutral)
sqrt(1-proc_obj_neutral$ss)


# goodness of fit
H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo

y_neutral <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.985 
#Correlation in a symmetric Procrustes rotation: 0.1223 
#Significance:  0.001 


#

neutral_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
neutral_random$gof

# Proc
proc_obj_neutral_random <- neutral_random$proc
str(proc_obj_neutral_random)
sqrt(1-proc_obj_neutral_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_neutral_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9878 
#Correlation in a symmetric Procrustes rotation: 0.1106 
#Significance:  0.037 



```


# PACO test non-neutral otus

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- ps_nonneutral_genus %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix


paco_data <- prepare_paco_data(H = host_dist, P = microb_dist_nonneutral, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist_nonneutral, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names


```

PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# r0 method => the assumption that the column group tracks the evolution of the row group; therefore, maintaining the degree of the species in the row group

nonneutral_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
nonneutral_test$gof

# Proc
proc_obj_nonneutral <- nonneutral_test$proc
str(proc_obj_nonneutral)
sqrt(1-proc_obj_nonneutral$ss)

# goodness of fit
H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo

y_nonneutral <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9952 
#Correlation in a symmetric Procrustes rotation: 0.06906 
#Significance:  0.001 


nonneutral_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
nonneutral_random$gof

# Proc
proc_obj_nonneutral_random <- nonneutral_random$proc
str(proc_obj_nonneutral_random)
sqrt(1-proc_obj_nonneutral_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_nonneutral_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)

#Procrustes Sum of Squares (m12 squared):        0.997 
#Correlation in a symmetric Procrustes rotation: 0.05513 
#Significance:  1 


```


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

procr_p_host <- c(y_overlap$signif, y_neutral$signif, y_nonneutral$signif) 
p_adj_bonferroni <- round(p.adjust(as.numeric(procr_p),method = "bonferroni"),3)

procr_p_host <- as.data.frame(cbind(procr_p_host, p_adj_bonferroni))
rownames(procr_p_host) <- c("all", "neutral", "non-neutral")

procr_p_random <- c(y_overlap_random$signif, y_neutral_random$signif, y_nonneutral_random$signif) 
p_adj_bonferroni <- round(p.adjust(as.numeric(procr_p_random),method = "bonferroni"),3)

procr_p_random <- as.data.frame(cbind(procr_p_random, p_adj_bonferroni))
rownames(procr_p_random) <- c("all", "neutral", "non-neutral")


```


test on Genus level


Microbiome 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr_fix <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr <- speedyseq::tax_glom(microb_fr_fix, "Genus",  NArm = FALSE)

# Add TipID microbiome to match stuff later on
sample_data(microb_fr)$TipID_microbiome <- paste(sample_data(microb_fr)$Lab_Code, sample_data(microb_fr)$Species_Label, sep = "_")

microb_fr_meta <- microb_fr%>% sample_data %>% data.frame 
nrow(microb_fr_meta)


```

# Overlap WGS data

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_overlap <- microb_fr %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap <- biomeUtils::removeZeros(microb_overlap)
microb_overlap_meta <- microb_overlap%>% sample_data %>% data.frame 

```

#subset based on neutral expectations (neutral vs non-neutral)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


neutral <- readRDS(paste0(dataDir, "/ps_overlap_neutral_speciesFix.rds" ))
#otu_table()   OTU Table:          [ 1883 taxa and 209 samples ]:
#sample_data() Sample Data:        [ 209 samples by 50 sample variables ]:

nonneutral <- readRDS(paste0(dataDir, "/ps_overlap_nonneutral_speciesFix.rds" ))
#otu_table()   OTU Table:          [ 558 taxa and 209 samples ]:
#sample_data() Sample Data:        [ 209 samples by 50 sample variables ]:


tax_table(neutral)
neutral_species <- neutral %>% tax_table %>% data.frame
keep_neutral <- rownames(neutral_species)
keep_neutral_species <- neutral_species %>% dplyr::select(Species)
keep_neutral_species$prediction <- paste("neutral") 
ps_neutral_species <- prune_taxa(keep_neutral, microb_overlap)
neutral_meta <- ps_neutral_species %>% sample_data %>% data.frame 

nonneutral_species <- nonneutral %>% tax_table %>% data.frame
keep_nonneutral <- rownames(nonneutral_species)
keep_nonneutral_species <- nonneutral_species %>% dplyr::select(Species)
keep_nonneutral_species$prediction <- paste("non-neutral") 
ps_nonneutral_species <- prune_taxa(keep_nonneutral, microb_overlap)
nonneutral_meta <- ps_nonneutral_species %>% sample_data %>% data.frame 

df_species <- rbind(keep_neutral_species, keep_nonneutral_species) 
str(df_species)

```


Distance matrix

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# overlap
microb_dist <- microb_overlap %>% phy_tree %>% cophenetic %>%  as.matrix
str(microb_dist)

# neutral
microb_dist_neutral <- ps_neutral_species %>% phy_tree %>% cophenetic %>%  as.matrix

# nonneutral
microb_dist_nonneutral <- ps_nonneutral_species %>% phy_tree %>% cophenetic %>%  as.matrix

```

# PACO test core across overlap species

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- microb_overlap %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix

setdiff(rownames(otu), rownames(host_dist))
setdiff(rownames(core_across_meta), rownames(otu))
setdiff(rownames(core_across_meta), rownames(host_dist))

paco_data <- prepare_paco_data(H = host_dist, P = microb_dist, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names

```

PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

overlap_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
overlap_test$gof

# Proc
proc_obj_across <- overlap_test$proc
str(proc_obj_across)
sqrt(1-proc_obj_across$ss)

# goodness of fit

H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo

y_overlap <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
++++++++++++++++++
#Procrustes Sum of Squares (m12 squared):        0.9974 
#Correlation in a symmetric Procrustes rotation: 0.05114 
#Significance:  0.001 



overlap_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
overlap_random$gof

# Proc
proc_obj_overlap_random <- overlap_random$proc
str(proc_obj_overlap_random)
sqrt(1-proc_obj_overlap_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_overlap_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9986 
#Correlation in a symmetric Procrustes rotation: 0.0376 
#Significance:  0.001 

```

# PACO test neutral otus

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- ps_neutral_otu %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix


setdiff(rownames(otu), rownames(host_dist))
setdiff(rownames(microb_overlap_meta), rownames(otu))
setdiff(rownames(microb_overlap_meta), rownames(host_dist))

paco_data <- prepare_paco_data(H = host_dist, P = microb_dist_neutral, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist_neutral, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names

```


PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# r0 method => the assumption that the column group tracks the evolution of the row group; therefore, maintaining the degree of the species in the row group

neutral_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
neutral_test$gof

# Proc
proc_obj_neutral <- neutral_test$proc
str(proc_obj_neutral)
sqrt(1-proc_obj_neutral$ss)


# goodness of fit
H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo

y_neutral <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.985 
#Correlation in a symmetric Procrustes rotation: 0.1223 
#Significance:  0.001 


#

neutral_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
neutral_random$gof

# Proc
proc_obj_neutral_random <- neutral_random$proc
str(proc_obj_neutral_random)
sqrt(1-proc_obj_neutral_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_neutral_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9878 
#Correlation in a symmetric Procrustes rotation: 0.1106 
#Significance:  0.037 



```


# PACO test non-neutral otus

Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu <- ps_nonneutral_genus %>% otu_table %>% as.data.frame %>% as.matrix 

# convert to binary
otu <- otu %>% apply(2, function(x) ifelse(x > 0, 1, 0)) %>% as.matrix


paco_data <- prepare_paco_data(H = host_dist, P = microb_dist_nonneutral, HP = otu)
paco_data %>% str

paco_data = add_pcoord(paco_data, correction = "cailliez")
paco_data %>% names

paco_random <-  prepare_paco_data(H = random_host_dist, P = microb_dist_nonneutral, HP = otu)

paco_random = add_pcoord(paco_random, correction = "cailliez")
paco_random %>% names


```

PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# r0 method => the assumption that the column group tracks the evolution of the row group; therefore, maintaining the degree of the species in the row group

nonneutral_test <- PACo(paco_data, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
nonneutral_test$gof

# Proc
proc_obj_nonneutral <- nonneutral_test$proc
str(proc_obj_nonneutral)
sqrt(1-proc_obj_nonneutral$ss)

# goodness of fit
H_PCo <- paco_data$H_PCo
P_PCo <- paco_data$P_PCo

y_nonneutral <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)
#Procrustes Sum of Squares (m12 squared):        0.9952 
#Correlation in a symmetric Procrustes rotation: 0.06906 
#Significance:  0.001 


nonneutral_random <- PACo(paco_random, nperm = 10, seed = 05062024, method = 'quasiswap', symmetric = TRUE) 
nonneutral_random$gof

# Proc
proc_obj_nonneutral_random <- nonneutral_random$proc
str(proc_obj_nonneutral_random)
sqrt(1-proc_obj_nonneutral_random$ss)

# goodness of fit

H_PCo <- paco_random$H_PCo
P_PCo <- paco_random$P_PCo

y_nonneutral_random <- vegan::protest(X = H_PCo, Y = P_PCo, symmetric = TRUE)

#Procrustes Sum of Squares (m12 squared):        0.997 
#Correlation in a symmetric Procrustes rotation: 0.05513 
#Significance:  1 


```


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

procr_p_host <- c(y_overlap$signif, y_neutral$signif, y_nonneutral$signif) 
p_adj_bonferroni <- round(p.adjust(as.numeric(procr_p),method = "bonferroni"),3)

procr_p_host <- as.data.frame(cbind(procr_p_host, p_adj_bonferroni))
rownames(procr_p_host) <- c("all", "neutral", "non-neutral")

procr_p_random <- c(y_overlap_random$signif, y_neutral_random$signif, y_nonneutral_random$signif) 
p_adj_bonferroni <- round(p.adjust(as.numeric(procr_p_random),method = "bonferroni"),3)

procr_p_random <- as.data.frame(cbind(procr_p_random, p_adj_bonferroni))
rownames(procr_p_random) <- c("all", "neutral", "non-neutral")


```
Test on core families


## read top taxa list 


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_list <- readRDS(paste0(dataDir, "/list_ps_coreFamily_taxglomSp.rds"))

#ps_list <- readRDS(paste0(dataDir, "/list_ps_coreFamily_taxglomOTU.rds"))


ps_list[[1]]
```

count_taxa

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

count_tax <- list()

for (i in 1:length(ps_list)) {
    count_tax[[i]] <- ntaxa(ps_list[[i]])
    names(count_tax)[i] <- names(ps_list)[i]
}

count_tax_df<- as.data.frame(t(data.frame(lapply(count_tax, type.convert), stringsAsFactors=FALSE)))
names(count_tax_df) <- "n_taxa"

check <- count_tax_df %>% data.frame %>% dplyr::filter(n_taxa < 5)
nrow(check)

remove <- rownames(check)

# remove from phyloseq list
ps_list_filter <- within(ps_list, rm(list=remove))
length(ps_list_filter)
#[1] 27
```

distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

dist_l <- list()

for (i in 1:length(ps_list_filter)) {
    dist_l[[i]] <- cophenetic(phy_tree(ps_list_filter[[i]]))
    names(dist_l)[i] <- paste0("microb_dist_", str_remove(names(ps_list_filter)[i], "ps_"))
}

names(dist_l)

```

otu

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

otu_l <- list()

for (i in 1:length(ps_list_filter)) {

    otu_l[[i]] <- as.matrix(as.data.frame(otu_table(ps_list_filter[[i]])))
    otu_l[[i]] <- as.matrix(apply(otu_l[[i]], 2,function(x) ifelse(x > 0, 1, 0)))
    names(otu_l)[i] <- paste0("otu_", str_remove(names(ps_list_filter)[i], "ps_"))
}

setdiff(rownames(otu_l[[1]]), rownames(host_dist))
setdiff(rownames(microb_overlap_meta), rownames(otu_l[[1]]))
setdiff(rownames(microb_overlap_meta), rownames(host_dist))

```


Prepare data for PACo Test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

paco_data_l <- list()

for (i in 1:length(ps_list_filter)) {
    paco_data_l[[i]] <- prepare_paco_data(H = host_dist, P = dist_l[[i]], HP = otu_l[[i]])
    paco_data_l[[i]] = add_pcoord(paco_data_l[[i]], correction = "cailliez")
    names(paco_data_l)[i] <- paste0("paco_data_", str_remove(names(ps_list_filter)[i], "ps_"))
}

names(paco_data_l)


paco_data_l[[1]]

H_PCo <- paco_data_l[[1]]$H_PCo
P_PCo <- paco_data_l[[1]]$P_PCo


paco_data_random_l <- list()

for (i in 1:length(ps_list_filter)) {
    paco_data_random_l[[i]] <- prepare_paco_data(H = random_host_dist, P = dist_l[[i]], HP = otu_l[[i]])
    paco_data_random_l[[i]] = add_pcoord(paco_data_random_l[[i]], correction = "cailliez")
    names(paco_data_random_l)[i] <- paste0("paco_data_", str_remove(names(ps_list_filter)[i], "ps_"))
}

names(paco_data_random_l)


paco_data_random_l[[1]]

H_PCo_random <- paco_data_random_l[[1]]$H_PCo
P_PCo_random <- paco_data_random_l[[1]]$P_PCo


```

paco test

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

paco_each = function(D, nperm=99, seed=3843){
    PACo(D, nperm=nperm, seed=seed, method='quasiswap', symmetric=TRUE)
}

paco_res_l = future_lapply(paco_data_l, paco_each, nperm=99, future.packages=c('paco'))
paco_res_l %>% length
paco_res_l[[1]]


# goodness of fit
gof = lapply(paco_res_l, function(x) as.vector(x$gof)) %>% 
        do.call(rbind, .) %>%
        as.data.frame

gof$ss %>% as.numeric 
class(gof)
head(gof)
gof$R2 <- sqrt(1-as.numeric(gof$ss))

gof_df <- data.frame(Test= rownames(gof), p = as.numeric(gof$p), ss = as.numeric(gof$ss), n = as.numeric(gof$n), R2 = as.numeric(gof$R2))

gof_df$Genus_edit <- str_remove(gof_df$Test, "paco_data_")

#add number of OTU column 
count_tax_df$Genus_edit <- str_remove(rownames(count_tax_df), "ps_")
gof_df <- left_join(gof_df, count_tax_df, by = "Genus_edit")
head(gof_df)



# get p values from protest 
procrustes_each = function(D){
    protest(X = D$H_PCo, Y = D$P_PCo, symmetric=TRUE, seed = 3999)
}
procr_res_l = future_lapply(paco_data_l, procrustes_each,  future.packages=c('vegan', future.seed=TRUE))
procr_res_l[[1]]

# for loading results instead of re-running
#saveRDS(procr_res_l, paste0(dataDir, "/procr_res_l.RDS"))

# goodness of fit
procr_ss = lapply(procr_res_l, function(x) as.vector(x$ss)) %>% 
        do.call(rbind, .) %>%
        as.data.frame
procr_p = lapply(procr_res_l, function(x) as.vector(x$signif)) %>% 
        do.call(rbind, .) %>%
        as.data.frame
procr_df <- cbind(procr_ss, procr_p)
names(procr_df) <- c("ss", "p")
procr_df$p_adj_bonferroni <- round(p.adjust(as.numeric(procr_df$p),method = "bonferroni"),3)
procr_df$p_adj_fdr <- round(p.adjust(as.numeric(procr_df$p),method = "fdr"),3)

head(procr_df)
procr_df$Test <- rownames(procr_df)
names(procr_df) <- c("procr_ss", "procr_p", "p_adj_bonferroni", "p_adj_fdr", "Test")

sort(procr_df$p_adj_bonferroni) 
sort(procr_df$p_adj_fdr) 

gof$ss %>% as.numeric 
class(gof)
head(gof)
gof$R2 <- sqrt(1-as.numeric(gof$ss))

gof_df <- data.frame(Test= rownames(gof), p = as.numeric(gof$p), ss = as.numeric(gof$ss), n = as.numeric(gof$n), R2 = as.numeric(gof$R2))
head(gof_df)

out_merge_df <- left_join(gof_df, procr_df, by = "Test")
head(out_merge_df)

write.csv(out_merge_df, paste0(dataDir, "/paco_output_Sp.csv"))
#write.csv(out_merge_df, paste0(dataDir, "/paco_output_OTU.csv"))

```

paco test random

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

paco_each = function(D, nperm=99, seed=3843){
    PACo(D, nperm=nperm, seed=seed, method='quasiswap', symmetric=TRUE)
}

paco_res_random_l = future_lapply(paco_data_random_l, paco_each, nperm=99, future.packages=c('paco'))
paco_res_random_l %>% length
paco_res_random_l[[1]]


# goodness of fit
gof_random = lapply(paco_res_random_l, function(x) as.vector(x$gof)) %>% 
        do.call(rbind, .) %>%
        as.data.frame

gof_random$ss %>% as.numeric 
class(gof_random)
head(gof_random)
gof_random$R2 <- sqrt(1-as.numeric(gof_random$ss))

gof_random_df <- data.frame(Test= rownames(gof_random), p = as.numeric(gof_random$p), ss = as.numeric(gof_random$ss), n = as.numeric(gof_random$n), R2 = as.numeric(gof_random$R2))

gof_random_df$Genus_edit <- str_remove(gof_random_df$Test, "paco_data_")

#add number of OTU column 
count_tax_df$Genus_edit <- str_remove(rownames(count_tax_df), "ps_")
gof_random_df <- left_join(gof_random_df, count_tax_df, by = "Genus_edit")
head(gof_random_df)

sort(gof_random_df$p)

procrustes_each = function(D){
    protest(X = D$H_PCo, Y = D$P_PCo, symmetric=TRUE, nperm=99, seed = 3999)
}
procr_res_random_l = future_lapply(paco_data_random_l, procrustes_each,  future.packages=c('vegan', future.seed=TRUE))
procr_res_random_l[[1]]


# goodness of fit
procr_ss = lapply(procr_res_random_l, function(x) as.vector(x$ss)) %>% 
        do.call(rbind, .) %>%
        as.data.frame
procr_p = lapply(procr_res_random_l, function(x) as.vector(x$signif)) %>% 
        do.call(rbind, .) %>%
        as.data.frame
procr_df <- cbind(procr_ss, procr_p)
names(procr_df) <- c("ss", "p")
procr_df$p_adj_bonferroni <- round(p.adjust(as.numeric(procr_df$p),method = "bonferroni"),3)
procr_df$p_adj_fdr <- round(p.adjust(as.numeric(procr_df$p),method = "fdr"),3)

head(procr_df)
procr_df$Test <- rownames(procr_df)
names(procr_df) <- c("procr_ss", "procr_p", "p_adj_bonferroni", "p_adj_fdr", "Test")

sort(procr_df$p_adj_bonferroni) 
sort(procr_df$p_adj_fdr) 

gof_random$ss %>% as.numeric 
class(gof_random)
head(gof_random)
gof_random$R2 <- sqrt(1-as.numeric(gof_random$ss))

gof_random_df <- data.frame(Test= rownames(gof_random), p = as.numeric(gof_random$p), ss = as.numeric(gof_random$ss), n = as.numeric(gof_random$n), R2 = as.numeric(gof_random$R2))
head(gof_random_df)

out_merge_random_df <- left_join(gof_random_df, procr_df, by = "Test")
head(out_merge_random_df)


write.csv(out_merge_random_df, paste0(dataDir, "/paco_output_Sp_random.csv"))
#write.csv(out_merge_random_df, paste0(dataDir, "/paco_output_OTU_random.csv"))


```