---
title: 'Shared Taxa Analysis: Sympatric Chaffinch Microbiomes'
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute:
    code_folding: hide
    toc_depth: 2
---

# Setup

## Load Libraries

```{r load-libraries, message=FALSE, warning=FALSE}
rm(list = ls())

# Core packages
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(magrittr)

# Microbiome analysis
library(phyloseq)
library(vegan)
library(speedyseq)
library(microbiome)
library(biomeUtils)

# Statistics
library(MKinfer)  # For permutation t-tests

# Visualization
library(ggplot2)
library(ggpubr)
library(patchwork)
library(ggtext)

# Utilities
library(knitr)
library(kableExtra)
library(details)
```

## Working Directories

```{r source-directories, message=FALSE, warning=FALSE, results="hide"}
source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")
```

## Define Consistent Plot Aesthetics

```{r define-aesthetics}

comparison_colors <- c(
 "diff_Island_diff_SpComplex" = "#26547c",
 "diff_Island_same_SpComplex" = "#ef476f",
 "same_Island_diff_SpComplex" = "#ffd166",
 "same_Island_same_SpComplex" = "#06d6a0"
)

comparison_labels <- c(
 "diff_Island_diff_SpComplex" = "Different island, different species complex",
 "diff_Island_same_SpComplex" = "Different island, same species complex",
 "same_Island_diff_SpComplex" = "Same island, different species complex",
 "same_Island_same_SpComplex" = "Same island, same species complex"
)

# Factor level order for consistent x-axis
comparison_levels <- c(
 "diff_Island_diff_SpComplex",
 "diff_Island_same_SpComplex",
 "same_Island_diff_SpComplex",
 "same_Island_same_SpComplex"
)

# NMDS plot colors for host species
host_colors <- c(
 "Gran_Canaria_F_c_bakeri"     = "#31A354",
 "Tenerife_F_c_canariensis"    = "#708238",
 "Gran_Canaria_F_polatzeki"    = "#1F78B4",
 "Tenerife_F_teydea"           = "#A6CEE3"
)

host_shapes <- c(
 "Gran_Canaria_F_c_bakeri"     = 23,
 "Tenerife_F_c_canariensis"    = 22,
 "Gran_Canaria_F_polatzeki"    = 21,
 "Tenerife_F_teydea"           = 25
)

host_labels <- c(
 "Gran_Canaria_F_c_bakeri"     = "Gran Canaria: *F.c. bakeri*",
 "Tenerife_F_c_canariensis"    = "Tenerife: *F.c. canariensis*",
 "Gran_Canaria_F_polatzeki"    = "Gran Canaria: *F. polatzeki*",
 "Tenerife_F_teydea"           = "Tenerife: *F. teydea*"
)

# =============================================================================
# Theme for violin/boxplots
# =============================================================================

theme_violin <- theme_bw() +
 theme(
   axis.text.x        = element_blank(),
   axis.ticks.x       = element_blank(),
   axis.line.x        = element_blank(),
   panel.grid.major.x = element_blank(),
   legend.position    = "bottom",
   plot.title         = element_text(face = "bold", size = 12)
 )

# =============================================================================
# Theme for NMDS plots
# =============================================================================

theme_nmds <- theme_bw() +
 theme(
   legend.text  = element_markdown(),
   plot.title   = element_text(face = "bold", size = 12),
   axis.title   = element_text(size = 10)
 )
```

# Data Loading and Preparation

## Load Microbiome Data

```{r load-microbiome-data, message=FALSE, warning=FALSE}
microb_fr <- readRDS(paste0(dataDir, "/microb_overlap_diet_corrected.rds"))
microb_fr <- biomeUtils::removeZeros(microb_fr)
```

## Taxonomic Agglomeration

```{r tax-glom, message=FALSE, warning=FALSE}
# Create phyloseq objects at each taxonomic level
microb_fr_otu     <- microb_fr
microb_fr_species <- speedyseq::tax_glom(microb_fr, "Species", NArm = FALSE)
microb_fr_genus   <- speedyseq::tax_glom(microb_fr, "Genus",   NArm = FALSE)
microb_fr_family  <- speedyseq::tax_glom(microb_fr, "Family",  NArm = FALSE)
microb_fr_order   <- speedyseq::tax_glom(microb_fr, "Order",   NArm = FALSE)
microb_fr_class   <- speedyseq::tax_glom(microb_fr, "Class",   NArm = FALSE)
microb_fr_phylum  <- speedyseq::tax_glom(microb_fr, "Phylum",  NArm = FALSE)
```

## Update Taxonomy Tables with OTU IDs

```{r update-tax-tables, message=FALSE, warning=FALSE}
# Species
tax_table(microb_fr_species) <- tax_table(microb_fr_species)[, 1:7]
tax_table(microb_fr_species) <- cbind(tax_table(microb_fr_species), 
                                       rownames(tax_table(microb_fr_species)))
colnames(tax_table(microb_fr_species)) <- c("Kingdom", "Phylum", "Class", "Order", 
                                             "Family", "Genus", "Species", "OTU_ID")

# Genus
tax_table(microb_fr_genus) <- tax_table(microb_fr_genus)[, 1:6]
tax_table(microb_fr_genus) <- cbind(tax_table(microb_fr_genus), 
                                     rownames(tax_table(microb_fr_genus)))
colnames(tax_table(microb_fr_genus)) <- c("Kingdom", "Phylum", "Class", "Order", 
                                           "Family", "Genus", "OTU_ID")

# Family
tax_table(microb_fr_family) <- tax_table(microb_fr_family)[, 1:5]
tax_table(microb_fr_family) <- cbind(tax_table(microb_fr_family), 
                                      rownames(tax_table(microb_fr_family)))
colnames(tax_table(microb_fr_family)) <- c("Kingdom", "Phylum", "Class", "Order", 
                                            "Family", "OTU_ID")

# Order
tax_table(microb_fr_order) <- tax_table(microb_fr_order)[, 1:4]
tax_table(microb_fr_order) <- cbind(tax_table(microb_fr_order), 
                                     rownames(tax_table(microb_fr_order)))
colnames(tax_table(microb_fr_order)) <- c("Kingdom", "Phylum", "Class", "Order", "OTU_ID")

# Class
tax_table(microb_fr_class) <- tax_table(microb_fr_class)[, 1:3]
tax_table(microb_fr_class) <- cbind(tax_table(microb_fr_class), 
                                     rownames(tax_table(microb_fr_class)))
colnames(tax_table(microb_fr_class)) <- c("Kingdom", "Phylum", "Class", "OTU_ID")

# Phylum
tax_table(microb_fr_phylum) <- tax_table(microb_fr_phylum)[, 1:2]
tax_table(microb_fr_phylum) <- cbind(tax_table(microb_fr_phylum), 
                                      rownames(tax_table(microb_fr_phylum)))
colnames(tax_table(microb_fr_phylum)) <- c("Kingdom", "Phylum", "OTU_ID")
```

## Extract Metadata

```{r extract-metadata, message=FALSE, warning=FALSE}
microb_fr_meta_otu     <- sample_data(microb_fr_otu)     %>% data.frame()
microb_fr_meta_species <- sample_data(microb_fr_species) %>% data.frame()
microb_fr_meta_genus   <- sample_data(microb_fr_genus)   %>% data.frame()
microb_fr_meta_family  <- sample_data(microb_fr_family)  %>% data.frame()
microb_fr_meta_order   <- sample_data(microb_fr_order)   %>% data.frame()
microb_fr_meta_class   <- sample_data(microb_fr_class)   %>% data.frame()
microb_fr_meta_phylum  <- sample_data(microb_fr_phylum)  %>% data.frame()

# Add sympatric indicator to all metadata
microb_fr_meta_otu$sympatric     <- if_else(microb_fr_meta_otu$Island_Mainland %in% c("Tenerife", "Gran Canaria"), "sympatric", "allopatric")
microb_fr_meta_species$sympatric <- if_else(microb_fr_meta_species$Island_Mainland %in% c("Tenerife", "Gran Canaria"), "sympatric", "allopatric")
microb_fr_meta_genus$sympatric   <- if_else(microb_fr_meta_genus$Island_Mainland %in% c("Tenerife", "Gran Canaria"), "sympatric", "allopatric")
microb_fr_meta_family$sympatric  <- if_else(microb_fr_meta_family$Island_Mainland %in% c("Tenerife", "Gran Canaria"), "sympatric", "allopatric")
microb_fr_meta_order$sympatric   <- if_else(microb_fr_meta_order$Island_Mainland %in% c("Tenerife", "Gran Canaria"), "sympatric", "allopatric")
microb_fr_meta_class$sympatric   <- if_else(microb_fr_meta_class$Island_Mainland %in% c("Tenerife", "Gran Canaria"), "sympatric", "allopatric")
microb_fr_meta_phylum$sympatric  <- if_else(microb_fr_meta_phylum$Island_Mainland %in% c("Tenerife", "Gran Canaria"), "sympatric", "allopatric")
```

# Jaccard Similarity Analysis (Proportion of Shared Taxa)

## Calculate Pairwise Jaccard Distances

```{r jaccard-distances, message=FALSE, warning=FALSE}
# Helper to process distance matrices into pairwise comparison dataframes
process_pairwise_distances <- function(phyloseq_obj, meta_df) {
 
 dist <- phyloseq::distance(phyloseq_obj, method = "jaccard", binary = TRUE)
 distM <- as.matrix(dist)
 
 df1 <- data.frame(t(combn(colnames(distM), 2)), 
                   dist = t(distM)[lower.tri(distM)])
 colnames(df1) <- c("ind1", "ind2", "dist")
 
 site_species <- meta_df %>%
   dplyr::select(Species_Label, Species, Island_Mainland, 
                 Chaff_Blue, Archipelago_Mainland, sympatric) %>%
   rownames_to_column(var = "Sampleid")
 
 df2 <- df1 %>%
   left_join(site_species, by = c("ind1" = "Sampleid")) %>%
   left_join(site_species, by = c("ind2" = "Sampleid"), suffix = c(".ind1", ".ind2"))
 
 df_all <- df2 %>%
   mutate(
     Label_comp           = if_else(Species_Label.ind1 == Species_Label.ind2, "same_Label", "diff_Label"),
     Species_comp         = if_else(Species.ind1 == Species.ind2, "same_Species", "diff_Species"),
     Island_comp          = if_else(Island_Mainland.ind1 == Island_Mainland.ind2, "same_Island", "diff_Island"),
     Chaff_Blue_comp      = if_else(Chaff_Blue.ind1 == Chaff_Blue.ind2, "same_SpComplex", "diff_SpComplex"),
     Archipelago_comp     = if_else(Archipelago_Mainland.ind1 == Archipelago_Mainland.ind2, "same_Archipelago", "diff_Archipelago"),
     Archipelago_Island_comp = paste(Archipelago_comp, Island_comp, sep = "_"),
     Island_Species_comp  = paste(Island_comp, Species_comp, sep = "_"),
     Island_SpComplex_comp = paste(Island_comp, Chaff_Blue_comp, sep = "_"),
     similarity           = 1 - dist
   )
 
 return(df_all)
}

# Calculate for all taxonomic levels
df_all_otu     <- process_pairwise_distances(microb_fr_otu,     microb_fr_meta_otu)
df_all_species <- process_pairwise_distances(microb_fr_species, microb_fr_meta_species)
df_all_genus   <- process_pairwise_distances(microb_fr_genus,   microb_fr_meta_genus)
df_all_family  <- process_pairwise_distances(microb_fr_family,  microb_fr_meta_family)
df_all_order   <- process_pairwise_distances(microb_fr_order,   microb_fr_meta_order)
df_all_class   <- process_pairwise_distances(microb_fr_class,   microb_fr_meta_class)
df_all_phylum  <- process_pairwise_distances(microb_fr_phylum,  microb_fr_meta_phylum)

# Filter to sympatric populations only
df_sympatric_otu_jac     <- df_all_otu     %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_species_jac <- df_all_species %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_genus_jac   <- df_all_genus   %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_family_jac  <- df_all_family  %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_order_jac   <- df_all_order   %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_class_jac   <- df_all_class   %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_phylum_jac  <- df_all_phylum  %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")

# Set factor levels for consistent ordering
df_sympatric_otu_jac$Island_SpComplex_comp     <- factor(df_sympatric_otu_jac$Island_SpComplex_comp,     levels = comparison_levels)
df_sympatric_species_jac$Island_SpComplex_comp <- factor(df_sympatric_species_jac$Island_SpComplex_comp, levels = comparison_levels)
df_sympatric_genus_jac$Island_SpComplex_comp   <- factor(df_sympatric_genus_jac$Island_SpComplex_comp,   levels = comparison_levels)
df_sympatric_family_jac$Island_SpComplex_comp  <- factor(df_sympatric_family_jac$Island_SpComplex_comp,  levels = comparison_levels)
df_sympatric_order_jac$Island_SpComplex_comp   <- factor(df_sympatric_order_jac$Island_SpComplex_comp,   levels = comparison_levels)
df_sympatric_class_jac$Island_SpComplex_comp   <- factor(df_sympatric_class_jac$Island_SpComplex_comp,   levels = comparison_levels)
df_sympatric_phylum_jac$Island_SpComplex_comp  <- factor(df_sympatric_phylum_jac$Island_SpComplex_comp,  levels = comparison_levels)
```

## Jaccard: Plots

```{r jaccard-plots, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
# Helper to create consistent plots
create_similarity_plot <- function(df, title, ylab) {
 
 ggplot(df, aes(x = Island_SpComplex_comp, y = similarity)) +
   # Violin layer (background)
   geom_violin(aes(fill = Island_SpComplex_comp), 
               alpha = 0.3, color = NA, width = 0.9) +
   # Boxplot layer
   geom_boxplot(aes(fill = Island_SpComplex_comp), 
                width = 0.25, outlier.shape = NA, 
                alpha = 0.8, color = "black", linewidth = 0.5) +
   # Jittered points
   geom_jitter(aes(color = Island_SpComplex_comp), 
               width = 0.08, alpha = 0.4, size = 0.8) +
   # Scales
   scale_fill_manual(name = NULL, values = comparison_colors, labels = comparison_labels) +
   scale_color_manual(name = NULL, values = comparison_colors, labels = comparison_labels) +
   labs(title = title, y = ylab, x = NULL) +
   guides(fill = guide_legend(ncol = 2, override.aes = list(alpha = 0.8)), color = "none") +
   theme_violin
}

# Create plots for each taxonomic level
plot_jac_otu     <- create_similarity_plot(df_sympatric_otu_jac,     "OTU",     "Jaccard similarity")
plot_jac_species <- create_similarity_plot(df_sympatric_species_jac, "Species", "Jaccard similarity")
plot_jac_genus   <- create_similarity_plot(df_sympatric_genus_jac,   "Genus",   "Jaccard similarity")
plot_jac_family  <- create_similarity_plot(df_sympatric_family_jac,  "Family",  "Jaccard similarity")
plot_jac_order   <- create_similarity_plot(df_sympatric_order_jac,   "Order",   "Jaccard similarity")
plot_jac_class   <- create_similarity_plot(df_sympatric_class_jac,   "Class",   "Jaccard similarity")
plot_jac_phylum  <- create_similarity_plot(df_sympatric_phylum_jac,  "Phylum",  "Jaccard similarity")

# Combine all levels
fig_jaccard_all <- ggarrange(
 plot_jac_otu, plot_jac_species, plot_jac_genus, plot_jac_family,
 plot_jac_order, plot_jac_class, plot_jac_phylum,
 common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 2
)

fig_jaccard_all

# Save
png(paste0(outputDir, "/Figures/SharedTaxa_AllLevels_Jaccard.png"), 
   width = 2400, height = 1600, res = 200)
print(fig_jaccard_all)
dev.off()
```

## Jaccard: Permutation t-tests

```{r jaccard-permutation-tests, message=FALSE, warning=FALSE}
# =============================================================================
# First, subset to sympatric populations (Tenerife and Gran Canaria only)
# =============================================================================

microb_fr_otu_symp     <- microb_fr_otu     %>% subset_samples(Island_Mainland %in% c("Tenerife", "Gran Canaria"))
microb_fr_species_symp <- microb_fr_species %>% subset_samples(Island_Mainland %in% c("Tenerife", "Gran Canaria"))
microb_fr_genus_symp   <- microb_fr_genus   %>% subset_samples(Island_Mainland %in% c("Tenerife", "Gran Canaria"))
microb_fr_family_symp  <- microb_fr_family  %>% subset_samples(Island_Mainland %in% c("Tenerife", "Gran Canaria"))
microb_fr_order_symp   <- microb_fr_order   %>% subset_samples(Island_Mainland %in% c("Tenerife", "Gran Canaria"))
microb_fr_class_symp   <- microb_fr_class   %>% subset_samples(Island_Mainland %in% c("Tenerife", "Gran Canaria"))
microb_fr_phylum_symp  <- microb_fr_phylum  %>% subset_samples(Island_Mainland %in% c("Tenerife", "Gran Canaria"))

cat("Sympatric population sample sizes:\n")
table(sample_data(microb_fr_otu_symp)$Species_Label)

# =============================================================================
# Permutation t-tests for Jaccard similarity
# =============================================================================

# OTU level
perm_ttest_jac_otu <- function(group1, group2) {
  group1_values <- df_sympatric_otu_jac$dist[df_sympatric_otu_jac$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_otu_jac$dist[df_sympatric_otu_jac$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_jac_otu <- data.frame(rbind(
  perm_ttest_jac_otu("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_jac_otu("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_otu("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_otu("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_otu("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_otu("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_jac_otu <- stats_jac_otu %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_jac_otu$p_adj_bonferroni <- round(p.adjust(stats_jac_otu$p.value, method = "bonferroni"), 4)
stats_jac_otu$p.value <- round(stats_jac_otu$p.value, 4)
stats_jac_otu$tax <- "OTU99"

# Species level
perm_ttest_jac_species <- function(group1, group2) {
  group1_values <- df_sympatric_species_jac$dist[df_sympatric_species_jac$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_species_jac$dist[df_sympatric_species_jac$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_jac_species <- data.frame(rbind(
  perm_ttest_jac_species("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_jac_species("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_species("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_species("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_species("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_species("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_jac_species <- stats_jac_species %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_jac_species$p_adj_bonferroni <- round(p.adjust(stats_jac_species$p.value, method = "bonferroni"), 4)
stats_jac_species$p.value <- round(stats_jac_species$p.value, 4)
stats_jac_species$tax <- "Species"

# Genus level
perm_ttest_jac_genus <- function(group1, group2) {
  group1_values <- df_sympatric_genus_jac$dist[df_sympatric_genus_jac$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_genus_jac$dist[df_sympatric_genus_jac$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_jac_genus <- data.frame(rbind(
  perm_ttest_jac_genus("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_jac_genus("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_genus("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_genus("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_genus("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_genus("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_jac_genus <- stats_jac_genus %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_jac_genus$p_adj_bonferroni <- round(p.adjust(stats_jac_genus$p.value, method = "bonferroni"), 4)
stats_jac_genus$p.value <- round(stats_jac_genus$p.value, 4)
stats_jac_genus$tax <- "Genus"

# Family level
perm_ttest_jac_family <- function(group1, group2) {
  group1_values <- df_sympatric_family_jac$dist[df_sympatric_family_jac$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_family_jac$dist[df_sympatric_family_jac$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_jac_family <- data.frame(rbind(
  perm_ttest_jac_family("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_jac_family("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_family("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_family("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_family("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_family("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_jac_family <- stats_jac_family %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_jac_family$p_adj_bonferroni <- round(p.adjust(stats_jac_family$p.value, method = "bonferroni"), 4)
stats_jac_family$p.value <- round(stats_jac_family$p.value, 4)
stats_jac_family$tax <- "Family"

# Order level
perm_ttest_jac_order <- function(group1, group2) {
  group1_values <- df_sympatric_order_jac$dist[df_sympatric_order_jac$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_order_jac$dist[df_sympatric_order_jac$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_jac_order <- data.frame(rbind(
  perm_ttest_jac_order("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_jac_order("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_order("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_order("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_order("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_order("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_jac_order <- stats_jac_order %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_jac_order$p_adj_bonferroni <- round(p.adjust(stats_jac_order$p.value, method = "bonferroni"), 4)
stats_jac_order$p.value <- round(stats_jac_order$p.value, 4)
stats_jac_order$tax <- "Order"

# Class level
perm_ttest_jac_class <- function(group1, group2) {
  group1_values <- df_sympatric_class_jac$dist[df_sympatric_class_jac$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_class_jac$dist[df_sympatric_class_jac$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_jac_class <- data.frame(rbind(
  perm_ttest_jac_class("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_jac_class("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_class("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_class("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_class("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_class("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_jac_class <- stats_jac_class %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_jac_class$p_adj_bonferroni <- round(p.adjust(stats_jac_class$p.value, method = "bonferroni"), 4)
stats_jac_class$p.value <- round(stats_jac_class$p.value, 4)
stats_jac_class$tax <- "Class"

# Phylum level
perm_ttest_jac_phylum <- function(group1, group2) {
  group1_values <- df_sympatric_phylum_jac$dist[df_sympatric_phylum_jac$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_phylum_jac$dist[df_sympatric_phylum_jac$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_jac_phylum <- data.frame(rbind(
  perm_ttest_jac_phylum("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_jac_phylum("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_phylum("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_phylum("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_jac_phylum("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_jac_phylum("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_jac_phylum <- stats_jac_phylum %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_jac_phylum$p_adj_bonferroni <- round(p.adjust(stats_jac_phylum$p.value, method = "bonferroni"), 4)
stats_jac_phylum$p.value <- round(stats_jac_phylum$p.value, 4)
stats_jac_phylum$tax <- "Phylum"

# Combine all results
stats_jac_all <- rbind(
  stats_jac_otu, stats_jac_species, stats_jac_genus, stats_jac_family,
  stats_jac_order, stats_jac_class, stats_jac_phylum
)

# Display species-level results
kable(stats_jac_species, 
     caption = "Permutation t-test results: Jaccard distance (Species level)") %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

# Save
write.csv(stats_jac_all, paste0(dataDir, "/out_permtest_Jaccard.csv"), row.names = FALSE)
```

# Unweighted UniFrac Analysis

## Calculate Pairwise Unweighted UniFrac Distances

```{r unifrac-unweighted-distances, message=FALSE, warning=FALSE}
# Helper to process UniFrac distances
process_unifrac_distances <- function(phyloseq_obj, meta_df, weighted = FALSE) {
 
 dist <- UniFrac(phyloseq_obj, weighted = weighted)
 distM <- as.matrix(dist)
 
 df1 <- data.frame(t(combn(colnames(distM), 2)), 
                   dist = t(distM)[lower.tri(distM)])
 colnames(df1) <- c("ind1", "ind2", "dist")
 
 site_species <- meta_df %>%
   dplyr::select(Species_Label, Species, Island_Mainland, 
                 Chaff_Blue, Archipelago_Mainland, sympatric) %>%
   rownames_to_column(var = "Sampleid")
 
 df2 <- df1 %>%
   left_join(site_species, by = c("ind1" = "Sampleid")) %>%
   left_join(site_species, by = c("ind2" = "Sampleid"), suffix = c(".ind1", ".ind2"))
 
 df_all <- df2 %>%
   mutate(
     Label_comp           = if_else(Species_Label.ind1 == Species_Label.ind2, "same_Label", "diff_Label"),
     Species_comp         = if_else(Species.ind1 == Species.ind2, "same_Species", "diff_Species"),
     Island_comp          = if_else(Island_Mainland.ind1 == Island_Mainland.ind2, "same_Island", "diff_Island"),
     Chaff_Blue_comp      = if_else(Chaff_Blue.ind1 == Chaff_Blue.ind2, "same_SpComplex", "diff_SpComplex"),
     Archipelago_comp     = if_else(Archipelago_Mainland.ind1 == Archipelago_Mainland.ind2, "same_Archipelago", "diff_Archipelago"),
     Archipelago_Island_comp = paste(Archipelago_comp, Island_comp, sep = "_"),
     Island_Species_comp  = paste(Island_comp, Species_comp, sep = "_"),
     Island_SpComplex_comp = paste(Island_comp, Chaff_Blue_comp, sep = "_"),
     similarity           = 1 - dist
   )
 
 return(df_all)
}

# Calculate unweighted UniFrac for all taxonomic levels
df_all_otu_uw     <- process_unifrac_distances(microb_fr_otu,     microb_fr_meta_otu,     weighted = FALSE)
df_all_species_uw <- process_unifrac_distances(microb_fr_species, microb_fr_meta_species, weighted = FALSE)
df_all_genus_uw   <- process_unifrac_distances(microb_fr_genus,   microb_fr_meta_genus,   weighted = FALSE)
df_all_family_uw  <- process_unifrac_distances(microb_fr_family,  microb_fr_meta_family,  weighted = FALSE)
df_all_order_uw   <- process_unifrac_distances(microb_fr_order,   microb_fr_meta_order,   weighted = FALSE)
df_all_class_uw   <- process_unifrac_distances(microb_fr_class,   microb_fr_meta_class,   weighted = FALSE)
df_all_phylum_uw  <- process_unifrac_distances(microb_fr_phylum,  microb_fr_meta_phylum,  weighted = FALSE)

# Filter to sympatric populations
df_sympatric_otu_uw     <- df_all_otu_uw     %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_species_uw <- df_all_species_uw %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_genus_uw   <- df_all_genus_uw   %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_family_uw  <- df_all_family_uw  %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_order_uw   <- df_all_order_uw   %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_class_uw   <- df_all_class_uw   %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")
df_sympatric_phylum_uw  <- df_all_phylum_uw  %>% filter(sympatric.ind1 == "sympatric" & sympatric.ind2 == "sympatric")

# Set factor levels
df_sympatric_otu_uw$Island_SpComplex_comp     <- factor(df_sympatric_otu_uw$Island_SpComplex_comp,     levels = comparison_levels)
df_sympatric_species_uw$Island_SpComplex_comp <- factor(df_sympatric_species_uw$Island_SpComplex_comp, levels = comparison_levels)
df_sympatric_genus_uw$Island_SpComplex_comp   <- factor(df_sympatric_genus_uw$Island_SpComplex_comp,   levels = comparison_levels)
df_sympatric_family_uw$Island_SpComplex_comp  <- factor(df_sympatric_family_uw$Island_SpComplex_comp,  levels = comparison_levels)
df_sympatric_order_uw$Island_SpComplex_comp   <- factor(df_sympatric_order_uw$Island_SpComplex_comp,   levels = comparison_levels)
df_sympatric_class_uw$Island_SpComplex_comp   <- factor(df_sympatric_class_uw$Island_SpComplex_comp,   levels = comparison_levels)
df_sympatric_phylum_uw$Island_SpComplex_comp  <- factor(df_sympatric_phylum_uw$Island_SpComplex_comp,  levels = comparison_levels)
```

## Unweighted UniFrac: Plots

```{r unifrac-unweighted-plots, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}

plot_uw_otu     <- create_similarity_plot(df_sympatric_otu_uw,     "OTU",     "1 − Unweighted UniFrac")
plot_uw_species <- create_similarity_plot(df_sympatric_species_uw, "Species", "1 − Unweighted UniFrac")
plot_uw_genus   <- create_similarity_plot(df_sympatric_genus_uw,   "Genus",   "1 − Unweighted UniFrac")
plot_uw_family  <- create_similarity_plot(df_sympatric_family_uw,  "Family",  "1 − Unweighted UniFrac")
plot_uw_order   <- create_similarity_plot(df_sympatric_order_uw,   "Order",   "1 − Unweighted UniFrac")
plot_uw_class   <- create_similarity_plot(df_sympatric_class_uw,   "Class",   "1 − Unweighted UniFrac")
plot_uw_phylum  <- create_similarity_plot(df_sympatric_phylum_uw,  "Phylum",  "1 − Unweighted UniFrac")

fig_uw_all <- ggarrange(
 plot_uw_otu, plot_uw_species, plot_uw_genus, plot_uw_family,
 plot_uw_order, plot_uw_class, plot_uw_phylum,
 common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 2
)

fig_uw_all

png(paste0(outputDir, "/Figures/SharedTaxa_AllLevels_UW_UniFrac.png"), 
   width = 2400, height = 1600, res = 200)
print(fig_uw_all)
dev.off()
```

## Unweighted UniFrac: Permutation t-tests

```{r unifrac-unweighted-permutation-tests, message=FALSE, warning=FALSE}

# OTU level
perm_ttest_uw_otu <- function(group1, group2) {
  group1_values <- df_sympatric_otu_uw$dist[df_sympatric_otu_uw$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_otu_uw$dist[df_sympatric_otu_uw$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_uw_otu <- data.frame(rbind(
  perm_ttest_uw_otu("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_uw_otu("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_otu("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_otu("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_otu("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_otu("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_uw_otu <- stats_uw_otu %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_uw_otu$p_adj_bonferroni <- round(p.adjust(stats_uw_otu$p.value, method = "bonferroni"), 4)
stats_uw_otu$p.value <- round(stats_uw_otu$p.value, 4)
stats_uw_otu$tax <- "OTU99"

# Species level
perm_ttest_uw_species <- function(group1, group2) {
  group1_values <- df_sympatric_species_uw$dist[df_sympatric_species_uw$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_species_uw$dist[df_sympatric_species_uw$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_uw_species <- data.frame(rbind(
  perm_ttest_uw_species("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_uw_species("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_species("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_species("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_species("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_species("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_uw_species <- stats_uw_species %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_uw_species$p_adj_bonferroni <- round(p.adjust(stats_uw_species$p.value, method = "bonferroni"), 4)
stats_uw_species$p.value <- round(stats_uw_species$p.value, 4)
stats_uw_species$tax <- "Species"

# Genus level
perm_ttest_uw_genus <- function(group1, group2) {
  group1_values <- df_sympatric_genus_uw$dist[df_sympatric_genus_uw$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_genus_uw$dist[df_sympatric_genus_uw$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_uw_genus <- data.frame(rbind(
  perm_ttest_uw_genus("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_uw_genus("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_genus("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_genus("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_genus("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_genus("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_uw_genus <- stats_uw_genus %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_uw_genus$p_adj_bonferroni <- round(p.adjust(stats_uw_genus$p.value, method = "bonferroni"), 4)
stats_uw_genus$p.value <- round(stats_uw_genus$p.value, 4)
stats_uw_genus$tax <- "Genus"

# Family level
perm_ttest_uw_family <- function(group1, group2) {
  group1_values <- df_sympatric_family_uw$dist[df_sympatric_family_uw$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_family_uw$dist[df_sympatric_family_uw$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_uw_family <- data.frame(rbind(
  perm_ttest_uw_family("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_uw_family("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_family("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_family("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_family("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_family("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_uw_family <- stats_uw_family %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_uw_family$p_adj_bonferroni <- round(p.adjust(stats_uw_family$p.value, method = "bonferroni"), 4)
stats_uw_family$p.value <- round(stats_uw_family$p.value, 4)
stats_uw_family$tax <- "Family"

# Order level
perm_ttest_uw_order <- function(group1, group2) {
  group1_values <- df_sympatric_order_uw$dist[df_sympatric_order_uw$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_order_uw$dist[df_sympatric_order_uw$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_uw_order <- data.frame(rbind(
  perm_ttest_uw_order("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_uw_order("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_order("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_order("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_order("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_order("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_uw_order <- stats_uw_order %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_uw_order$p_adj_bonferroni <- round(p.adjust(stats_uw_order$p.value, method = "bonferroni"), 4)
stats_uw_order$p.value <- round(stats_uw_order$p.value, 4)
stats_uw_order$tax <- "Order"

# Class level
perm_ttest_uw_class <- function(group1, group2) {
  group1_values <- df_sympatric_class_uw$dist[df_sympatric_class_uw$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_class_uw$dist[df_sympatric_class_uw$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_uw_class <- data.frame(rbind(
  perm_ttest_uw_class("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_uw_class("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_class("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_class("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_class("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_class("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_uw_class <- stats_uw_class %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_uw_class$p_adj_bonferroni <- round(p.adjust(stats_uw_class$p.value, method = "bonferroni"), 4)
stats_uw_class$p.value <- round(stats_uw_class$p.value, 4)
stats_uw_class$tax <- "Class"

# Phylum level
perm_ttest_uw_phylum <- function(group1, group2) {
  group1_values <- df_sympatric_phylum_uw$dist[df_sympatric_phylum_uw$Island_SpComplex_comp == group1]
  group2_values <- df_sympatric_phylum_uw$dist[df_sympatric_phylum_uw$Island_SpComplex_comp == group2]
  ttest <- perm.t.test(group1_values, group2_values)
  res <- c(group1, group2, length(group1_values), length(group2_values),
           ttest$estimate[1], abs(ttest$estimate[2]), ttest$statistic, ttest$p.value)
  names(res) <- c('group1', 'group2', 'group1_n', 'group2_n',
                  'group1_mean', 'group2_mean', 'statistic', 'p.value')
  return(res)
}

stats_uw_phylum <- data.frame(rbind(
  perm_ttest_uw_phylum("diff_Island_diff_SpComplex", "diff_Island_same_SpComplex"),
  perm_ttest_uw_phylum("diff_Island_diff_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_phylum("diff_Island_diff_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_phylum("diff_Island_same_SpComplex", "same_Island_diff_SpComplex"),
  perm_ttest_uw_phylum("diff_Island_same_SpComplex", "same_Island_same_SpComplex"),
  perm_ttest_uw_phylum("same_Island_diff_SpComplex", "same_Island_same_SpComplex")
))
stats_uw_phylum <- stats_uw_phylum %>%
  mutate(p.value = as.numeric(p.value),
         statistic = round(as.numeric(statistic), 1),
         group1_mean = round(as.numeric(group1_mean), 4),
         group2_mean = round(as.numeric(group2_mean), 4))
stats_uw_phylum$p_adj_bonferroni <- round(p.adjust(stats_uw_phylum$p.value, method = "bonferroni"), 4)
stats_uw_phylum$p.value <- round(stats_uw_phylum$p.value, 4)
stats_uw_phylum$tax <- "Phylum"

# Combine
stats_uw_all <- rbind(
  stats_uw_otu, stats_uw_species, stats_uw_genus, stats_uw_family,
  stats_uw_order, stats_uw_class, stats_uw_phylum
)

# Display Species level
kable(stats_uw_species, 
     caption = "Permutation t-test results: Unweighted UniFrac distance (Species level)") %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


write.csv(stats_uw_all, paste0(dataDir, "/out_permtest_UW_UniFrac.csv"), row.names = FALSE)
```
# Combined Jaccard + Unweighted UniFrac (Species Level)

```{r publication-figure, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}

add_significance_annotations <- function(p, stats_df, y_base = 0.46, y_step = 0.04) {
 
 # Filter for significant comparisons (Bonferroni-adjusted)
 sig_comps <- stats_df %>% filter(p_adj_bonferroni < 0.05)
 
 if (nrow(sig_comps) == 0) return(p)
 
 # Define x positions for each category
 x_pos <- c(
   "diff_Island_diff_SpComplex" = 1,
   "diff_Island_same_SpComplex" = 2,
   "same_Island_diff_SpComplex" = 3,
   "same_Island_same_SpComplex" = 4
 )
 
 # Add annotations for each significant comparison
 for (i in 1:nrow(sig_comps)) {
   g1 <- sig_comps$group1[i]
   g2 <- sig_comps$group2[i]
   pval <- sig_comps$p_adj_bonferroni[i]
   
   # Get x positions
   x1 <- x_pos[g1] + 0.1
   x2 <- x_pos[g2] - 0.1
   
   # Ensure x1 < x2 for proper bracket direction
   if (x1 > x2) {
     tmp <- x1; x1 <- x2; x2 <- tmp
   }
   
   y <- y_base + (i - 1) * y_step
   
   # Determine significance symbol: * <0.05, ** <0.01, *** <0.001
   sig_symbol <- ifelse(pval < 0.001, "***", ifelse(pval < 0.01, "**", "*"))
   
   p <- p +
     annotate("segment", x = x1, xend = x2, y = y, yend = y, linewidth = 0.4) +
     annotate("text", x = (x1 + x2) / 2, y = y + 0.008, label = sig_symbol, size = 5)
 }
 
 return(p)
}


panel_A <- ggplot(df_sympatric_species_jac, 
                 aes(x = Island_SpComplex_comp, y = similarity)) +
 geom_violin(aes(fill = Island_SpComplex_comp), alpha = 0.3, color = NA, width = 0.9) +
 geom_boxplot(aes(fill = Island_SpComplex_comp), width = 0.25, outlier.shape = NA, 
              alpha = 0.8, color = "black", linewidth = 0.5) +
 geom_jitter(aes(color = Island_SpComplex_comp), width = 0.08, alpha = 0.4, size = 1) +
 scale_fill_manual(name = NULL, values = comparison_colors, labels = comparison_labels) +
 scale_color_manual(name = NULL, values = comparison_colors, labels = comparison_labels) +
 coord_cartesian(ylim = c(0, 0.65)) +
 labs(title = "A", y = "Jaccard similarity\n(proportion of shared species)", x = NULL) +
 guides(fill = guide_legend(ncol = 2, override.aes = list(alpha = 0.8)), color = "none") +
 theme_violin + theme(plot.title = element_text(size = 18, face = "bold"))

panel_A <- add_significance_annotations(panel_A, stats_jac_species, y_base = 0.46, y_step = 0.04)


panel_B <- ggplot(df_sympatric_species_uw, 
                 aes(x = Island_SpComplex_comp, y = similarity)) +
 geom_violin(aes(fill = Island_SpComplex_comp), alpha = 0.3, color = NA, width = 0.9) +
 geom_boxplot(aes(fill = Island_SpComplex_comp), width = 0.25, outlier.shape = NA, 
              alpha = 0.8, color = "black", linewidth = 0.5) +
 geom_jitter(aes(color = Island_SpComplex_comp), width = 0.08, alpha = 0.4, size = 1) +
 scale_fill_manual(name = NULL, values = comparison_colors, labels = comparison_labels) +
 scale_color_manual(name = NULL, values = comparison_colors, labels = comparison_labels) +
 coord_cartesian(ylim = c(0, 0.65)) +
 labs(title = "B", y = "1 − Unweighted UniFrac\n(phylogenetic similarity)", x = NULL) +
 guides(fill = guide_legend(ncol = 2, override.aes = list(alpha = 0.8)), color = "none") +
 theme_violin + theme(plot.title = element_text(size = 18, face = "bold"))

# Add significance annotations for Species-level UniFrac
panel_B <- add_significance_annotations(panel_B, stats_uw_species, y_base = 0.46, y_step = 0.04)


fig_publication <- panel_A + panel_B +
 plot_layout(guides = "collect") &
 theme(legend.position = "bottom")

fig_publication


png(paste0(outputDir, "/Figures/Fig_SharedSpecies_UWUniFrac_Combined.png"),
   width = 3600, height = 2200, res = 300)
print(fig_publication)
dev.off()
```

# Summary Statistics

```{r summary-stats, message=FALSE, warning=FALSE}
# Calculate mean similarity values for manuscript text
summary_stats <- df_sympatric_species_jac %>%
 group_by(Island_SpComplex_comp) %>%
 summarise(
   n = n(),
   mean_similarity = round(mean(similarity) * 100, 1),
   sd_similarity = round(sd(similarity) * 100, 1),
   .groups = "drop"
 )

kable(summary_stats, 
     caption = "Proportion of shared microbial species (Jaccard similarity, Species level)",
     col.names = c("Comparison", "N pairs", "Mean (%)", "SD (%)")) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

# Key values for results text
cat("\n=== Key values for manuscript text ===\n")
cat("\nDifferent island, same species complex:", 
   summary_stats$mean_similarity[summary_stats$Island_SpComplex_comp == "diff_Island_same_SpComplex"], "%\n")
cat("Different island, different species complex:", 
   summary_stats$mean_similarity[summary_stats$Island_SpComplex_comp == "diff_Island_diff_SpComplex"], "%\n")
cat("Same island, same species complex:", 
   summary_stats$mean_similarity[summary_stats$Island_SpComplex_comp == "same_Island_same_SpComplex"], "%\n")
cat("Same island, different species complex:", 
   summary_stats$mean_similarity[summary_stats$Island_SpComplex_comp == "same_Island_diff_SpComplex"], "%\n")
```

# NMDS Ordination and PERMANOVA

## Helper Function: NMDS Plot

```{r nmds-helper, message=FALSE, warning=FALSE}
create_nmds_plot <- function(phyloseq_obj, distance_method = "jaccard", binary = TRUE, 
                             weighted_unifrac = FALSE, title = "NMDS", seed = 306202) {
 
 set.seed(seed)
 
 # Calculate distance and ordination
 if (distance_method == "unifrac") {
   dist_matrix <- UniFrac(phyloseq_obj, weighted = weighted_unifrac)
   ord <- metaMDS(dist_matrix, k = 4)
 } else {
   ord <- metaMDS(otu_table(phyloseq_obj), k = 4, distance = distance_method, binary = binary)
 }
 
 # Extract scores
 scrs <- as.data.frame(scores(ord, display = "sites"))
 scrs$Species_Label <- sample_data(phyloseq_obj)$Species_Label
 scrs$Island_Mainland <- sample_data(phyloseq_obj)$Island_Mainland
 scrs$Chaff_Blue <- sample_data(phyloseq_obj)$Chaff_Blue
 
 # Calculate centroids
 cent <- aggregate(cbind(NMDS1, NMDS2) ~ Species_Label, data = scrs, FUN = mean)
 
 # Create plot
 p <- ggplot(scrs, aes(x = NMDS1, y = NMDS2, shape = Species_Label, fill = Species_Label)) +
   geom_point(size = 4, alpha = 0.7) +
   geom_point(data = cent, aes(y = NMDS2, x = NMDS1, fill = Species_Label), 
              size = 8, stroke = 1.2, color = "black") +
   scale_fill_manual(name = "Host species", values = host_colors, labels = host_labels) +
   scale_shape_manual(name = "Host species", values = host_shapes, labels = host_labels) +
   scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
   scale_x_continuous(labels = scales::number_format(accuracy = 0.01)) +
   labs(title = title) +
   guides(color = "none") +
   theme_nmds
 
 return(list(plot = p, stress = ord$stress))
}
```

## Jaccard NMDS (Sympatric Populations)

```{r nmds-jaccard, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
nmds_jac_otu     <- create_nmds_plot(microb_fr_otu_symp,     "jaccard", TRUE, FALSE, "OTU")
nmds_jac_species <- create_nmds_plot(microb_fr_species_symp, "jaccard", TRUE, FALSE, "Species")
nmds_jac_genus   <- create_nmds_plot(microb_fr_genus_symp,   "jaccard", TRUE, FALSE, "Genus")
nmds_jac_family  <- create_nmds_plot(microb_fr_family_symp,  "jaccard", TRUE, FALSE, "Family")
nmds_jac_order   <- create_nmds_plot(microb_fr_order_symp,   "jaccard", TRUE, FALSE, "Order")
nmds_jac_class   <- create_nmds_plot(microb_fr_class_symp,   "jaccard", TRUE, FALSE, "Class")
nmds_jac_phylum  <- create_nmds_plot(microb_fr_phylum_symp,  "jaccard", TRUE, FALSE, "Phylum")

# Combine
fig_nmds_jac <- ggarrange(
 nmds_jac_otu$plot, nmds_jac_species$plot, nmds_jac_genus$plot, nmds_jac_family$plot,
 nmds_jac_order$plot, nmds_jac_class$plot, nmds_jac_phylum$plot,
 common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 2
)

fig_nmds_jac

png(paste0(outputDir, "/Figures/NMDSSymp_AllLevels_Jaccard.png"), 
   width = 2400, height = 1600, res = 150)
print(fig_nmds_jac)
dev.off()

cat("Jaccard NMDS Stress values:\n")
cat("OTU:", round(nmds_jac_otu$stress, 4), "\n")
cat("Species:", round(nmds_jac_species$stress, 4), "\n")
cat("Genus:", round(nmds_jac_genus$stress, 4), "\n")
cat("Family:", round(nmds_jac_family$stress, 4), "\n")
cat("Order:", round(nmds_jac_order$stress, 4), "\n")
cat("Class:", round(nmds_jac_class$stress, 4), "\n")
cat("Phylum:", round(nmds_jac_phylum$stress, 4), "\n")
```

## Jaccard PERMANOVA

```{r permanova-jaccard, message=FALSE, warning=FALSE}
jac_dist_otu     <- vegdist(otu_table(microb_fr_otu_symp),     method = "jaccard", binary = TRUE)
jac_dist_species <- vegdist(otu_table(microb_fr_species_symp), method = "jaccard", binary = TRUE)
jac_dist_genus   <- vegdist(otu_table(microb_fr_genus_symp),   method = "jaccard", binary = TRUE)
jac_dist_family  <- vegdist(otu_table(microb_fr_family_symp),  method = "jaccard", binary = TRUE)
jac_dist_order   <- vegdist(otu_table(microb_fr_order_symp),   method = "jaccard", binary = TRUE)
jac_dist_class   <- vegdist(otu_table(microb_fr_class_symp),   method = "jaccard", binary = TRUE)
jac_dist_phylum  <- vegdist(otu_table(microb_fr_phylum_symp),  method = "jaccard", binary = TRUE)

cat("=== Jaccard PERMANOVA Results ===\n\n")
cat("OTU level:\n")
print(adonis2(jac_dist_otu ~ Species_Label, data = meta(microb_fr_otu_symp), by = "margin", permutations = 9999))

cat("\nSpecies level:\n")
print(adonis2(jac_dist_species ~ Species_Label, data = meta(microb_fr_species_symp), by = "margin", permutations = 9999))

cat("\nGenus level:\n")
print(adonis2(jac_dist_genus ~ Species_Label, data = meta(microb_fr_genus_symp), by = "margin", permutations = 9999))

cat("\nFamily level:\n")
print(adonis2(jac_dist_family ~ Species_Label, data = meta(microb_fr_family_symp), by = "margin", permutations = 9999))

cat("\nOrder level:\n")
print(adonis2(jac_dist_order ~ Species_Label, data = meta(microb_fr_order_symp), by = "margin", permutations = 9999))

cat("\nClass level:\n")
print(adonis2(jac_dist_class ~ Species_Label, data = meta(microb_fr_class_symp), by = "margin", permutations = 9999))

cat("\nPhylum level:\n")
print(adonis2(jac_dist_phylum ~ Species_Label, data = meta(microb_fr_phylum_symp), by = "margin", permutations = 9999))
```

## Bray-Curtis NMDS (Sympatric Populations)

```{r nmds-bray, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
nmds_bray_otu     <- create_nmds_plot(microb_fr_otu_symp,     "bray", FALSE, FALSE, "OTU")
nmds_bray_species <- create_nmds_plot(microb_fr_species_symp, "bray", FALSE, FALSE, "Species")
nmds_bray_genus   <- create_nmds_plot(microb_fr_genus_symp,   "bray", FALSE, FALSE, "Genus")
nmds_bray_family  <- create_nmds_plot(microb_fr_family_symp,  "bray", FALSE, FALSE, "Family")
nmds_bray_order   <- create_nmds_plot(microb_fr_order_symp,   "bray", FALSE, FALSE, "Order")
nmds_bray_class   <- create_nmds_plot(microb_fr_class_symp,   "bray", FALSE, FALSE, "Class")
nmds_bray_phylum  <- create_nmds_plot(microb_fr_phylum_symp,  "bray", FALSE, FALSE, "Phylum")

fig_nmds_bray <- ggarrange(
 nmds_bray_otu$plot, nmds_bray_species$plot, nmds_bray_genus$plot, nmds_bray_family$plot,
 nmds_bray_order$plot, nmds_bray_class$plot, nmds_bray_phylum$plot,
 common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 2
)

fig_nmds_bray

png(paste0(outputDir, "/Figures/NMDSSymp_AllLevels_Bray.png"), 
   width = 2400, height = 1600, res = 150)
print(fig_nmds_bray)
dev.off()
```

## Bray-Curtis PERMANOVA

```{r permanova-bray, message=FALSE, warning=FALSE}
bray_dist_otu     <- vegdist(otu_table(microb_fr_otu_symp),     method = "bray", binary = FALSE)
bray_dist_species <- vegdist(otu_table(microb_fr_species_symp), method = "bray", binary = FALSE)
bray_dist_genus   <- vegdist(otu_table(microb_fr_genus_symp),   method = "bray", binary = FALSE)
bray_dist_family  <- vegdist(otu_table(microb_fr_family_symp),  method = "bray", binary = FALSE)
bray_dist_order   <- vegdist(otu_table(microb_fr_order_symp),   method = "bray", binary = FALSE)
bray_dist_class   <- vegdist(otu_table(microb_fr_class_symp),   method = "bray", binary = FALSE)
bray_dist_phylum  <- vegdist(otu_table(microb_fr_phylum_symp),  method = "bray", binary = FALSE)

cat("=== Bray-Curtis PERMANOVA Results ===\n\n")
cat("OTU level:\n")
print(adonis2(bray_dist_otu ~ Species_Label, data = meta(microb_fr_otu_symp), by = "margin", permutations = 9999))

cat("\nSpecies level:\n")
print(adonis2(bray_dist_species ~ Species_Label, data = meta(microb_fr_species_symp), by = "margin", permutations = 9999))

cat("\nGenus level:\n")
print(adonis2(bray_dist_genus ~ Species_Label, data = meta(microb_fr_genus_symp), by = "margin", permutations = 9999))

cat("\nFamily level:\n")
print(adonis2(bray_dist_family ~ Species_Label, data = meta(microb_fr_family_symp), by = "margin", permutations = 9999))

cat("\nOrder level:\n")
print(adonis2(bray_dist_order ~ Species_Label, data = meta(microb_fr_order_symp), by = "margin", permutations = 9999))

cat("\nClass level:\n")
print(adonis2(bray_dist_class ~ Species_Label, data = meta(microb_fr_class_symp), by = "margin", permutations = 9999))

cat("\nPhylum level:\n")
print(adonis2(bray_dist_phylum ~ Species_Label, data = meta(microb_fr_phylum_symp), by = "margin", permutations = 9999))
```

## Unweighted UniFrac NMDS

```{r nmds-unifrac-unweighted, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
nmds_uw_otu     <- create_nmds_plot(microb_fr_otu_symp,     "unifrac", TRUE, FALSE, "OTU")
nmds_uw_species <- create_nmds_plot(microb_fr_species_symp, "unifrac", TRUE, FALSE, "Species")
nmds_uw_genus   <- create_nmds_plot(microb_fr_genus_symp,   "unifrac", TRUE, FALSE, "Genus")
nmds_uw_family  <- create_nmds_plot(microb_fr_family_symp,  "unifrac", TRUE, FALSE, "Family")
nmds_uw_order   <- create_nmds_plot(microb_fr_order_symp,   "unifrac", TRUE, FALSE, "Order")
nmds_uw_class   <- create_nmds_plot(microb_fr_class_symp,   "unifrac", TRUE, FALSE, "Class")
nmds_uw_phylum  <- create_nmds_plot(microb_fr_phylum_symp,  "unifrac", TRUE, FALSE, "Phylum")

# Combine
fig_nmds_uw <- ggarrange(
 nmds_uw_otu$plot, nmds_uw_species$plot, nmds_uw_genus$plot, nmds_uw_family$plot,
 nmds_uw_order$plot, nmds_uw_class$plot, nmds_uw_phylum$plot,
 common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 2
)

fig_nmds_uw


png(paste0(outputDir, "/Figures/NMDSSymp_AllLevels_UW_UniFrac.png"), 
   width = 2400, height = 1600, res = 150)
print(fig_nmds_uw)
dev.off()
```

## Unweighted UniFrac PERMANOVA

```{r permanova-unifrac-unweighted, message=FALSE, warning=FALSE}
uw_dist_otu     <- UniFrac(microb_fr_otu_symp,     weighted = FALSE)
uw_dist_species <- UniFrac(microb_fr_species_symp, weighted = FALSE)
uw_dist_genus   <- UniFrac(microb_fr_genus_symp,   weighted = FALSE)
uw_dist_family  <- UniFrac(microb_fr_family_symp,  weighted = FALSE)
uw_dist_order   <- UniFrac(microb_fr_order_symp,   weighted = FALSE)
uw_dist_class   <- UniFrac(microb_fr_class_symp,   weighted = FALSE)
uw_dist_phylum  <- UniFrac(microb_fr_phylum_symp,  weighted = FALSE)

cat("=== Unweighted UniFrac PERMANOVA Results ===\n\n")
cat("OTU level:\n")
print(adonis2(uw_dist_otu ~ Species_Label, data = meta(microb_fr_otu_symp), by = "margin", permutations = 9999))

cat("\nSpecies level:\n")
print(adonis2(uw_dist_species ~ Species_Label, data = meta(microb_fr_species_symp), by = "margin", permutations = 9999))

cat("\nGenus level:\n")
print(adonis2(uw_dist_genus ~ Species_Label, data = meta(microb_fr_genus_symp), by = "margin", permutations = 9999))

cat("\nFamily level:\n")
print(adonis2(uw_dist_family ~ Species_Label, data = meta(microb_fr_family_symp), by = "margin", permutations = 9999))

cat("\nOrder level:\n")
print(adonis2(uw_dist_order ~ Species_Label, data = meta(microb_fr_order_symp), by = "margin", permutations = 9999))

cat("\nClass level:\n")
print(adonis2(uw_dist_class ~ Species_Label, data = meta(microb_fr_class_symp), by = "margin", permutations = 9999))

cat("\nPhylum level:\n")
print(adonis2(uw_dist_phylum ~ Species_Label, data = meta(microb_fr_phylum_symp), by = "margin", permutations = 9999))
```

## Weighted UniFrac NMDS

```{r nmds-unifrac-weighted, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
nmds_w_otu     <- create_nmds_plot(microb_fr_otu_symp,     "unifrac", TRUE, TRUE, "OTU")
nmds_w_species <- create_nmds_plot(microb_fr_species_symp, "unifrac", TRUE, TRUE, "Species")
nmds_w_genus   <- create_nmds_plot(microb_fr_genus_symp,   "unifrac", TRUE, TRUE, "Genus")
nmds_w_family  <- create_nmds_plot(microb_fr_family_symp,  "unifrac", TRUE, TRUE, "Family")
nmds_w_order   <- create_nmds_plot(microb_fr_order_symp,   "unifrac", TRUE, TRUE, "Order")
nmds_w_class   <- create_nmds_plot(microb_fr_class_symp,   "unifrac", TRUE, TRUE, "Class")
nmds_w_phylum  <- create_nmds_plot(microb_fr_phylum_symp,  "unifrac", TRUE, TRUE, "Phylum")


fig_nmds_w <- ggarrange(
 nmds_w_otu$plot, nmds_w_species$plot, nmds_w_genus$plot, nmds_w_family$plot,
 nmds_w_order$plot, nmds_w_class$plot, nmds_w_phylum$plot,
 common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 2
)

fig_nmds_w


png(paste0(outputDir, "/Figures/NMDSSymp_AllLevels_W_UniFrac.png"), 
   width = 2400, height = 1600, res = 150)
print(fig_nmds_w)
dev.off()
```

## Weighted UniFrac PERMANOVA

```{r permanova-unifrac-weighted, message=FALSE, warning=FALSE}
w_dist_otu     <- UniFrac(microb_fr_otu_symp,     weighted = TRUE)
w_dist_species <- UniFrac(microb_fr_species_symp, weighted = TRUE)
w_dist_genus   <- UniFrac(microb_fr_genus_symp,   weighted = TRUE)
w_dist_family  <- UniFrac(microb_fr_family_symp,  weighted = TRUE)
w_dist_order   <- UniFrac(microb_fr_order_symp,   weighted = TRUE)
w_dist_class   <- UniFrac(microb_fr_class_symp,   weighted = TRUE)
w_dist_phylum  <- UniFrac(microb_fr_phylum_symp,  weighted = TRUE)

cat("=== Weighted UniFrac PERMANOVA Results ===\n\n")
cat("OTU level:\n")
print(adonis2(w_dist_otu ~ Species_Label, data = meta(microb_fr_otu_symp), by = "margin", permutations = 9999))

cat("\nSpecies level:\n")
print(adonis2(w_dist_species ~ Species_Label, data = meta(microb_fr_species_symp), by = "margin", permutations = 9999))

cat("\nGenus level:\n")
print(adonis2(w_dist_genus ~ Species_Label, data = meta(microb_fr_genus_symp), by = "margin", permutations = 9999))

cat("\nFamily level:\n")
print(adonis2(w_dist_family ~ Species_Label, data = meta(microb_fr_family_symp), by = "margin", permutations = 9999))

cat("\nOrder level:\n")
print(adonis2(w_dist_order ~ Species_Label, data = meta(microb_fr_order_symp), by = "margin", permutations = 9999))

cat("\nClass level:\n")
print(adonis2(w_dist_class ~ Species_Label, data = meta(microb_fr_class_symp), by = "margin", permutations = 9999))

cat("\nPhylum level:\n")
print(adonis2(w_dist_phylum ~ Species_Label, data = meta(microb_fr_phylum_symp), by = "margin", permutations = 9999))
```


# Session Info

```{r session-info, message=FALSE, warning=FALSE}
sessioninfo::session_info() %>%
 details::details(summary = "Current session info", open = TRUE)
```
