---
title: 'Fst '
author: "B. K. Montero and M. Gillingham"
date: "`r Sys.Date()`"
output: html_document

---


## LOAD LIBRARIES

```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

rm(list = ls())

# basics
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(vegan)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(MicEco)
library(rlist) 
library(grid)
library(tsiR) 
library(ape)
library(ggh4x)
# phyloseq section
library(phyloseq)
library(microViz)
library(biomeUtils)
library(patchwork)

# dist
library(treedataverse)
library(geosphere)
library(treedata.table)
library(sf)


# MRM 
library(ecodist)
library(modEvA)
library(eulerr)


```


## WORKING DIRECTORIES

```{r source, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
source("~/Dropbox/FringillaMicrobiome/05_Scripts/CleanScripts/graphics_param.R")

```


## LOAD DATA

Rarefied 16S

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_fr_otu <- readRDS(paste0(dataDir, "/microb_overlap_OTU99_fr_taxFix.rds"))
microb_fr_species <- speedyseq::tax_glom(microb_fr_otu, "Species")
microb_fr_genus <- speedyseq::tax_glom(microb_fr_otu, "Genus")
microb_fr_family <- speedyseq::tax_glom(microb_fr_otu, "Family")
microb_fr_order <- speedyseq::tax_glom(microb_fr_otu, "Order")
microb_fr_class <- speedyseq::tax_glom(microb_fr_otu, "Class")
microb_fr_phylum <- speedyseq::tax_glom(microb_fr_otu, "Phylum")

# check for NAs
tax_micro <- tax_table(microb_fr) %>% data.frame
table(is.na(tax_micro$Genus))

# TipId 
sample_data(microb_fr_otu)$TipID_microbiome <- paste(sample_data(microb_fr_otu)$Lab_Code, sample_data(microb_fr_otu)$Species_Label, sep = "_")
sample_data(microb_fr_species)$TipID_microbiome <- paste(sample_data(microb_fr_species)$Lab_Code, sample_data(microb_fr_species)$Species_Label, sep = "_")
sample_data(microb_fr_genus)$TipID_microbiome <- paste(sample_data(microb_fr_genus)$Lab_Code, sample_data(microb_fr_genus)$Species_Label, sep = "_")
sample_data(microb_fr_family)$TipID_microbiome <- paste(sample_data(microb_fr_family)$Lab_Code, sample_data(microb_fr_family)$Species_Label, sep = "_")
sample_data(microb_fr_order)$TipID_microbiome <- paste(sample_data(microb_fr_order)$Lab_Code, sample_data(microb_fr_order)$Species_Label, sep = "_")
sample_data(microb_fr_class)$TipID_microbiome <- paste(sample_data(microb_fr_class)$Lab_Code, sample_data(microb_fr_class)$Species_Label, sep = "_")
sample_data(microb_fr_phylum)$TipID_microbiome <- paste(sample_data(microb_fr_phylum)$Lab_Code, sample_data(microb_fr_phylum)$Species_Label, sep = "_")

# meta
microb_fr_meta_otu <- sample_data(microb_fr_otu) %>% data.frame
microb_fr_meta_species <- sample_data(microb_fr_species) %>% data.frame
microb_fr_meta_genus <- sample_data(microb_fr_genus) %>% data.frame
microb_fr_meta_family <- sample_data(microb_fr_family) %>% data.frame
microb_fr_meta_order <- sample_data(microb_fr_order) %>% data.frame
microb_fr_meta_class <- sample_data(microb_fr_class) %>% data.frame
microb_fr_meta_phylum <- sample_data(microb_fr_phylum) %>% data.frame

```


Diet data

```{r psObject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

diet_fr <- readRDS(paste0(dataDir, "/diet_overlap_OTU99_fr.rds"))

```


## Overlap all datasets

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_overlap_otu <- microb_fr_otu %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_species <- microb_fr_species %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_genus <- microb_fr_genus %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_family <- microb_fr_family %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_class <- microb_fr_class %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_order <- microb_fr_order %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")
microb_overlap_phylum <- microb_fr_phylum %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89")

microb_overlap_meta <- microb_overlap_otu %>% sample_data %>% data.frame 
names(microb_overlap_meta)


```

## Overlap chaffinch subset

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

microb_overlap_chaff_otu <- microb_fr_otu %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89" & Chaff_Blue == "Chaffinch")
microb_overlap_chaff_species <- microb_fr_species %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89" & Chaff_Blue == "Chaffinch")
microb_overlap_chaff_genus <- microb_fr_genus %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89" & Chaff_Blue == "Chaffinch")
microb_overlap_chaff_family <- microb_fr_family %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89" & Chaff_Blue == "Chaffinch")
microb_overlap_chaff_class <- microb_fr_class %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89" & Chaff_Blue == "Chaffinch")
microb_overlap_chaff_order <- microb_fr_order %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89" & Chaff_Blue == "Chaffinch")
microb_overlap_chaff_phylum <- microb_fr_phylum %>% subset_samples(Genus == "Fringilla" & WGS_data == 1 & Lab_Code != "132" & Lab_Code != "252" & Lab_Code != "89" & Chaff_Blue == "Chaffinch")

microb_overlap_chaff_meta <- microb_overlap_chaff_otu %>% sample_data %>% data.frame 
nrow(microb_overlap_chaff_meta)


```



# TEST Fst with merged phyloseq objects per population 


Centroids of geographic distance

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# all finches
# lat long data
my_points <- data.frame(longitude = sample_data(microb_fr_otu)$Xdec, latitude = sample_data(microb_fr_otu)$Ydec, Island_Mainland = sample_data(microb_fr_otu)$Species_Label, row.names = sample_names(microb_fr_otu))
str(my_points)
table(my_points$Island_Mainland)
df_sf <- st_as_sf(my_points, coords = c('longitude', 'latitude'))

# get centroids per island / mainland 
centroids_sf <- df_sf %>%
  group_by(Island_Mainland) %>% 
  summarize(geometry = st_union(geometry)) %>% 
  st_centroid %>%   
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])
str(centroids_sf)
rownames(centroids_sf) <- centroids_sf$Island_Mainland

# dist
dist_mat <- distm(cbind(centroids_sf$lon, centroids_sf$lat), fun = distGeo)  # Vector of distances in meters
dist_mat <-dist_mat/1000 #Vector of distances in km
rownames(dist_mat) <- rownames(centroids_sf)
colnames(dist_mat) <- rownames(centroids_sf)
geo_dist <- as.dist(dist_mat)
str(dist_mat)

# chaffinch subset
my_points_subset <- my_points %>% filter(Island_Mainland != "Tenerife_F_teydea" & Island_Mainland != "Gran_Canaria_F_polatzeki")
table(my_points_subset$Island_Mainland)

df_sf <- st_as_sf(my_points_subset, coords = c('longitude', 'latitude'))

# get centroids per island / mainland 
centroids_sf <- df_sf %>%
  group_by(Island_Mainland) %>% 
  summarize(geometry = st_union(geometry)) %>% 
  st_centroid %>%   
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])
str(centroids_sf)
rownames(centroids_sf) <- centroids_sf$Island_Mainland

# dist
dist_mat_chaff <- distm(cbind(centroids_sf$lon, centroids_sf$lat), fun = distGeo)  # Vector of distances in meters
dist_mat_chaff <-dist_mat_chaff/1000 #Vector of distances in km
rownames(dist_mat_chaff) <- rownames(centroids_sf)
colnames(dist_mat_chaff) <- rownames(centroids_sf)
geo_dist_chaff <- as.dist(dist_mat_chaff)
str(dist_mat_chaff)
```

Fst values

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

#fst <- read.csv(paste0(dataDir, "/MeanFst_Fr_estimates.csv"), header = T, row.names = 1)
fst <- read.csv(paste0(dataDir, "/MeanFst_Fr_weighted_estimates.csv"), header = T, row.names = 1)

str(fst)

fst <- fst %>% as.matrix
fst1 <- fst[rownames(dist_mat),colnames(dist_mat)] %>% as.matrix
rownames(fst1)
fst_dist <- as.dist(fst1)
str(fst_dist)

#fst <- read.csv(paste0(dataDir, "/MeanFst_Fr_estimates_chaff.csv"), header = T, row.names = 1)
fst1_chaff <- fst[rownames(dist_mat_chaff),colnames(dist_mat_chaff)] %>% as.matrix
fst_dist_chaff <- as.dist(fst1_chaff)
str(fst_dist_chaff)

library(gdata)
mat <- data.frame(geo=unmatrix(dist_mat), host=unmatrix(fst1))
#write.csv(mat, paste0(dataDir, "/fst_dist_df.csv"))
mat_edit <- read.csv(paste0(dataDir, "/fst_dist_df_edit.csv"))
head(mat_edit)
str(mat_edit)
library(ggrepel)

mat_unique <- mat_edit %>% distinct(geo, .keep_all = TRUE) %>% filter(geo != 0 | host != 0)
head(mat_unique)
#png(paste0(outputDir, "/01_Exploration/fst_dist.png"), width = 400, height = 250, units = "mm", res = 200)
(mm_hostdist <- ggplot(mat_unique, aes(x = geo, y = host, label=X, col = Region)) + geom_point(size = 4, alpha = 0.2, show.legend = FALSE) + labs(x = "Geographic distance", y = "Fst") + geom_smooth(method = "lm", colour = "red", alpha = 0.1,lwd=2)+ geom_label_repel(hjust=0, vjust=0, max.overlaps=20, show.legend = FALSE) +theme_classic(base_size = 13)+theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+ RegionDyad.color)
#dev.off()


```

subsample microbiome equal sample size


```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

sample_ps <- function(ps, FUN = sample, ...){
  ids <- sample_names(ps)
  sampled_ids <- FUN(ids, ...)
  ps <- prune_samples(sampled_ids, ps)
  return(ps)
}

levels(factor(sample_data(microb_fr_otu)$Species_Label))

# otu 
(ElHierro <- microb_fr_otu %>% subset_samples(Species_Label=="El_Hierro_F_c_ombriosa" ))
(Flores <- microb_fr_otu %>% subset_samples(Species_Label=="Flores_F_moreletti" ))
(GranCanaria <- microb_fr_otu %>% subset_samples(Species_Label=="Gran_Canaria_F_c_bakeri" ))
(Polatzeki <- microb_fr_otu %>% subset_samples(Species_Label=="Gran_Canaria_F_polatzeki" ))
(LaGomera <- microb_fr_otu %>% subset_samples(Species_Label=="La_Gomera_F_c_canariensis" ))
(LaPalma <- microb_fr_otu %>% subset_samples(Species_Label=="La_Palma_F_c_palmae" ))
(Madeira <- microb_fr_otu %>% subset_samples(Species_Label=="Madeira_F_maderensis" ))
(Mainland <- microb_fr_otu %>% subset_samples(Species_Label=="Mainland_F_coelebs" ))
(SaoMiguel <- microb_fr_otu %>% subset_samples(Species_Label=="Sao_Miguel_F_moreletti" ))
(Tenerife <- microb_fr_otu %>% subset_samples(Species_Label=="Tenerife_F_c_canariensis" ))
(Teydea <- microb_fr_otu %>% subset_samples(Species_Label=="Tenerife_F_teydea" ))

(ElHierro <- sample_ps(ElHierro, size = 9))
(Flores <- sample_ps(Flores, size = 9))
(GranCanaria <- sample_ps(GranCanaria, size = 9))
(Polatzeki <- sample_ps(Polatzeki, size = 9))
(LaGomera <- sample_ps(LaGomera, size = 9))
(LaPalma <- sample_ps(LaPalma, size = 9))
(Madeira <- sample_ps(Madeira, size = 9))
(Mainland <- sample_ps(Mainland, size = 9))
(SaoMiguel <- sample_ps(SaoMiguel, size = 9))
(Tenerife <- sample_ps(Tenerife, size = 9))
(Teydea <- sample_ps(Teydea, size = 9))


(microb_fr_subsample_otu <- merge_phyloseq(ElHierro, Flores, GranCanaria, Polatzeki, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife, Teydea) %>% removeZeros())


(microb_fr_subsample_otu_chaff <- merge_phyloseq(ElHierro, Flores, GranCanaria, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife) %>% removeZeros())


# species
(ElHierro <- microb_fr_species %>% subset_samples(Species_Label=="El_Hierro_F_c_ombriosa" ))
(Flores <- microb_fr_species %>% subset_samples(Species_Label=="Flores_F_moreletti" ))
(GranCanaria <- microb_fr_species %>% subset_samples(Species_Label=="Gran_Canaria_F_c_bakeri" ))
(Polatzeki <- microb_fr_species %>% subset_samples(Species_Label=="Gran_Canaria_F_polatzeki" ))
(LaGomera <- microb_fr_species %>% subset_samples(Species_Label=="La_Gomera_F_c_canariensis" ))
(LaPalma <- microb_fr_species %>% subset_samples(Species_Label=="La_Palma_F_c_palmae" ))
(Madeira <- microb_fr_species %>% subset_samples(Species_Label=="Madeira_F_maderensis" ))
(Mainland <- microb_fr_species %>% subset_samples(Species_Label=="Mainland_F_coelebs" ))
(SaoMiguel <- microb_fr_species %>% subset_samples(Species_Label=="Sao_Miguel_F_moreletti" ))
(Tenerife <- microb_fr_species %>% subset_samples(Species_Label=="Tenerife_F_c_canariensis" ))
(Teydea <- microb_fr_species %>% subset_samples(Species_Label=="Tenerife_F_teydea" ))

(ElHierro <- sample_ps(ElHierro, size = 9))
(Flores <- sample_ps(Flores, size = 9))
(GranCanaria <- sample_ps(GranCanaria, size = 9))
(Polatzeki <- sample_ps(Polatzeki, size = 9))
(LaGomera <- sample_ps(LaGomera, size = 9))
(LaPalma <- sample_ps(LaPalma, size = 9))
(Madeira <- sample_ps(Madeira, size = 9))
(Mainland <- sample_ps(Mainland, size = 9))
(SaoMiguel <- sample_ps(SaoMiguel, size = 9))
(Tenerife <- sample_ps(Tenerife, size = 9))
(Teydea <- sample_ps(Teydea, size = 9))


(microb_fr_subsample_species <- merge_phyloseq(ElHierro, Flores, GranCanaria, Polatzeki, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife, Teydea) %>% removeZeros())
(microb_fr_subsample_species_chaff <- merge_phyloseq(ElHierro, Flores, GranCanaria, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife) %>% removeZeros())


# genus
(ElHierro <- microb_fr_genus %>% subset_samples(Species_Label=="El_Hierro_F_c_ombriosa" ))
(Flores <- microb_fr_genus %>% subset_samples(Species_Label=="Flores_F_moreletti" ))
(GranCanaria <- microb_fr_genus %>% subset_samples(Species_Label=="Gran_Canaria_F_c_bakeri" ))
(Polatzeki <- microb_fr_genus %>% subset_samples(Species_Label=="Gran_Canaria_F_polatzeki" ))
(LaGomera <- microb_fr_genus %>% subset_samples(Species_Label=="La_Gomera_F_c_canariensis" ))
(LaPalma <- microb_fr_genus %>% subset_samples(Species_Label=="La_Palma_F_c_palmae" ))
(Madeira <- microb_fr_genus %>% subset_samples(Species_Label=="Madeira_F_maderensis" ))
(Mainland <- microb_fr_genus %>% subset_samples(Species_Label=="Mainland_F_coelebs" ))
(SaoMiguel <- microb_fr_genus %>% subset_samples(Species_Label=="Sao_Miguel_F_moreletti" ))
(Tenerife <- microb_fr_genus %>% subset_samples(Species_Label=="Tenerife_F_c_canariensis" ))
(Teydea <- microb_fr_genus %>% subset_samples(Species_Label=="Tenerife_F_teydea" ))

(ElHierro <- sample_ps(ElHierro, size = 9))
(Flores <- sample_ps(Flores, size = 9))
(GranCanaria <- sample_ps(GranCanaria, size = 9))
(Polatzeki <- sample_ps(Polatzeki, size = 9))
(LaGomera <- sample_ps(LaGomera, size = 9))
(LaPalma <- sample_ps(LaPalma, size = 9))
(Madeira <- sample_ps(Madeira, size = 9))
(Mainland <- sample_ps(Mainland, size = 9))
(SaoMiguel <- sample_ps(SaoMiguel, size = 9))
(Tenerife <- sample_ps(Tenerife, size = 9))
(Teydea <- sample_ps(Teydea, size = 9))


(microb_fr_subsample_genus <- merge_phyloseq(ElHierro, Flores, GranCanaria, Polatzeki, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife, Teydea) %>% removeZeros())
(microb_fr_subsample_genus_chaff <- merge_phyloseq(ElHierro, Flores, GranCanaria, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife) %>% removeZeros())

# family
(ElHierro <- microb_fr_family %>% subset_samples(Species_Label=="El_Hierro_F_c_ombriosa" ))
(Flores <- microb_fr_family %>% subset_samples(Species_Label=="Flores_F_moreletti" ))
(GranCanaria <- microb_fr_family %>% subset_samples(Species_Label=="Gran_Canaria_F_c_bakeri" ))
(Polatzeki <- microb_fr_family %>% subset_samples(Species_Label=="Gran_Canaria_F_polatzeki" ))
(LaGomera <- microb_fr_family %>% subset_samples(Species_Label=="La_Gomera_F_c_canariensis" ))
(LaPalma <- microb_fr_family %>% subset_samples(Species_Label=="La_Palma_F_c_palmae" ))
(Madeira <- microb_fr_family %>% subset_samples(Species_Label=="Madeira_F_maderensis" ))
(Mainland <- microb_fr_family %>% subset_samples(Species_Label=="Mainland_F_coelebs" ))
(SaoMiguel <- microb_fr_family %>% subset_samples(Species_Label=="Sao_Miguel_F_moreletti" ))
(Tenerife <- microb_fr_family %>% subset_samples(Species_Label=="Tenerife_F_c_canariensis" ))
(Teydea <- microb_fr_family %>% subset_samples(Species_Label=="Tenerife_F_teydea" ))

(ElHierro <- sample_ps(ElHierro, size = 9))
(Flores <- sample_ps(Flores, size = 9))
(GranCanaria <- sample_ps(GranCanaria, size = 9))
(Polatzeki <- sample_ps(Polatzeki, size = 9))
(LaGomera <- sample_ps(LaGomera, size = 9))
(LaPalma <- sample_ps(LaPalma, size = 9))
(Madeira <- sample_ps(Madeira, size = 9))
(Mainland <- sample_ps(Mainland, size = 9))
(SaoMiguel <- sample_ps(SaoMiguel, size = 9))
(Tenerife <- sample_ps(Tenerife, size = 9))
(Teydea <- sample_ps(Teydea, size = 9))

(microb_fr_subsample_family <- merge_phyloseq(ElHierro, Flores, GranCanaria, Polatzeki, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife, Teydea) %>% removeZeros())
(microb_fr_subsample_family_chaff <- merge_phyloseq(ElHierro, Flores, GranCanaria, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife) %>% removeZeros())

# order
(ElHierro <- microb_fr_order %>% subset_samples(Species_Label=="El_Hierro_F_c_ombriosa" ))
(Flores <- microb_fr_order %>% subset_samples(Species_Label=="Flores_F_moreletti" ))
(GranCanaria <- microb_fr_order %>% subset_samples(Species_Label=="Gran_Canaria_F_c_bakeri" ))
(Polatzeki <- microb_fr_order %>% subset_samples(Species_Label=="Gran_Canaria_F_polatzeki" ))
(LaGomera <- microb_fr_order %>% subset_samples(Species_Label=="La_Gomera_F_c_canariensis" ))
(LaPalma <- microb_fr_order %>% subset_samples(Species_Label=="La_Palma_F_c_palmae" ))
(Madeira <- microb_fr_order %>% subset_samples(Species_Label=="Madeira_F_maderensis" ))
(Mainland <- microb_fr_order %>% subset_samples(Species_Label=="Mainland_F_coelebs" ))
(SaoMiguel <- microb_fr_order %>% subset_samples(Species_Label=="Sao_Miguel_F_moreletti" ))
(Tenerife <- microb_fr_order %>% subset_samples(Species_Label=="Tenerife_F_c_canariensis" ))
(Teydea <- microb_fr_order %>% subset_samples(Species_Label=="Tenerife_F_teydea" ))

(ElHierro <- sample_ps(ElHierro, size = 9))
(Flores <- sample_ps(Flores, size = 9))
(GranCanaria <- sample_ps(GranCanaria, size = 9))
(Polatzeki <- sample_ps(Polatzeki, size = 9))
(LaGomera <- sample_ps(LaGomera, size = 9))
(LaPalma <- sample_ps(LaPalma, size = 9))
(Madeira <- sample_ps(Madeira, size = 9))
(Mainland <- sample_ps(Mainland, size = 9))
(SaoMiguel <- sample_ps(SaoMiguel, size = 9))
(Tenerife <- sample_ps(Tenerife, size = 9))
(Teydea <- sample_ps(Teydea, size = 9))


(microb_fr_subsample_order <- merge_phyloseq(ElHierro, Flores, GranCanaria, Polatzeki, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife, Teydea) %>% removeZeros())
(microb_fr_subsample_order_chaff <- merge_phyloseq(ElHierro, Flores, GranCanaria, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife) %>% removeZeros())

# class
(ElHierro <- microb_fr_class %>% subset_samples(Species_Label=="El_Hierro_F_c_ombriosa" ))
(Flores <- microb_fr_class %>% subset_samples(Species_Label=="Flores_F_moreletti" ))
(GranCanaria <- microb_fr_class %>% subset_samples(Species_Label=="Gran_Canaria_F_c_bakeri" ))
(Polatzeki <- microb_fr_class %>% subset_samples(Species_Label=="Gran_Canaria_F_polatzeki" ))
(LaGomera <- microb_fr_class %>% subset_samples(Species_Label=="La_Gomera_F_c_canariensis" ))
(LaPalma <- microb_fr_class %>% subset_samples(Species_Label=="La_Palma_F_c_palmae" ))
(Madeira <- microb_fr_class %>% subset_samples(Species_Label=="Madeira_F_maderensis" ))
(Mainland <- microb_fr_class %>% subset_samples(Species_Label=="Mainland_F_coelebs" ))
(SaoMiguel <- microb_fr_class %>% subset_samples(Species_Label=="Sao_Miguel_F_moreletti" ))
(Tenerife <- microb_fr_class %>% subset_samples(Species_Label=="Tenerife_F_c_canariensis" ))
(Teydea <- microb_fr_class %>% subset_samples(Species_Label=="Tenerife_F_teydea" ))

(ElHierro <- sample_ps(ElHierro, size = 9))
(Flores <- sample_ps(Flores, size = 9))
(GranCanaria <- sample_ps(GranCanaria, size = 9))
(Polatzeki <- sample_ps(Polatzeki, size = 9))
(LaGomera <- sample_ps(LaGomera, size = 9))
(LaPalma <- sample_ps(LaPalma, size = 9))
(Madeira <- sample_ps(Madeira, size = 9))
(Mainland <- sample_ps(Mainland, size = 9))
(SaoMiguel <- sample_ps(SaoMiguel, size = 9))
(Tenerife <- sample_ps(Tenerife, size = 9))
(Teydea <- sample_ps(Teydea, size = 9))


(microb_fr_subsample_class <- merge_phyloseq(ElHierro, Flores, GranCanaria, Polatzeki, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife, Teydea) %>% removeZeros())
(microb_fr_subsample_class_chaff <- merge_phyloseq(ElHierro, Flores, GranCanaria, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife) %>% removeZeros())

# phylum
(ElHierro <- microb_fr_phylum %>% subset_samples(Species_Label=="El_Hierro_F_c_ombriosa" ))
(Flores <- microb_fr_phylum %>% subset_samples(Species_Label=="Flores_F_moreletti" ))
(GranCanaria <- microb_fr_phylum %>% subset_samples(Species_Label=="Gran_Canaria_F_c_bakeri" ))
(Polatzeki <- microb_fr_phylum %>% subset_samples(Species_Label=="Gran_Canaria_F_polatzeki" ))
(LaGomera <- microb_fr_phylum %>% subset_samples(Species_Label=="La_Gomera_F_c_canariensis" ))
(LaPalma <- microb_fr_phylum %>% subset_samples(Species_Label=="La_Palma_F_c_palmae" ))
(Madeira <- microb_fr_phylum %>% subset_samples(Species_Label=="Madeira_F_maderensis" ))
(Mainland <- microb_fr_phylum %>% subset_samples(Species_Label=="Mainland_F_coelebs" ))
(SaoMiguel <- microb_fr_phylum %>% subset_samples(Species_Label=="Sao_Miguel_F_moreletti" ))
(Tenerife <- microb_fr_phylum %>% subset_samples(Species_Label=="Tenerife_F_c_canariensis" ))
(Teydea <- microb_fr_phylum %>% subset_samples(Species_Label=="Tenerife_F_teydea" ))

(ElHierro <- sample_ps(ElHierro, size = 9))
(Flores <- sample_ps(Flores, size = 9))
(GranCanaria <- sample_ps(GranCanaria, size = 9))
(Polatzeki <- sample_ps(Polatzeki, size = 9))
(LaGomera <- sample_ps(LaGomera, size = 9))
(LaPalma <- sample_ps(LaPalma, size = 9))
(Madeira <- sample_ps(Madeira, size = 9))
(Mainland <- sample_ps(Mainland, size = 9))
(SaoMiguel <- sample_ps(SaoMiguel, size = 9))
(Tenerife <- sample_ps(Tenerife, size = 9))
(Teydea <- sample_ps(Teydea, size = 9))


(microb_fr_subsample_phylum <- merge_phyloseq(ElHierro, Flores, GranCanaria, Polatzeki, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife, Teydea) %>% removeZeros())
(microb_fr_subsample_phylum_chaff <- merge_phyloseq(ElHierro, Flores, GranCanaria, LaGomera, LaPalma, Madeira, Mainland, SaoMiguel, Tenerife) %>% removeZeros())

```
 
merge
 
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# merge samples per island / mainland : species
ps_merged_otu <- phyloseq::merge_samples(microb_fr_subsample_otu, "Species_Label", fun = sum)
ps_merged_species <- phyloseq::merge_samples(microb_fr_subsample_species, "Species_Label", fun = sum)
ps_merged_genus <- phyloseq::merge_samples(microb_fr_subsample_genus, "Species_Label", fun = sum)
ps_merged_family <- phyloseq::merge_samples(microb_fr_subsample_family, "Species_Label", fun = sum)
ps_merged_class <- phyloseq::merge_samples(microb_fr_subsample_class, "Species_Label", fun = sum)
ps_merged_order <- phyloseq::merge_samples(microb_fr_subsample_order, "Species_Label", fun = sum)
ps_merged_phylum <- phyloseq::merge_samples(microb_fr_subsample_phylum, "Species_Label", fun = sum)


ps_merged_otu_chaff <- phyloseq::merge_samples(microb_fr_subsample_otu_chaff, "Species_Label", fun = sum)
ps_merged_species_chaff <- phyloseq::merge_samples(microb_fr_subsample_species_chaff, "Species_Label", fun = sum)
ps_merged_genus_chaff <- phyloseq::merge_samples(microb_fr_subsample_genus_chaff, "Species_Label", fun = sum)
ps_merged_family_chaff <- phyloseq::merge_samples(microb_fr_subsample_family_chaff, "Species_Label", fun = sum)
ps_merged_class_chaff <- phyloseq::merge_samples(microb_fr_subsample_class_chaff, "Species_Label", fun = sum)
ps_merged_order_chaff <- phyloseq::merge_samples(microb_fr_subsample_order_chaff, "Species_Label", fun = sum)
ps_merged_phylum_chaff <- phyloseq::merge_samples(microb_fr_subsample_phylum_chaff, "Species_Label", fun = sum)
```

metadata
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

meta_merged_otu <- ps_merged_otu %>% sample_data %>% data.frame %>% dplyr::select(!Species_Label) %>% rownames_to_column("Species_Label")

rownames(meta_merged_otu) <- meta_merged_otu[,"Species_Label"]

```

microbiome distance matrix

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

## OTU
# Bray
microb_otu_tab <- otu_table(ps_merged_otu)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_otu <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_otu)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_otu <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## species
# Bray
microb_otu_tab <- otu_table(ps_merged_species)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_species <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_species)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_species <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)


## genus
# Bray
microb_otu_tab <- otu_table(ps_merged_genus)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_genus <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_genus)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_genus <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## family
# Bray
microb_otu_tab <- otu_table(ps_merged_family)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_family <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_family)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_family <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## order
# Bray
microb_otu_tab <- otu_table(ps_merged_order)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_order <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_order)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_order <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## class
# Bray
microb_otu_tab <- otu_table(ps_merged_class)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_class <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_class)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_class <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## phylum
# Bray
microb_otu_tab <- otu_table(ps_merged_phylum)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
bray_dist_phylum <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_phylum)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat),]
jaccard_dist_phylum <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

```


microbiome dist of chaff subset 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

## OTU
# Bray
microb_otu_tab <- otu_table(ps_merged_otu_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
bray_dist_otu_chaff <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_otu_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
jaccard_dist_otu_chaff <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)


## genus
# Bray
microb_otu_tab <- otu_table(ps_merged_species_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
bray_dist_species_chaff <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_species_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
jaccard_dist_species_chaff <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)


## genus
# Bray
microb_otu_tab <- otu_table(ps_merged_genus_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
bray_dist_genus_chaff <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_genus_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
jaccard_dist_genus_chaff <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## family
# Bray
microb_otu_tab <- otu_table(ps_merged_family_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
bray_dist_family_chaff <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_family_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
jaccard_dist_family_chaff <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## order
# Bray
microb_otu_tab <- otu_table(ps_merged_order_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
bray_dist_order_chaff <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_order_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
jaccard_dist_order_chaff <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## class
# Bray
microb_otu_tab <- otu_table(ps_merged_class_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
bray_dist_class_chaff <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_class_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
jaccard_dist_class_chaff <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

## phylum
# Bray
microb_otu_tab <- otu_table(ps_merged_phylum_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
bray_dist_phylum_chaff <- vegdist(microb_otu_tab_order, method = "bray")

# Jaccard 
microb_otu_tab <- otu_table(ps_merged_phylum_chaff)
microb_otu_tab_order <- microb_otu_tab[rownames(dist_mat_chaff),]
jaccard_dist_phylum_chaff <- vegdist(microb_otu_tab_order, method = "jaccard", binary = TRUE)

```

 merge diet according to subsamble 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

diet_wgs_overlap_otu <-  prune_samples(sample_names(diet_fr) %in% sample_names(microb_fr_subsample_otu), diet_fr) 
diet_wgs_overlap_species <-  prune_samples(sample_names(diet_fr) %in% sample_names(microb_fr_subsample_species), diet_fr) 
diet_wgs_overlap_genus <-  prune_samples(sample_names(diet_fr) %in% sample_names(microb_fr_subsample_genus), diet_fr) 
diet_wgs_overlap_family <-  prune_samples(sample_names(diet_fr) %in% sample_names(microb_fr_subsample_family), diet_fr) 
diet_wgs_overlap_order <-  prune_samples(sample_names(diet_fr) %in% sample_names(microb_fr_subsample_order), diet_fr) 
diet_wgs_overlap_class <-  prune_samples(sample_names(diet_fr) %in% sample_names(microb_fr_subsample_class), diet_fr) 
diet_wgs_overlap_phylum <-  prune_samples(sample_names(diet_fr) %in% sample_names(microb_fr_subsample_phylum), diet_fr) 

diet_merged_otu <- phyloseq::merge_samples(diet_wgs_overlap_otu, "Species_Label", fun = sum)
diet_merged_species <- phyloseq::merge_samples(diet_wgs_overlap_species, "Species_Label", fun = sum)
diet_merged_genus <- phyloseq::merge_samples(diet_wgs_overlap_genus, "Species_Label", fun = sum)
diet_merged_family <- phyloseq::merge_samples(diet_wgs_overlap_family, "Species_Label", fun = sum)
diet_merged_order <- phyloseq::merge_samples(diet_wgs_overlap_order, "Species_Label", fun = sum)
diet_merged_class <- phyloseq::merge_samples(diet_wgs_overlap_class, "Species_Label", fun = sum)
diet_merged_phylum <- phyloseq::merge_samples(diet_wgs_overlap_otu, "Species_Label", fun = sum)

# OTU subset 
# bray
diet_otu_tab <- otu_table(diet_merged_otu)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_bray_dist_otu <- vegdist(diet_otu_tab_order, method = "bray")

# jaccard bray
diet_otu_tab <- otu_table(diet_merged_otu)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_jaccard_dist_otu <- vegdist(diet_otu_tab_order, method = "jaccard", binary = TRUE)

# species subset 
# bray
diet_otu_tab <- otu_table(diet_merged_species)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_bray_dist_species <- vegdist(diet_otu_tab_order, method = "bray")

# jaccard bray
diet_otu_tab <- otu_table(diet_merged_species)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_jaccard_dist_species <- vegdist(diet_otu_tab_order, method = "jaccard", binary = TRUE)


# genus subset 
# bray
diet_otu_tab <- otu_table(diet_merged_genus)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_bray_dist_genus <- vegdist(diet_otu_tab_order, method = "bray")

# jaccard bray
diet_otu_tab <- otu_table(diet_merged_genus)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_jaccard_dist_genus <- vegdist(diet_otu_tab_order, method = "jaccard", binary = TRUE)

# family subset 
# bray
diet_otu_tab <- otu_table(diet_merged_family)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_bray_dist_family <- vegdist(diet_otu_tab_order, method = "bray")

# jaccard bray
diet_otu_tab <- otu_table(diet_merged_family)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_jaccard_dist_family <- vegdist(diet_otu_tab_order, method = "jaccard", binary = TRUE)

# class subset 
# bray
diet_otu_tab <- otu_table(diet_merged_class)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_bray_dist_class <- vegdist(diet_otu_tab_order, method = "bray")

# jaccard bray
diet_otu_tab <- otu_table(diet_merged_class)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_jaccard_dist_class <- vegdist(diet_otu_tab_order, method = "jaccard", binary = TRUE)

# phylum subset 
# bray
diet_otu_tab <- otu_table(diet_merged_phylum)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_bray_dist_phylum <- vegdist(diet_otu_tab_order, method = "bray")

# jaccard bray
diet_otu_tab <- otu_table(diet_merged_phylum)
diet_otu_tab_order <- diet_otu_tab[rownames(dist_mat),]
diet_jaccard_dist_phylum <- vegdist(diet_otu_tab_order, method = "jaccard", binary = TRUE)


```



### Mantel tests

Mantel Bray-Curtis 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_bray_fst_otu <- vegan::mantel(bray_dist_otu, fst_dist, method = "pearson", permutations = 9999))
str(mantel_bray_fst_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_fst_otu$statistic, p=mantel_bray_fst_otu$signif)
#Mantel statistic based on Pearson's product-moment correlation 
#
#Call:
#vegan::mantel(xdis = bray_dist_otu, ydis = fst_dist, method = "pearson",      #permutations = 9999) 
#
#Mantel statistic r: 0.311 
#      Significance: 0.0288 
#
#Upper quantiles of permutations (null model):
#  90%   95% 97.5%   99% 
#0.224 0.275 0.318 0.373 
#Permutation: free
#Number of permutations: 9999


(mantel_bray_fst_otu_chaff <- vegan::mantel(bray_dist_otu_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))
df_otu_chaff <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_fst_otu$statistic, p=mantel_bray_fst_otu$signif)



(mantel_bray_geo_otu  <- vegan::mantel.partial(bray_dist_otu, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))
df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)

#Partial Mantel statistic based on Pearson's product-moment correlation 
#
#Call:
#vegan::mantel.partial(xdis = bray_dist_otu, ydis = fst_dist,      zdis = as.dist(#log10(geo_dist + 1)), method = "pearson",      permutations = 9999) 
#
#Mantel statistic r: 0.3216 
#      Significance: 0.0309 
#
#Upper quantiles of permutations (null model):
#  90%   95% 97.5%   99% 
#0.226 0.282 0.338 0.397 
#Permutation: free
#Number of permutations: 9999



(mantel_bray_geo_otu_chaff  <- vegan::mantel.partial(bray_dist_otu_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))
df_otu_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu_chaff$statistic, p=mantel_bray_geo_otu_chaff$signif)



vegan::mantel(fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999)

#Mantel statistic r: -0.1602 
#      Significance: 0.7297 



vegan::mantel(fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999)
#Mantel statistic r: 0.6181 
#      Significance: 4e-04 
#
#Upper quantiles of permutations (null model):
#  90%   95% 97.5%   99% 
#0.295 0.349 0.398 0.457 
#Permutation: free
#Number of permutations: 9999

bray = as.vector(bray_dist_otu)
jacc = as.vector(jaccard_dist_otu)
fst = as.vector(fst_dist)
geo = as.vector(geo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(bray, jacc, fst, geo)
#mat_edit <- read.csv(paste0(dataDir, "/fst_dist_d


(p1 <- ggplot(mat, aes(y = bray, x = fst)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Fst", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_fst_bray$statistic, 3), ";p-value=", round(mantel_fst_bray$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

(p2 <- ggplot(mat, aes(y = fst, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "Fst") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))


bray = as.vector(bray_dist_otu_chaff)
jacc = as.vector(jaccard_dist_otu_chaff)
fst = as.vector(fst_dist_chaff)
geo = as.vector(geo_dist_chaff)
#new data frame with vectorized distance matrices
mat_chaff = data.frame(bray, jacc, fst, geo)
(p3 <- ggplot(mat_chaff, aes(y = bray, x = fst)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Fst", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_fst_bray$statistic, 3), ";p-value=", round(mantel_fst_bray$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

png(paste0(outputDir, "/01_Exploration/fst_dist_chaff_speciesFix.png"), width = 400, height = 250, units = "mm", res = 200)
(p4 <- ggplot(mat_chaff, aes(y = fst, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "Fst") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))
dev.off()



# species
(mantel_bray_fst_species <- vegan::mantel(bray_dist_species, fst_dist, method = "pearson", permutations = 9999))

(mantel_bray_fst_species_chaff <- vegan::mantel(bray_dist_species_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_bray_geo_species  <- vegan::mantel.partial(bray_dist_species, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_bray_geo_species_chaff  <- vegan::mantel.partial(bray_dist_species_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))



# genus
(mantel_bray_fst_genus <- vegan::mantel(bray_dist_genus, fst_dist, method = "pearson", permutations = 9999))

(mantel_bray_fst_genus_chaff <- vegan::mantel(bray_dist_genus_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_bray_geo_genus  <- vegan::mantel.partial(bray_dist_genus, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_bray_geo_genus_chaff  <- vegan::mantel.partial(bray_dist_genus_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))

# family
(mantel_bray_fst_family <- vegan::mantel(bray_dist_family, fst_dist, method = "pearson", permutations = 9999))

(mantel_bray_fst_family_chaff <- vegan::mantel(bray_dist_family_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_bray_geo_family  <- vegan::mantel.partial(bray_dist_family, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_bray_geo_family_chaff  <- vegan::mantel.partial(bray_dist_family_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))


# order
(mantel_bray_fst_order <- vegan::mantel(bray_dist_order, fst_dist, method = "pearson", permutations = 9999))

(mantel_bray_fst_order_chaff <- vegan::mantel(bray_dist_order_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_bray_geo_order  <- vegan::mantel.partial(bray_dist_order, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_bray_geo_order_chaff  <- vegan::mantel.partial(bray_dist_order_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))

# class
(mantel_bray_fst_class <- vegan::mantel(bray_dist_class, fst_dist, method = "pearson", permutations = 9999))

(mantel_bray_fst_class_chaff <- vegan::mantel(bray_dist_class_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_bray_geo_class  <- vegan::mantel.partial(bray_dist_class, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_bray_geo_class_chaff  <- vegan::mantel.partial(bray_dist_class_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))

# phylum
(mantel_bray_fst_phylum <- vegan::mantel(bray_dist_phylum, fst_dist, method = "pearson", permutations = 9999))

(mantel_bray_fst_phylum_chaff <- vegan::mantel(bray_dist_phylum_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_bray_geo_phylum  <- vegan::mantel.partial(bray_dist_phylum, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_bray_geo_phylum_chaff  <- vegan::mantel.partial(bray_dist_phylum_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))


# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_fst_otu$statistic, p=mantel_bray_fst_otu$signif)

df_otu_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_fst_otu$statistic, p=mantel_bray_fst_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu$statistic, p=mantel_bray_geo_otu$signif)

df_otu_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_bray_geo_otu_chaff$statistic, p=mantel_bray_geo_otu_chaff$signif)


# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_fst_species$statistic, p=mantel_bray_fst_species$signif)

df_species_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_fst_species_chaff$statistic, p=mantel_bray_fst_species_chaff$signif)

df_species_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_geo_species$statistic, p=mantel_bray_geo_species$signif)

df_species_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Species", dist="Bray-Curtis", r=mantel_bray_geo_species_chaff$statistic, p=mantel_bray_geo_species_chaff$signif)



# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_fst_genus$statistic, p=mantel_bray_fst_genus$signif)

df_genus_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_fst_genus_chaff$statistic, p=mantel_bray_fst_genus_chaff$signif)

df_genus_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_geo_genus$statistic, p=mantel_bray_geo_genus$signif)

df_genus_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Genus", dist="Bray-Curtis", r=mantel_bray_geo_genus_chaff$statistic, p=mantel_bray_geo_genus_chaff$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_fst_family$statistic, p=mantel_bray_fst_family$signif)

df_family_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_fst_family_chaff$statistic, p=mantel_bray_fst_family_chaff$signif)

df_family_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_geo_family$statistic, p=mantel_bray_geo_family$signif)

df_family_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Family", dist="Bray-Curtis", r=mantel_bray_geo_family_chaff$statistic, p=mantel_bray_geo_family_chaff$signif)

# order

df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_fst_order$statistic, p=mantel_bray_fst_order$signif)

df_order_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_fst_order_chaff$statistic, p=mantel_bray_fst_order_chaff$signif)

df_order_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_geo_order$statistic, p=mantel_bray_geo_order$signif)

df_order_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Order", dist="Bray-Curtis", r=mantel_bray_geo_order_chaff$statistic, p=mantel_bray_geo_order_chaff$signif)


# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_fst_class$statistic, p=mantel_bray_fst_class$signif)

df_class_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_fst_class_chaff$statistic, p=mantel_bray_fst_class_chaff$signif)

df_class_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_geo_class$statistic, p=mantel_bray_geo_class$signif)

df_class_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Class", dist="Bray-Curtis", r=mantel_bray_geo_class_chaff$statistic, p=mantel_bray_geo_class_chaff$signif)


# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_fst_phylum$statistic, p=mantel_bray_fst_phylum$signif)

df_phylum_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_fst_phylum_chaff$statistic, p=mantel_bray_fst_phylum_chaff$signif)

df_phylum_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_geo_phylum$statistic, p=mantel_bray_geo_phylum$signif)

df_phylum_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Phylum", dist="Bray-Curtis", r=mantel_bray_geo_phylum_chaff$statistic, p=mantel_bray_geo_phylum_chaff$signif)



mantel_out_bray <- rbind(df_otu, df_otu_chaff, df_otu_partial, df_otu_chaff_partial, df_species, df_species_chaff, df_species_partial, df_species_chaff_partial, df_genus, df_genus_chaff, df_genus_partial, df_genus_chaff_partial, df_family, df_family_chaff, df_family_partial, df_family_chaff_partial, df_order, df_order_chaff, df_order_partial, df_order_chaff_partial, df_class, df_class_chaff, df_class_partial, df_class_chaff_partial, df_phylum, df_phylum_chaff, df_phylum_partial, df_phylum_chaff_partial)



```


Mantel Jaccard
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(mantel_jaccard_fst_otu <- vegan::mantel(jaccard_dist_otu, fst_dist, method = "pearson", permutations = 9999))
str(mantel_jaccard_fst_otu)
df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_fst_otu$statistic, p=mantel_jaccard_fst_otu$signif)
#Call:
#vegan::mantel(xdis = jaccard_dist_otu, ydis = fst_dist, #method = "pearson",      permutations = 9999) 
#
#Mantel statistic r: -0.1776 
#      Significance: 0.7905 
#
#Upper quantiles of permutations (null model):
#  90%   95% 97.5%   99% 
#0.272 0.356 0.442 0.544 

(mantel_jaccard_fst_otu_chaff <- vegan::mantel(jaccard_dist_otu_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))
df_otu_chaff <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_fst_otu$statistic, p=mantel_jaccard_fst_otu$signif)

#vegan::mantel(xdis = jaccard_dist_otu_chaff, ydis = #fst_dist_chaff,      method = "pearson", permutations = #9999) 
#
#Mantel statistic r: 0.186 
#      Significance: 0.2027 
#
#Upper quantiles of permutations (null model):
#  90%   95% 97.5%   99% 
#0.287 0.360 0.424 0.486 


(mantel_jaccard_geo_otu  <- vegan::mantel.partial(jaccard_dist_otu, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))
df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

#vegan::mantel.partial(xdis = jaccard_dist_otu, ydis = #fst_dist,      zdis = as.dist(log10(geo_dist + 1)), method# = "pearson",      permutations = 9999) 
#
#Mantel statistic r: -0.04104 
#      Significance: 0.5445 
#
#Upper quantiles of permutations (null model):
#  90%   95% 97.5%   99% 
#0.268 0.349 0.438 0.537  

(mantel_jaccard_geo_otu_chaff  <- vegan::mantel.partial(jaccard_dist_otu_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))
df_otu_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "OTU99", dist="Bray-Curtis", r=mantel_jaccard_geo_otu_chaff$statistic, p=mantel_jaccard_geo_otu_chaff$signif)
#vegan::mantel.partial(xdis = jaccard_dist_otu_chaff, ydis = #fst_dist_chaff,      zdis = as.dist(log10(geo_dist_chaff +# 1)), method = "pearson",      permutations = 9999) 
#
#Mantel statistic r: -0.3766 
#      Significance: 0.9204 
#
#Upper quantiles of permutations (null model):
#  90%   95% 97.5%   99% 
#0.320 0.390 0.440 0.481 


jacc = as.vector(jaccard_dist_otu)
fst = as.vector(fst_dist)
geo = as.vector(geo_dist)
#new data frame with vectorized distance matrices
mat = data.frame(jacc, fst, geo)
#mat_edit <- read.csv(paste0(dataDir, "/fst_dist_d


(p1 <- ggplot(mat, aes(y = jacc, x = fst)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Fst", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle("")) 
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_fst_jaccard$statistic, 3), ";p-value=", round(mantel_fst_jaccard$signif, 3)), parse=F,x=0.4,y=0.4,size=10))

(p2 <- ggplot(mat, aes(y = fst, x = geo)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Geographic distance", y = "Fst") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))


jacc = as.vector(jaccard_dist_otu_chaff)
fst = as.vector(fst_dist_chaff)
geo = as.vector(geo_dist_chaff)
#new data frame with vectorized distance matrices
mat_chaff = data.frame(jacc, fst, geo)
(p3 <- ggplot(mat_chaff, aes(y = jacc, x = fst)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Fst", y = "Beta diversity (Bray distance)") +
        geom_smooth(method = "lm", colour = "red", alpha = 0.2,lwd=3) +
        theme_classic(base_size = 31)+
        theme(plot.title = element_text(face = "bold"), panel.border = element_rect(fill = NA, colour = "black"),plot.margin=grid::unit(c(5,5,5,5), "mm"))+
        ggtitle(""))
        #+ 
        #annotate(geom="text",label=paste0("Mantel statistic r =", round(mantel_fst_jaccard$statistic, 3), ";p-value=", round(mantel_fst_jaccard$signif, 3)), parse=F,x=0.4,y=0.4,size=10))



# species
(mantel_jaccard_fst_species <- vegan::mantel(jaccard_dist_species, fst_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_fst_species_chaff <- vegan::mantel(jaccard_dist_species_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_species  <- vegan::mantel.partial(jaccard_dist_species, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_jaccard_geo_species_chaff  <- vegan::mantel.partial(jaccard_dist_species_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))



# genus
(mantel_jaccard_fst_genus <- vegan::mantel(jaccard_dist_genus, fst_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_fst_genus_chaff <- vegan::mantel(jaccard_dist_genus_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_genus  <- vegan::mantel.partial(jaccard_dist_genus, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_jaccard_geo_genus_chaff  <- vegan::mantel.partial(jaccard_dist_genus_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))

# family
(mantel_jaccard_fst_family <- vegan::mantel(jaccard_dist_family, fst_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_fst_family_chaff <- vegan::mantel(jaccard_dist_family_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_family  <- vegan::mantel.partial(jaccard_dist_family, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_jaccard_geo_family_chaff  <- vegan::mantel.partial(jaccard_dist_family_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))


# class
(mantel_jaccard_fst_class <- vegan::mantel(jaccard_dist_class, fst_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_fst_class_chaff <- vegan::mantel(jaccard_dist_class_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_class  <- vegan::mantel.partial(jaccard_dist_class, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_jaccard_geo_class_chaff  <- vegan::mantel.partial(jaccard_dist_class_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))


# order
(mantel_jaccard_fst_order <- vegan::mantel(jaccard_dist_order, fst_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_fst_order_chaff <- vegan::mantel(jaccard_dist_order_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_order  <- vegan::mantel.partial(jaccard_dist_order, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_jaccard_geo_order_chaff  <- vegan::mantel.partial(jaccard_dist_order_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))

# phylum
(mantel_jaccard_fst_phylum <- vegan::mantel(jaccard_dist_phylum, fst_dist, method = "pearson", permutations = 9999))

(mantel_jaccard_fst_phylum_chaff <- vegan::mantel(jaccard_dist_phylum_chaff, fst_dist_chaff, method = "pearson", permutations = 9999))

(mantel_jaccard_geo_phylum  <- vegan::mantel.partial(jaccard_dist_phylum, fst_dist, as.dist(log10(geo_dist+1)), method = "pearson", permutations = 9999))

(mantel_jaccard_geo_phylum_chaff  <- vegan::mantel.partial(jaccard_dist_phylum_chaff, fst_dist_chaff, as.dist(log10(geo_dist_chaff+1)), method = "pearson", permutations = 9999))


# output df 

# OTU99

df_otu <- data.frame(test="mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_fst_otu$statistic, p=mantel_jaccard_fst_otu$signif)

df_otu_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_fst_otu$statistic, p=mantel_jaccard_fst_otu$signif)

df_otu_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_geo_otu$statistic, p=mantel_jaccard_geo_otu$signif)

df_otu_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "OTU99", dist="Jaccard", r=mantel_jaccard_geo_otu_chaff$statistic, p=mantel_jaccard_geo_otu_chaff$signif)


# species
df_species <- data.frame(test="mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_fst_species$statistic, p=mantel_jaccard_fst_species$signif)

df_species_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_fst_species_chaff$statistic, p=mantel_jaccard_fst_species_chaff$signif)

df_species_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_geo_species$statistic, p=mantel_jaccard_geo_species$signif)

df_species_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Species", dist="Jaccard", r=mantel_jaccard_geo_species_chaff$statistic, p=mantel_jaccard_geo_species_chaff$signif)


# genus
df_genus <- data.frame(test="mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_fst_genus$statistic, p=mantel_jaccard_fst_genus$signif)

df_genus_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_fst_genus_chaff$statistic, p=mantel_jaccard_fst_genus_chaff$signif)

df_genus_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_geo_genus$statistic, p=mantel_jaccard_geo_genus$signif)

df_genus_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Genus", dist="Jaccard", r=mantel_jaccard_geo_genus_chaff$statistic, p=mantel_jaccard_geo_genus_chaff$signif)

# family
df_family <- data.frame(test="mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_fst_family$statistic, p=mantel_jaccard_fst_family$signif)

df_family_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_fst_family_chaff$statistic, p=mantel_jaccard_fst_family_chaff$signif)

df_family_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_geo_family$statistic, p=mantel_jaccard_geo_family$signif)

df_family_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Family", dist="Jaccard", r=mantel_jaccard_geo_family_chaff$statistic, p=mantel_jaccard_geo_family_chaff$signif)

# order

df_order <- data.frame(test="mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_fst_order$statistic, p=mantel_jaccard_fst_order$signif)

df_order_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_fst_order_chaff$statistic, p=mantel_jaccard_fst_order_chaff$signif)

df_order_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_geo_order$statistic, p=mantel_jaccard_geo_order$signif)

df_order_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Order", dist="Jaccard", r=mantel_jaccard_geo_order_chaff$statistic, p=mantel_jaccard_geo_order_chaff$signif)


# class
df_class <- data.frame(test="mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_fst_class$statistic, p=mantel_jaccard_fst_class$signif)

df_class_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_fst_class_chaff$statistic, p=mantel_jaccard_fst_class_chaff$signif)

df_class_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_geo_class$statistic, p=mantel_jaccard_geo_class$signif)

df_class_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Class", dist="Jaccard", r=mantel_jaccard_geo_class_chaff$statistic, p=mantel_jaccard_geo_class_chaff$signif)


# phylum
df_phylum <- data.frame(test="mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_fst_phylum$statistic, p=mantel_jaccard_fst_phylum$signif)

df_phylum_chaff <- data.frame(test="mantel", data = "chaffinch", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_fst_phylum_chaff$statistic, p=mantel_jaccard_fst_phylum_chaff$signif)

df_phylum_partial <- data.frame(test="partial-mantel", data = "all", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_geo_phylum$statistic, p=mantel_jaccard_geo_phylum$signif)

df_phylum_chaff_partial <- data.frame(test="partial-mantel", data = "chaffinch", taxonomy = "Phylum", dist="Jaccard", r=mantel_jaccard_geo_phylum_chaff$statistic, p=mantel_jaccard_geo_phylum_chaff$signif)



mantel_out_jaccard <- rbind(df_otu, df_otu_chaff, df_otu_partial, df_otu_chaff_partial, df_species, df_species_chaff, df_species_partial, df_species_chaff_partial, df_genus, df_genus_chaff, df_genus_partial, df_genus_chaff_partial, df_family, df_family_chaff, df_family_partial, df_family_chaff_partial, df_order, df_order_chaff, df_order_partial, df_order_chaff_partial, df_class, df_class_chaff, df_class_partial, df_class_chaff_partial, df_phylum, df_phylum_chaff, df_phylum_partial, df_phylum_chaff_partial)


```

Join bray and jaccard output

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

mantel_out <- rbind(mantel_out_bray, mantel_out_jaccard)
head(mantel_out)

write.csv(mantel_out, paste0(dataDir, "/output_mantel_weightedfst.csv"), row.names=FALSE)

```

## Calculate divergence

microbiome distance matrix

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

## OTU
# Bray
microb_otu_tab <- otu_table(microb_overlap_otu)
bray_dist_ind_otu <- vegdist(microb_otu_tab, method = "bray")
str(bray_dist_ind_otu)
bray_divergence_otu <- betadisper(bray_dist_ind_otu, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_otu)[2] <- "b.beta_median"


# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_otu)
jaccard_dist_ind_otu <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_otu <- betadisper(jaccard_dist_ind_otu, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_otu)[2] <- "j.beta_median"


## species
# Bray
microb_otu_tab <- otu_table(microb_overlap_species)
bray_dist_ind_species <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_species <- betadisper(bray_dist_ind_species, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_species)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_species)
jaccard_dist_ind_species <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_species <- betadisper(jaccard_dist_ind_species, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_species)[2] <- "j.beta_median"



## genus
# Bray
microb_otu_tab <- otu_table(microb_overlap_genus)
bray_dist_ind_genus <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_genus <- betadisper(bray_dist_ind_genus, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_genus)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_genus)
jaccard_dist_ind_genus <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_genus <- betadisper(jaccard_dist_ind_genus, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_genus)[2] <- "j.beta_median"

## family
# Bray
microb_otu_tab <- otu_table(microb_overlap_family)
bray_dist_ind_family <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_family <- betadisper(bray_dist_ind_family, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_family)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_family)
jaccard_dist_ind_family <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_family <- betadisper(jaccard_dist_ind_family, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_family)[2] <- "j.beta_median"

## order
# Bray
microb_otu_tab <- otu_table(microb_overlap_order)
bray_dist_ind_order <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_order <- betadisper(bray_dist_ind_order, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_order)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_order)
jaccard_dist_ind_order <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_order <- betadisper(jaccard_dist_ind_order, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_order)[2] <- "j.beta_median"

## class
# Bray
microb_otu_tab <- otu_table(microb_overlap_class)
bray_dist_ind_class <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_class <- betadisper(bray_dist_ind_class, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_class)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_class)
jaccard_dist_ind_class <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_class <- betadisper(jaccard_dist_ind_class, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_class)[2] <- "j.beta_median"

## phylum
# Bray
microb_otu_tab <- otu_table(microb_overlap_phylum)
bray_dist_ind_phylum <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_phylum <- betadisper(bray_dist_ind_phylum, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_phylum)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_phylum)
jaccard_dist_ind_phylum <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_phylum <- betadisper(jaccard_dist_ind_phylum, microb_overlap_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_phylum)[2] <- "j.beta_median"

```


microbiome dist of chaff subset 

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

## OTU
# Bray
microb_otu_tab <- otu_table(microb_overlap_chaff_otu)
bray_dist_ind_chaff_otu <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_chaff_otu <- betadisper(bray_dist_ind_chaff_otu, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_chaff_otu)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_chaff_otu)
jaccard_dist_ind_chaff_otu <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_chaff_otu <- betadisper(jaccard_dist_ind_chaff_otu, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_chaff_otu)[2] <- "j.beta_median"


## species
# Bray
microb_otu_tab <- otu_table(microb_overlap_chaff_species)
bray_dist_ind_chaff_species <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_chaff_species <- betadisper(bray_dist_ind_chaff_species, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_chaff_species)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_chaff_species)
jaccard_dist_ind_chaff_species <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_chaff_species <- betadisper(jaccard_dist_ind_chaff_species, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_chaff_species)[2] <- "j.beta_median"



## genus
# Bray
microb_otu_tab <- otu_table(microb_overlap_chaff_genus)
bray_dist_ind_chaff_genus <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_chaff_genus <- betadisper(bray_dist_ind_chaff_genus, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_chaff_genus)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_chaff_genus)
jaccard_dist_ind_chaff_genus <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_chaff_genus <- betadisper(jaccard_dist_ind_chaff_genus, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_chaff_genus)[2] <- "j.beta_median"

## family
# Bray
microb_otu_tab <- otu_table(microb_overlap_chaff_family)
bray_dist_ind_chaff_family <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_chaff_family <- betadisper(bray_dist_ind_chaff_family, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_chaff_family)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_chaff_family)
jaccard_dist_ind_chaff_family <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_chaff_family <- betadisper(jaccard_dist_ind_chaff_family, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_chaff_family)[2] <- "j.beta_median"

## order
# Bray
microb_otu_tab <- otu_table(microb_overlap_chaff_order)
bray_dist_ind_chaff_order <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_chaff_order <- betadisper(bray_dist_ind_chaff_order, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_chaff_order)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_chaff_order)
jaccard_dist_ind_chaff_order <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_chaff_order <- betadisper(jaccard_dist_ind_chaff_order, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_chaff_order)[2] <- "j.beta_median"

## class
# Bray
microb_otu_tab <- otu_table(microb_overlap_chaff_class)
bray_dist_ind_chaff_class <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_chaff_class <- betadisper(bray_dist_ind_chaff_class, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_chaff_class)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_chaff_class)
jaccard_dist_ind_chaff_class <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_chaff_class <- betadisper(jaccard_dist_ind_chaff_class, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_chaff_class)[2] <- "j.beta_median"

## phylum
# Bray
microb_otu_tab <- otu_table(microb_overlap_chaff_phylum)
bray_dist_ind_chaff_phylum <- vegdist(microb_otu_tab, method = "bray")
bray_divergence_chaff_phylum <- betadisper(bray_dist_ind_chaff_phylum, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(bray_divergence_chaff_phylum)[2] <- "b.beta_median"

# Jaccard 
microb_otu_tab <- otu_table(microb_overlap_chaff_phylum)
jaccard_dist_ind_chaff_phylum <- vegdist(microb_otu_tab, method = "jaccard", binary = TRUE)
jaccard_divergence_chaff_phylum <- betadisper(jaccard_dist_ind_chaff_phylum, microb_overlap_chaff_meta$Species_Label, type = "median")$distance %>% data.frame %>% rownames_to_column("TipID_microbiome")
names(jaccard_divergence_chaff_phylum)[2] <- "j.beta_median"

```


Heterozygosity data

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

metadata <- read_delim(paste0(dataDir, "/WGS_Fringilla_samples.tsv"), delim = ",")

ind_het <- read_delim(paste0(dataDir, "/fringilla_variants_microbiome.het"), delim = "\t",
           col_names = c("TipID_microbiome","ho", "he", "nsites", "f"), skip = 1)

head(metadata$TipID_microbiome, 30)
names(metadata) 
ind_het

ind_het_meta <- left_join(ind_het, metadata, by ="TipID_microbiome")
ind_het_meta %>% data.frame %>% head


#froh
froh <- read_delim(paste0(dataDir, "/fringilla_microbiome_roh_rg_filtered.csv"), delim = ",")
names(froh) <- c("TipID_microbiome", "total_length_bp", "froh")


froh_meta <- left_join(froh, metadata, by ="TipID_microbiome")
froh_meta %>% data.frame %>% head


```


Join df

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

froh_meta %>% data.frame %>% head(2)
froh_meta$Region_Species

# OTU
roh_bray_otu <- left_join(froh_meta, bray_divergence_otu, by = "TipID_microbiome")
head(roh_bray_otu)
roh_bray_otu <- subset(roh_bray_otu, subset = !is.na(roh_bray_otu$b.beta_median))

roh_jacc_otu <- left_join(froh_meta, jaccard_divergence_otu, by = "TipID_microbiome")
roh_jacc_otu <- subset(roh_jacc_otu, subset = !is.na(roh_jacc_otu$j.beta_median))


# species
roh_bray_species <- left_join(froh_meta, bray_divergence_species, by = "TipID_microbiome")
roh_bray_species <- subset(roh_bray_species, subset = !is.na(roh_bray_species$b.beta_median))

roh_jacc_species <- left_join(froh_meta, jaccard_divergence_species, by = "TipID_microbiome")
roh_jacc_species <- subset(roh_jacc_species, subset = !is.na(roh_jacc_species$j.beta_median))



# genus
roh_bray_genus <- left_join(froh_meta, bray_divergence_genus, by = "TipID_microbiome")
roh_bray_genus <- subset(roh_bray_genus, subset = !is.na(roh_bray_genus$b.beta_median))

roh_jacc_genus <- left_join(froh_meta, jaccard_divergence_genus, by = "TipID_microbiome")
roh_jacc_genus <- subset(roh_jacc_genus, subset = !is.na(roh_jacc_genus$j.beta_median))

# family
roh_bray_family <- left_join(froh_meta, bray_divergence_family, by = "TipID_microbiome")
roh_bray_family <- subset(roh_bray_family, subset = !is.na(roh_bray_family$b.beta_median))

roh_jacc_family <- left_join(froh_meta, jaccard_divergence_family, by = "TipID_microbiome")
roh_jacc_family <- subset(roh_jacc_family, subset = !is.na(roh_jacc_family$j.beta_median))

# order
roh_bray_order <- left_join(froh_meta, bray_divergence_order, by = "TipID_microbiome")
roh_bray_order <- subset(roh_bray_order, subset = !is.na(roh_bray_order$b.beta_median))

roh_jacc_order <- left_join(froh_meta, jaccard_divergence_order, by = "TipID_microbiome")
roh_jacc_order <- subset(roh_jacc_order, subset = !is.na(roh_jacc_order$j.beta_median))

# class
roh_bray_class <- left_join(froh_meta, bray_divergence_class, by = "TipID_microbiome")
roh_bray_class <- subset(roh_bray_class, subset = !is.na(roh_bray_class$b.beta_median))

roh_jacc_class <- left_join(froh_meta, jaccard_divergence_class, by = "TipID_microbiome")
roh_jacc_class <- subset(roh_jacc_class, subset = !is.na(roh_jacc_class$j.beta_median))

# phylum
roh_bray_phylum <- left_join(froh_meta, bray_divergence_phylum, by = "TipID_microbiome")
roh_bray_phylum <- subset(roh_bray_phylum, subset = !is.na(roh_bray_phylum$b.beta_median))

roh_jacc_phylum <- left_join(froh_meta, jaccard_divergence_phylum, by = "TipID_microbiome")
roh_jacc_phylum <- subset(roh_jacc_phylum, subset = !is.na(roh_jacc_phylum$j.beta_median))

```

Join chaff df

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# OTU
roh_bray_chaff_otu <- left_join(froh_meta, bray_divergence_chaff_otu, by = "TipID_microbiome")
roh_bray_chaff_otu <- subset(roh_bray_chaff_otu, subset = !is.na(roh_bray_chaff_otu$b.beta_median))

roh_jacc_chaff_otu <- left_join(froh_meta, jaccard_divergence_chaff_otu, by = "TipID_microbiome")
roh_jacc_chaff_otu <- subset(roh_jacc_chaff_otu, subset = !is.na(roh_jacc_chaff_otu$j.beta_median))


# species
roh_bray_chaff_species <- left_join(froh_meta, bray_divergence_chaff_species, by = "TipID_microbiome")
roh_bray_chaff_species <- subset(roh_bray_chaff_species, subset = !is.na(roh_bray_chaff_species$b.beta_median))

roh_jacc_chaff_species <- left_join(froh_meta, jaccard_divergence_chaff_species, by = "TipID_microbiome")
roh_jacc_chaff_species <- subset(roh_jacc_chaff_species, subset = !is.na(roh_jacc_chaff_species$j.beta_median))



# genus
roh_bray_chaff_genus <- left_join(froh_meta, bray_divergence_chaff_genus, by = "TipID_microbiome")
roh_bray_chaff_genus <- subset(roh_bray_chaff_genus, subset = !is.na(roh_bray_chaff_genus$b.beta_median))

roh_jacc_chaff_genus <- left_join(froh_meta, jaccard_divergence_chaff_genus, by = "TipID_microbiome")
roh_jacc_chaff_genus <- subset(roh_jacc_chaff_genus, subset = !is.na(roh_jacc_chaff_genus$j.beta_median))

# family
roh_bray_chaff_family <- left_join(froh_meta, bray_divergence_chaff_family, by = "TipID_microbiome")
roh_bray_chaff_family <- subset(roh_bray_chaff_family, subset = !is.na(roh_bray_chaff_family$b.beta_median))

roh_jacc_chaff_family <- left_join(froh_meta, jaccard_divergence_chaff_family, by = "TipID_microbiome")
roh_jacc_chaff_family <- subset(roh_jacc_chaff_family, subset = !is.na(roh_jacc_chaff_family$j.beta_median))

# order
roh_bray_chaff_order <- left_join(froh_meta, bray_divergence_chaff_order, by = "TipID_microbiome")
roh_bray_chaff_order <- subset(roh_bray_chaff_order, subset = !is.na(roh_bray_chaff_order$b.beta_median))

roh_jacc_chaff_order <- left_join(froh_meta, jaccard_divergence_chaff_order, by = "TipID_microbiome")
roh_jacc_chaff_order <- subset(roh_jacc_chaff_order, subset = !is.na(roh_jacc_chaff_order$j.beta_median))

# class
roh_bray_chaff_class <- left_join(froh_meta, bray_divergence_chaff_class, by = "TipID_microbiome")
roh_bray_chaff_class <- subset(roh_bray_chaff_class, subset = !is.na(roh_bray_chaff_class$b.beta_median))

roh_jacc_chaff_class <- left_join(froh_meta, jaccard_divergence_chaff_class, by = "TipID_microbiome")
roh_jacc_chaff_class <- subset(roh_jacc_chaff_class, subset = !is.na(roh_jacc_chaff_class$j.beta_median))

# phylum
roh_bray_chaff_phylum <- left_join(froh_meta, bray_divergence_chaff_phylum, by = "TipID_microbiome")
roh_bray_chaff_phylum <- subset(roh_bray_chaff_phylum, subset = !is.na(roh_bray_chaff_phylum$b.beta_median))

roh_jacc_chaff_phylum <- left_join(froh_meta, jaccard_divergence_chaff_phylum, by = "TipID_microbiome")
roh_jacc_chaff_phylum <- subset(roh_jacc_chaff_phylum, subset = !is.na(roh_jacc_chaff_phylum$j.beta_median))

```

glmm test bray

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
library(glmmTMB)
library(car)

# OTU
glmm_bray_otu <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_otu, na.action = na.fail, family = gaussian)
summary(glmm_bray_otu)
confint(glmm_bray_otu)

glmm_bray_chaff_otu <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_chaff_otu, na.action = na.fail, family = gaussian)
confint(glmm_bray_chaff_otu)


# species
glmm_bray_species <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_species, na.action = na.fail, family = gaussian)
confint(glmm_bray_species)

glmm_bray_chaff_species <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_chaff_species, na.action = na.fail, family = gaussian)
confint(glmm_bray_chaff_species)


# genus
glmm_bray_genus <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_genus, na.action = na.fail, family = gaussian)
confint(glmm_bray_genus)

glmm_bray_chaff_genus <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_chaff_genus, na.action = na.fail, family = gaussian)
confint(glmm_bray_chaff_genus)

# family
glmm_bray_family <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_family, na.action = na.fail, family = gaussian)
confint(glmm_bray_family)

glmm_bray_chaff_family <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_chaff_family, na.action = na.fail, family = gaussian)
confint(glmm_bray_chaff_family)

# order
glmm_bray_order <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_order, na.action = na.fail, family = gaussian)
confint(glmm_bray_order)

glmm_bray_chaff_order <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_chaff_order, na.action = na.fail, family = gaussian)
confint(glmm_bray_chaff_order)

# class
glmm_bray_class <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_class, na.action = na.fail, family = gaussian)
confint(glmm_bray_class)

glmm_bray_chaff_class <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_chaff_class, na.action = na.fail, family = gaussian)
confint(glmm_bray_chaff_class)

# phylum
glmm_bray_phylum <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_phylum, na.action = na.fail, family = gaussian)
confint(glmm_bray_phylum)

glmm_bray_chaff_phylum <- glmmTMB(b.beta_median ~  scale(froh) + (1|Species_Label), data = roh_bray_chaff_phylum, na.action = na.fail, family = gaussian)
confint(glmm_bray_chaff_phylum)

```

glmm test jaccard

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# OTU
glmm_jacc_otu <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_otu, na.action = na.fail, family = gaussian)
confint(glmm_jacc_otu)

glmm_jacc_chaff_otu <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_chaff_otu, na.action = na.fail, family = gaussian)
confint(glmm_jacc_chaff_otu)


# species
glmm_jacc_species <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_species, na.action = na.fail, family = gaussian)
confint(glmm_jacc_species)

glmm_jacc_chaff_species <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_chaff_species, na.action = na.fail, family = gaussian)
confint(glmm_jacc_chaff_species)



# genus
glmm_jacc_genus <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_genus, na.action = na.fail, family = gaussian)
confint(glmm_jacc_genus)

glmm_jacc_chaff_genus <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_chaff_genus, na.action = na.fail, family = gaussian)
confint(glmm_jacc_chaff_genus)

# family
glmm_jacc_family <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_family, na.action = na.fail, family = gaussian)
confint(glmm_jacc_family)

glmm_jacc_chaff_family <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_chaff_family, na.action = na.fail, family = gaussian)
confint(glmm_jacc_chaff_family)

# order
glmm_jacc_order <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_order, na.action = na.fail, family = gaussian)
confint(glmm_jacc_order)

glmm_jacc_chaff_order <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_chaff_order, na.action = na.fail, family = gaussian)
confint(glmm_jacc_chaff_order)

# class
glmm_jacc_class <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_class, na.action = na.fail, family = gaussian)
confint(glmm_jacc_class)

glmm_jacc_chaff_class <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_chaff_class, na.action = na.fail, family = gaussian)
confint(glmm_jacc_chaff_class)

# phylum
glmm_jacc_phylum <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_phylum, na.action = na.fail, family = gaussian)
confint(glmm_jacc_phylum)

glmm_jacc_chaff_phylum <- glmmTMB(j.beta_median ~  scale(froh) + (1|Species_Label), data = roh_jacc_chaff_phylum, na.action = na.fail, family = gaussian)
confint(glmm_jacc_chaff_phylum)

```

summary bray

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# OTU99
summ_glmm_bray_otu <- as.data.frame(cbind(summary(glmm_bray_otu)$coefficients[[1]], confint(glmm_bray_otu)[1:2,1:2]))
summ_glmm_bray_otu$data <- "all"
summ_glmm_bray_otu$taxonomy <- "OTU99"
summ_glmm_bray_otu$dist <- "Bray-Curtis"

summ_glmm_bray_chaff_otu <-as.data.frame(cbind(summary(glmm_bray_chaff_otu)$coefficients[[1]], confint(glmm_bray_chaff_otu)[1:2,1:2]))
summ_glmm_bray_chaff_otu$data <- "chaffinch"
summ_glmm_bray_chaff_otu$taxonomy <- "OTU99"
summ_glmm_bray_chaff_otu$dist <- "Bray-Curtis"


# species
summ_glmm_bray_species <- as.data.frame(cbind(summary(glmm_bray_species)$coefficients[[1]], confint(glmm_bray_species)[1:2,1:2]))
summ_glmm_bray_species$data <- "all"
summ_glmm_bray_species$taxonomy <- "species"
summ_glmm_bray_species$dist <- "Bray-Curtis"

summ_glmm_bray_chaff_species <- as.data.frame(cbind(summary(glmm_bray_chaff_species)$coefficients[[1]], confint(glmm_bray_chaff_species)[1:2,1:2]))
summ_glmm_bray_chaff_species$data <- "chaffinch"
summ_glmm_bray_chaff_species$taxonomy <- "species"
summ_glmm_bray_chaff_species$dist <- "Bray-Curtis"


# genus
summ_glmm_bray_genus <- as.data.frame(cbind(summary(glmm_bray_genus)$coefficients[[1]], confint(glmm_bray_genus)[1:2,1:2]))
summ_glmm_bray_genus$data <- "all"
summ_glmm_bray_genus$taxonomy <- "genus"
summ_glmm_bray_genus$dist <- "Bray-Curtis"

summ_glmm_bray_chaff_genus <- as.data.frame(cbind(summary(glmm_bray_chaff_genus)$coefficients[[1]], confint(glmm_bray_chaff_genus)[1:2,1:2]))
summ_glmm_bray_chaff_genus$data <- "chaffinch"
summ_glmm_bray_chaff_genus$taxonomy <- "genus"
summ_glmm_bray_chaff_genus$dist <- "Bray-Curtis"

# family
summ_glmm_bray_family <- as.data.frame(cbind(summary(glmm_bray_family)$coefficients[[1]], confint(glmm_bray_family)[1:2,1:2]))
summ_glmm_bray_family$data <- "all"
summ_glmm_bray_family$taxonomy <- "family"
summ_glmm_bray_family$dist <- "Bray-Curtis"

summ_glmm_bray_chaff_family <- as.data.frame(cbind(summary(glmm_bray_chaff_family)$coefficients[[1]], confint(glmm_bray_chaff_family)[1:2,1:2]))
summ_glmm_bray_chaff_family$data <- "chaffinch"
summ_glmm_bray_chaff_family$taxonomy <- "family"
summ_glmm_bray_chaff_family$dist <- "Bray-Curtis"

# order
summ_glmm_bray_order <- as.data.frame(cbind(summary(glmm_bray_order)$coefficients[[1]], confint(glmm_bray_order)[1:2,1:2]))
summ_glmm_bray_order$data <- "all"
summ_glmm_bray_order$taxonomy <- "order"
summ_glmm_bray_order$dist <- "Bray-Curtis"

summ_glmm_bray_chaff_order <-as.data.frame(cbind(summary(glmm_bray_chaff_order)$coefficients[[1]], confint(glmm_bray_chaff_order)[1:2,1:2]))
summ_glmm_bray_chaff_order$data <- "chaffinch"
summ_glmm_bray_chaff_order$taxonomy <- "order"
summ_glmm_bray_chaff_order$dist <- "Bray-Curtis"

# class
summ_glmm_bray_class <- as.data.frame(cbind(summary(glmm_bray_class)$coefficients[[1]], confint(glmm_bray_class)[1:2,1:2]))
summ_glmm_bray_class$data <- "all"
summ_glmm_bray_class$taxonomy <- "class"
summ_glmm_bray_class$dist <- "Bray-Curtis"

summ_glmm_bray_chaff_class <- as.data.frame(cbind(summary(glmm_bray_chaff_class)$coefficients[[1]], confint(glmm_bray_chaff_class)[1:2,1:2]))
summ_glmm_bray_chaff_class$data <- "chaffinch"
summ_glmm_bray_chaff_class$taxonomy <- "class"
summ_glmm_bray_chaff_class$dist <- "Bray-Curtis"

# phylum
summ_glmm_bray_phylum <- as.data.frame(cbind(summary(glmm_bray_phylum)$coefficients[[1]], confint(glmm_bray_phylum)[1:2,1:2]))
summ_glmm_bray_phylum$data <- "all"
summ_glmm_bray_phylum$taxonomy <- "phylum"
summ_glmm_bray_phylum$dist <- "Bray-Curtis"

summ_glmm_bray_chaff_phylum <- as.data.frame(cbind(summary(glmm_bray_chaff_phylum)$coefficients[[1]], confint(glmm_bray_chaff_phylum)[1:2,1:2]))
summ_glmm_bray_chaff_phylum$data <- "chaffinch"
summ_glmm_bray_chaff_phylum$taxonomy <- "phylum"
summ_glmm_bray_chaff_phylum$dist <- "Bray-Curtis"

#merge

summ_glmm_bray <- rbind(summ_glmm_bray_otu,summ_glmm_bray_species,summ_glmm_bray_genus, summ_glmm_bray_family, summ_glmm_bray_order, summ_glmm_bray_class, summ_glmm_bray_phylum, summ_glmm_bray_chaff_otu,summ_glmm_bray_chaff_genus, summ_glmm_bray_chaff_family, summ_glmm_bray_chaff_order, summ_glmm_bray_chaff_class, summ_glmm_bray_chaff_phylum)

```

summary jaccard

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# OTU99
summ_glmm_jacc_otu <- as.data.frame(cbind(summary(glmm_jacc_otu)$coefficients[[1]], confint(glmm_jacc_otu)[1:2,1:2]))
summ_glmm_jacc_otu$data <- "all"
summ_glmm_jacc_otu$taxonomy <- "OTU99"
summ_glmm_jacc_otu$dist <- "Jaccard"

summ_glmm_jacc_chaff_otu <-as.data.frame(cbind(summary(glmm_jacc_chaff_otu)$coefficients[[1]], confint(glmm_jacc_chaff_otu)[1:2,1:2]))
summ_glmm_jacc_chaff_otu$data <- "chaffinch"
summ_glmm_jacc_chaff_otu$taxonomy <- "OTU99"
summ_glmm_jacc_chaff_otu$dist <- "Jaccard"


# species
summ_glmm_jacc_species <- as.data.frame(cbind(summary(glmm_jacc_species)$coefficients[[1]], confint(glmm_jacc_species)[1:2,1:2]))
summ_glmm_jacc_species$data <- "all"
summ_glmm_jacc_species$taxonomy <- "species"
summ_glmm_jacc_species$dist <- "Jaccard"

summ_glmm_jacc_chaff_species <- as.data.frame(cbind(summary(glmm_jacc_chaff_species)$coefficients[[1]], confint(glmm_jacc_chaff_species)[1:2,1:2]))
summ_glmm_jacc_chaff_species$data <- "chaffinch"
summ_glmm_jacc_chaff_species$taxonomy <- "species"
summ_glmm_jacc_chaff_species$dist <- "Jaccard"


# genus
summ_glmm_jacc_genus <- as.data.frame(cbind(summary(glmm_jacc_genus)$coefficients[[1]], confint(glmm_jacc_genus)[1:2,1:2]))
summ_glmm_jacc_genus$data <- "all"
summ_glmm_jacc_genus$taxonomy <- "genus"
summ_glmm_jacc_genus$dist <- "Jaccard"

summ_glmm_jacc_chaff_genus <- as.data.frame(cbind(summary(glmm_jacc_chaff_genus)$coefficients[[1]], confint(glmm_jacc_chaff_genus)[1:2,1:2]))
summ_glmm_jacc_chaff_genus$data <- "chaffinch"
summ_glmm_jacc_chaff_genus$taxonomy <- "genus"
summ_glmm_jacc_chaff_genus$dist <- "Jaccard"

# family
summ_glmm_jacc_family <- as.data.frame(cbind(summary(glmm_jacc_family)$coefficients[[1]], confint(glmm_jacc_family)[1:2,1:2]))
summ_glmm_jacc_family$data <- "all"
summ_glmm_jacc_family$taxonomy <- "family"
summ_glmm_jacc_family$dist <- "Jaccard"

summ_glmm_jacc_chaff_family <- as.data.frame(cbind(summary(glmm_jacc_chaff_family)$coefficients[[1]], confint(glmm_jacc_chaff_family)[1:2,1:2]))
summ_glmm_jacc_chaff_family$data <- "chaffinch"
summ_glmm_jacc_chaff_family$taxonomy <- "family"
summ_glmm_jacc_chaff_family$dist <- "Jaccard"

# order
summ_glmm_jacc_order <- as.data.frame(cbind(summary(glmm_jacc_order)$coefficients[[1]], confint(glmm_jacc_order)[1:2,1:2]))
summ_glmm_jacc_order$data <- "all"
summ_glmm_jacc_order$taxonomy <- "order"
summ_glmm_jacc_order$dist <- "Jaccard"

summ_glmm_jacc_chaff_order <-as.data.frame(cbind(summary(glmm_jacc_chaff_order)$coefficients[[1]], confint(glmm_jacc_chaff_order)[1:2,1:2]))
summ_glmm_jacc_chaff_order$data <- "chaffinch"
summ_glmm_jacc_chaff_order$taxonomy <- "order"
summ_glmm_jacc_chaff_order$dist <- "Jaccard"

# class
summ_glmm_jacc_class <- as.data.frame(cbind(summary(glmm_jacc_class)$coefficients[[1]], confint(glmm_jacc_class)[1:2,1:2]))
summ_glmm_jacc_class$data <- "all"
summ_glmm_jacc_class$taxonomy <- "class"
summ_glmm_jacc_class$dist <- "Jaccard"

summ_glmm_jacc_chaff_class <- as.data.frame(cbind(summary(glmm_jacc_chaff_class)$coefficients[[1]], confint(glmm_jacc_chaff_class)[1:2,1:2]))
summ_glmm_jacc_chaff_class$data <- "chaffinch"
summ_glmm_jacc_chaff_class$taxonomy <- "class"
summ_glmm_jacc_chaff_class$dist <- "Jaccard"

# phylum
summ_glmm_jacc_phylum <- as.data.frame(cbind(summary(glmm_jacc_phylum)$coefficients[[1]], confint(glmm_jacc_phylum)[1:2,1:2]))
summ_glmm_jacc_phylum$data <- "all"
summ_glmm_jacc_phylum$taxonomy <- "phylum"
summ_glmm_jacc_phylum$dist <- "Jaccard"

summ_glmm_jacc_chaff_phylum <- as.data.frame(cbind(summary(glmm_jacc_chaff_phylum)$coefficients[[1]], confint(glmm_jacc_chaff_phylum)[1:2,1:2]))
summ_glmm_jacc_chaff_phylum$data <- "chaffinch"
summ_glmm_jacc_chaff_phylum$taxonomy <- "phylum"
summ_glmm_jacc_chaff_phylum$dist <- "Jaccard"

#merge

summ_glmm_jacc <- rbind(summ_glmm_jacc_otu,summ_glmm_jacc_species,summ_glmm_jacc_genus, summ_glmm_jacc_family, summ_glmm_jacc_order, summ_glmm_jacc_class, summ_glmm_jacc_phylum, summ_glmm_jacc_chaff_otu,summ_glmm_jacc_chaff_species,summ_glmm_jacc_chaff_genus, summ_glmm_jacc_chaff_family, summ_glmm_jacc_chaff_order, summ_glmm_jacc_chaff_class, summ_glmm_jacc_chaff_phylum)

```

merge bray and jaccard

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

glmm_out <- rbind(summ_glmm_bray, summ_glmm_jacc)

write.csv(glmm_out, paste0(dataDir, "/glmm_fst_roh_output_miss0.5_mac3_noZ_speciesFix.csv"))
```


# plots

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Fst scatter plot
bray = as.vector(bray_dist_genus)
jacc = as.vector(jaccard_dist_genus)
fst = as.vector(fst_dist)
geo = as.vector(geo_dist)

#new data frame with vectorized distance matrices
mat = data.frame(bray, jacc, fst, geo)

# Mantel r 
str(mantel_out)
mantel_out$signif <- ifelse(mantel_out$p < 0.05 & mantel_out$p > 0.005, "< 0.05", ifelse(mantel_out$p < 0.005 & mantel_out$p > 0.0005 , "< 0.005", ifelse(mantel_out$p < 0.0005, "< 0.0005", "n.s")))
table(mantel_out$signif)
check <- mantel_out %>% filter(p < 0.05)

mantel_out$order <- factor(mantel_out$taxonomy, levels = c("OTU99","Species","Genus", "Family", "Order", "Class", "Phylum"))


```



```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

(p_bray_fst <- ggplot(mat, aes(y = bray, x = fst)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Fst", y = "Bray-Curtis distance") +
        geom_smooth(method = "lm", colour = "darkred", alpha = 0.2,lwd=3) +scatter_theme + ggtitle("") + sl_theme +theme(legend.position="bottom", 
          axis.text.x = element_text(angle = 0,  size = 36), axis.text.y = element_text(angle = 0,  size = 36),
          axis.title = element_text(size = 38), 
          legend.title=element_text(size=38), 
          legend.text=element_text(size=36), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank(),
          axis.line = element_line(colour = 'black', size = 2)))

        #+  annotate(geom="text",label=paste0("Mantel r =", round(mantel_bray_fst_genus$statistic, 2), " \n ;p-value=", round(mantel_bray_fst_genus$signif, 3)), parse=F,x=0.3,y=.8,size=10)) 
        #+scale_y_continuous(breaks = seq(0, 200, by=25), limits=c(0,190)))

(p_jacc_fst <- ggplot(mat, aes(y = jacc, x = fst)) + 
        geom_point(size = 5, alpha = 0.2) + 
        labs(x = "Fst", y = "Jaccard distance") +
       geom_smooth(method = "lm", colour = "black", alpha = 0.2,lwd=3) +
       scatter_theme + 
       sl_theme +
       theme(legend.position="bottom", 
          axis.text.x = element_text(angle = 0,  size = 36), axis.text.y = element_text(angle = 0,  size = 36),
          axis.title = element_text(size = 38), 
          legend.title=element_text(size=38), 
          legend.text=element_text(size=36), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank(),
          axis.line = element_line(colour = 'black', size = 2)))


mantel_out$signifFill <- as.factor(mantel_out$signif)
levels(mantel_out$signifFill)[levels(mantel_out$signifFill)=="< 0.05"] <- "#999999"
levels(mantel_out$signifFill)[levels(mantel_out$signifFill)=="< 0.005"] <- "#333333"
levels(mantel_out$signifFill)[levels(mantel_out$signifFill)=="n.s"] <- "white"




(p_mantel <- mantel_out %>% filter(data == "all", test == "partial-mantel") %>% 
    ggplot(aes(y = r, x = order, fill = signif, color = signif)) + 
        geom_bar(stat = "identity", linewidth = 1.5) +
        scale_fill_manual("p-value", values = levels(mantel_out$signifFill), limits=names(mantel_out$signifFill))+ 
        scale_color_manual("p-value", values = c("black", "black","black", "black","black"), limits=names(mantel_out$signifFill)) +
        ylab(expression("Mantel"~r^2)) + xlab("")+
        barplot_theme + 
         facet_wrap(.~factor(dist), strip.position = "top") +
        guides(color = guide_legend(override.aes = list(linewidth = 25))) +
        theme(legend.position="bottom", 
          axis.text.x = element_text(angle = 35,  size = 36,  hjust=1), 
          legend.title=element_text(size=38), 
          legend.text=element_text(size=36), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))

(p_bray_het <- ggplot(roh_bray_species, aes(y = b.beta_median, x = froh)) + 
        geom_point(size = 5, alpha = 0.2) + 
        xlab("FROH") +
        ylab("Divergence (Bray-Curtis)") + 
        geom_smooth(method = "lm", colour = "darkred", alpha = 0.2,lwd=3) + scatter_theme + ggtitle("") + sl_theme +
        theme(legend.position="bottom", 
          axis.text.x = element_text(angle = 0,  size = 36), axis.text.y = element_text(angle = 0,  size = 36),
          axis.title = element_text(size = 38), 
          legend.title=element_text(size=38), 
          legend.text=element_text(size=36), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank(),
          axis.line = element_line(colour = 'black', size = 2)))

(p_jacc_het <- ggplot(roh_jacc_species, aes(y = j.beta_median, x = froh)) + 
        geom_point(size = 5, alpha = 0.2) + 
        xlab("FROH") +
        ylab("Divergence (Jaccard)") + 
        geom_smooth(method = "lm", colour = "black", alpha = 0.2,lwd=3) + scatter_theme + ggtitle("") + 
        sl_theme +
        theme(legend.position="bottom", 
          axis.text.x = element_text(angle = 0,  size = 36), axis.text.y = element_text(angle = 0,  size = 36),
          axis.title = element_text(size = 38), 
          legend.title=element_text(size=38), 
          legend.text=element_text(size=36), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank(),
          axis.line = element_line(colour = 'black', size = 2)))

glmm_out <- read.csv(paste0(dataDir, "/glmm_fst_roh_output_miss0.5_mac3_noZ_speciesFix.csv"))
head(glmm_out)
glmm_out_het <- read.csv(paste0(dataDir, "/glmm_fst_het_output_miss0.5_mac3_noZ_speciesFix.csv"))
head(glmm_out_het)

colnames(glmm_out) <- c("term","Estimate","Std_Error","z_value","P","lower","upper","data","taxonomy","dist")
head(glmm_out)

glmm_out$order <- factor(glmm_out$taxonomy, levels = c("OTU99", "species" ,"genus", "family", "order", "class", "phylum"), labels = c("OTU99", "Species" , "Genus", "Family", "Order", "Class", "Phylum"))

glmm_out$signif <- ifelse(glmm_out$P < 0.05 & glmm_out$P > 0.005, "< 0.05", ifelse(glmm_out$P < 0.005 & glmm_out$P > 0.0005 , "< 0.005", ifelse(glmm_out$P < 0.0005, "< 0.0005", "n.s")))
table(glmm_out$signif)

head(glmm_out)
check <- glmm_out %>% filter(P < 0.05)

glmm_out$signifFill <- as.factor(glmm_out$signif)
levels(glmm_out$signifFill)[levels(glmm_out$signifFill)=="< 0.05"] <- "#999999"
levels(glmm_out$signifFill)[levels(glmm_out$signifFill)=="< 0.005"] <- "#333333"
levels(glmm_out$signifFill)[levels(glmm_out$signifFill)=="< 0.0005"] <- "#000000"
levels(glmm_out$signifFill)[levels(glmm_out$signifFill)=="n.s"] <- "white"

glmm_out$term <- gsub('[[:digit:]]+', '', glmm_out$term)

(p_glmm <- glmm_out %>% filter(term == "scale(froh)", data == "all") %>% 
    ggplot(aes(x = Estimate, y = order)) + 
      geom_linerange(aes(xmin = lower, xmax = upper), linewidth = 2.5, color = "#000000") +
      geom_point(aes(fill = signif, color = signif), pch =21, size = 8) +
      geom_vline(xintercept = 0, linetype = "longdash") + 
      scale_fill_manual("p-value", values = levels(glmm_out$signifFill), limits=names(glmm_out$signifFill))+ 
        scale_color_manual("p-value", values = c("black", "black","black", "black","black"), limits=names(glmm_out$signifFill)) +
        xlab("Estimate") + ylab("")+
        barplot_theme + 
        facet_wrap(.~factor(dist), strip.position = "top") +
        guides(color = guide_legend(override.aes = list(linewidth = 25))) +
        theme(legend.position="bottom", 
          axis.text.x = element_text(angle = 0,  size = 36), 
          legend.title=element_text(size=38), 
          legend.text=element_text(size=36), 
          panel.spacing = unit(2, "lines"),
          strip.background = element_blank(),
          strip.text.x = element_blank()))


png(paste0(outputDir, "/Figures/Fst_roh_speciesFix.png"),width=2800, height=4100,res=150)

pushViewport(viewport(layout = grid.layout(40, 10)))
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
print(p_bray_fst, vp = vplayout(1:10, 1:5))
print(p_jacc_fst, vp = vplayout(1:10, 6:10))
print(p_mantel + theme(legend.position="none"), vp = vplayout(11:20, 1:10))

print(p_bray_het, vp = vplayout(20:30, 1:5))
print(p_jacc_het, vp = vplayout(20:30, 6:10))
print(p_glmm + theme(legend.position="none"), vp = vplayout(31:40, 1:10))

dev.off()

```



Adding individual inbreeding coefficients to microbiome metadata
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
names(froh_meta)
names(sample_data(microb_overlap_otu))

froh_meta$Lab_Code <- as.character(froh_meta$"16S_Lab_Code")

f_df <- data.frame(cbind(Lab_Code = froh_meta$Lab_Code, f = as.numeric(froh_meta$f)))
f_df <- f_df %>% drop_na()
f_df$f <- as.numeric(f_df$f)
f_df$Lab_Code
sample_data(microb_overlap_otu)$Lab_Code

Sample_df <- data.frame(sample_data(microb_overlap_otu))
Sample_df <- merge(Sample_df,f_df ,by = "Lab_Code")

Sample_df$Lab_Code == sample_data(microb_overlap_otu)$Lab_Code
Sample_df$Lab_Code == sample_data(microb_overlap_genus)$Lab_Code
Sample_df$Lab_Code == sample_data(microb_overlap_family)$Lab_Code
Sample_df$Lab_Code == sample_data(microb_overlap_class)$Lab_Code
Sample_df$Lab_Code == sample_data(microb_overlap_order)$Lab_Code
Sample_df$Lab_Code == sample_data(microb_overlap_phylum)$Lab_Code

sample_data(microb_overlap_otu)$f <- Sample_df$f
sample_data(microb_overlap_genus)$f <- Sample_df$f
sample_data(microb_overlap_family)$f <- Sample_df$f
sample_data(microb_overlap_class)$f <- Sample_df$f
sample_data(microb_overlap_order)$f <- Sample_df$f
sample_data(microb_overlap_phylum)$f <- Sample_df$f

```

