---
title: 'Processing the rbcL output files from the Dubois et al. (2022) pipeline'
author: "Mark Gillingham and B. Karina Montero"
date: "`r Sys.Date()`"
#output: pdf_document
output: html_document
---

## Loading libraries
```{r libraries, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
rm(list = ls())

# basics
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(ggpubr)

# phyloseq section
library(phyloseq)
library(decontam)

```

## WORKING DIRECTORIES
```{r directories, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}
source("~/Dropbox/FringillaMicrobiome/05_Scripts/workingDirectories.R")
```


## I. LOAD & CHECK DATA

### 1. metadata
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
fringilla <- import_qiime_sample_data(mapfilename = paste0(dataDir, "/metadata_fringilla_diet_03.06.23.csv"))

```
### 2. Directory containing the fastq files
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
rbcL_path <- "/home/kari/Documents/Raw_data"

paste0(metabarcodingDir, "rbcL_raw_data")  
list.files(rbcL_path)
rbcL_files <- sort(list.files(rbcL_path, pattern="rbcL", full.names = TRUE))
```

### 3. Inspect read quality profiles

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
plotQualityProfile(rbcL_files[c(45,36,33,38,8,46,10,18,17)])
```
## II. REMOVE PRIMERS


Set patterns to discriminate your forward and reverse read files
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
file_pattern <- c("F" = "_R1_001.fastq", "R" = "_R2_001.fastq")
fas_Fs_raw <- sort(list.files(rbcL_path, pattern = file_pattern["F"], full.names = TRUE))
fas_Rs_raw <- sort(list.files(rbcL_path, pattern = file_pattern["R"], full.names = TRUE))
```
Primer sequences

For plant library preparation, a fragment of the rbcL gene of around 184 bp was amplified using the following primers::

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
FWD_rbcL <- "TTGGCAGCATTYCGAGTAACTCC" # rbcL1 (Palmieri, Bozza, and Giongo 2009).
REV_rbcL <-  "AACCYTCTTCAAAAAGGTC" # rbcLB (Palmieri, Bozza, and Giongo 2009).
```

To ensure we have the right primers, and the correct orientation of the primers on the reads, we will verify the presence and orientation of these primers in the data.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
allOrients <- function(primer) {
    # Create all orientations of the input sequence
    require(Biostrings)
    dna <- DNAString(primer)  # The Biostrings works w/ DNAString objects rather than character vectors
    orients <- c(Forward = dna, Complement = complement(dna), Reverse = reverse(dna), 
        RevComp = reverseComplement(dna))
    return(sapply(orients, toString))  # Convert back to character vector
}

FWD_orients_rbcL <- allOrients(FWD_rbcL)
REV_orients_rbcL <- allOrients(REV_rbcL)

```

Count the number of times the primers appear in the forward and reverse read, while considering all possible primer orientations.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
primerHits <- function(primer, fn) {
    # Counts number of reads in which the primer is found
    nhits <- vcountPattern(primer, sread(readFastq(fn)), fixed = FALSE)
    return(sum(nhits > 0))
}

primer_count <- rbind(FWD.ForwardReads_rbcL = sapply(FWD_orients_rbcL, primerHits, fn = fas_Fs_raw[[1]]), 
    FWD.ReverseReads_rbcL = sapply(FWD_orients_rbcL, primerHits, fn = fas_Rs_raw[[1]]), 
    REV.ForwardReads_rbcL = sapply(REV_orients_rbcL, primerHits, fn = fas_Fs_raw[[1]]), 
    REV.ReverseReads_rbcL = sapply(REV_orients_rbcL, primerHits, fn = fas_Rs_raw[[1]]))
primer_count
```
As expected, the FWD primer is found in the forward reads in its forward orientation, and in some of the reverse reads in its reverse-complement orientation (due to read-through when the ITS region is short). Similarly the REV primer is found with its expected orientations.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
FWD <- c("TTGGCAGCATTYCGAGTAACTCC")
REV <- c("AACCYTCTTCAAAAAGGTC")

FWD_RC <- dada2:::rc(FWD)
REV_RC <- dada2:::rc(REV)

path_cut <- file.path(rbcL_path, "cutadapt")
if(!dir.exists(path_cut)) dir.create(path_cut)

#path_cut <- paste0(barcodingDir, "/cutadapt")

fas_Fs_cut <- file.path(path_cut, basename(fas_Fs_raw))
fas_Rs_cut <- file.path(path_cut, basename(fas_Rs_raw))

R1_flags <- paste(paste("-g", FWD, collapse = " "), paste("-a", REV_RC, collapse = " "))
R2_flags <- paste(paste("-G", REV, collapse = " "), paste("-A", FWD_RC, collapse = " "))

cutadapt <- "/home/fringilla/miniconda3/envs/cutadaptenv/bin/cutadapt" # Path to the executable

for(i in seq_along(fas_Fs_raw)) {
  cat("Processing", "-----------", i, "/", length(fas_Fs_raw), "-----------\n")
  system2(cutadapt, args = c(R1_flags, R2_flags,
                             "-m 50",
                             "-M 200",
                             "--discard-untrimmed",
                             "--max-n 0",
                             "--cores 0", # Use 0 to auto-detect.
                             "-n 2", # required to remove FWD and REV from read
                             # Optional strong constraint on expected length
                             #paste0("-m ", 250-nchar(FWD)[1], ":", 250-nchar(REV)[1]), 
                             #paste0("-M ", 250-nchar(FWD)[1], ":", 250-nchar(REV)[1]), 
                             "-o", fas_Fs_cut[i], "-p", fas_Rs_cut[i],
                             fas_Fs_raw[i], fas_Rs_raw[i]))
}


primer_count <- rbind(FWD.ForwardReads_rbcL = sapply(FWD_orients_rbcL, primerHits, fn = fas_Fs_cut[[1]]), 
    FWD.ReverseReads_rbcL = sapply(FWD_orients_rbcL, primerHits, fn = fas_Rs_cut[[1]]), 
    REV.ForwardReads_rbcL = sapply(REV_orients_rbcL, primerHits, fn = fas_Fs_cut[[1]]), 
    REV.ReverseReads_rbcL = sapply(REV_orients_rbcL, primerHits, fn = fas_Rs_cut[[1]]))
primer_count

```

## III. FILTER AND TRIM

Remove 284_rbcL empty file

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
path_cut <- paste0(metabarcodingDir, "rbcL_raw_data/cutadapt") # If you skipped primers removal, provide the path to your sequences here
list.files(path_cut)

fas_Fs_cut <- sort(list.files(path_cut, pattern = file_pattern["F"], full.names = TRUE))
fas_Rs_cut <- sort(list.files(path_cut, pattern = file_pattern["R"], full.names = TRUE))

sample_names <- sapply(strsplit(basename(fas_Fs_cut), "_"), function(x) x[1])

rbcL_cut <- sort(list.files(path_cut, pattern="rbcL", full.names = TRUE))
```

Inspect read quality profiles

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

plotQualityProfile(rbcL_cut[c(45,36,33,38,8,46,10,18,17)])
#ggsave(paste(outputDir, "/02_DietMetabarcoding/01_Exploration/01_qualityProfile/rbcL_cut_plotQualityProfile.png", sep = "/"))

fas_Fs_filtered <- file.path(rbcL_path, "filtered", basename(fas_Fs_cut))
fas_Rs_filtered <- file.path(rbcL_path, "filtered", basename(fas_Rs_cut))
all.equal(basename(fas_Fs_raw), basename(fas_Fs_filtered))

names(fas_Fs_filtered) <- sample_names
names(fas_Rs_filtered) <- sample_names

```

Standard filtering paraments: maxN=0 (DADA2 requires sequences contain no Ns), truncQ = 2, rm.phix = TRUE and maxEE=2. The maxEE parameter sets the maximum number of “expected errors” allowed in a read, which is a better filter than simply averaging quality scores.

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
filter_trim <- filterAndTrim(fas_Fs_cut, fas_Fs_filtered, fas_Rs_cut, fas_Rs_filtered,
                       #truncLen = c(170, 100), 
                       minLen = 174,
                       maxLen = 194, 
                       maxN = 0, maxEE = c(2, 2), truncQ = 2,
                       rm.phix = TRUE, multithread = TRUE, verbose = FALSE)
head(filter_trim)
```
## IV. DADA 2

Learn error rates
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
error_F <- learnErrors(fas_Fs_filtered, multithread = TRUE, randomize = TRUE)
#101903502 total bases in 565302 reads from 112 samples will be used for learning the error rates.
saveRDS(error_F, paste0(outputDir, "/02_DietMetabarcoding/02_dada2/rbcL_error_F.rds"))
#err <- readRDS(paste0(outputDir, "/02_Dada2/err.rds"))

plotErrors(error_F, nominalQ=TRUE)
ggsave(paste(outputDir, "/02_DietMetabarcoding/02_dada2/plotErrors_rbcL_F.png", sep = "/"))

error_R <- learnErrors(fas_Rs_filtered, multithread = TRUE, randomize = TRUE)
#100633957 total bases in 560007 reads from 137 samples will be used for learning the error rates.
saveRDS(error_R, paste0(outputDir, "/02_DietMetabarcoding/02_dada2/rbcL_error_R.rds"))
#err <- readRDS(paste0(outputDir, "/02_Dada2/err.rds"))

plotErrors(error_R, nominalQ=TRUE)
ggsave(paste(outputDir, "/02_DietMetabarcoding/02_dada2/plotErrors_rbcL_R.png", sep = "/"))
```


### DEREPLICATION, SAMPLE INFERENCE & MERGE PAIRED READS
from https://github.com/fkeck/DADA2_diatoms_pipeline/blob/master/pipeline.R

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
merged_list <- vector("list", length(sample_names))
names(merged_list) <- sample_names

for(i in sample_names){
  cat("Processing -------", which(sample_names == i), "/", length(sample_names), "-------", i, "\n")
  derep_Fs <- derepFastq(fas_Fs_filtered[[i]], verbose = TRUE)
  derep_Rs <- derepFastq(fas_Rs_filtered[[i]], verbose = TRUE)
  dds_Fs <- dada(derep_Fs, err = error_F, multithread = TRUE, verbose = TRUE)
  dds_Rs <- dada(derep_Rs, err = error_R, multithread = TRUE, verbose = TRUE)
  merged_list[[i]] <- mergePairs(dds_Fs, derep_Fs, dds_Rs, derep_Rs, verbose = TRUE)
}
```

### CONSTRUCT SEQUENCE TABLE

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
seqtab <- makeSequenceTable(merged_list)
dim(seqtab)
table(nchar(getSequences(seqtab)))
```

### REMOVE CHIMERA

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
seqtab_nochim <- removeBimeraDenovo(seqtab, method = "consensus", multithread = TRUE, verbose = TRUE)
#Identified 232 bimeras out of 3275 input sequences.
dim(seqtab_nochim)
table(nchar(getSequences(seqtab_nochim)))
```


### Export files for use in QIIME2
https://forum.qiime2.org/t/exporting-dada2-r-package-results-to-work-with-qiime2/2573/3
```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
write.table(t(seqtab_nochim), paste0(outputDir, "/02_DietMetabarcoding/02_dada2/rbcLQIIME2seqtab.txt"), sep="\t", row.names=TRUE, col.names=NA, quote=FALSE)
uniquesToFasta(seqtab_nochim, fout=paste0(outputDir, "/02_DietMetabarcoding/02_dada2/rbcL-rep-seqs.fna"), ids=colnames(seqtab_nochim))

```

### TAXONOMY ASSIGNMENT (CESGA cluster)

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

qiime2run qiime feature-classifier classify-sklearn \
  --i-classifier /mnt/lustre/scratch/nlsas/home/csic/dde/pla/FringillaProject/06_rbcL/NCBI_rbcL_Viridiplantae_classifier_extracted_2022_09_22_derep.qza \
  --i-reads /mnt/lustre/scratch/nlsas/home/csic/dde/pla/FringillaProject/06_rbcL/rbcL-rep-seqs.qza \
  --o-classification /mnt/lustre/scratch/nlsas/home/csic/dde/pla/FringillaProject/06_rbcL/rbcL_4_derep_restricted.qza

```

# CLEAN rbcL data 

The best classifier was trained uniq dereplication once (high taxonomic assignment at the Family level: ps_rbcL_4_derep_restricted_filt). We will use this classifier for all down stream analyses. 


```{r psobject, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}

rbcL_4_derep_restricted_biom_data <- import_biom(BIOMfilename = paste0(dataDir,"/classifiers/rbcL_4_derep_restricted.biom"))

rep.seqs <- Biostrings::readDNAStringSet(paste0(outputDir,"/02_DietMetabarcoding/02_dada2/rbcL-rep-seqs.fna"), format = "fasta")

ps_rbcL_4_derep_restricted <- merge_phyloseq(t(otu_table(rbcL_4_derep_restricted_biom_data)),tax_table(rbcL_4_derep_restricted_biom_data), fringilla,rep.seqs)

taxa_names(ps_rbcL_4_derep_restricted) <- paste0("ASVrbcL", seq(ntaxa(ps_rbcL_4_derep_restricted)))

tax_rbcL_4_derep_restricted <- data.frame(tax_table(ps_rbcL_4_derep_restricted))

tax_rbcL_4_derep_restricted.clean <- data.frame(row.names = row.names(tax_rbcL_4_derep_restricted),Kingdom = str_replace(tax_rbcL_4_derep_restricted[,1], "k__",""),Phylum = str_replace(tax_rbcL_4_derep_restricted[,2], "p__",""),Class = str_replace(tax_rbcL_4_derep_restricted[,3], "c__",""),Order = str_replace(tax_rbcL_4_derep_restricted[,4], "o__",""),Family = str_replace(tax_rbcL_4_derep_restricted[,5], "f__",""),Genus = str_replace(tax_rbcL_4_derep_restricted[,6], "g__",""),Species = str_replace(tax_rbcL_4_derep_restricted[,7], "s__",""),stringsAsFactors = FALSE)

tax_rbcL_4_derep_restricted.clean [tax_rbcL_4_derep_restricted.clean  == ''] <- NA

tax_table(ps_rbcL_4_derep_restricted) <- as.matrix(tax_rbcL_4_derep_restricted.clean)


ps_rbcL_4_derep_restricted_filt <- ps_rbcL_4_derep_restricted %>%
  subset_taxa(
    Kingdom == "Viridiplantae" &
    Phylum == "Streptophyta")


```


## Removing contaminants using PCR controls

We remove the contaminant ASVs using the DECONTAM package:

(https://bioconductor.org/packages/release/bioc/vignettes/decontam/inst/doc/decontam_intro.html)

We first look at the library sizes (i.e. the number of reads) in each sample, as a function of whether that sample was a true positive sample or a negative control
```{r librarysizes, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE }


df <- as.data.frame(sample_data(ps_rbcL_4_derep_restricted_filt)) 
df$LibrarySize <- sample_sums(ps_rbcL_4_derep_restricted_filt)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))

#png(paste0(outputDir, "/02_DietMetabarcoding/temp.png"),width=4000, height=2000,res=300)
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_or_Control)) + geom_point() + theme_bw()
##dev.off()


```

Second we investigate the prevalence of ASVs in control samples and so called "true samples". Then we remove identified contaminants from our dataset, 4 were identified as contaminants
```{r prevalenceDecontam, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE }

sample_data(ps_rbcL_4_derep_restricted_filt)$is.neg <- sample_data(ps_rbcL_4_derep_restricted_filt)$Sample_or_Control == "Control"

contamdf.prev <- isContaminant(ps_rbcL_4_derep_restricted_filt, method="prevalence", neg="is.neg")
table(contamdf.prev$contaminant)

contamdf.prev05 <- isContaminant(ps_rbcL_4_derep_restricted_filt, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev05$contaminant)

ps_rbcL_4_derep_restricted_filt.pa <- transform_sample_counts(ps_rbcL_4_derep_restricted_filt, function(abund) 1*(abund>0))

ps_rbcL_4_derep_restricted_filt.pa.neg <- prune_samples(sample_data(ps_rbcL_4_derep_restricted_filt.pa)$Sample_or_Control == "Control", ps_rbcL_4_derep_restricted_filt.pa)

ps_rbcL_4_derep_restricted_filt.pa.pos <- prune_samples(sample_data(ps_rbcL_4_derep_restricted_filt.pa)$Sample_or_Control == "Sample", ps_rbcL_4_derep_restricted_filt.pa)

df.pa <- data.frame(pa.pos=taxa_sums(ps_rbcL_4_derep_restricted_filt.pa.pos), pa.neg=taxa_sums(ps_rbcL_4_derep_restricted_filt.pa.neg),contaminant=contamdf.prev$contaminant)

ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)") + theme_bw()

tax_table(ps_rbcL_4_derep_restricted_filt)[contamdf.prev05$contaminant,]

ps_rbcL_4_derep_restricted_filt.contam <- prune_taxa(contamdf.prev05$contaminant, ps_rbcL_4_derep_restricted_filt)
ps_rbcL_4_derep_restricted_filt

ps_rbcL_4_derep_restricted_filt.noncontam <- prune_taxa(!contamdf.prev05$contaminant, ps_rbcL_4_derep_restricted_filt)
ps_rbcL_4_derep_restricted_filt.noncontam

### Four taxa are identified as contaminants

ps_rbcL_4_derep_restricted_filt<-ps_rbcL_4_derep_restricted_filt.noncontam



```


We remove the PCR control samples from the phyloseq object
```{r removeControls, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE ,results="hide"}

subset_samples(ps_rbcL_4_derep_restricted_filt, Sample_or_Control %in% Samples_toRemove)

subset_samples(ps_rbcL_4_derep_restricted_filt, !(Sample_or_Control %in% Samples_toRemove))

ps_rbcL_4_derep_restricted_filt<-subset_samples(ps_rbcL_4_derep_restricted_filt, !(Sample_or_Control %in% Samples_toRemove))


```

### Fraction of **reads** that can be assigned to taxonomic level

```{r fractionAssigned, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
readsPerSample = rowSums(otu_table(ps_rbcL_4_derep_restricted_filt))
fractionReadsAssigned = sapply(colnames(tax_table(ps_rbcL_4_derep_restricted_filt)), function(x){
  rowSums(otu_table(ps_rbcL_4_derep_restricted_filt)[, !is.na(tax_table(ps_rbcL_4_derep_restricted_filt))[,x]]) / readsPerSample
})

fractionReadsAssigned = data.frame(SampleID = rownames(fractionReadsAssigned), fractionReadsAssigned)
fractionReadsAssigned.L = pivot_longer(fractionReadsAssigned, 
    cols=colnames(tax_table(ps_rbcL_4_derep_restricted_filt)), names_to="taxlevel", values_to="fractionReadsAssigned")
fractionReadsAssigned.L = data.frame(fractionReadsAssigned.L, sample_data(ps_rbcL_4_derep_restricted_filt)[fractionReadsAssigned.L$SampleID,])

fractionReadsAssigned.L$taxlevelf = factor(fractionReadsAssigned.L$taxlevel, levels=c("Kingdom","Phylum","Class","Order","Family","Genus","Species"))

# Boxplot, fraction assigned by SampleType
#png(paste0(outputDir, "/02_DietMetabarcoding/01_Exploration/rbcLReads_Taxonomy_4_derep_restricted.png"),width=4000, height=2000,res=300)

p4<-ggplot(fractionReadsAssigned.L, aes(y = fractionReadsAssigned, x = taxlevelf)) +  theme_bw(base_size = 10) +  geom_boxplot(color='black', outlier.shape=NA) + geom_point() +  ggtitle("Fraction of reads identified by taxonomic level, NCBI classifier \n(primer trained and dereplicated 'uniq'; 1885 ASVs)") + ylab ("Fraction of reads assigned") + xlab ("Taxonomic rank")
##dev.off()

Table4<-round(tapply(fractionReadsAssigned.L$fractionReadsAssigned,fractionReadsAssigned.L$taxlevelf,mean),4)*100
c.names<-append(names(Table4),paste0("Classifier"))
Table4<-append(Table4,paste0("4_derep_restricted"))
Table4<-array(Table4,dimnames=list(c.names))


```

### Replace NAs with the highest possible rank in the taxonomic names
```{r ReplaceNAs, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE, results="hide"}


tax_rbcL_4_derep_restricted_filt<- data.frame(tax_table(ps_rbcL_4_derep_restricted_filt))

tax_rbcL_4_derep_restricted_filt[is.na(tax_rbcL_4_derep_restricted_filt)] <- ""

for (i in 1:7){ tax_rbcL_4_derep_restricted_filt[,i] <- as.character(tax_rbcL_4_derep_restricted_filt[,i])}
tax_rbcL_4_derep_restricted_filt[is.na(tax_rbcL_4_derep_restricted_filt)] <- ""

for (i in 1:nrow(tax_rbcL_4_derep_restricted_filt)){if (tax_rbcL_4_derep_restricted_filt[i,2] == "")
  {
    kingdom <- paste("Kingdom_", tax_rbcL_4_derep_restricted_filt[i,1], sep = "")
    tax_rbcL_4_derep_restricted_filt[i, 2:7] <- kingdom
  } else if (tax_rbcL_4_derep_restricted_filt[i,3] == ""){
    phylum <- paste("Phylum_", tax_rbcL_4_derep_restricted_filt[i,2], sep = "")
    tax_rbcL_4_derep_restricted_filt[i, 3:7] <- phylum
  } else if (tax_rbcL_4_derep_restricted_filt[i,4] == ""){
    class <- paste("Class_", tax_rbcL_4_derep_restricted_filt[i,3], sep = "")
    tax_rbcL_4_derep_restricted_filt[i, 4:7] <- class
  } else if (tax_rbcL_4_derep_restricted_filt[i,5] == ""){
    order <- paste("Order_", tax_rbcL_4_derep_restricted_filt[i,4], sep = "")
    tax_rbcL_4_derep_restricted_filt[i, 5:7] <- order
  } else if (tax_rbcL_4_derep_restricted_filt[i,6] == ""){
    family <- paste("Family_", tax_rbcL_4_derep_restricted_filt[i,5], sep = "")
    tax_rbcL_4_derep_restricted_filt[i, 6:7] <- family
  } else if (tax_rbcL_4_derep_restricted_filt[i,7] == ""){
    tax_rbcL_4_derep_restricted_filt$Species[i] <- paste("Genus",tax_rbcL_4_derep_restricted_filt$Genus[i], sep = "_")
  }
}

tax_rbcL_4_derep_restricted_filt

tax_table(ps_rbcL_4_derep_restricted_filt)<-as.matrix(tax_rbcL_4_derep_restricted_filt)
```

### Barplot of taxonomy at the class level according to the geographical local location of the samples (all sample types pooled):

```{r taxonomyClass, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}


#(sample_data(ps_rbcL_4_derep_restricted_filt))$Island_Mainland

Merged<-merge_samples(ps_rbcL_4_derep_restricted_filt,"Island_Mainland")
warnings() 

#names(sample_data(Merged))

physeq2 <- phyloseq::transform_sample_counts(Merged, function(x) x / sum(x) )
physeq2

ps_melt<- psmelt(physeq2)

ps_melt$Family[ps_melt$Abundance < 0.01] <- "< 1% abund."

p4 <- ggplot(data=ps_melt, aes(x=Sample, y=Abundance, fill=Family))
p4 <-p4 + geom_bar(aes(), stat="identity", position="stack") +scale_fill_manual(values = c("darkblue", "#8b5b5b", "darkorchid", "lightskyblue", "darkgreen", "#4a3b2c", "deeppink", "khaki2","#aa6c39", "darkorange1", "#E0FFFF", "darksalmon", "#eee7e1", "black", "darkolivegreen1","pink","darkseagreen","#E41A1C","#0000FF","#00FF00","#FFA500","#E7298A","#00FFFF","#008000","#800080","#008080","#999999","#FF4500","#FFFF00","#4DAF4A","#A65628","#F781BF","#1B9E77","#808000","#377EB8", "#CBC3E3","#460046","#AA336A","#d2691e")) + guides(fill=guide_legend(title="Taxanomic assignment (Family or higher)",ncol=3)) +theme(text = element_text(size = 10),legend.position="right",axis.text.x = element_text(angle = 45, hjust = 1),panel.background = element_blank(),plot.margin = unit(c(1,1,1,1), "cm"))+ggtitle("primer trained and dereplicated 'uniq'; 1885 ASVs")+
  xlab("")


```

## Explore sequencing depth and number of ASVs of

```{r, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Add sequencing depth column 
sample_data(ps_rbcL_4_derep_restricted_filt)$Seq_depth_rbcL <- sample_sums(ps_rbcL_4_derep_restricted_filt)

df_explore_filt <- ps_rbcL_4_derep_restricted_filt %>% subset_samples(Genus == "Fringilla") %>%
    sample_data %>%
    as.data.frame
table(df_explore_filt$Seq_depth_rbcL == 0)

asv_df_explore_filt <- ps_rbcL_4_derep_restricted_filt %>%subset_samples(Genus == "Fringilla") %>%
    t %>%
    otu_table

ASV_number <- colSums(asv_df_explore_filt != 0)

df_explore_filt$ASV_number <- ASV_number

table(df_explore_filt$ASV_number == 0)

(p_run <- ggplot(df_explore_filt) + geom_boxplot(aes(x = Species_Label, y = Seq_depth_rbcL)) +
    ylab("Sequencing depth") + scale_y_log10() + theme_linedraw() + theme(panel.grid = element_line(colour = "darkgrey"),
    legend.position = "none"))

# ASV number per species

(p_ASVs <- ggplot(df_explore_filt) + geom_boxplot(aes(x = Species_Label, y = ASV_number)) +
    ylab("Observed ASVs") + scale_y_log10() + theme_linedraw() + theme(panel.grid = element_line(colour = "darkgrey"),
    legend.position = "none"))

(p_ASVs <- ggplot(df_explore_filt) + geom_boxplot(aes(x = Month, y = ASV_number)) +
    ylab("Observed ASVs") + scale_y_log10() + theme_linedraw() + theme(panel.grid = element_line(colour = "darkgrey"),
    legend.position = "none"))

table(df_explore_filt$Species_Label, df_explore_filt$Month)

```

## EXPORT rbcL PHYLOSEQ OBJECT

Export ps_rbcL_4_derep_restricted_filt and use this phyloseq object for further downstream analyses on rbcL.

Update metadata Fringilla project

```{r export, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

# Add sequencing depth column 
sample_data(ps_rbcL_4_derep_restricted_filt)$Seq_depth_rbcL <- sample_sums(ps_rbcL_4_derep_restricted_filt)

# remove unnecessary columns from metadata
df <-  data.frame(sample_data(ps_rbcL_4_derep_restricted_filt)) %>%  select("sampleid", "Lab_Code", "Plate", "Seq_depth_rbcL", "Ring","Genus" , "Species", "Subspecies", "Species_Label", "Month", "Year", "Locality", "Habitat", "Xdec", "Ydec", "Archipelago_Mainland", "Island_Mainland",  "Area_km2", "Distance_mainland", "Age_ma", "Age", "Sex", , "Weight", "Wing", "X8_P", "Tail", "Nar", "P_Cr", "Height", "Width", "Cran", "Tars")

sample_data(ps_rbcL_4_derep_restricted_filt) <- df

# Replace sampleid with Lab_Code to match sample names across datasets 

sample_names(ps_rbcL_4_derep_restricted_filt) <- paste0("X",sample_data(ps_rbcL_4_derep_restricted_filt)$Lab_Code)

```
Save phyloseq object
```{r exportpartII saveObjects, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}

ps_rbcL_4_derep_restricted_filt
ps_rbcL_4_derep_restricted_filt %>% sample_data %>% names

saveRDS(ps_rbcL_4_derep_restricted_filt, paste0(dataDir, "/ps.clean_rbcL.rds"))

```

## Session Info

```{r sessioninfo, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=45), paged.print=FALSE}
sessioninfo::session_info()%>%
  details::details(summary = 'Current session info', open = FALSE)
```